### 2014-07-07 tex/brief_model_description.tex
Added all files to new repository

\section*{Integrated modelling framework}

The key idea of our approach is to constrain the parameters of a correlative meta-model using predictions made by one or several mechanistic sub-models.
The role of the meta-model is to integrate data at the same ecological scale as predictions. 
As an example, the meta-model could take the form of a simple correlative \ac{SDM}.
The output of sub-models models should provide additional predictions at a comparable ecological scale at which the meta-model operates, although it is not necessary for the sub-models to actually operate at the same scale.
Ideally, each submodel will incorporate different data and a different set of hypotheses that specify underlying ecological mechanisms contributing to the model output.
For sub-models operating at a substantially different scale than the meta-model, scaling functions will be required; however, as we will show in the first example, these functions can be simple while still contributing useful information to the meta-model. 
All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the meta-model (see Box 1, Fig. \ref{fig:diagram}).
We present the complete framework in a hypothetical example in the following section.
A more formal presentation of the framework appears in Box 1.

\begin{figure}
\input{diagram}
\caption{The parameters of a correlative meta-model model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
The meta-model ($\theta_M$) operates at a single ecological scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the meta-model via a scaling function \(g(\theta_S, D_S)\).
The final prediction can be obtained by applying the integrated model to the original explanatory variables (as shown) or, if projection is desired, on some new set of explanatory variables (e.g., future climate).
This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).
}
\label{fig:diagram}
\end{figure}

### 2014-07-07 tex/discussion.tex
Added all files to new repository

\section*{Discussion}

\subsection*{Comparison with other methods}
\subsubsection*{Paragraph about hybrid models}

Our approach is an original application of Bayesian multilevel modelling, and can be considered a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales and using several models simultaneously. 
In particular, this development has certain similarities with Bayesian model averaging, Bayesian calibration of process-based models and hierarchical Bayesian modeling. 
Bayesian model averaging aims to combine several alternative models to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.][]{Wintle2003, Link2006}. 
Bayesian model averaging considers models operating at the same scale and the posterior distribution is usually determined with Gibbs sampling. 
Bayesian calibration of process-based model focuses on uncertainty of the parameter values in the process-based models, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, VanOijen2011}. 
In contrast with the Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of a generally more correlative meta-model.
Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005, Efendiev2005a}, plant physiology \citep{Ogle2008, Ogle2009}, and climate and atmospheric modelling \citep{Zimmerman2005, Mcmillan2010, Kang2012}.

\subsection*{Advantages of Model Integration}
Models are important tools that are being increasingly used by land managers to face challenges associated to decision-making. 
A growing number of models are available, but they can provide diverging answers to very similar questions, due to the differences in their assumptions and methodology. 
This can create confusion and even some mistrust towards models, and as a consequence some managers may be discouraged to incorporate any model at all in their management plans. 
By integrating different types of models---and their outputs---into a common framework, we believe that our approach can contribute to overcome the gap between modelers and practitioners and thus to promote a widespread use of models to support decision-making.

Our approach also allows for a clear, transparent identification of uncertainties and how they are transmitted throughout the modelling process. 
This is one of the main strengths of this approach, and offers several advantages. 
For instance, transparency in uncertainty can be considered as a sort of sensitivity analysis, in which areas with large uncertainty can be detected and new experimental research or additional data collection can be designed (e.g., Example 1, Figs. \ref{fig:ex1_sampling}, \ref{fig:ex1_precip})
The new knowledge resulting from this research can be readily incorporated into the metamodel, allowing for an iterative learning process that will undoubtedly contribute to reduce uncertainty in predictions for a wide range of environmental conditions. 
Moreover, our framework embraces change as a fundamental process and is able to adapt and respond to it. 
The ease of incorporation of new knowledge to the modeling framework (including new theory or the result of management and experimental efforts), will allow for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans. 
In an era of continuous and unprecedented change, adaptive approaches such as the one presented here are often highlighted as a pressing need in order to develop strategies to promote ecosystems that are both feasible and resilient \citep[][Fig. \ref{fig:adaptive_management}]{Seastedt2008, Mori2013}.

%==================
% FIGURE

\begin{figure}[tb]
	\includegraphics{adaptive_management.pdf}
	\caption{Sample workflow for applying the models presented in the first example in an adaptive management context.
	Critical steps include specifying the meta-model, identifying additional sources of information to be used as constraints on the meta-model, then using the integrated prediction for decision-making.
	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the meta-model.}
	\label{fig:adaptive_management}
\end{figure}

% ERUGIF
%==================

\subsection*{Challenges} 
Although our approach is highly flexible and can be applied in a number of situations, there are some challenges to successfully using the framework.
As with any modelling effort, good model specification with strong links to theory are essential \citep{Austin2007}.
Poorly specified models will produce outputs that are uninformative or misleading, and model integration is not a cure for these problems.
In the worst case, constraining a meta-model with a poor sub-model can result in outcomes that are worse in terms of bias and uncertainty than those produced by a naive model.
We expect model selection will play an important role in applications of our framework, and such schemes can be incorporated relatively easily \citep{Madigan1995, Wasserman2000, Tenan2014}.

In a similar fashion, the quality and availability of data imposes a significant constraint on the number and type of models that can be implemented.
The capacity to implement a model is very low if data requirements are prohibitive. 
Adequate coverage of explanatory variables is a significant obstacle, and exploratory analyses can be a significant aid in understanding how data coverage impacts resulting predictions \citep{Mckenney2002}.
Integration can ameliorate these issues to some extent by using supplemental information (and conceptual advances) in additional sub-models where coverage is weak (e.g., Example 1, Figs. \ref{fig:ex1_precip}, \ref{fig:ex1_map}).
For example, \citet{Freckleton2009} estimated that data is too scant to successfully develop highly mechanistic models predicting weed population numbers. 
A strong asset of our proposed method is that it can be used without the full suite of data that would be required to run a fully mechanistic model. 
Since the meta-model is correlative, the rough structure of our approach only needs e.g. presence-absence data to be effective. 
Then any additional mechanistic data that is available will enhance predictions by constraining outputs of the meta-model.

Finally, determining the functions to use to express the likelihood of the sub-models given the meta-model (i.e., Eqs. \ref{eq:ex1_integrated}, \ref{eq:integrated2}) will remain a significant challenge.
We have shown in our examples that even simple scaling functions can provide reasonable constraints on the meta-model. 
However, it is likely that with very large differences in scales (e.g., comparing cellular processes to landscapes), simple functions will be inadequate, requiring more traditional upscaling methods that combine explicit processes with additional data (and, necessarily, additional parameters) before meaningful model integration will be possible.

\subsection*{Applicability of the method}
\defcitealias{TERN2013}{TERN, 2013} %% this is needed to make the organizational citation in the following paragraph work correctly
Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organisations around the world \citep[e.g.][\citetalias{TERN2013}]{Bernier2013}. 
Incorporating information from multiple sources, particularly with respect to uncertainty,  fosters a connection between scientifically-generated knowledge and policy, and is therefore an important tool for adaptive management \citep[][Fig. \ref{fig:adaptive_management}]{Rehme2011}.
Such approaches are needed in formulating management plans for vulnerable species and ecosystems to avoid basing decisions on too-narrow subsets of the available information \citep{Dawson2011}.
The flexibility of our approach may also represent an advantage for management by easing the integration of specific decision making criteria (e.g. desired grain and extent of outputs) into model development. 
Successful use of an integrated modelling approach will always remain dependent on an intimate understanding of the decision-making process by modellers, emphasizing the importance of close collaboration between with practitioners at all stages of model development \citep{Guisan2013}.

The transparency of our approach is also a clear advantage. 
To be useful, models should be transparent analytical tools, not black boxes \citep{Addison2013}. 
A key criteria for model applicability is the explicit and detailed communication of specific model objectives, characteristics, limits, and uncertainties, as well as its ecological foundation \citep{Guisan2013}.
Integrating submodels allows for clear specification of desired model outputs (via the specification of the metamodel) while easily retaining important ecological objectives (via careful specification of submodels).
Insufficiently documented models are difficult to assess in real world situation, and therefore only having limited relevance  for decision support \citep{Guisan2013}.
Model workflow documentation becomes more crucial in the case of integrated modelling approaches, which incorporates information from various scales and resolutions within a number of aggregated sub-models. 
A transparent and well documented workflow describing the process of model integration ensures reproducibility and facilitates its use in an applied context (Fig. \ref{fig:adaptive_management}). 
The issue of uncertainty also limits the use of models in guiding decisions. 
Model outputs are considered too uncertain for decision-making \citep{Addison2013}. 
The transparent estimation of the uncertainty provided by this modelling approach may be a significant advantage.

\subsection*{Accessibility}
Results of models predicting species ranges are increasingly used by resource managers, conservation biologists, or forest ecologists for formulating or adjusting recommendations. 
The utility of such models depends on their ability to help evaluate events beyond the bounds of the available data, including in future situations constrained by climate change. 
In this context, practionners judge model usefulness on two criteria : Does the model provide reliable predictions at the needed time and space scales, and can the model be implemented given the available data and technical expertise? 
Since most decision makers work at local space scales and follow specific time frames, modelers face the obvious challenge of providing detailed outputs while preserving implementation capacity.
In terms of model outputs, our framework is transparent in terms of uncertainty by virtue of providing easily interpretable probability distributions for model parameters and predictions.
However, developing the models requires careful model specification, knowledge of applied Bayesian methods, and, in some cases, extensive programming.
In many cases, off-the-shelf software \citep[e.g.,][]{R, RJAGS} can adequately express the model likelihoods with minimal programming, but more complicated models will require the development of custom software.
Thus, this framework will require significant investment in developing customized code for the samplers in order to actually estimate parameters.
No doubt, the implementation of our proposed method is currently too technically challenging for most practionners. 
However, the same was true of all statistical techniques or modeling approaches when they were initially developed. 
In addition, some of the computation costs that were associated with many techniques have now vanished, and even the conceptual challenges associated Bayesian statistics are being reduced as they gain popularity in the scientific literature. 
We therefore argue that our proposed approach as a strong potential for direct use in our real world where climate is quickly changing and conservation practices must be adjusted accordingly.

\subsection*{Future directions}
\begin{itemize}
	\item incorporation of spatially explicit models into the framework? \citep[e.g.,][]{Fortin2012}
	\item more sophisticated meta-models (e.g., state-transition models)
	\item others?
\end{itemize}

### 2014-07-07 tex/examples.tex
Added all files to new repository

\section*{Examples}
Our approach to model integration is best explained with concrete examples. 
In this section we first present how experimental results could be introduced as prior knowledge of the fundamental niche in a species distribution model (\ac{SDM}), and in the following we apply our approach to integrating phenological information with occurrence data for a widespread North American tree.

%---------------------------------------------
\subsection*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}

In this first hypothetical example, we build a meta-model relating the distribution (i.e. occurrence probability) of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
The meta-model here captures the realized distribution of a species, thereby encapsulating in a single correlative model all physiological constraints and ecological processes constraining a species distribution. 
However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the envelope of where a species could potentially be found. 
Prior information of the physiological constraints affecting species distribution is sometimes available, but usually too incomplete to perform a reasonable comparison with the realized distribution.
For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature. 
We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty. 
We focus here on the specification of the modelling framework; complete procedural details and code for generating the data sets and executing the model are provided as supplemental information.

We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation. 
We will consider \(X_M\) (see Box 1, Fig. \ref{fig:diagram}) to be a vector of $n$ observations that takes the value of one to indicate presence and zero for absence:
\begin{equation}
X_M = \{x_{M,1}, x_{M,2}, \ldots, x_{M,n}\}
\end{equation}
We further assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
The model is evaluated by relating these observations to environmental data \((D_M)\), which for the sake of the example consists of mean annual temperature $T_M$ and total precipitation $P_M$. 
For simplicity, we use a simple logistic model for the meta-model \((\theta_M)\) with a second order effect of temperature and precipitation. 
This naive model (i.e., the meta-model with no constraints from sub-models) predicts  occurrence probability (\(\psi_N\)) as a function of the environment:
%-----
\begin{equation}
\begin{aligned}
	\psi_N &= f\left(\theta_M, D_M \right) \\
	&= p \left (X_M = 1 \mid \theta_M, T_M, P_M \right) \\
	&=\text{logit}^{-1}\left( \mathbf{\Theta_M} \mathbf{D_M} \right)
\end{aligned}
\end{equation}
%-----
where \(\mathbf{\Theta_M}\) is the parameter vector of the model \(\theta_M\), \(\mathbf{D_M} \) is the covariate matrix (i.e., \(T_M, P_M\), with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
We can fit this model in a Bayesian framework to allow for easy integration of models (as we will show later).
In this context, the goal of modelling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, we recommend][for a concise introduction]{Link2010}:
%-----
\begin{equation}
\label{eq:ex1_bayes}
	p\left (\theta_M \mid X_M,T_M,P_M \right ) \propto 
	p \left(X_M \mid \theta_M, T_M, P_M \right)
	p \left(\theta_M \right)
\end{equation}
%-----
where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.

%==================
% FIGURE

\begin{figure}[tb]
	\includegraphics{ex1_sampling.pdf}
	\caption{Two simulated data sets used to illustrate the model integration framework.
	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation.
	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.}
	\label{fig:ex1_sampling}
\end{figure}

% ERUGIF
%==================

Thus far, we have considered only a single source of information to fit this model, and therefore the prior distribution \(p\left(\theta_M \right)\) from Eq. \ref{eq:ex1_bayes} is uninformative.
As a secondary source of information, we will consider an experiment relating the population growth rate of the plant to manipulations to the precipitation regime, with results (but no raw data) available from the literature (Fig. \ref{fig:ex1_sampling}b). 
Furthermore, no information is available regarding the temperature regime for the experiment.
Transplant experiments that evaluate performance beyond the range of a species are common, and represent a plausible scenario for model integration \citep{Hargreaves2014}.
According to niche theory \citep{Holt2009}, the fundamental niche corresponds to the set of environmental conditions where the per capita intrinsic growth rate $r$ is positive.
This concept gives us a reasonable model to fit the scaling function $g$ for our sub-mode (Fig. \ref{fig:diagram}).
If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, the for observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
\begin{equation}
	\psi_{S,i} = \int_0^\infty N \left(X_{S,i}, \sigma_{S,i} \right)
\end{equation}
where \(N\) is the Normal density function.
We can then estimate the posterior distribution for the sub-model by fitting the relationship between \(\psi_S\) and precipitation \( \left( P_S \right) \) using Bayesian beta regression \citep{Ferrari2004}:
%-----
\begin{equation}
\label{eq:ex1_thetas}
	p\left (\theta_S \mid \psi_S,P_S \right ) \propto 
	p \left(\psi_S \mid \theta_S, P_S \right)
	p \left(\theta_S \right) \\
\end{equation}
%-----

Although the two data sets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the meta-model second (i.e., the probability of presence at a given precipitation). 
This scaling is quite simplistic, and would never be used to predict a species' range using this mechanistic submodel in isolation.
Despite the strong theoretical foundation (i.e., niche theory) linking the observations to presence-absence at large scales, the upscaled sub-model (i.e., the transformation of \(X_S\) and \(\sigma_S\) to \(\psi_S\)) is built on the strong hypothesis that the species will be present when $r>0$, neglecting other population dynamics constraints such as Allee effects and metapopulation dynamics. 
The fundamental niche is also incomplete, as we lack environmental variables such as temperature, as well as all other ecological processes responsible for the realized distribution.
As such, it would be unwise to expect predictions from this model alone to resemble the actual distribution of the species; as a mechanistic model, it is simply too incomplete to predict distribution.
As we will show, however, the information from this sub-model, when applied as a constraint on the meta-model, can result in improved predictions that incorporate the information within each model.

We accomplish model integration by treating the posterior of \(\theta_S\) as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
%-----
\begin{equation}
	\label{eq:ex1_integrated}
	\overbrace{p(\theta_M \mid X_M, T_M, P_M, \theta_S, \psi_S, P_S)}^\text{integrated posterior}
	\propto
	\overbrace{p\left (\psi_S \mid \theta_S,P_S \right )}^{\substack{\text{new information} \\ \text{from sub-model}}}
	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) P \left(\theta_M \right)}^{\substack{\text{naive meta-model} \\ \text{posterior}}}
	\overbrace{p \left(\theta_S \right)}^{\substack{\text{prior for} \\ \text{sub-model}}}	
\end{equation}
%-----
As before, the meta-model \(\theta_M\) can be used to predict the species distribution \((\psi_I)\).
However, these predictions will not reflect only the presence-absence samples in \(X_M\), but will also reflect the information from \(\theta_S\), including all of the data sources used to produce this sub-model.
In other words, with this evaluation, the set of parameters $\theta_M$ is most likely when it agrees with both the original data and predictions from the sub-model. 
Note that in this particular example, because the sub-model is only contingent on precipitation, there is no integration on temperature. 
Finally, we note the presence of marginal distributions for both models (i.e., \(p(\theta_M)\) and \(p(\theta_S)\)).
These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
For purposes of this example, we used these terms to reduce the precision of the second model to reflect the the uncertainty of the modelling process (due to the lack of additional mechanisms in the second model, as mentioned above).

We performed this procedure with our hypothetical data sets.
When comparing the three models (naive meta-model, mechanistic sub-model, and integrated meta-model), we observed extreme uncertainty in the first model when projecting beyond the range of the original data.
Unsurprisingly given the quality of the data, the sub-model was highly precise, providing a fairly strong constraint when producing the integrated model.
The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
At the scale of the meta-model, considering both temperature and precipitation, we observed similar results, with reduced uncertainty in the predictions over the domain not covered by the presence-absence data (Fig. \ref{fig:ex1_map}).

%==================
% FIGURE

\begin{figure}[tb]
	\includegraphics{ex1_precip.pdf}
	\caption{Results of integration showing the probability of presence (\(\psi\)) when considering only precipitation.
	(a) Naive model, using only presence-absence data. Uncertainty increases dramatically when attempting to project beyond the scope of the source data (represented by the horizontal line above the bottom axis).
	(b) Mechanistic model, using observations of an experiment to infer probability of presence. Predictions are highly precise due to high quality source data, but are likely to be overly optimistic given the simplicity in assuming only precipitation determines probability of presence (see text).
	(c) Integrated model, showing predictions that are intermediate between the two submodels and uncertainty that is reduced compared to (a).
	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.}
	\label{fig:ex1_precip}
\end{figure}
% ERUGIF
%==================

%==================
% FIGURE

\begin{figure}[tb]
	\includegraphics[width=5.25in]{ex1_map.pdf}
	\caption{Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.}
	\label{fig:ex1_map}
\end{figure}
%
% ERUGIF
%==================

Once the technical challenges of evaluating the parameters are overcome, the integration has several advantages over the both the naive parameterization of the meta-model and the mechanistic sub-model. 
In the situation where both models strongly agree on the relationship between water availability and occurrence probability, then we should expect significantly reduced uncertainty in that domain (Fig. \ref{fig:ex1_precip}, \ref{fig:ex1_map}).
Because the meta-model includes both temperature and precipitation, any reduction in the uncertainly with respect to precipitation might also influence the evaluation of the relationship with temperature (reducing bias in the case where there is a correlation between T and P in the data). 
However, uncertainty could increase in the case of model disagreement.
The relationship between precipitation in the naive meta-model could be caused by a spurious relationship (e.g. spatial autocorrelation caused by historical contingencies), leading to a false confidence in the species distribution over precipitation gradients. 
The information from the sub-model will thus bring back the right amount of uncertainty in the relationship.

\subsection*{Example 2: Constraining an SDM using phenological information}
For the second example, we consider the problem of predicting the future distribution of a species following climate change.
There is considerable interest in comparing correlative and mechanistic projections with respect to climate change \citep{Morin2009}, and correctly characterizing uncertainty is a critical aspect of this problem \citep{Cheaib2012}.
Despite being a relatively common application of \ac{SDM}s \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios is problematic for a number of reasons previously discussed \citep{Araujo2006, Austin2011}.
We used our method to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
We use data from \citet{Morin2009} for sugar maple (\emph{Acer saccharum}), an economically and ecologically important species occurring in eastern North America.
Here we describe briefly the dataset, methods, and the results of the analysis.
Complete details, including all data and scripts to reproduce the analysis, are included as supplemental information.

\defcitealias{IPCC2001}{IPCC, 2001} %% this is needed to make the organizational citation in the following paragraph work correctly
For the meta-model, we used a dataset of 1013 sugar maple presence records at 0.5 degree resolution, with 498 corresponding pseudo-absence records sampled from North America.
We used a binomial GLM to relate the suitability for sugar maple to 5 climate variables, scaled to the same resolution (see SI for a description of the climate variables used).
For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} with the A2 climate scenario \citepalias{IPCC2001}, and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
Phenofit produces predictions at continental scales; thus the scaling function \(g\) was incorporated within Phenofit itself and we performed model integration directly on the outputs.
To perform the integration, we used \ac{MCMC} to condition the parameters of the meta-model simultaneously on two likelihood functions.
The first is the naive SDM, where the likelihood of a ``success'' in the presence/pseudo-absence dataset was a Bernoulli density with the (logit-transformed) probability expressed as a linear function of the climate predictors.
The sub-model likelihood was the Beta density of the observed Phenofit prediction for the present climate given the predicted suitability under the naive model. 
Thus, the posterior probability of the meta-model was:
\begin{equation}
\label{eq:integrated2}
	p( \theta_M \mid X_M, D_M, X_S, \theta_S )
	\propto 
	p( X_S \mid \theta_M, \theta_S )
	p( Y_M \mid \theta_M, D_M ) 
	p( \theta_M )
	p( \theta_S )
\end{equation}
where \(\theta_M\) is the meta-model, 
\(X_M\) is the vector of presences and pseudo-absences, 
\(D_M\) is the matrix of climate predictors,
\(X_S\) is the prediction from Phenofit (in this case a single deterministic value, because we did not incorporate errors from the Phenofit modeling),
and \(\theta_S\) is the model relating the Phenofit predictions to the meta-model.
Note that here, the integration is performed on the \emph{predictions} of the sub-model directly, rather than on compatible parameters as in Example 1. Thus, the meta-model must appear in the sub-model likelihood.

The naive model, despite being relatively simple (xx climate variables with xx total parameters), was highly fit to the data and consequently showed very high suitabilitiy scores near the core of the contemporary range with very little uncertainty (Fig \ref{fig:ex2}C, E).
Similar patters were repeated for analogous climates under the future scenario (Fig \ref{fig:ex2}D, F).
However, standard errors were extreme for the range margins where suitability scores were intermediate.
In comparison, Phenofit suitability scores were more moderate, and Phenofit predicted less northward movement of sugar maple under the future scenario \citep[Fig. \ref{fig:ex2}A, B,][]{Morin2009}.
The integrated model demonstrated generally lower suitability scores throughout both the present and projected ranges, and thus represents lower confidence in the presence of sugar maple at any given location in its future projected range (Fig \ref{fig:ex2}G, H).
Standard errors at range margins were significantly lower (as these are regions where both the naive model and phenofit were in agreement about intermediate suitability scores), whereas standard errors were slightly higher at the core range (where Phenofit predicted lower suitability than the naive model) (Fig \ref{fig:ex2}I, J).

%==================
% FIGURE

\begin{figure}[tb]
	\includegraphics[width=6in]{ex2.pdf}
	\caption{Results of model integration for sugar maple, \emph{Acer saccharum}.
	The mechanistic submodel Phenofit predicts present (A) and future (B) suitability using phenological information, but does not provide estimates of uncertainty.
	Predictions from a naive metal-model for the present (C) were quite similar to the sub-model, but future predictions (D) were somewhat different, and uncertainty was extreme at the predicted range margins for both present (E) and future (F).
	Applying the integration results in reduced suitabilities overall, possibly due to overfitting regions where the models disagree, but patterns in relative suitability largely reflected information from both models for both present (G) and future (H) climate.
	Integrated model uncertainty for both present (I) and future (J) climates was greatly reduced near range margins but slightly increased in areas where uncertainty was low in the naive model.
	}
	\label{fig:ex2}
\end{figure}
%
% ERUGIF
%==================

For modeling the contemporary range of sugar maple, both Phenofit and the naive model perform very well relative to the integrated model, espeically when capturing the range margin (Fig \ref{fig:ex2}A, C, G).
Conventional SDMs are quite good at producing empirical models of species distributions under present climatic regimes, and the naive model here could be further improved with the addition of more parameters or the inclusion of ensemble forecasting (See SI for model-fitting details).
However, Phenofit and the naive model differ substantially in their predictions of the future distribution of the species.
The integrated model presents a reasonable compromise between the two predictions. 
The core range of the species under the integrated model generally reflects the regions of highest overlap between Phenofit and the naive model.
Regions surrounding the core range generally show low suitabilities, but higher than those under either submodel, demonstrating reduced certainty about where the species will be absent in the future.

### 2014-07-07 tex/introduction.tex
Added all files to new repository

\section*{Introduction}

Models of species range limits have wide applications and play a large role in conservation biology, where they are used as decision-making tools in biodiversity management \citep{Rosenzweig2008, Thuiller2008, Rodenhouse2009}.
Due to large temporal and spatial scales and the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating range limits \citep{Wu1995, Levin1998}. 
Thus, models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution and hybrid zone dynamics \citep{Engler2013}, and the impacts of climate change  \citep{Rosenzweig2008, Thuiller2008, Milad2011, Blois2013, Loyola2013}. 
However, despite having recognized potential to support decision making, the contributions of models to conservation and applied ecology remain limited \citep{Dawson2011, Guisan2013}.

An important constraint on the modelling process is the availability of data and appropriate ecological theory to  link data to modelling objectives.
Poor data or poorly conceived models, will result in predictions with unacceptably high uncertainty, or, more dangerously, precise but biased predictions.
As a result of data accumulation from various sources and scales and conceptual advances in ecology, modelling techniques have proliferated to take advantage of what is available.
Model diversification also arises due to the diverse processes generating species ranges \cite{Soberon2007}.
In the context of forecasting, which is the focus hereafter, mechanistic and correlative approaches represent quite different modes of inference, but both have strong limitations. 
Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, may perform poorly when applied to species ranges.
For instance, biotic interactions (including trophic interactions) are usually not modelled mechanistically at the scale of species distributions because they are not well-known, or have not been recorded, despite being considered a key determinant of range limits \citep{Soberon2007, Roux2012, Guo2013, Pigot2013}. 
On the other hand, coarse-scale correlative models statically relating species distributions to associated distributions of other variables at the same scale have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, which implies that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}. 
Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
Furthermore, correlative models often do not explicitly incorporate theory into their formulations, and when applied at coarse-scales, they may miss important fine-scale processes, making a multi-scale approach desirable when accurate forecasting is required \citep{Austin2007, Austin2011, Soranno2014}.

\subsection*{Toward an integrated approach for modelling ranges}
A key problem with the diversity of approaches is that multiple models often produce differing predictions for the same organism, with no simple way to reconcile these predictions.
We present an application of hierarchical Bayesian methods for integrating the predictions of multiple models; our approach is an alternative to other approaches to drawing inference from multiple models (summarized below) that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
For instance, correlative models are often criticized because they assume that the selected explanatory variables limit (directly of indirectly) the species range, and, as mentioned above, that the relationship between the selected variables and the species occurrences will not change in the prediction context \citep{Araujo2006, Berteaux2006, Braunisch2013}.
However, additional information (in the form of both data and theory) is often available on the species of interest (e.g., growth rates in different climatic conditions, phenology, interactions with other species, sensitivity to disturbances) \citep{Holt2009, Thuiller2013}. 
The challenge is to account for all this available knowledge while producing a single prediction.
Integrating multiple models to provide a single prediction would allow more knowledge to be used, potentially mitigating the limitations inherent in each modelling approach and contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.

One approach is to integrate multiple models in a single modelling framework \citep{Buckley2010}. 
Some researchers have explicitely reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Catterall2012, Strigul2012, Stewart-Koster2013, Soranno2014}. 
Recent literature has also focused on the development of hybrid models, which allow for flexible combinations between mechanistic and phenomenological sub-models \citep{Gallien2010, Franklin2010, Thuiller2013}. 
For instance, correlative models are used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while mechanistic approaches can include space and time dynamics and therefore account for dispersal processes \citep{Kearney2008}. 
This complementarity has been used to merge the two types of models into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Boulangeat2014}. 
However, the link between different sub-models is based on additional assumptions that are not easy to test \citep{Gallien2010}. 
Moreover, this approach is not suitable for merging models that are based on the same processes but with different underlying hypotheses.
A second approach is to directly combine predictions.
If models operate at the same scales have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting][]{Araujo2007, Diniz-Filho2009}. 
If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Serra-Diaz2013}.
However, this approach is limited; for example, it is not possible to evaluate the effects of convergent predictions on total uncertainty of outcomes.
One of the greatest challenges inherent in both approaches above is estimating uncertainty, which is of utmost importance in a prediction context.
A complete understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying parameters that make large contributions to the total uncertainty in the model predictions \citep{McMahon2011}. 
In models that encompass multiple scales or levels of organization, it is particularly challenging to evaluate all sources of uncertainty, especially those related to model specification (\citealt{Calder2003} but see \citealt{Conlisk2013}).
Clearly, an approach that integrates predictions and uncertainty from multiple model types is needed \citep{Beck2012, Thuiller2013}. 
Such approach would incorporate multiple modes of inference and integrate data from multiple sources and scales \citep{Levin1992, Peters2004, Thuiller2013}.

Here, we present a framework for modelling range dynamics that aims to account for all the information available on a given question, from a variety of sources and scales.
Contrary to hybrid and hierarchical models, the aim is not to link differents sub-models into a single model, but to condition the predictions of a meta-model at the target scale (e.g., an entire species' range) with information from sub-models at a variety of scales, allowing as much flexibility as possible regarding the type of information included. 
We use the power and generality of a hierarchical Bayesian framework, which allows us to include multiple data sources and modes of inference \citep{Clark2005, VanOijen2005, Clark2006, Hobbs2011, Hartig2012}. 
Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to intertpret and reflects uncertainty at all levels of organization \citep{Clark2005, Cressie2009, Hobbs2011}. 
We use simulated data to present a complete example of the application of the framework to multiple information sources that are relevant at very different scales.
Finally, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}) to climate with the predictions of a phenological model, Phenofit, with the goal of improving uncertainty estimation when projecting model predictions to future climate.

### 2014-07-07 tex/manuscript.tex
Added all files to new repository

\documentclass[11pt]{article}

%% ----------------------------------
%
%     Packages to be used in the final manuscript
%
%% ----------------------------------

\usepackage[margin=1in]{geometry}

% use proper unicode fonts
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{amsmath} % for better display of equations

\usepackage{natbib}

\usepackage{titling} % controls the way the title information is displayed
\pretitle{\begin{flushleft}\Large}
\posttitle{\end{flushleft}}
\predate{}
\postdate{}
\preauthor{\begin{flushleft}}
\postauthor{\end{flushleft}}
\setlength{\droptitle}{-3em}

\usepackage{authblk} % adds some nice options for displaying the author list

\usepackage{acronym} % set up some acronyms
\acrodef{SDM}{species distribution model}
\acrodef{MCMC}{Markov Chain Monte Carlo}

%% graphics packages
\usepackage{graphicx}
\usepackage{pdflscape}   % for figures in landscape orientation

% Tikz libraries for building the diagram
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc, shapes}
\usetikzlibrary{shapes.geometric,shapes.arrows,decorations.pathmorphing}
\usetikzlibrary{matrix,chains,scopes,positioning,arrows,fit}

%% ----------------------------------
%
%     Packages to be used for drafts and editing
%     Remove from the final manuscript
%
%% ----------------------------------
%\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
%
%\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
%
%% available colors:
%% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
%
%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
%\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}}
%\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
%\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
%\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
%\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
%\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
%\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
%\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
%\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
%\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
%\newcommand{\wt}[1]{\todo[color=violet!40]{\textsuperscript{Wilfried T}#1}}
%\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}

% uncomment to disable individual authors
% \renewcommand{\mt}[1]{\relax}%
% \renewcommand{\ib}[1]{\relax}%
% \renewcommand{\aa}[1]{\relax}%
% \renewcommand{\ia}[1]{\relax}%
% \renewcommand{\db}[1]{\relax}%
% \renewcommand{\rd}[1]{\relax}%
% \renewcommand{\mjf}[1]{\relax}%
% \renewcommand{\tf}[1]{\relax}%
% \renewcommand{\dm}[1]{\relax}%
% \renewcommand{\ks}[1]{\relax}%
% \renewcommand{\ns}[1]{\relax}%
% \renewcommand{\wt}[1]{\relax}%
% \renewcommand{\dg}[1]{\relax}%

%% ----------------------------------
%
%     Title and authorship information
%
%% ----------------------------------

\title{Cross-scale integration of data and knowledge for predicting species ranges}
\date{}
\author[1,2]{Matthew V. Talluto}
\author[1,2]{Isabelle Boulangeat}
\author[3]{Aitor Ameztegui}
\author[4]{Isabelle Aubin}
\author[1,2,5]{Dominique Berteaux}
\author[1,2]{Alyssa Butler}
\author[10]{Dominique Cyr} %??
\author[10]{Fred Doyon} %??
\author[6]{C. Ronnie Drever}
\author[7]{Marie-Josée Fortin}
\author[1]{Tony Franceschini}
\author[8]{Jean Liénard}
\author[4]{Dan McKenney}
\author[2,3]{Kevin A. Solarik}
\author[8]{Nick Strigul}
\author[9]{Wilfried Thuiller}
\author[1,2]{Dominique Gravel}
\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[2]{Quebec Centre for Biodiversity Science, Montreal, Quebec, Canada}
\affil[3]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
\affil[4]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
\affil[5]{Centre for Northern Studies, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[6]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
\affil[7]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
\affil[8]{Department of Mathematics, Washington State University Vancouver, Washington, USA}
\affil[9]{Laboratorie d'Ecologie Alpine, Unité Mixte de Recherche, Université de Grenoble, Grenoble, France}
\affil[10]{Affiliations to be filled in}

%% ----------------------------------
%
%     END PREAMBLE
%
%% ----------------------------------

\begin{document}

%% ----------------------------------
%
%     TITLE PAGE
%
%% ----------------------------------

%\listoftodos
\maketitle

\begin{flushleft}
\textbf{Keywords:} Uncertainty, scaling, range dynamics, trees, disturbances, patterns and processes, predict or understand, climate change, biotic interactions, spatial ecology, decision making, metamodelling, species distribution modeling
\end{flushleft}

\begin{abstract}
There is great interest in modeling species ranges, and such models are useful in guiding conservation and management decisions.
There presently exists a great diversity of modeling approaches, and individual models often use only a small subset of the total available information about a species.
For example, a broad-scale correlative model might use climate variables to predict presence or absence, but ignore what is known about smaller-scale processes such as the effect of local climate on growth, fecundity, and dispersal rates.
A central problem with the diversity of approaches is that multiple models produce differing predictions for the same organism, with no simple way to reconcile these predictions.
We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
The resulting meta-model produces probabilistic estimates of species presence with uncertainty that reflects error from all sub-models and data sources. 
These predictions reflect all of the information used as input for the original sub-models. 
We illustrate the approach through two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of some of the original data sources.
Finally, we discuss the application of our method and its accessibility to conservation biologists and land managers.
Although the method requires extensive statistical programming experience, we anticipate the results will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.
\end{abstract}

%% ----------------------------------
%
%     REST OF DOCUMENT
%
%% ----------------------------------

\input{introduction}

%---------------------------------------------
%---------------------------------------------

\input{brief_model_description}

%---------------------------------------------
%---------------------------------------------

\input{examples}

%---------------------------------------------
%---------------------------------------------

\input{discussion}

%---------------------------------------------
%---------------------------------------------

\input{model_description}

%---------------------------------------------
%---------------------------------------------

%% ----------------------------------
%
%     REFERENCES
%
%% ----------------------------------

\renewcommand\refname{Literature Cited}
\bibliography{model_integration}{}
\bibliographystyle{model_integration}

%% ----------------------------------
%
%     FIGURES
%
%% ----------------------------------

\end{document}

### 2014-07-07 tex/model_description.tex
Added all files to new repository

\hrule
\vspace{2em}

\section*{Box 1: Model description}

We begin with a simple correlative model (i.e. calibrated with data provided at the same ecological scale as predictions), characterized in a Bayesian framework to allow further integration. 
We refer to this model (with associated parameters) as the meta-model, \(\theta_M\), and it is parameterized with information on the species' presence, \(X_M\), along with associated covariates (e.g., climate, presence/absence of interacting species, etc), collectively referred to as \(D_M\). 
The naive (i.e., not integrated with information from sub-models) probability that a species is present, \(\psi_N\), is a deterministic function of the meta-model and its covariates::
%-----
\begin{equation}
\label{eq:sdm1}
	\psi_N = f(\theta_M, D_M)
\end{equation}
%-----
The goal of model fitting is to estimate the posterior probability distribution for the parameters of \(\theta_M\), given the observations (\(X_M\)) and covariates (\(D_M\)):
%-----
\begin{equation}
\label{eq:sdm2}
	p( \theta_M \mid X_M, D_M)
\end{equation}
%-----

Using Bayes' theorem, we find the posterior probability of \(\theta_M\) as follows:
%-----
\begin{equation}
\label{eq:sdm3}
	p( \theta_M \mid X_M, D_M ) = \frac{ p( X_M \mid \theta_M, D_M ) p( \theta_M ) } { p(X_M, D_M) }
\end{equation}
%-----
where \(p( X_M \mid \theta_M, D_M )\) is referred to as the likelihood of the observations, \(p(\theta_M )\) is the prior probability of the model, and \(p(X_M, D_M)\) is the normalization constant.
The normalization constant involves computing integrals that are often impossible to solve analytically. 
In practice, simulation techniques such as \ac{MCMC} can be used to sample directly from the posterior distribution, making such computations unnecessary.
Thus, we use the proportional form of Bayes' Theorem:
%-----
\begin{equation}
\label{eq:sdm4}
	p( \theta_M \mid X_M, D_M ) \propto p( X_M \mid \theta_M, D_M ) p( \theta_M ) 
\end{equation}
%-----

The role of the meta-model model is to integrate data at the same ecological scale of predictions.
Additional sub-models require outputs that should be comparable to this given ecological scale (e.g. constraining presence or absence on the landscape).
Formally, we will add an additional model, \(\theta_S\), that is based on a different set of hypotheses and that makes predictions \((\psi_S)\) based on an additional dataset \((X_S, D_S)\):
%-----
\begin{equation}
\label{eq:model2-1}
	\psi_S = g(\theta_S, X_S, D_S)
\end{equation}
%-----

For integration to be successful, it must be possible to compute the likelihood of this prediction given the meta-model, \(p \left( \psi_S \mid \theta_M ) = \theta_S \mid \theta_M, X_S, D_S \right)\), thus, we refer to $g$ as the \emph{scaling function}. 
In many cases, computing this likelihood will be challenging.
The simplest case is when the parameters of \(\theta_S\) are directly comparable to those in \(\theta_M\) (see Example 1), or when the scaling is performed directly within the sub-model (see Example 2), but other solutions are possible.
This new likelihood, which integrates the information from the second model into the first, can be treated as new ``data'' to evaluate the parameters of \(\theta_M\).
We simply use the posterior distribution of $\theta_M$ from the previous step (Eq. \ref{eq:sdm4}) as the new prior probability of $\theta_M$, and evaluate the likelihood of $\theta_M$ in light of the new data and the prior knowledge, as before:
%-----
\begin{equation}
\label{eq:integrated2}
	\overbrace{p( \theta_M \mid X_M, D_M, \theta_S, X_S, D_S )}^\text{integrated posterior}
	\propto 
	\overbrace{p( \theta_S \mid \theta_M, X_S, D_S )}^\text{likelihood}
	\overbrace{p( X_M \mid D_M, \theta_M ) P( \theta_M )}^{\text{prior for } \theta_M}
	\overbrace{p(\theta_S)}^{\text{prior for } \theta_S}
\end{equation}
%-----

This procedure may be repeated for an arbitrary number of models (limited only by available data and computational power). 
The models may be implemented simultaneously, when multiple data sets are available and can be evaluated under different underlying models, or they may be run sequentially, updating the posterior distribution of $\theta_M$ as new information becomes available. 
Note that $\theta_S$ itself need not be implemented in a Bayesian framework and the output need not be identical in form to the output of \(\theta_M\); it is enough to produce an output that can be evaluated as a likelihood under the first model and to have some assessment of the confidence in the parameter estimates of $\theta_S$ to use as a prior.
\vspace{2em}
\hrule

### 2014-07-18 tex/brief_model_description.tex
Added comments from coauthors

@@
 \section*{Integrated modelling framework}
+\aa{Integrated modelling framework and examples: I would just merge these two sections into: 'Examples of the integrated modelling framework'}
 
 The key idea of our approach is to constrain the parameters of a correlative meta-model using predictions made by one or several mechanistic sub-models.
+\aa{Does the meta-model need to be correlative? Because later on you say that it 'could take the form of a correlative model' as implying that it could take other forms.}
 The role of the meta-model is to integrate data at the same ecological scale as predictions. 
 As an example, the meta-model could take the form of a simple correlative \ac{SDM}.
-The output of sub-models models should provide additional predictions at a comparable ecological scale at which the meta-model operates, although it is not necessary for the sub-models to actually operate at the same scale.
-Ideally, each submodel will incorporate different data and a different set of hypotheses that specify underlying ecological mechanisms contributing to the model output.
+Ideally, each submodel incorporates different data and hypotheses that specify underlying ecological mechanisms contributing to the model output.
+\tf{Check in examples if hypothesis are given}
+\mt{they are there, though perhaps not so explicit}
+The output of sub-models should provide additional predictions at a comparable ecological scale at which the meta-model operates, although it is not necessary for the sub-models to actually operate at the same scale.
 For sub-models operating at a substantially different scale than the meta-model, scaling functions will be required; however, as we will show in the first example, these functions can be simple while still contributing useful information to the meta-model.
 All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the meta-model (see Box 1, Fig. \ref{fig:diagram}).
-We present the complete framework in a hypothetical example in the following section.
+We present the complete framework in a hypothetical example where experimental results are  introduced as prior knowledge of the fundamental niche in a species distribution model (SDM).
 A more formal presentation of the framework appears in Box 1.
+We follow with a second example where we apply our approach to integrate phenological information with occurrence data for a widespread North American tree.
 

### 2014-07-18 tex/discussion.tex
Added comments from coauthors

@@
 \section*{Discussion}
+\wt{Honestly, I do not really think you need this last part (future work) given space limitation (watch out the total number of words and references).
+Instead I would rather develop some of the already written part that remains a bit vague and imprecise to me.
+I added few things to see what I mean.
+}
+\aa{Discussion: the section ‘Advantages of model integration’, which I wrote, and ‘Applicability of the method’ have several points in common.
+Maybe they could be merged into a single section: ‘Applicability and advantages of model integration’.
+Maybe, it should be discussed more extensively how to face the situation in which predictions of correlative model and sub-models are so different that the integrated model performs poorer than the two (I mean, does that mean that the approach is not valid in those cases?)
+I liked a lot the idea of adding Figure 6.
+The link between this approach and adaptive management seems to me one of its biggest advantages.
+The box in the right could maybe just say: 'monitoring of management actions’?
+}
 
 \subsection*{Comparison with other methods}
-\subsubsection*{Paragraph about hybrid models}
+\ab{The methods provided here expand upon the motivation of hybrid models to develop a more robust and unifying approach for ecological models in a predictive context while striving to overcome some of the limitations characteristic of other integrated approaches.
+Attempting to incorporate the underlying ecological processes so that SDMs are not oversimplified by combining different model parameters or scales can pose a challenge to these hybrid approaches. 
+In particular, it can often be difficult to identify parameters that can be used to connect different modelling frameworks and what parameters can produce purposeful responses \citep{Thuiller2013}. 
+The methods presented in this study aim to overcome some of these limitations through the use of Bayesian statistics. 
+Bayesian methods allow for the inclusion of prior information and produce results as an entire distribution instead of a single point aiding in the reduction of error propagation \citep{Kery2010}. 
+The estimation of a response as an entire distribution helps to decrease over parameterization that is a concern for hybrid models and thereby also provides a more comprehensible response for reporting results.
+Bayesian frameworks provide a natural framework for the incorporation of prior information. 
+The inability or difficulty to include information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Thuiller2008, Smolik2010}. 
+Even when a link can be formed to include this information, it often is oversimplified to simplify computing \citep{Gallien2010}. 
+Bayesian methods therefore provide a reasonable alternative to allow for the easy unification of previous experimental studies or other prior information.
+Finally, Bayesian methods inherently allow for feedbacks or interactions between sub models, which may portray a more realistic response to many environmental dynamics where factors may simultaneously influence the change of one another.}
+\mt{Maybe change Kery citation to Link \& Barker}
 
-Our approach is an original application of Bayesian multilevel modelling, and can be considered a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales and using several models simultaneously.
-In particular, this development has certain similarities with Bayesian model averaging, Bayesian calibration of process-based models and hierarchical Bayesian modeling. 
+Our approach is an original application of Bayesian multilevel modelling, and can be considered a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales while using several models simultaneously.
+In particular, this development has certain similarities with Bayesian model averaging, Bayesian calibration of process-based models and hierarchical Bayesian modelling. 
 Bayesian model averaging aims to combine several alternative models to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.][]{Wintle2003, Link2006}.
@@
 Bayesian calibration of process-based model focuses on uncertainty of the parameter values in the process-based models, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, VanOijen2011}.
-In contrast with the Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of a generally more correlative meta-model.
+\wt{Hartig, F., Dyke, J., Hickler, T., Higgins, S.I., O'Hara, R.B., Scheiter, S.
+\& Huth, A.
+(2012) Connecting dynamic vegetation models to data - an inverse perspective.
+Journal of Biogeography, 39, 2240-2252.}
+In contrast with Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of a generally more correlative meta-model.
 Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005, Efendiev2005a}, plant physiology \citep{Ogle2008, Ogle2009}, and climate and atmospheric modelling \citep{Zimmerman2005, Mcmillan2010, Kang2012}.
+\wt{Why not being more precise and exploratory here.
+You may actually run a model averaging around your approach, may not you?
+Indeed, if we think to Ex 1 you could have several native meta-model posterior distributions, right?
+Or even a direct model averaging from them.
+As said, you could actually run something like biomod2 to get the distribution of values for each pixel given the models/algorithms and data and use them directly as a posterior into the integration.
+.
+The integration could also be done in the Ex 2 as far as I could see it. 
+It might nicely link your approach to the on-going move of SDM people towards ensemble modeling and forecasting. It also lonk to Bayesian model averaging. 
+If I was meant, I would say that a simple use of binomial GLM is a 10-years way back for most people where more 
+}
 
 \subsection*{Advantages of Model Integration}
-Models are important tools that are being increasingly used by land managers to face challenges associated to decision-making. 
-A growing number of models are available, but they can provide diverging answers to very similar questions, due to the differences in their assumptions and methodology. 
+Models are important tools that are increasingly used by land managers to face challenges associated to decision-making. 
+A growing number of models are available, but they can provide diverging answers to very similar questions due to the differences in their assumptions and methodology. 
 This can create confusion and even some mistrust towards models, and as a consequence some managers may be discouraged to incorporate any model at all in their management plans. 
-By integrating different types of models---and their outputs---into a common framework, we believe that our approach can contribute to overcome the gap between modelers and practitioners and thus to promote a widespread use of models to support decision-making.
+By integrating different types of models---and their outputs---into a common framework, we believe that our approach can contribute to overcome the gap between modelers and practitioners and thus promote more widespread use of models to support decision-making.
+\ia{This is repetitive with the applicability section. I suggest that you stick here on the modeling advantage of your approach and keep the applied advantage for the other section}
 
+\ia{suggest adding here:\\
+Model uncertainties is another key factor affecting applicability of model outputs (Addison et al., 2013). }
 Our approach also allows for a clear, transparent identification of uncertainties and how they are transmitted throughout the modelling process. 
+\tf{Really? I thought that only the uncertainty in the predictions are available and not the uncertainty from submodels.}
 This is one of the main strengths of this approach, and offers several advantages. 
 For instance, transparency in uncertainty can be considered as a sort of sensitivity analysis, in which areas with large uncertainty can be detected and new experimental research or additional data collection can be designed (e.g., Example 1, Figs.
-\ref{fig:ex1_sampling}, \ref{fig:ex1_precip})
+\ref{fig:ex1_sampling}, \ref{fig:ex1_precip}).
 The new knowledge resulting from this research can be readily incorporated into the metamodel, allowing for an iterative learning process that will undoubtedly contribute to reduce uncertainty in predictions for a wide range of environmental conditions.
+\tf{The question is also: which model do we aim to incorporate. At a point, incorporating a model maybe useless. }
 Moreover, our framework embraces change as a fundamental process and is able to adapt and respond to it. 
-The ease of incorporation of new knowledge to the modeling framework (including new theory or the result of management and experimental efforts), will allow for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans.
-In an era of continuous and unprecedented change, adaptive approaches such as the one presented here are often highlighted as a pressing need in order to develop strategies to promote ecosystems that are both feasible and resilient \citep[][Fig.
-\ref{fig:adaptive_management}]{Seastedt2008, Mori2013}.
-
+The ease of incorporating new knowledge to the modelling framework (including new theory or the result of management and experimental efforts), will allow for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans.
+In an era of continuous and unprecedented change, adaptive approaches such as the one presented here are often highlighted as a pressing need in order to develop strategies to promote ecosystems that are both feasible and resilient \citep[Fig.
+\ref{fig:adaptive_management};][]{Seastedt2008, Mori2013}.
+\wt{
+Perhaps we could also be more precise here.
+Give examples.
+Since you are working with forest data, an easy step forward would also be to calibrate growth curve of most tree species in function of both temp and precipitation, even a competition model (see Kunstler et al.
+2011 J.
+of Ecology), that can then be integrated into SDMs.
+Similarly you may also discuss about the NEON and LTER long term sites that could certainly give similar information than the one used in Ex 1.
+I think we should strenghen the different possibilities to make sure users find the approach useful.
+It remains to vague to me at this point.
+If you think about the Cheaib et al. paper, you may also suggest that your approach could be run over all different model outputs to integrate thema all together. 
+In a way that should re-inforce the link between modellers and long term experimental ecologists that did not really discuss so far since it was difficult to integrate the two informations. 
+}
 %==================
@@
 	\caption{Sample workflow for applying the models presented in the first example in an adaptive management context.
-	Critical steps include specifying the meta-model, identifying additional sources of information to be used as constraints on the meta-model, then using the integrated prediction for decision-making.
+	Critical steps include specifying the meta-model, identifying additional sources of information to be used as constraints on the meta-model, and using the integrated prediction for decision-making.
 	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the meta-model.}
@@
 For example, \citet{Freckleton2009} estimated that data is too scant to successfully develop highly mechanistic models predicting weed population numbers. 
-A strong asset of our proposed method is that it can be used without the full suite of data that would be required to run a fully mechanistic model. 
-Since the meta-model is correlative, the rough structure of our approach only needs e.g. presence-absence data to be effective. 
+A strong asset of our approach is that it can be used without the full suite of data that would be required to run a fully mechanistic model. 
+Since the meta-model is correlative, it can be effectively implemented with, e.g., just  presence-absence. 
 Then any additional mechanistic data that is available will enhance predictions by constraining outputs of the meta-model.
 
+\tf{Revised this paragraph to the following}
+\tf{Finally, determining the functions to use to express the likelihood of the sub-models given the meta-model (i.e., Eqs.
+6, 7) is a critical point.
+In the context of integration models, the challenge is three fold: (i) which spatial and time scales, i.e.
+which processes, are to be considered, (ii) how to build scaling function in order to be consistent with the meta-model and (iii) how error and uncertainty are propagated from the finest sub-models scale to the largest meta-model scale.
+We argue that our proposed framework is able to easily deal with very different scale and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered.
+The building of scaling function deserve more investigation.
+We have shown in our examples that even simple scaling functions can provide reasonable constraints on the meta-model.
+However, with the objective to take into account all know processes and models, it is likely that the modelling process proposed may concern more than one scaling function and not necessarily from the same scale.
+Indeed, species distribution is a function of e.g.
+species growth rate (Sykes et al.
+1996, Guisan \& Zimmerman 2000) which involves processes at the individual (e.g.
+competition) or cellular (e.g.
+photosynthesis) scales.
+Such very large differences in spatial scales imply that simple functions will be inadequate requiring more traditional upscaling methods.
+Therefore, intermediate models and processes corresponding to additionnal data will be necessary: additionnal parameters inducing additionnal uncertainty have to be estimated before meaningful model integration will be possible.
+In addition, not only the spatial scale is to be considered, but also temporal scales.
+Most of time, this scale is taken into account in the spatial scale, as there is a strong concordance between them: the most rapid processes corresponded to the finest scale (e.g.
+photosynthesis is related to daylight) while longest processes are related to largest spatial scale (e.g.
+species niche).
+}
+\tf{Sykes, M.T., Prentice, I.C., Cramer, W., 1996.
+A Bioclimatic Model for the Potential Distributions of North European Tree Species Under Present and Future Climates.
+Journal of Biogeography 23, 203–233.}
 Finally, determining the functions to use to express the likelihood of the sub-models given the meta-model (i.e., Eqs.
@@
 Successful use of an integrated modelling approach will always remain dependent on an intimate understanding of the decision-making process by modellers, emphasizing the importance of close collaboration between with practitioners at all stages of model development \citep{Guisan2013}.
+\tf{Just a stupid question. In the co-authors of this manuscript, is there any policy-maker and/or practitioners?}
 
@@
 Integrating submodels allows for clear specification of desired model outputs (via the specification of the metamodel) while easily retaining important ecological objectives (via careful specification of submodels).
-Insufficiently documented models are difficult to assess in real world situation, and therefore only having limited relevance  for decision support \citep{Guisan2013}.
+Insufficiently documented models are difficult to assess in real world situation, and therefore only have limited relevance for decision support \citep{Guisan2013}.
 Model workflow documentation becomes more crucial in the case of integrated modelling approaches, which incorporates information from various scales and resolutions within a number of aggregated sub-models.
-A transparent and well documented workflow describing the process of model integration ensures reproducibility and facilitates its use in an applied context (Fig. \ref{fig:adaptive_management}). 
+A transparent and well documented workflow describing the process of model integration ensures reproducibility and applicability (Fig. \ref{fig:adaptive_management}). 
 The issue of uncertainty also limits the use of models in guiding decisions. 
-Model outputs are considered too uncertain for decision-making \citep{Addison2013}. 
+Model outputs may be considered too uncertain for decision-making \citep{Addison2013}. 
+\tf{repetitive?}
+\ia{last 2 sentences are repetitive with ?? something was supposed to go here...}
 The transparent estimation of the uncertainty provided by this modelling approach may be a significant advantage.
@@
 The utility of such models depends on their ability to help evaluate events beyond the bounds of the available data, including in future situations constrained by climate change. 
-In this context, practionners judge model usefulness on two criteria : Does the model provide reliable predictions at the needed time and space scales, and can the model be implemented given the available data and technical expertise?
+In this context, practionners judge model usefulness on two criteria : (1) Does the model provide reliable predictions at the needed time and space scales, and (2) can the model be implemented given the available data and technical expertise?
+\tf{Practitioners may first want to validate models, no? This needs that practitioners are faithful to scientific results, which is not always the case.}
 Since most decision makers work at local space scales and follow specific time frames, modelers face the obvious challenge of providing detailed outputs while preserving implementation capacity.
@@
 However, developing the models requires careful model specification, knowledge of applied Bayesian methods, and, in some cases, extensive programming.
+\tf{This is needed to develop models. But not for using models?}
 In many cases, off-the-shelf software \citep[e.g.,][]{R, RJAGS} can adequately express the model likelihoods with minimal programming, but more complicated models will require the development of custom software.
@@
 No doubt, the implementation of our proposed method is currently too technically challenging for most practionners. 
-However, the same was true of all statistical techniques or modeling approaches when they were initially developed. 
-In addition, some of the computation costs that were associated with many techniques have now vanished, and even the conceptual challenges associated Bayesian statistics are being reduced as they gain popularity in the scientific literature.
+However, the same was true of all statistical techniques or modelling approaches when they were initially developed. 
+In addition, some of the computation costs that were associated with many techniques have now vanished, and even the conceptual challenges associated Bayesian statistics are being reduced as they gain recognition in the scientific literature.
+\tf{I find this argument poor. This is not because a method is popular that it is a good one. Maybe recognition is better.}
 We therefore argue that our proposed approach as a strong potential for direct use in our real world where climate is quickly changing and conservation practices must be adjusted accordingly.
+\ia{This would be possible with a strengthen collaboration between modelers and practitioners}
 
 \subsection*{Future directions}
+\wt{I have done a first round of comments on the MS.
+It looks very good to me.
+I have added some comments on the discussion.
+Honestly, I do not really think you need this last part (future work) given space limitation (watch out the total number of words and references).
+Instead I would rather develop some of the already written part that remains a bit vague and imprecise to me.
+I added few things to see what I mean.
+}
 \begin{itemize}

### 2014-07-18 tex/examples.tex
Added comments from coauthors

@@
 \section*{Examples}
+\aa{Example 1: nothing to add, I think it is great as is. Figures 3 and 4 are very informative now. Great job! 
+Example 2: it is nicely exposed, although the results don't seem too promising. You said you got new, better results, so that’s good news. ;-)}
+
 Our approach to model integration is best explained with concrete examples. 
 In this section we first present how experimental results could be introduced as prior knowledge of the fundamental niche in a species distribution model (\ac{SDM}), and in the following we apply our approach to integrating phenological information with occurrence data for a widespread North American tree.
+\ia{delete the above paragraph}
+\mt{I tend to agree}
 
@@
 
-In this first hypothetical example, we build a meta-model relating the distribution (i.e.
+In this hypothetical example, we build a meta-model relating the distribution (i.e.
 occurrence probability) of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
 The meta-model here captures the realized distribution of a species, thereby encapsulating in a single correlative model all physiological constraints and ecological processes constraining a species distribution.
+\tf{Well, I understand that this is an objective of the method, but in this example, all physiological and ecological constraints are not taken into account.
+Plus, it is impossible to know all physiological and ecological constraints. The correct expression would be “all known physiological and ecological constraints”}
 However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the envelope of where a species could potentially be found.
@@
 %-----
-where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
+where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modelling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
 
@@
 	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
-	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation.
+	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., \(-\)1--1).
 	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
@@
 According to niche theory \citep{Holt2009}, the fundamental niche corresponds to the set of environmental conditions where the per capita intrinsic growth rate $r$ is positive.
-This concept gives us a reasonable model to fit the scaling function $g$ for our sub-mode (Fig. \ref{fig:diagram}).
-If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, the for observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
+This concept gives us a reasonable model to fit the scaling function $g$ for our sub-model (Fig. \ref{fig:diagram}).
+If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, then for observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
 \begin{equation}
@@
 
-Although the two data sets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the meta-model second (i.e., the probability of presence at a given precipitation).
+Although the two data sets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the meta-model (i.e., the probability of presence at a given precipitation).
 This scaling is quite simplistic, and would never be used to predict a species' range using this mechanistic submodel in isolation.
@@
 These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
-For purposes of this example, we used these terms to reduce the precision of the second model to reflect the the uncertainty of the modelling process (due to the lack of additional mechanisms in the second model, as mentioned above).
+For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modelling process (due to the lack of additional mechanisms in the second model, as mentioned above).
 
@@
 \ref{fig:ex1_precip}, \ref{fig:ex1_map}).
-Because the meta-model includes both temperature and precipitation, any reduction in the uncertainly with respect to precipitation might also influence the evaluation of the relationship with temperature (reducing bias in the case where there is a correlation between T and P in the data).
+Because the meta-model includes both temperature and precipitation, any reduction in the uncertainty with respect to precipitation might also influence the evaluation of the relationship with temperature (reducing bias in the case where there is a correlation between T and P in the data).
 However, uncertainty could increase in the case of model disagreement.
-The relationship between precipitation in the naive meta-model could be caused by a spurious relationship (e.g.
+For instance, the relationship between precipitation and presence in the naive meta-model could be caused by a spurious relationship (e.g.
 spatial autocorrelation caused by historical contingencies), leading to a false confidence in the species distribution over precipitation gradients.
@@
 For the meta-model, we used a dataset of 1013 sugar maple presence records at 0.5 degree resolution, with 498 corresponding pseudo-absence records sampled from North America.
+\mt{update numbers with final versions, and indicate that they are true absences}
 We used a binomial GLM to relate the suitability for sugar maple to 5 climate variables, scaled to the same resolution (see SI for a description of the climate variables used).
-For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} with the A2 climate scenario \citepalias{IPCC2001}, and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
+\wt{We could also have used something more complex such as a GAM or boosted regression trees. }
+\mt{Perhaps note this; we selected this model for simplicity of explanation and implementation, but more flexible functions are possible}
+For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario (REF--Nakicenovic), and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
+\wt{citation:
+Nakicenovic, N. \& Swart, R. (ed.\^eds) (2000) Emissions Scenarios: A Special Report of Working Group III of the Intergovernmental Panel on Climate Change. Cambridge University Press, Cambridge.}
+\tf{Is it an optimistic or pessimistic or medium scenario? 
+Would it be interesting to compare two extreme scenarios? Would it change something to the demonstration of this manuscript?}
 Phenofit produces predictions at continental scales; thus the scaling function \(g\) was incorporated within Phenofit itself and we performed model integration directly on the outputs.
@@
 \(D_M\) is the matrix of climate predictors,
-\(X_S\) is the prediction from Phenofit (in this case a single deterministic value, because we did not incorporate errors from the Phenofit modeling),
+\(X_S\) is the prediction from Phenofit (in this case a single deterministic value, because we did not incorporate errors from the Phenofit modelling),
 and \(\theta_S\) is the model relating the Phenofit predictions to the meta-model.
@@
 %==================
+\tf{Fig 5: Is it realistic to make predictions as South as Mexico or In Cuba? 
+A question arises for me: to which spatial enveloppe should these models be used? Worldwide? Smaller enveloppe?
+This opens questions about species interactions.
+The predictions made are for sugar maple only without taking into account other speices.
+Sugr maple is present in EastAmrica.
+Is there another species which is only present in WesternAmerica ?
+Do we known how both species interact and who will “win” the competition game?
+I guess that the approach proposed is dedicated to answer these questions and that the figure are only here to present the distribution of sugar maple when only climate is changing.
+}
 
-For modeling the contemporary range of sugar maple, both Phenofit and the naive model perform very well relative to the integrated model, espeically when capturing the range margin (Fig \ref{fig:ex2}A, C, G).
+For modelling the contemporary range of sugar maple, both Phenofit and the naive model perform very well relative to the integrated model, espeically when capturing the range margin (Fig \ref{fig:ex2}A, C, G).
 Conventional SDMs are quite good at producing empirical models of species distributions under present climatic regimes, and the naive model here could be further improved with the addition of more parameters or the inclusion of ensemble forecasting (See SI for model-fitting details).

### 2014-07-18 tex/introduction.tex
Added comments from coauthors

@@
 \section*{Introduction}
+\aa{I like it very much now, it's good and concise.
+The only thing is that the first and last sentences in the first paragraph seem contradictory.
+Do models of species range limits play a large role on conservation biology or do their contributions remain limited?
+Maybe in the first sentence you wanted to say they ‘can’ play a large role?
+}
+\aa{In addition, you may want to add this reference: Snell, R.S., Huth, A., Nabel, J.E.M.S., Bocedi, G., Travis, J.M.J., Gravel, D., Bugmann, H., Gutiérrez, A.G., Hickler, T., Higgins, S.I., Scherstjanoi, M., Reineking, B., Zurbriggen, N.
+\& Lischke, H.
+(2014) Using dynamic vegetation models to simulate plant range shifts.
+Ecography, In press.}
 
 Models of species range limits have wide applications and play a large role in conservation biology, where they are used as decision-making tools in biodiversity management \citep{Rosenzweig2008, Thuiller2008, Rodenhouse2009}.
+\wt{I would just put Guisan et al. 2013 here. }
 Due to large temporal and spatial scales and the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating range limits \citep{Wu1995, Levin1998}.
-Thus, models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution and hybrid zone dynamics \citep{Engler2013}, and the impacts of climate change  \citep{Rosenzweig2008, Thuiller2008, Milad2011, Blois2013, Loyola2013}.
-However, despite having recognized potential to support decision making, the contributions of models to conservation and applied ecology remain limited \citep{Dawson2011, Guisan2013}.
+Thus, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution, hybrid zone dynamics \citep{Engler2013}, and the impacts of climate change  \citep{Rosenzweig2008, Thuiller2008, Milad2011, Blois2013, Loyola2013}.
+\wt{after evolution above: Jay, F., Manel, S., Alvarez, N., Durand, E., Thuiller, W., Holderregger, R., Taberlet, P.
+\& François, O.
+(2012) Forecasting changes in population genetic structure of Alpine plants in response to global warming.
+Molecular Ecology, 21, 2354-2368.}
+\wt{The above mixes up subfields vs applications}
+\wt{for climate change citations:
+Thuiller, W., Lavergne, S., Roquet, C., Boulangeat, I. \& Araujo, M.B. (2011) Consequences of climate change on the Tree of Life in Europe. Nature, 470, 531-534. \\
+Thuiller, W., Pironon, S., Psomas, A., Barbet-Massin, M., Jiguet, F., S., L., Pearman, P.B., Renaud, J., Zupan, L.
+\& Zimmermann, N.E.
+(2014) The European functional tree of bird life in face of global change.
+Nature Communications, 5, 3118, DOI: 10.1038/ncomms4118.
+}
+However, despite having recognized potential to support decision making, the contributions of these models to conservation and applied ecology remain limited \citep{Dawson2011, Guisan2013}.
+\wt{remove Dawson}
 
 An important constraint on the modelling process is the availability of data and appropriate ecological theory to  link data to modelling objectives.
-Poor data or poorly conceived models, will result in predictions with unacceptably high uncertainty, or, more dangerously, precise but biased predictions.
+Poor data or poorly conceived models will result in predictions with unacceptably high uncertainty, or, more dangerously, precise but biased predictions.
+\tf{Why is it dangerous? 
+It depends also of the importance of the bias.
+Some regression technique exists in order to deal multicolinearities and their basis is to produce precise and biased estimates.
+In this case, it is a way to reduce uncertainty.
+See e.g.
+ridge regression.
+With this kind of regression, there is no “danger”, and the estimates are known to be biased (and therefore should be interpreted and used as biased estimates).
+Plus, I feel that the word dangerous is not the correct word. There is ho danger in this, no harm. 
+}
 As a result of data accumulation from various sources and scales and conceptual advances in ecology, modelling techniques have proliferated to take advantage of what is available.
-Model diversification also arises due to the diverse processes generating species ranges \cite{Soberon2007}.
+Model diversification also arises due to the diverse processes generating species ranges \citep{Soberon2007}.
 In the context of forecasting, which is the focus hereafter, mechanistic and correlative approaches represent quite different modes of inference, but both have strong limitations. 
@@
 For instance, biotic interactions (including trophic interactions) are usually not modelled mechanistically at the scale of species distributions because they are not well-known, or have not been recorded, despite being considered a key determinant of range limits \citep{Soberon2007, Roux2012, Guo2013, Pigot2013}.
-On the other hand, coarse-scale correlative models statically relating species distributions to associated distributions of other variables at the same scale have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
-However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, which implies that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}.
+On the other hand, coarse-scale correlative models statistically relating species occurrences to associated distributions of other variables at the same scale have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
+However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}.
 Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
 Furthermore, correlative models often do not explicitly incorporate theory into their formulations, and when applied at coarse-scales, they may miss important fine-scale processes, making a multi-scale approach desirable when accurate forecasting is required \citep{Austin2007, Austin2011, Soranno2014}.
+\wt{remove all 3 cites, add \\
+Thuiller et al. 2013 ELE}
 
@@
 A key problem with the diversity of approaches is that multiple models often produce differing predictions for the same organism, with no simple way to reconcile these predictions.
-We present an application of hierarchical Bayesian methods for integrating the predictions of multiple models; our approach is an alternative to other approaches to drawing inference from multiple models (summarized below) that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
-For instance, correlative models are often criticized because they assume that the selected explanatory variables limit (directly of indirectly) the species range, and, as mentioned above, that the relationship between the selected variables and the species occurrences will not change in the prediction context \citep{Araujo2006, Berteaux2006, Braunisch2013}.
+\aa{Change to: \\
+In addition, an assortment of information (in the form of both data and theory)...}
+\aa{from here}
+We present an application of hierarchical Bayesian methods for integrating the predictions of multiple models.
+Our approach is an alternative to other approaches to drawing inference from multiple models that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
+\ia{this is repetitive with the last paragraph of this section. \\
+delete everything up and and including "For instance..."}
+For instance, correlative models are often criticized because they assume that the selected explanatory variables limit (directly or indirectly) the species range, and, as mentioned above, that the relationship between the selected variables and the species occurrences will not change in the prediction context \citep{Araujo2006, Berteaux2006, Braunisch2013}.
+\aa{to here}
 However, additional information (in the form of both data and theory) is often available on the species of interest (e.g., growth rates in different climatic conditions, phenology, interactions with other species, sensitivity to disturbances) \citep{Holt2009, Thuiller2013}.
@@
 However, the link between different sub-models is based on additional assumptions that are not easy to test \citep{Gallien2010}. 
+\aa{which assumptions? not clear}
+\ia{a bit vague here. Give an example?}
 Moreover, this approach is not suitable for merging models that are based on the same processes but with different underlying hypotheses.
+\ia{an example?}
+
 A second approach is to directly combine predictions.
-If models operate at the same scales have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting][]{Araujo2007, Diniz-Filho2009}.
+If models operate at the same scales and have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Araujo2007, Diniz-Filho2009}.
+\wt{add: Thuiller, W. (2004) Patterns and uncertainties of species' range shifts under climate change. Global Change Biology, 10, 2020-2027.\\
+remove: Diniz-Filho2009}
 If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Serra-Diaz2013}.
@@
 Clearly, an approach that integrates predictions and uncertainty from multiple model types is needed \citep{Beck2012, Thuiller2013}. 
-Such approach would incorporate multiple modes of inference and integrate data from multiple sources and scales \citep{Levin1992, Peters2004, Thuiller2013}.
+Such an approach would incorporate multiple modes of inference and integrate data from multiple sources and scales \citep{Levin1992, Peters2004, Thuiller2013}.
 
@@
 Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to intertpret and reflects uncertainty at all levels of organization \citep{Clark2005, Cressie2009, Hobbs2011}. 
-We use simulated data to present a complete example of the application of the framework to multiple information sources that are relevant at very different scales.
-Finally, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}) to climate with the predictions of a phenological model, Phenofit, with the goal of improving uncertainty estimation when projecting model predictions to future climate.
+We illustrate our approach with two examples.
+First, we use simulated data to present a complete example of the application of the framework to multiple information sources that are relevant at very different scales.
+In a second example, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}) to climate with the predictions of a phenological model, Phenofit, with the goal of improving uncertainty estimation when projecting model predictions to future climate.

### 2014-07-18 tex/manuscript.tex
Added comments from coauthors

@@
 
-\usepackage[margin=1in]{geometry}
+%\usepackage[margin=1in]{geometry}
 
@@
 %% ----------------------------------
-%\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
+\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
+
+\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
+
+% available colors:
+% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
 %
-%\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
-%
-%% available colors:
-%% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
-%
-%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
-%\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}}
-%\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
-%\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
-%\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
+\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
+\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
+\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
+\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
+\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
+\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
 %\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
 %\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
-%\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
+\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
 %\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
 %\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
-%\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
-%\newcommand{\wt}[1]{\todo[color=violet!40]{\textsuperscript{Wilfried T}#1}}
-%\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}
+\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
+\newcommand{\wt}[1]{\todo[color=orange!40]{\textsuperscript{Wilfried T}#1}}
+\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}
+
+%% most recent draft - nothing yet
+% IB
+% Dominique B
+% Ronnie D
+% MJF
+% Dan M
+% JL/NS
+% KS
+% Fred D
+
+% discussion - nothing yet
+% Ronnie D
+% MJF
+% Dan M?
+% Fred D
+% Kevin S
 
@@
 % \renewcommand{\db}[1]{\relax}%
+% \renewcommand{\ab}[1]{\relax}%
 % \renewcommand{\rd}[1]{\relax}%
@@
 \author[1,2]{Alyssa Butler}
-\author[10]{Dominique Cyr} %??
-\author[10]{Fred Doyon} %??
+%\author[10]{Fred Doyon} %??
 \author[6]{C. Ronnie Drever}
@@
 \author[8]{Nick Strigul}
-\author[9]{Wilfried Thuiller}
+\author[9, 10]{Wilfried Thuiller}
 \author[1,2]{Dominique Gravel}
@@
 \affil[8]{Department of Mathematics, Washington State University Vancouver, Washington, USA}
-\affil[9]{Laboratorie d'Ecologie Alpine, Unité Mixte de Recherche, Université de Grenoble, Grenoble, France}
-\affil[10]{Affiliations to be filled in}
+\affil[9]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
+\affil[10]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
 
@@
 \maketitle
+\aa{although you eliminated the sentence in which you defined knowledge as data+theory - which contradicted the title - if I play the role of a reviewer I would still ask you why you differentiate between data and knowledge and how would you define knowledge.
+Nothing too critical, though.}
 
 \begin{flushleft}
-\textbf{Keywords:} Uncertainty, scaling, range dynamics, trees, disturbances, patterns and processes, predict or understand, climate change, biotic interactions, spatial ecology, decision making, metamodelling, species distribution modeling
+\textbf{Keywords:} Uncertainty, scaling, range dynamics, trees, disturbances, patterns and processes, predict or understand, climate change, biotic interactions, spatial ecology, decision making, metamodelling, species distribution modelling
 \end{flushleft}
@@
 \begin{abstract}
-There is great interest in modeling species ranges, and such models are useful in guiding conservation and management decisions.
-There presently exists a great diversity of modeling approaches, and individual models often use only a small subset of the total available information about a species.
+There is great interest in modelling species ranges, and such models are useful in guiding conservation and management decisions.
+There presently exists a great diversity of modelling approaches, and individual models often use only a small subset of the total available information about a species.
+\ia{Revision for first 2 sentences:\\
+Modelling species ranges, is of great interest in guiding conservation and management decisions.
+A wide diversity of modeling approaches are currently used, with individual models using only a small subset of the total available information about a species according to the scale and the specification of the model.
+}
 For example, a broad-scale correlative model might use climate variables to predict presence or absence, but ignore what is known about smaller-scale processes such as the effect of local climate on growth, fecundity, and dispersal rates.
+\ia{I would not put an example in the abstract}
 A central problem with the diversity of approaches is that multiple models produce differing predictions for the same organism, with no simple way to reconcile these predictions.
+\ia{This may not be seen as a problem as each model are focussing on a different aspect/have different objective/work at different scale. Maybe reword here?}
 We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
@@
 Finally, we discuss the application of our method and its accessibility to conservation biologists and land managers.
-Although the method requires extensive statistical programming experience, we anticipate the results will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.
+Although it requires extensive statistical programming experience, we anticipate the method will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.
+\ia{need rewording here}
 \end{abstract}
@@
 \input{brief_model_description}
-
 %---------------------------------------------
 %---------------------------------------------
-
 \input{examples}

### 2014-08-18 tex/brief_model_description.tex
Added second batch of comments

@@
 
-The key idea of our approach is to constrain the parameters of a correlative meta-model using predictions made by one or several mechanistic sub-models.
+The key idea of our approach is to constrain the parameters of a correlative meta-model using predictions made by one or several mechanistic sub-models from other spatial scales.
 \aa{Does the meta-model need to be correlative? Because later on you say that it 'could take the form of a correlative model' as implying that it could take other forms.}
-The role of the meta-model is to integrate data at the same ecological scale as predictions. 
-As an example, the meta-model could take the form of a simple correlative \ac{SDM}.
-Ideally, each submodel incorporates different data and hypotheses that specify underlying ecological mechanisms contributing to the model output.
+The role of the metamodel is to integrate data at the same ecological scale as predictions. 
+As an example, the metamodel could take the form of a simple correlative \ac{SDM}.
+Ideally, each sub-model incorporates different data types and hypotheses that specify underlying ecological mechanisms contributing to the model output.
 \tf{Check in examples if hypothesis are given}
 \mt{they are there, though perhaps not so explicit}
-The output of sub-models should provide additional predictions at a comparable ecological scale at which the meta-model operates, although it is not necessary for the sub-models to actually operate at the same scale.
-For sub-models operating at a substantially different scale than the meta-model, scaling functions will be required; however, as we will show in the first example, these functions can be simple while still contributing useful information to the meta-model.
-All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the meta-model (see Box 1, Fig. \ref{fig:diagram}).
+The output of sub-models should provide additional predictions at a comparable ecological scale at which the metamodel operates, although it is not necessary for the sub-models to actually operate at the same scale.
+For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, as we will show in the first example, these functions can be simple while still contributing useful information to the metamodel.
+All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Box 1, Fig. \ref{fig:diagram}).
 We present the complete framework in a hypothetical example where experimental results are  introduced as prior knowledge of the fundamental niche in a species distribution model (SDM).
-A more formal presentation of the framework appears in Box 1.
+A more formal presentation of the modelling framework appears in Box 1.
 We follow with a second example where we apply our approach to integrate phenological information with occurrence data for a widespread North American tree.
@@
 \input{diagram}
-\caption{The parameters of a correlative meta-model model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
-The meta-model ($\theta_M$) operates at a single ecological scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
+\caption{The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
+The metamodel ($\theta_M$) operates at a single ecological scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
 The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
 The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
-This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the meta-model via a scaling function \(g(\theta_S, D_S)\).
+This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
 The final prediction can be obtained by applying the integrated model to the original explanatory variables (as shown) or, if projection is desired, on some new set of explanatory variables (e.g., future climate).

### 2014-08-18 tex/discussion.tex
Added second batch of comments

@@
 Journal of Biogeography, 39, 2240-2252.}
-In contrast with Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of a generally more correlative meta-model.
+In contrast with Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of a generally more correlative metamodel.
 Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005, Efendiev2005a}, plant physiology \citep{Ogle2008, Ogle2009}, and climate and atmospheric modelling \citep{Zimmerman2005, Mcmillan2010, Kang2012}.
@@
 You may actually run a model averaging around your approach, may not you?
-Indeed, if we think to Ex 1 you could have several native meta-model posterior distributions, right?
+Indeed, if we think to Ex 1 you could have several native metamodel posterior distributions, right?
 Or even a direct model averaging from them.
@@
 \subsection*{Advantages of Model Integration}
-Models are important tools that are increasingly used by land managers to face challenges associated to decision-making. 
+Models are important tools that are increasingly used by land managers to face challenges associated to decision-making \citep{Guisan2013}
 A growing number of models are available, but they can provide diverging answers to very similar questions due to the differences in their assumptions and methodology. 
-This can create confusion and even some mistrust towards models, and as a consequence some managers may be discouraged to incorporate any model at all in their management plans. 
-By integrating different types of models---and their outputs---into a common framework, we believe that our approach can contribute to overcome the gap between modelers and practitioners and thus promote more widespread use of models to support decision-making.
+This can create confusion and even some mistrust towards models, and as a consequence some managers may be discouraged to incorporate model results in their management plans. 
+By integrating different types of models---and their outputs---into a common framework, we believe that our approach can contribute to overcome the gap between modellers and practitioners and thus promote wider use of models to support decision-making.
 \ia{This is repetitive with the applicability section. I suggest that you stick here on the modeling advantage of your approach and keep the applied advantage for the other section}
@@
 	\caption{Sample workflow for applying the models presented in the first example in an adaptive management context.
-	Critical steps include specifying the meta-model, identifying additional sources of information to be used as constraints on the meta-model, and using the integrated prediction for decision-making.
-	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the meta-model.}
+	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
+	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the metamodel.}
 	\label{fig:adaptive_management}
 \end{figure}
+\dm{Could write out the word Objectives in the Figure}
 
@@
 Poorly specified models will produce outputs that are uninformative or misleading, and model integration is not a cure for these problems.
-In the worst case, constraining a meta-model with a poor sub-model can result in outcomes that are worse in terms of bias and uncertainty than those produced by a naive model.
+In the worst case, constraining a metamodel with a poor sub-model can result in outcomes that are worse in terms of bias and uncertainty than those produced by a naive model.
 We expect model selection will play an important role in applications of our framework, and such schemes can be incorporated relatively easily \citep{Madigan1995, Wasserman2000, Tenan2014}.
 
-In a similar fashion, the quality and availability of data imposes a significant constraint on the number and type of models that can be implemented.
+In a similar fashion, the quality and availability of data impose a significant constraint on the number and type of models that can be implemented.
 The capacity to implement a model is very low if data requirements are prohibitive. 
@@
 \ref{fig:ex1_precip}, \ref{fig:ex1_map}).
-For example, \citet{Freckleton2009} estimated that data is too scant to successfully develop highly mechanistic models predicting weed population numbers. 
+For example, \citet{Freckleton2009} estimated that data are too scant to successfully develop highly mechanistic models predicting weed population numbers. 
 A strong asset of our approach is that it can be used without the full suite of data that would be required to run a fully mechanistic model. 
-Since the meta-model is correlative, it can be effectively implemented with, e.g., just  presence-absence. 
-Then any additional mechanistic data that is available will enhance predictions by constraining outputs of the meta-model.
+Given that the metamodel is correlative, it can be effectively implemented with, e.g., only presence-absence data. 
+Then any additional mechanistic data that are available will enhance predictions by constraining outputs of the metamodel. 
+\dm{add: We note however that absence data can be problematic in some cases and naïve application of any modeling framework can be misleading.}
 
 \tf{Revised this paragraph to the following}
-\tf{Finally, determining the functions to use to express the likelihood of the sub-models given the meta-model (i.e., Eqs.
+\tf{Finally, determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eqs.
 6, 7) is a critical point.
 In the context of integration models, the challenge is three fold: (i) which spatial and time scales, i.e.
-which processes, are to be considered, (ii) how to build scaling function in order to be consistent with the meta-model and (iii) how error and uncertainty are propagated from the finest sub-models scale to the largest meta-model scale.
+which processes, are to be considered, (ii) how to build scaling function in order to be consistent with the metamodel and (iii) how error and uncertainty are propagated from the finest sub-models scale to the largest metamodel scale.
 We argue that our proposed framework is able to easily deal with very different scale and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered.
 The building of scaling function deserve more investigation.
-We have shown in our examples that even simple scaling functions can provide reasonable constraints on the meta-model.
+We have shown in our examples that even simple scaling functions can provide reasonable constraints on the metamodel.
 However, with the objective to take into account all know processes and models, it is likely that the modelling process proposed may concern more than one scaling function and not necessarily from the same scale.
@@
 Journal of Biogeography 23, 203–233.}
-Finally, determining the functions to use to express the likelihood of the sub-models given the meta-model (i.e., Eqs.
+Finally, determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eqs.
 \ref{eq:ex1_integrated}, \ref{eq:integrated2}) will remain a significant challenge.
-We have shown in our examples that even simple scaling functions can provide reasonable constraints on the meta-model. 
+We have shown in our examples that even simple scaling functions can provide reasonable constraints on the metamodel. 
 However, it is likely that with very large differences in scales (e.g., comparing cellular processes to landscapes), simple functions will be inadequate, requiring more traditional upscaling methods that combine explicit processes with additional data (and, necessarily, additional parameters) before meaningful model integration will be possible.
@@
 \defcitealias{TERN2013}{TERN, 2013} %% this is needed to make the organizational citation in the following paragraph work correctly
-Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organisations around the world \citep[e.g.][\citetalias{TERN2013}]{Bernier2013}.
+Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organizations around the world \citep[e.g.][\citetalias{TERN2013}]{Bernier2013}.
 Incorporating information from multiple sources, particularly with respect to uncertainty,  fosters a connection between scientifically-generated knowledge and policy, and is therefore an important tool for adaptive management \citep[][Fig.
-\ref{fig:adaptive_management}]{Rehme2011}.
+\ref{fig:adaptive_management}; ]{Rehme2011}.
 Such approaches are needed in formulating management plans for vulnerable species and ecosystems to avoid basing decisions on too-narrow subsets of the available information \citep{Dawson2011}.
@@
 To be useful, models should be transparent analytical tools, not black boxes \citep{Addison2013}. 
-A key criteria for model applicability is the explicit and detailed communication of specific model objectives, characteristics, limits, and uncertainties, as well as its ecological foundation \citep{Guisan2013}.
-Integrating submodels allows for clear specification of desired model outputs (via the specification of the metamodel) while easily retaining important ecological objectives (via careful specification of submodels).
+A key criterion for model applicability is the explicit and detailed communication of specific model objectives, characteristics, limits, and uncertainties, as well as its ecological foundation \citep{Guisan2013}.
+Integrating sub-models allows for clear specification of desired model outputs (via the specification of the metamodel) while easily retaining important ecological objectives (via careful specification of sub-models).
 Insufficiently documented models are difficult to assess in real world situation, and therefore only have limited relevance for decision support \citep{Guisan2013}.
@@
 \subsection*{Accessibility}
+\mjf{Accessibility of what? Modeling? Data?}
 Results of models predicting species ranges are increasingly used by resource managers, conservation biologists, or forest ecologists for formulating or adjusting recommendations. 
 The utility of such models depends on their ability to help evaluate events beyond the bounds of the available data, including in future situations constrained by climate change. 
-In this context, practionners judge model usefulness on two criteria : (1) Does the model provide reliable predictions at the needed time and space scales, and (2) can the model be implemented given the available data and technical expertise?
+In this context, practitioners may judge model usefulness on two criteria: (1) Does the model provide reliable predictions at the needed time and space scales, and (2) can the model be implemented given the available data and technical expertise?
 \tf{Practitioners may first want to validate models, no? This needs that practitioners are faithful to scientific results, which is not always the case.}
@@
 In terms of model outputs, our framework is transparent in terms of uncertainty by virtue of providing easily interpretable probability distributions for model parameters and predictions.
-However, developing the models requires careful model specification, knowledge of applied Bayesian methods, and, in some cases, extensive programming.
+However, developing the models requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming.
 \tf{This is needed to develop models. But not for using models?}
@@
 Thus, this framework will require significant investment in developing customized code for the samplers in order to actually estimate parameters.
-No doubt, the implementation of our proposed method is currently too technically challenging for most practionners. 
+No doubt, the implementation of our proposed method is currently too technically challenging for many practitioners. 
+\mt{This seems a bit negative to me... maybe nicen it up a bit}
+\dm{There is a difference between practioners and Decision-makers.
+You might want to work something in like: Adoption will be no doubt influenced by the percieved complexity and and usefulness to both practitioners (species distribution modelers) and decision-makers (those who would use to results to support operational or policy-related decisions).
+}
 However, the same was true of all statistical techniques or modelling approaches when they were initially developed. 
@@
 	\item incorporation of spatially explicit models into the framework? \citep[e.g.,][]{Fortin2012}
-	\item more sophisticated meta-models (e.g., state-transition models)
+	\item more sophisticated metamodels (e.g., state-transition models)
 	\item others?
 \end{itemize}
+\dm{Applications to demonstrate relevance; Use for productivity and abundance modeling rather than range delineation; Data development is also important}

### 2014-08-18 tex/examples.tex
Added second batch of comments

@@
 
-Our approach to model integration is best explained with concrete examples. 
+Our modelling framework is best explained with concrete examples. 
 In this section we first present how experimental results could be introduced as prior knowledge of the fundamental niche in a species distribution model (\ac{SDM}), and in the following we apply our approach to integrating phenological information with occurrence data for a widespread North American tree.
@@
 
-In this hypothetical example, we build a meta-model relating the distribution (i.e.
+In this hypothetical example, we build a metamodel relating the distribution (i.e.
 occurrence probability) of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
-The meta-model here captures the realized distribution of a species, thereby encapsulating in a single correlative model all physiological constraints and ecological processes constraining a species distribution.
-\tf{Well, I understand that this is an objective of the method, but in this example, all physiological and ecological constraints are not taken into account.
-Plus, it is impossible to know all physiological and ecological constraints. The correct expression would be “all known physiological and ecological constraints”}
-However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the envelope of where a species could potentially be found.
+The metamodel attempts to capture the realized distribution of a species, thereby encapsulating in a single correlative model the major physiological constraints and ecological processes constraining a species distribution.
+\mjf{How does it capture these constraints if it is a correlative model?}
+However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the climatic envelope of where a species may be found in a natural setting.
 Prior information of the physiological constraints affecting species distribution is sometimes available, but usually too incomplete to perform a reasonable comparison with the realized distribution.
 For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature.
+\dm{May want to provide a bit more justification as to why this particular hypothetical example is useful?
+When I think of experimental data I think you are meaning lab-controlled data which is quite rare I think?}
 We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty. 
-We focus here on the specification of the modelling framework; complete procedural details and code for generating the data sets and executing the model are provided as supplemental information.
+Here, we focus only on the specification of the modelling framework; complete procedural details and code for generating the data sets and executing the model are provided as supplemental information.
 
@@
 \end{equation}
+\dm{Just noting my initial comment that even absence data can be of course tricky eg structural zeros; a cryptic species; looked at the wrong time of day; some disturbance happened that year….etc….etc.
+I note your comment in the next sentence which is good}
 We further assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
-The model is evaluated by relating these observations to environmental data \((D_M)\), which for the sake of the example consists of mean annual temperature $T_M$ and total precipitation $P_M$. 
-For simplicity, we use a simple logistic model for the meta-model \((\theta_M)\) with a second order effect of temperature and precipitation. 
-This naive model (i.e., the meta-model with no constraints from sub-models) predicts  occurrence probability (\(\psi_N\)) as a function of the environment:
+The model is evaluated by relating these observations to environmental data \((D_M)\), which for the sake of the example consists of mean annual temperature $T_M$ and annual precipitation $P_M$. 
+For simplicity, we use a simple logistic model for the metamodel \((\theta_M)\) with a second order effect of temperature and precipitation. 
+This naive model (i.e., the metamodel with no constraints from sub-models) predicts  occurrence probability (\(\psi_N\)) as a function of the environment:
 %-----
@@
 In this context, the goal of modelling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, we recommend][for a concise introduction]{Link2010}:
+\dm{add citation: McCarthy M.A. 2007 Bayesian methods for Ecology. Cambridge University Press 296p}
+\dm{Mick McCarthy might be a good suggestion for a reviewer (I know him – an Aussie)}
 %-----
@@
 
-Although the two data sets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the meta-model (i.e., the probability of presence at a given precipitation).
-This scaling is quite simplistic, and would never be used to predict a species' range using this mechanistic submodel in isolation.
+Although the two data sets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation).
+This scaling is quite simplistic, and would never be used to predict a species' range using this mechanistic sub-model in isolation.
 Despite the strong theoretical foundation (i.e., niche theory) linking the observations to presence-absence at large scales, the upscaled sub-model (i.e., the transformation of \(X_S\) and \(\sigma_S\) to \(\psi_S\)) is built on the strong hypothesis that the species will be present when $r>0$, neglecting other population dynamics constraints such as Allee effects and metapopulation dynamics.
@@
 As such, it would be unwise to expect predictions from this model alone to resemble the actual distribution of the species; as a mechanistic model, it is simply too incomplete to predict distribution.
-As we will show, however, the information from this sub-model, when applied as a constraint on the meta-model, can result in improved predictions that incorporate the information within each model.
+As we will show, however, the information from this sub-model, when applied as a constraint on the metamodel, can result in improved predictions that incorporate the information within each model.
 
@@
 	\overbrace{p\left (\psi_S \mid \theta_S,P_S \right )}^{\substack{\text{new information} \\ \text{from sub-model}}}
-	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) P \left(\theta_M \right)}^{\substack{\text{naive meta-model} \\ \text{posterior}}}
+	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) P \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
 	\overbrace{p \left(\theta_S \right)}^{\substack{\text{prior for} \\ \text{sub-model}}}	
@@
 %-----
-As before, the meta-model \(\theta_M\) can be used to predict the species distribution \((\psi_I)\).
+As before, the metamodel \(\theta_M\) can be used to predict the species distribution \((\psi_I)\).
 However, these predictions will not reflect only the presence-absence samples in \(X_M\), but will also reflect the information from \(\theta_S\), including all of the data sources used to produce this sub-model.
@@
 We performed this procedure with our hypothetical data sets.
-When comparing the three models (naive meta-model, mechanistic sub-model, and integrated meta-model), we observed extreme uncertainty in the first model when projecting beyond the range of the original data.
+When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data.
 Unsurprisingly given the quality of the data, the sub-model was highly precise, providing a fairly strong constraint when producing the integrated model.
 The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
-At the scale of the meta-model, considering both temperature and precipitation, we observed similar results, with reduced uncertainty in the predictions over the domain not covered by the presence-absence data (Fig.
+At the scale of the metamodel, considering both temperature and precipitation, we observed similar results, with reduced uncertainty in the predictions over the domain not covered by the presence-absence data (Fig.
 \ref{fig:ex1_map}).
@@
 Predictions are highly precise due to high quality source data, but are likely to be overly optimistic given the simplicity in assuming only precipitation determines probability of presence (see text).
-	(c) Integrated model, showing predictions that are intermediate between the two submodels and uncertainty that is reduced compared to (a).
+	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
 	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.}
@@
 %==================
+\dm{FIG 4: Again just a comment …one could discuss what does this probability of presence really mean?
+Presence with a self sustaining population.
+Someone could plant a specimen outside its historic natural range and expect it to live.
+I woder if rather than naïve the term "simple" would be better?
+I have substituted for that word in some places …Aren’t we all naïve ??!!}
 
-Once the technical challenges of evaluating the parameters are overcome, the integration has several advantages over the both the naive parameterization of the meta-model and the mechanistic sub-model.
-In the situation where both models strongly agree on the relationship between water availability and occurrence probability, then we should expect significantly reduced uncertainty in that domain (Fig.
+Once the technical challenges of evaluating the parameters are overcome, the integration has several advantages over the both the naive parameterization of the metamodel and the mechanistic sub-model.
+In the situation where both models strongly agree on the relationship between water availability and occurrence probability, then we should expect significantly reduced uncertainty in that domain (Figs.
 \ref{fig:ex1_precip}, \ref{fig:ex1_map}).
-Because the meta-model includes both temperature and precipitation, any reduction in the uncertainty with respect to precipitation might also influence the evaluation of the relationship with temperature (reducing bias in the case where there is a correlation between T and P in the data).
+Because the metamodel includes both temperature and precipitation, any reduction in the uncertainty with respect to precipitation might also influence the evaluation of the relationship with temperature (reducing bias in the case where there is a correlation between T and P in the data).
 However, uncertainty could increase in the case of model disagreement.
-For instance, the relationship between precipitation and presence in the naive meta-model could be caused by a spurious relationship (e.g.
+For instance, the relationship between precipitation and presence in the naive metamodel could be caused by a spurious relationship (e.g.
 spatial autocorrelation caused by historical contingencies), leading to a false confidence in the species distribution over precipitation gradients.
@@
 Despite being a relatively common application of \ac{SDM}s \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios is problematic for a number of reasons previously discussed \citep{Araujo2006, Austin2011}.
-We used our method to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
+We used our metamodeling framework to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
 We use data from \citet{Morin2009} for sugar maple (\emph{Acer saccharum}), an economically and ecologically important species occurring in eastern North America.
@@
 \defcitealias{IPCC2001}{IPCC, 2001} %% this is needed to make the organizational citation in the following paragraph work correctly
-For the meta-model, we used a dataset of 1013 sugar maple presence records at 0.5 degree resolution, with 498 corresponding pseudo-absence records sampled from North America.
+For the metamodel, we used a dataset of 1013 sugar maple presence records at 0.5 degree resolution, with 498 corresponding pseudo-absence records sampled from North America.
 \mt{update numbers with final versions, and indicate that they are true absences}
@@
 Phenofit produces predictions at continental scales; thus the scaling function \(g\) was incorporated within Phenofit itself and we performed model integration directly on the outputs.
-To perform the integration, we used \ac{MCMC} to condition the parameters of the meta-model simultaneously on two likelihood functions.
+To perform the integration, we used \ac{MCMC} to condition the parameters of the metamodel simultaneously on two likelihood functions.
 The first is the naive SDM, where the likelihood of a ``success'' in the presence/pseudo-absence dataset was a Bernoulli density with the (logit-transformed) probability expressed as a linear function of the climate predictors.
 The sub-model likelihood was the Beta density of the observed Phenofit prediction for the present climate given the predicted suitability under the naive model. 
-Thus, the posterior probability of the meta-model was:
+Thus, the posterior probability of the metamodel was:
 \begin{equation}
@@
 \end{equation}
-where \(\theta_M\) is the meta-model, 
+where \(\theta_M\) is the metamodel, 
 \(X_M\) is the vector of presences and pseudo-absences, 
@@
 \(X_S\) is the prediction from Phenofit (in this case a single deterministic value, because we did not incorporate errors from the Phenofit modelling),
-and \(\theta_S\) is the model relating the Phenofit predictions to the meta-model.
+and \(\theta_S\) is the model relating the Phenofit predictions to the metamodel.
 Note that here, the integration is performed on the \emph{predictions} of the sub-model directly, rather than on compatible parameters as in Example 1.
-Thus, the meta-model must appear in the sub-model likelihood.
+Thus, the metamodel must appear in the sub-model likelihood.
 
-The naive model, despite being relatively simple (xx climate variables with xx total parameters), was highly fit to the data and consequently showed very high suitabilitiy scores near the core of the contemporary range with very little uncertainty (Fig \ref{fig:ex2}C, E).
-Similar patters were repeated for analogous climates under the future scenario (Fig \ref{fig:ex2}D, F).
+The naive model, despite being relatively simple (xx climate variables with xx total parameters), was highly fit to the data and consequently showed very high suitability scores near the core of the contemporary range with very little uncertainty (Fig \ref{fig:ex2}C, E).
+Similar patterns were repeated for analogous climates under the future scenario (Fig \ref{fig:ex2}D, F).
 However, standard errors were extreme for the range margins where suitability scores were intermediate.
-In comparison, Phenofit suitability scores were more moderate, and Phenofit predicted less northward movement of sugar maple under the future scenario \citep[Fig. \ref{fig:ex2}A, B,][]{Morin2009}.
+In comparison, Phenofit suitability scores were more moderate, and Phenofit predicted less northward movement of sugar maple under the future scenario \citep[Fig. \ref{fig:ex2}A, B;][]{Morin2009}.
 The integrated model demonstrated generally lower suitability scores throughout both the present and projected ranges, and thus represents lower confidence in the presence of sugar maple at any given location in its future projected range (Fig \ref{fig:ex2}G, H).
@@
 	\caption{Results of model integration for sugar maple, \emph{Acer saccharum}.
-	The mechanistic submodel Phenofit predicts present (A) and future (B) suitability using phenological information, but does not provide estimates of uncertainty.
+	The mechanistic sub-model Phenofit predicts present (A) and future (B) suitability using phenological information, but does not provide estimates of uncertainty.
 	Predictions from a naive metal-model for the present (C) were quite similar to the sub-model, but future predictions (D) were somewhat different, and uncertainty was extreme at the predicted range margins for both present (E) and future (F).
-	Applying the integration results in reduced suitabilities overall, possibly due to overfitting regions where the models disagree, but patterns in relative suitability largely reflected information from both models for both present (G) and future (H) climate.
-	Integrated model uncertainty for both present (I) and future (J) climates was greatly reduced near range margins but slightly increased in areas where uncertainty was low in the naive model.
+	Applying the integration results in reduced suitability overall, possibly due to overfitting regions where the models disagree, but patterns in relative suitability largely reflected information from both models for both present (G) and future (H) climate.
+	Integrated model uncertainty for both present (I) and future (J) climates is reduced near range margins but slightly increased in areas where uncertainty was low in the naive model.
 	}
@@
 
-For modelling the contemporary range of sugar maple, both Phenofit and the naive model perform very well relative to the integrated model, espeically when capturing the range margin (Fig \ref{fig:ex2}A, C, G).
+For modelling the contemporary range of sugar maple, both Phenofit and the naive model perform very well relative to the integrated model, espeicially when capturing the range margin (Fig \ref{fig:ex2}A, C, G).
 Conventional SDMs are quite good at producing empirical models of species distributions under present climatic regimes, and the naive model here could be further improved with the addition of more parameters or the inclusion of ensemble forecasting (See SI for model-fitting details).
@@
 The core range of the species under the integrated model generally reflects the regions of highest overlap between Phenofit and the naive model.
-Regions surrounding the core range generally show low suitabilities, but higher than those under either submodel, demonstrating reduced certainty about where the species will be absent in the future.
+Regions surrounding the core range generally show low suitabilities, but higher than those under either sub-model, demonstrating reduced certainty about where the species will be absent in the future.

### 2014-08-18 tex/introduction.tex
Added second batch of comments

@@
 
-Models of species range limits have wide applications and play a large role in conservation biology, where they are used as decision-making tools in biodiversity management \citep{Rosenzweig2008, Thuiller2008, Rodenhouse2009}.
+Models of species range limits have wide applications and play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Rosenzweig2008, Thuiller2008, Rodenhouse2009}.
 \wt{I would just put Guisan et al. 2013 here. }
-Due to large temporal and spatial scales and the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating range limits \citep{Wu1995, Levin1998}.
-Thus, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution, hybrid zone dynamics \citep{Engler2013}, and the impacts of climate change  \citep{Rosenzweig2008, Thuiller2008, Milad2011, Blois2013, Loyola2013}.
+\dm{Just a note to you that I still think people confuse or associate potential distribution (range?) modeling with actual distribution models.
+One could write a paper on the understanding, use and mis-use of these terms.
+I also think people often associate a purpose to a piece of work that mis-interprets the intended purpose }
+Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}.
+Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution, hybrid zone dynamics \citep{Engler2013}, and the impacts of climate change on species distributions \citep{Rosenzweig2008, Thuiller2008, Milad2011, Blois2013, Loyola2013}.
 \wt{after evolution above: Jay, F., Manel, S., Alvarez, N., Durand, E., Thuiller, W., Holderregger, R., Taberlet, P.
@@
 
-An important constraint on the modelling process is the availability of data and appropriate ecological theory to  link data to modelling objectives.
+An important constraint on modelling ecological processes is having the appropriate ecological theory needed to link data to modelling objectives. 
+Another constraint is the lacking of data over a while range of conditions.
 Poor data or poorly conceived models will result in predictions with unacceptably high uncertainty, or, more dangerously, precise but biased predictions.
@@
 }
-As a result of data accumulation from various sources and scales and conceptual advances in ecology, modelling techniques have proliferated to take advantage of what is available.
+In recent decades, however, modelling techniques have proliferated to take advantage of what is available in terms of both data and modeling platforms.
+\dm{Again just a note that I think people often do not realize the nature/problems with data.
+Weather station data is a classic example…lots of issues remain hence spatial clikmate models have problems; species observation data may have many issues from mis-identifcation to inaccurate geocoding.
+Maybe some of this this should be noted more explicitly later in the manuscript but I am not married to this … I don’t want to dwell on it}
 Model diversification also arises due to the diverse processes generating species ranges \citep{Soberon2007}.
-In the context of forecasting, which is the focus hereafter, mechanistic and correlative approaches represent quite different modes of inference, but both have strong limitations. 
-Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, may perform poorly when applied to species ranges.
+In the context of forecasting, which is the focus here, mechanistic and correlative approaches represent quite different modes of inference, but both have strong limitations. 
+Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, they may perform poorly when applied at the scale of species ranges.
 For instance, biotic interactions (including trophic interactions) are usually not modelled mechanistically at the scale of species distributions because they are not well-known, or have not been recorded, despite being considered a key determinant of range limits \citep{Soberon2007, Roux2012, Guo2013, Pigot2013}.
+\mjf{add citation: Holt et al 2005}
+\mjf{Holt, R. D., T.H. Keitt, M.A. Lewis, B.A. Maurer and M.L. Taper. 2005. Theoretical models of species’ borders: single species approaches. Oikos 108: 18-27.}
 On the other hand, coarse-scale correlative models statistically relating species occurrences to associated distributions of other variables at the same scale have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
@@
 Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
-Furthermore, correlative models often do not explicitly incorporate theory into their formulations, and when applied at coarse-scales, they may miss important fine-scale processes, making a multi-scale approach desirable when accurate forecasting is required \citep{Austin2007, Austin2011, Soranno2014}.
+Furthermore, correlative models often do not explicitly incorporate ecological processes into their formulations, and when applied at coarse scales, they may miss important fine-scale processes, making a multi-scale approach desirable when accurate forecasting is required \citep{Austin2007, Austin2011, Soranno2014}.
 \wt{remove all 3 cites, add \\
 Thuiller et al. 2013 ELE}
+\dm{add to end: On the other hand process-based models often require assumptions, many of which are poorly understood across space and time. }
 
@@
 A key problem with the diversity of approaches is that multiple models often produce differing predictions for the same organism, with no simple way to reconcile these predictions.
+\mjf{delete first sentence of this paragraph, it was already said above}
 \aa{Change to: \\
@@
 \aa{from here}
+\mjf{begin with: To integrate knowledge about species ranging from the stand to the landscape, we present an application...}
 We present an application of hierarchical Bayesian methods for integrating the predictions of multiple models.
@@
 The challenge is to account for all this available knowledge while producing a single prediction.
+\dm{change to: while producing predictions---I do have as bit of difficulty with this notion.
+Multiple predictions would seem to me to be quite reasonable given the range of understanding and issues with data.
+Reasonable people may differ if their understanding and beliefs hence multiple predictions that represent these may be desired.
+I am not sure my suggested wording in this sentence or the next is satisfactory.}
 Integrating multiple models to provide a single prediction would allow more knowledge to be used, potentially mitigating the limitations inherent in each modelling approach and contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
@@
 One approach is to integrate multiple models in a single modelling framework \citep{Buckley2010}. 
-Some researchers have explicitely reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Catterall2012, Strigul2012, Stewart-Koster2013, Soranno2014}.
+Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Catterall2012, Strigul2012, Stewart-Koster2013, Soranno2014}.
 Recent literature has also focused on the development of hybrid models, which allow for flexible combinations between mechanistic and phenomenological sub-models \citep{Gallien2010, Franklin2010, Thuiller2013}.
@@
 This complementarity has been used to merge the two types of models into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Boulangeat2014}. 
+\mjf{add citation: Naujokaitis-Lewis et al., 2013}
+\mjf{Naujokaitis-Lewis, I.R., J.M.R.
+Curtis, L.
+Tischendorf, D.
+Badzinski, K.
+Lyndsay, M.-J.
+Fortin.
+2013.
+Uncertainties in coupled species distribution–metapopulation dynamics models for risk assessments under climate change.
+Diversity and Distributions, 19: 541-554.}
 However, the link between different sub-models is based on additional assumptions that are not easy to test \citep{Gallien2010}. 
@@
 \ia{a bit vague here. Give an example?}
-Moreover, this approach is not suitable for merging models that are based on the same processes but with different underlying hypotheses.
+Moreover, hybrid approaches are not suitable for merging models that are based on the same processes but with different underlying hypotheses.
 \ia{an example?}
@@
 If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Serra-Diaz2013}.
-However, this approach is limited; for example, it is not possible to evaluate the effects of convergent predictions on total uncertainty of outcomes.
+However, the applicability of ensemble forecasts are limited; for example, it is not currently possible to evaluate the effects of convergent predictions on total uncertainty of outcomes.
 One of the greatest challenges inherent in both approaches above is estimating uncertainty, which is of utmost importance in a prediction context.
-A complete understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying parameters that make large contributions to the total uncertainty in the model predictions \citep{McMahon2011}.
+A more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying parameters that make large contributions to the total uncertainty in the model predictions \citep{McMahon2011}.
 In models that encompass multiple scales or levels of organization, it is particularly challenging to evaluate all sources of uncertainty, especially those related to model specification (\citealt{Calder2003} but see \citealt{Conlisk2013}).
@@
 
-Here, we present a framework for modelling range dynamics that aims to account for all the information available on a given question, from a variety of sources and scales.
-Contrary to hybrid and hierarchical models, the aim is not to link differents sub-models into a single model, but to condition the predictions of a meta-model at the target scale (e.g., an entire species' range) with information from sub-models at a variety of scales, allowing as much flexibility as possible regarding the type of information included.
+Here, we present a framework for modelling species range dynamics that aims to account for all the information available on a focal species, from a variety of sources and scales.
+Contrary to hybrid and hierarchical models, the aim is not to link different sub-models into a single model, but to condition the predictions of a meta-model at the target scale (e.g., an entire species' range) with information from sub-models at a variety of spatial scales, allowing as much flexibility as possible regarding the type of information included.
 We use the power and generality of a hierarchical Bayesian framework, which allows us to include multiple data sources and modes of inference \citep{Clark2005, VanOijen2005, Clark2006, Hobbs2011, Hartig2012}.
-Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to intertpret and reflects uncertainty at all levels of organization \citep{Clark2005, Cressie2009, Hobbs2011}. 
+Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to interpret and reflects uncertainty at all levels of organization \citep{Clark2005, Cressie2009, Hobbs2011}. 
 We illustrate our approach with two examples.

### 2014-08-18 tex/manuscript.tex
Added second batch of comments

@@
 %\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
-%\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
+\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
 \newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
-%\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
+\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
 %\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
@@
 \maketitle
+\mjf{Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework}
+\dm{I like this suggestion for a title change}
 \aa{although you eliminated the sentence in which you defined knowledge as data+theory - which contradicted the title - if I play the role of a reviewer I would still ask you why you differentiate between data and knowledge and how would you define knowledge.
@@
 We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
+\mjf{Multiple resolutions (same extent but different resolutions) and/or multiple scales (same resolution but different study extents).}
 The resulting meta-model produces probabilistic estimates of species presence with uncertainty that reflects error from all sub-models and data sources. 
@@
 \ia{need rewording here}
+
+\mjf{There is great interest in modeling species ranges resulting in a great diversity of modeling approaches.
+Yet each model uses only a small subset of the total available information about a species.
+For example, a broad-scale correlative model might use climate variables to predict presence or absence, but ignore what is known about smaller-scale processes such as the effect of local climate on growth, fecundity, and dispersal rates.
+Furthermore an inherent problem with the diversity of approaches is that different models often produce distinct predictions for the same species, with no simple way to reconcile these predictions.
+We present a flexible analytical framework for integrating models at multiple scales using hierarchical Bayesian methods.
+The resulting meta-model framework produces probabilistic estimates of species presence with uncertainty that reflects process-error from all sub-models as well as dataerror.
+These predictions reflect all of the information used as input for the original scale-specific sub-models.
+We illustrate the metamodeling approach using two forest modeling examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of some of the original data sources.
+We conclude by discussing the application of our metamodeling method and its accessibility to conservation biologists and land managers.
+Although this hierarchical Baysian method requires extensive statistical programming experience, we anticipate the results will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.}
+\dm{Again I like these edits; I think it provides a good tone. In the manuscript the tone at times felt a little too self confident ! I have made somme suggestions with this in mind.}
 \end{abstract}
@@
 %% ----------------------------------
-
+\dm{Attached are some thoughts as well.
+I think it is morphing into a nice paper and indeed a bit of review of sorts with the extent of literature cited.
+I did not have much to say (I liked Josee’s edits like the suggestion for the title change) but have made a few comments and editorial suggestions.
+One of my main thoughts (which perhaps is just a style suggestion) is that I think it needs to have a humble tone.
+Bayesian stuff is not as simple as I think you implying (for me anyway !!).
+Adoption will be challenging not just for researchers but also for decision-makers who might use the results.
+A bit more of a tip of the hat to its complexity is warranted (I acknowledge you do have a bit) is warranted in my opinion.
+At the end of the day you are the lead author so decisions like this will be your call.}
+\mjf{The introduction could be shorten to get to the point of the paper more quickly.
+While I like figure 6, I am not sure about the "adaptive management" part: is this figure really shows the links to management.}
 \input{introduction}

### 2014-08-18 tex/model_description.tex
Added second batch of comments

@@
 We begin with a simple correlative model (i.e. calibrated with data provided at the same ecological scale as predictions), characterized in a Bayesian framework to allow further integration. 
-We refer to this model (with associated parameters) as the meta-model, \(\theta_M\), and it is parameterized with information on the species' presence, \(X_M\), along with associated covariates (e.g., climate, presence/absence of interacting species, etc), collectively referred to as \(D_M\).
-The naive (i.e., not integrated with information from sub-models) probability that a species is present, \(\psi_N\), is a deterministic function of the meta-model and its covariates::
+We refer to this model (with associated parameters) as the metamodel, \(\theta_M\), and it is parameterized with information on the species' presence, \(X_M\), along with associated covariates (e.g., climate, presence/absence of interacting species, etc.), collectively referred to as \(D_M\).
+The naive (i.e., not integrated with information from sub-models) probability that a species is present, \(\psi_N\), is a deterministic function of the metamodel and its covariates::
 %-----
@@
 
-The role of the meta-model model is to integrate data at the same ecological scale of predictions.
+The role of the metamodel model is to integrate data at the same ecological scale of predictions.
 Additional sub-models require outputs that should be comparable to this given ecological scale (e.g. constraining presence or absence on the landscape).
@@
 
-For integration to be successful, it must be possible to compute the likelihood of this prediction given the meta-model, \(p \left( \psi_S \mid \theta_M ) = \theta_S \mid \theta_M, X_S, D_S \right)\), thus, we refer to $g$ as the \emph{scaling function}.
+For integration to be successful, it must be possible to compute the likelihood of this prediction given the metamodel, \(p \left( \psi_S \mid \theta_M ) = \theta_S \mid \theta_M, X_S, D_S \right)\), thus, we refer to $g$ as the \emph{scaling function}.
 In many cases, computing this likelihood will be challenging.

### 2014-08-19 tex/discussion.tex
Added last remaining comments

@@
 
+\fd{FD will work on this part: \\
+It would be good to compare this approach with the kind of one develop by Moorcroft et al.
+2001 : Moorcroft, P.
+R., Hurtt, G.
+C., \& Pacala, S.
+W.
+(2001).
+A method for scaling vegetation dynamics: the ecosystem demography model (ED).
+Ecological monographs, 71(4), 557-586.\\
+%Medvigy, D., Wofsy, S.
+C., Munger, J.
+W., Hollinger, D.
+Y., \& Moorcroft, P.
+R.
+(2009).
+Mechanistic scaling of ecosystem function and dynamics in space and time: Ecosystem Demography model version 2.
+Journal of Geophysical Research: Biogeosciences (2005–2012), 114(G1).\\
+Fisher, R., McDowell, N., Purves, D., Moorcroft, P., Sitch, S., Cox, P., \& Ian Woodward, F.
+(2010).
+Assessing uncertainties in a second-generation dynamic vegetation model caused by ecological scale limitations.
+New Phytologist, 187(3), 666-681.
+\\
+These models are multilevel and use scaling functions using statisitical moments series.
+}
+
 \subsection*{Advantages of Model Integration}
+\fd{I see that paragraph coming closer to the end of the paper in the accesibility section, maybe}
 Models are important tools that are increasingly used by land managers to face challenges associated to decision-making \citep{Guisan2013}
@@
 Model uncertainties is another key factor affecting applicability of model outputs (Addison et al., 2013). }
+\fd{I agree.
+That is A very important outcome that makes this approach more transparent than many other model approach.
+Eplxixitly demonstraiting model agreement and disagreement is an innovative feature.}
 Our approach also allows for a clear, transparent identification of uncertainties and how they are transmitted throughout the modelling process. 
@@
 The ease of incorporating new knowledge to the modelling framework (including new theory or the result of management and experimental efforts), will allow for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans.
+\fd{This framwork is in perfect accordance with the approach developed by Walters and Holling in Adaptive Management paradigm (Walters, C.
+J., \& Holling, C.
+S.
+(1990).
+Large-scale management experiments and learning by doing.
+Ecology, 71(6), 2060-2068.)\\
+Keith, D. A., Martin, T. G., McDonald-Madden, E., \& Walters, C. (2011). Uncertainty and adaptive management for biodiversity conservation. Biological Conservation, 144(4), 1175-1178.
+}
 In an era of continuous and unprecedented change, adaptive approaches such as the one presented here are often highlighted as a pressing need in order to develop strategies to promote ecosystems that are both feasible and resilient \citep[Fig.
@@
 }
+\fd{Moreover, overlapping scale domains of different sub-models showing divergent repsonse point out particular process/scale that require investigation for reconciling the sub-models or identify limits of use for each sub-models.}
 %==================
@@
 Model outputs may be considered too uncertain for decision-making \citep{Addison2013}. 
+\fd{That sentence seems incomplete or odd in regards of what has been said previously.}
 \tf{repetitive?}
@@
 \dm{Applications to demonstrate relevance; Use for productivity and abundance modeling rather than range delineation; Data development is also important}
+\fd{development of a post-hoc analytical procedure in multileveleld meta-models integration allowing to understand source of uncertainties and to test multiple intergration schemes.}

### 2014-08-19 tex/examples.tex
Added last remaining comments

@@
 In this context, the goal of modelling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, we recommend][for a concise introduction]{Link2010}:
+\fd{Bonne idée d'avoir introduit cette proposition ici.}
 \dm{add citation: McCarthy M.A. 2007 Bayesian methods for Ecology. Cambridge University Press 296p}
@@
 These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
+\fd{A word should be said about how the likelihood function balance the two sources of information for (Psi)M.  (How greater weight can be given to a more informative sub-model).}
 For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modelling process (due to the lack of additional mechanisms in the second model, as mentioned above).
+\fd{MAYBE NOT EXACTLY THERE:  "Moreover, convergent or divergent sub-models' response in ovelapping scale domains can  help in detecting crucial ecological processes."}
 
@@
 We used a binomial GLM to relate the suitability for sugar maple to 5 climate variables, scaled to the same resolution (see SI for a description of the climate variables used).
+fd{Present the abbreviation for SI= Supplemental Information earlier in the text (at the end of the previous paragraph, for example.)}
 \wt{We could also have used something more complex such as a GAM or boosted regression trees. }
@@
 and \(\theta_S\) is the model relating the Phenofit predictions to the metamodel.
+\fd{Shouldn't be Psi\_S here as we speak of prediction of the sub-model?}
 Note that here, the integration is performed on the \emph{predictions} of the sub-model directly, rather than on compatible parameters as in Example 1.
@@
 However, Phenofit and the naive model differ substantially in their predictions of the future distribution of the species.
-The integrated model presents a reasonable compromise between the two predictions. 
+The integrated model presents a compromise between the two predictions. 
 The core range of the species under the integrated model generally reflects the regions of highest overlap between Phenofit and the naive model.

### 2014-08-19 tex/introduction.tex
Added last remaining comments

@@
 An important constraint on modelling ecological processes is having the appropriate ecological theory needed to link data to modelling objectives.
+\fd{is maintaining the coherence between the data availability, the underlying ecological theory, the spatio-temporal scales and the applicability domain (while using it for forecasting).}
 Another constraint is the lacking of data over a while range of conditions.
 Poor data or poorly conceived models will result in predictions with unacceptably high uncertainty, or, more dangerously, precise but biased predictions.
+\fd{Incoherent model}
 \tf{Why is it dangerous? 
@@
 In a second example, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}) to climate with the predictions of a phenological model, Phenofit, with the goal of improving uncertainty estimation when projecting model predictions to future climate.
+
+\fd{figure 1 \\
+Shouldn't be:\\
+
+ "g(0s,Ds)" ther than g(0s, Xs, Ds) here?}

### 2014-08-19 tex/manuscript.tex
Added last remaining comments

@@
 \newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
+\newcommand{\fd}[1]{\todo[color=orange!30]{\textsuperscript{Frédérik D}#1}}
 %\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
@@
 \newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
-\newcommand{\wt}[1]{\todo[color=orange!40]{\textsuperscript{Wilfried T}#1}}
+\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
 \newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}
@@
 \author[1,2]{Alyssa Butler}
-%\author[10]{Fred Doyon} %??
-\author[6]{C. Ronnie Drever}
-\author[7]{Marie-Josée Fortin}
+\author[6,7]{Frédérik Doyon} %??
+\author[8]{C. Ronnie Drever}
+\author[9]{Marie-Josée Fortin}
 \author[1]{Tony Franceschini}
-\author[8]{Jean Liénard}
+\author[10]{Jean Liénard}
 \author[4]{Dan McKenney}
 \author[2,3]{Kevin A. Solarik}
-\author[8]{Nick Strigul}
-\author[9, 10]{Wilfried Thuiller}
+\author[10]{Nickolay Strigul}
+\author[11, 12]{Wilfried Thuiller}
 \author[1,2]{Dominique Gravel}
@@
 \affil[5]{Centre for Northern Studies, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
-\affil[6]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
-\affil[7]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
-\affil[8]{Department of Mathematics, Washington State University Vancouver, Washington, USA}
-\affil[9]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
-\affil[10]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
+\affil[6]{Université du Québec en Outaouais, Gatineau, Quebec, Canada}
+\affil[7]{Institut des Sciences de la Forêt Tempérée (ISFORT), Ripon, Quebec, Canada}
+\affil[8]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
+\affil[9]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
+\affil[10]{Department of Mathematics, Washington State University Vancouver, Washington, USA}
+\affil[11]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
+\affil[12]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
 
@@
 
+\section*{Acknowledgements}
+We acknowledge funding from the Quebec Centre for Biodiversity Science as well as NSERC strategic grant 430393-12.
+Comments from Charles Canham and William Godsoe improved a previous version of this manuscript.
+
 %% ----------------------------------

### 2014-08-19 tex/model_description.tex
Added last remaining comments

@@
 
-The role of the metamodel model is to integrate data at the same ecological scale of predictions.
+The role of the metamodel is to integrate data at the same ecological scale of predictions.
 Additional sub-models require outputs that should be comparable to this given ecological scale (e.g. constraining presence or absence on the landscape).
@@
 \ref{eq:sdm4}) as the new prior probability of $\theta_M$, and evaluate the likelihood of $\theta_M$ in light of the new data and the prior knowledge, as before:
+\fd{Shouldn't there be "Os" somewhere here?\\
+
+like:   "We simply use the posterior distibution of Om from the previous step (eq.11) as the new prior probability of Os, and evaluate the likelihood of Om in the ..."}
 %-----

### 2014-08-28 tex/discussion.tex
First round of edits for the next draft

@@
 \section*{Discussion}
-\wt{Honestly, I do not really think you need this last part (future work) given space limitation (watch out the total number of words and references).
-Instead I would rather develop some of the already written part that remains a bit vague and imprecise to me.
-I added few things to see what I mean.
-}
-\aa{Discussion: the section ‘Advantages of model integration’, which I wrote, and ‘Applicability of the method’ have several points in common.
-Maybe they could be merged into a single section: ‘Applicability and advantages of model integration’.
-Maybe, it should be discussed more extensively how to face the situation in which predictions of correlative model and sub-models are so different that the integrated model performs poorer than the two (I mean, does that mean that the approach is not valid in those cases?)
-I liked a lot the idea of adding Figure 6.
-The link between this approach and adaptive management seems to me one of its biggest advantages.
-The box in the right could maybe just say: 'monitoring of management actions’?
-}
 
 \subsection*{Comparison with other methods}
-\ab{The methods provided here expand upon the motivation of hybrid models to develop a more robust and unifying approach for ecological models in a predictive context while striving to overcome some of the limitations characteristic of other integrated approaches.
-Attempting to incorporate the underlying ecological processes so that SDMs are not oversimplified by combining different model parameters or scales can pose a challenge to these hybrid approaches. 
-In particular, it can often be difficult to identify parameters that can be used to connect different modelling frameworks and what parameters can produce purposeful responses \citep{Thuiller2013}. 
+The methods provided here expand upon the motivation of hybrid models to develop a more robust and unifying approach for ecological models in a predictive context while striving to overcome some of the limitations characteristic of other integrated approaches.
+Attempting to incorporate underlying ecological processes so that SDMs are not oversimplified by combining different model parameters or scales can be challenging using these hybrid approaches. 
+In particular, it can often be difficult to identify parameters that can be used to connect different modeling frameworks and what parameters can produce meaningful responses \citep{Thuiller2013}. 
 The methods presented in this study aim to overcome some of these limitations through the use of Bayesian statistics. 
-Bayesian methods allow for the inclusion of prior information and produce results as an entire distribution instead of a single point aiding in the reduction of error propagation \citep{Kery2010}. 
-The estimation of a response as an entire distribution helps to decrease over parameterization that is a concern for hybrid models and thereby also provides a more comprehensible response for reporting results.
-Bayesian frameworks provide a natural framework for the incorporation of prior information. 
+Bayesian methods produce results as an entire distribution instead of a single point, allowing for a comprehensive understanding of the uncertainty of estimates \citep{Link2010}. 
+The estimation of model parameters (and predictions) as distributions can also help decrease overparameterization that can be a concern for hybrid models. 
+Bayesian frameworks also provide a natural framework for the incorporation of prior information. 
 The inability or difficulty to include information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Thuiller2008, Smolik2010}. 
-Even when a link can be formed to include this information, it often is oversimplified to simplify computing \citep{Gallien2010}. 
-Bayesian methods therefore provide a reasonable alternative to allow for the easy unification of previous experimental studies or other prior information.
-Finally, Bayesian methods inherently allow for feedbacks or interactions between sub models, which may portray a more realistic response to many environmental dynamics where factors may simultaneously influence the change of one another.}
-\mt{Maybe change Kery citation to Link \& Barker}
+Even when a link can be formed to include this information, it often is oversimplified to simplify computation \citep{Gallien2010}. 
+Finally, Bayesian methods inherently allow for feedbacks or interactions between sub models, which may portray a more realistic response to many environmental dynamics where factors may simultaneously influence the change of one another.
 
-Our approach is an original application of Bayesian multilevel modelling, and can be considered a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales while using several models simultaneously.
-In particular, this development has certain similarities with Bayesian model averaging, Bayesian calibration of process-based models and hierarchical Bayesian modelling. 
+Our approach is an original application of Bayesian multilevel modeling, and can be considered a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales while using several models simultaneously.
+In particular, this development has certain similarities with Bayesian model averaging, Bayesian calibration of process-based models and hierarchical Bayesian modeling. 
 Bayesian model averaging aims to combine several alternative models to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.][]{Wintle2003, Link2006}.
 Bayesian model averaging considers models operating at the same scale and the posterior distribution is usually determined with Gibbs sampling. 
-Bayesian calibration of process-based model focuses on uncertainty of the parameter values in the process-based models, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, VanOijen2011}.
-\wt{Hartig, F., Dyke, J., Hickler, T., Higgins, S.I., O'Hara, R.B., Scheiter, S.
-\& Huth, A.
-(2012) Connecting dynamic vegetation models to data - an inverse perspective.
-Journal of Biogeography, 39, 2240-2252.}
+Bayesian calibration of process-based model focuses on uncertainty of the parameter values in the process-based models, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, VanOijen2011, Hartig2012}.
 In contrast with Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of a generally more correlative metamodel.
-Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005, Efendiev2005a}, plant physiology \citep{Ogle2008, Ogle2009}, and climate and atmospheric modelling \citep{Zimmerman2005, Mcmillan2010, Kang2012}.
+Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005, Efendiev2005a}, plant physiology \citep{Ogle2008, Ogle2009}, and climate and atmospheric modeling \citep{Zimmerman2005, Mcmillan2010, Kang2012}.
 \wt{Why not being more precise and exploratory here.
@@
 \subsection*{Advantages of Model Integration}
+\aa{Discussion: the section ‘Advantages of model integration’, which I wrote, and ‘Applicability of the method’ have several points in common.
+Maybe they could be merged into a single section: ‘Applicability and advantages of model integration’.
+Maybe, it should be discussed more extensively how to face the situation in which predictions of correlative model and sub-models are so different that the integrated model performs poorer than the two (I mean, does that mean that the approach is not valid in those cases?)}
 \fd{I see that paragraph coming closer to the end of the paper in the accesibility section, maybe}
@@
 This can create confusion and even some mistrust towards models, and as a consequence some managers may be discouraged to incorporate model results in their management plans. 
-By integrating different types of models---and their outputs---into a common framework, we believe that our approach can contribute to overcome the gap between modellers and practitioners and thus promote wider use of models to support decision-making.
+By integrating different types of models---and their outputs---into a common framework, we believe that our approach can contribute to overcome the gap between modelers and practitioners and thus promote wider use of models to support decision-making.
 \ia{This is repetitive with the applicability section. I suggest that you stick here on the modeling advantage of your approach and keep the applied advantage for the other section}
@@
 Model uncertainties is another key factor affecting applicability of model outputs (Addison et al., 2013). }
-\fd{I agree.
-That is A very important outcome that makes this approach more transparent than many other model approach.
-Eplxixitly demonstraiting model agreement and disagreement is an innovative feature.}
-Our approach also allows for a clear, transparent identification of uncertainties and how they are transmitted throughout the modelling process. 
-\tf{Really? I thought that only the uncertainty in the predictions are available and not the uncertainty from submodels.}
+Our approach also allows for a clear, transparent identification of uncertainties and how they are transmitted throughout the modeling process. 
 This is one of the main strengths of this approach, and offers several advantages. 
@@
 The new knowledge resulting from this research can be readily incorporated into the metamodel, allowing for an iterative learning process that will undoubtedly contribute to reduce uncertainty in predictions for a wide range of environmental conditions.
-\tf{The question is also: which model do we aim to incorporate. At a point, incorporating a model maybe useless. }
 Moreover, our framework embraces change as a fundamental process and is able to adapt and respond to it. 
-The ease of incorporating new knowledge to the modelling framework (including new theory or the result of management and experimental efforts), will allow for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans.
+The ease of incorporating new knowledge to the modeling framework (including new theory or the result of management and experimental efforts), will allow for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans.
 \fd{This framwork is in perfect accordance with the approach developed by Walters and Holling in Adaptive Management paradigm (Walters, C.
@@
 If you think about the Cheaib et al. paper, you may also suggest that your approach could be run over all different model outputs to integrate thema all together. 
-In a way that should re-inforce the link between modellers and long term experimental ecologists that did not really discuss so far since it was difficult to integrate the two informations. 
+In a way that should re-inforce the link between modelers and long term experimental ecologists that did not really discuss so far since it was difficult to integrate the two informations. 
 }
 \fd{Moreover, overlapping scale domains of different sub-models showing divergent repsonse point out particular process/scale that require investigation for reconciling the sub-models or identify limits of use for each sub-models.}
+
+\dm{Could write out the word Objectives in the Figure}
 %==================
@@
 \begin{figure}[tb]
-	\includegraphics{adaptive_management.pdf}
-	\caption{Sample workflow for applying the models presented in the first example in an adaptive management context.
+	\includegraphics{management.pdf}
+	\caption{Sample workflow for applying the models presented in the first example in a management context.
 	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
@@
 \end{figure}
-\dm{Could write out the word Objectives in the Figure}
 
@@
 %==================
+\aa{The box in the right could maybe just say: monitoring of management actions}
 
@@
 Although our approach is highly flexible and can be applied in a number of situations, there are some challenges to successfully using the framework.
-As with any modelling effort, good model specification with strong links to theory are essential \citep{Austin2007}.
+As with any modeling effort, good model specification with strong links to theory are essential \citep{Austin2007}.
 Poorly specified models will produce outputs that are uninformative or misleading, and model integration is not a cure for these problems.
@@
 A strong asset of our approach is that it can be used without the full suite of data that would be required to run a fully mechanistic model. 
-Given that the metamodel is correlative, it can be effectively implemented with, e.g., only presence-absence data. 
+Given that the metamodel is correlative, it can be effectively implemented with, e.g., only presence-absence data, or, in the case where true absences are difficult to obtain, with presences and pseudo absences (provided sufficient care is used in interpreting the results of such a model).
 Then any additional mechanistic data that are available will enhance predictions by constraining outputs of the metamodel.
-\dm{add: We note however that absence data can be problematic in some cases and naïve application of any modeling framework can be misleading.}
 
@@
 We have shown in our examples that even simple scaling functions can provide reasonable constraints on the metamodel.
-However, with the objective to take into account all know processes and models, it is likely that the modelling process proposed may concern more than one scaling function and not necessarily from the same scale.
+However, with the objective to take into account all know processes and models, it is likely that the modeling process proposed may concern more than one scaling function and not necessarily from the same scale.
 Indeed, species distribution is a function of e.g.
@@
 desired grain and extent of outputs) into model development.
-Successful use of an integrated modelling approach will always remain dependent on an intimate understanding of the decision-making process by modellers, emphasizing the importance of close collaboration between with practitioners at all stages of model development \citep{Guisan2013}.
-\tf{Just a stupid question. In the co-authors of this manuscript, is there any policy-maker and/or practitioners?}
+Successful use of an integrated modeling approach will always remain dependent on an intimate understanding of the decision-making process by modelers, emphasizing the importance of close collaboration between with practitioners at all stages of model development \citep{Guisan2013}.
 
@@
 Integrating sub-models allows for clear specification of desired model outputs (via the specification of the metamodel) while easily retaining important ecological objectives (via careful specification of sub-models).
-Insufficiently documented models are difficult to assess in real world situation, and therefore only have limited relevance for decision support \citep{Guisan2013}.
-Model workflow documentation becomes more crucial in the case of integrated modelling approaches, which incorporates information from various scales and resolutions within a number of aggregated sub-models.
+Insufficiently documented models are difficult to assess in real world situation, and therefore only have limited relevance for decision support \citep{Addison2013, Guisan2013}.
+Model workflow documentation becomes more crucial in the case of integrated modeling approaches, which incorporates information from various scales and resolutions within a number of aggregated sub-models.
 A transparent and well documented workflow describing the process of model integration ensures reproducibility and applicability (Fig. \ref{fig:adaptive_management}).
-The issue of uncertainty also limits the use of models in guiding decisions. 
-Model outputs may be considered too uncertain for decision-making \citep{Addison2013}. 
-\fd{That sentence seems incomplete or odd in regards of what has been said previously.}
-\tf{repetitive?}
-\ia{last 2 sentences are repetitive with ?? something was supposed to go here...}
-The transparent estimation of the uncertainty provided by this modelling approach may be a significant advantage.
 
@@
 }
-However, the same was true of all statistical techniques or modelling approaches when they were initially developed. 
+However, the same was true of all statistical techniques or modeling approaches when they were initially developed. 
 In addition, some of the computation costs that were associated with many techniques have now vanished, and even the conceptual challenges associated Bayesian statistics are being reduced as they gain recognition in the scientific literature.
@@
 \subsection*{Future directions}
-\wt{I have done a first round of comments on the MS.
-It looks very good to me.
-I have added some comments on the discussion.
-Honestly, I do not really think you need this last part (future work) given space limitation (watch out the total number of words and references).
-Instead I would rather develop some of the already written part that remains a bit vague and imprecise to me.
-I added few things to see what I mean.
-}
+\wt{Honestly, I do not really think you need this last part (future work) given space limitation (watch out the total number of words and references).}
+\mjf{Will work on this section}
 \begin{itemize}
@@
 	\item more sophisticated metamodels (e.g., state-transition models)
-	\item others?
+	\item Applications to demonstrate relevance; Use for productivity and abundance modeling rather than range delineation; Data development is also important
+	\item development of a post-hoc analytical procedure in multileveleld meta-models integration allowing to understand source of uncertainties and to test multiple intergration schemes.
 \end{itemize}
-\dm{Applications to demonstrate relevance; Use for productivity and abundance modeling rather than range delineation; Data development is also important}
-\fd{development of a post-hoc analytical procedure in multileveleld meta-models integration allowing to understand source of uncertainties and to test multiple intergration schemes.}

### 2014-08-28 tex/introduction.tex
First round of edits for the next draft

@@
 \section*{Introduction}
-\aa{I like it very much now, it's good and concise.
-The only thing is that the first and last sentences in the first paragraph seem contradictory.
-Do models of species range limits play a large role on conservation biology or do their contributions remain limited?
-Maybe in the first sentence you wanted to say they ‘can’ play a large role?
-}
-\aa{In addition, you may want to add this reference: Snell, R.S., Huth, A., Nabel, J.E.M.S., Bocedi, G., Travis, J.M.J., Gravel, D., Bugmann, H., Gutiérrez, A.G., Hickler, T., Higgins, S.I., Scherstjanoi, M., Reineking, B., Zurbriggen, N.
-\& Lischke, H.
-(2014) Using dynamic vegetation models to simulate plant range shifts.
-Ecography, In press.}
 
-Models of species range limits have wide applications and play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Rosenzweig2008, Thuiller2008, Rodenhouse2009}.
-\wt{I would just put Guisan et al. 2013 here. }
-\dm{Just a note to you that I still think people confuse or associate potential distribution (range?) modeling with actual distribution models.
-One could write a paper on the understanding, use and mis-use of these terms.
-I also think people often associate a purpose to a piece of work that mis-interprets the intended purpose }
+Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
 Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}.
-Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution, hybrid zone dynamics \citep{Engler2013}, and the impacts of climate change on species distributions \citep{Rosenzweig2008, Thuiller2008, Milad2011, Blois2013, Loyola2013}.
-\wt{after evolution above: Jay, F., Manel, S., Alvarez, N., Durand, E., Thuiller, W., Holderregger, R., Taberlet, P.
-\& François, O.
-(2012) Forecasting changes in population genetic structure of Alpine plants in response to global warming.
-Molecular Ecology, 21, 2354-2368.}
-\wt{The above mixes up subfields vs applications}
-\wt{for climate change citations:
-Thuiller, W., Lavergne, S., Roquet, C., Boulangeat, I. \& Araujo, M.B. (2011) Consequences of climate change on the Tree of Life in Europe. Nature, 470, 531-534. \\
-Thuiller, W., Pironon, S., Psomas, A., Barbet-Massin, M., Jiguet, F., S., L., Pearman, P.B., Renaud, J., Zupan, L.
-\& Zimmermann, N.E.
-(2014) The European functional tree of bird life in face of global change.
-Nature Communications, 5, 3118, DOI: 10.1038/ncomms4118.
-}
-However, despite having recognized potential to support decision making, the contributions of these models to conservation and applied ecology remain limited \citep{Dawson2011, Guisan2013}.
-\wt{remove Dawson}
+Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution \citep{Jay2012}, hybrid zone dynamics \citep{Engler2013}, and the impacts of climate change on species distributions \citep{Thuiller2011, Blois2013, Thuiller2014}.
+However, despite having recognized potential to support decision making, the contributions of these models to conservation and applied ecology remain limited \citep{Guisan2013}.
 
-An important constraint on modelling ecological processes is having the appropriate ecological theory needed to link data to modelling objectives.
-\fd{is maintaining the coherence between the data availability, the underlying ecological theory, the spatio-temporal scales and the applicability domain (while using it for forecasting).}
-Another constraint is the lacking of data over a while range of conditions.
-Poor data or poorly conceived models will result in predictions with unacceptably high uncertainty, or, more dangerously, precise but biased predictions.
-\fd{Incoherent model}
-\tf{Why is it dangerous? 
-It depends also of the importance of the bias.
-Some regression technique exists in order to deal multicolinearities and their basis is to produce precise and biased estimates.
-In this case, it is a way to reduce uncertainty.
-See e.g.
-ridge regression.
-With this kind of regression, there is no “danger”, and the estimates are known to be biased (and therefore should be interpreted and used as biased estimates).
-Plus, I feel that the word dangerous is not the correct word. There is ho danger in this, no harm. 
-}
-In recent decades, however, modelling techniques have proliferated to take advantage of what is available in terms of both data and modeling platforms.
-\dm{Again just a note that I think people often do not realize the nature/problems with data.
-Weather station data is a classic example…lots of issues remain hence spatial clikmate models have problems; species observation data may have many issues from mis-identifcation to inaccurate geocoding.
-Maybe some of this this should be noted more explicitly later in the manuscript but I am not married to this … I don’t want to dwell on it}
+An important constraint on modeling ecological processes is having the appropriate ecological theory needed to link data to modeling objectives, and further in maintaining coherence between the spatial and temporal scales of data and theory.
+Another constraint is a paucity of data over wide ranges of ecological conditions and scales.
+Poor data or incoherent models will result in predictions with unacceptably high uncertainty or precise but biased predictions.
+In recent decades, however, modeling techniques have proliferated to take advantage of what is available in terms of both data and modeling platforms.
 Model diversification also arises due to the diverse processes generating species ranges \citep{Soberon2007}.
@@
 Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, they may perform poorly when applied at the scale of species ranges.
-For instance, biotic interactions (including trophic interactions) are usually not modelled mechanistically at the scale of species distributions because they are not well-known, or have not been recorded, despite being considered a key determinant of range limits \citep{Soberon2007, Roux2012, Guo2013, Pigot2013}.
-\mjf{add citation: Holt et al 2005}
-\mjf{Holt, R. D., T.H. Keitt, M.A. Lewis, B.A. Maurer and M.L. Taper. 2005. Theoretical models of species’ borders: single species approaches. Oikos 108: 18-27.}
+For instance, biotic interactions (including trophic interactions) are usually not modeled mechanistically at the scale of species distributions because they are not well-known, or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Soberon2007, Roux2012, Guo2013, Pigot2013}.
 On the other hand, coarse-scale correlative models statistically relating species occurrences to associated distributions of other variables at the same scale have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
@@
 Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
-Furthermore, correlative models often do not explicitly incorporate ecological processes into their formulations, and when applied at coarse scales, they may miss important fine-scale processes, making a multi-scale approach desirable when accurate forecasting is required \citep{Austin2007, Austin2011, Soranno2014}.
-\wt{remove all 3 cites, add \\
-Thuiller et al. 2013 ELE}
-\dm{add to end: On the other hand process-based models often require assumptions, many of which are poorly understood across space and time. }
+Furthermore, correlative models often do not explicitly incorporate ecological processes into their formulations, and when applied at coarse scales, they may miss important fine-scale processes, making a multi-scale approach desirable when accurate forecasting is required \citep{Austin2011, Thuiller2013}.
+On the other hand process-based models often require a variety of assumptions which may be poorly understood across space and time.
 
-\subsection*{Toward an integrated approach for modelling ranges}
-A key problem with the diversity of approaches is that multiple models often produce differing predictions for the same organism, with no simple way to reconcile these predictions.
-\mjf{delete first sentence of this paragraph, it was already said above}
-\aa{Change to: \\
-In addition, an assortment of information (in the form of both data and theory)...}
-\aa{from here}
-\mjf{begin with: To integrate knowledge about species ranging from the stand to the landscape, we present an application...}
-We present an application of hierarchical Bayesian methods for integrating the predictions of multiple models.
+\subsection*{Toward an integrated approach for modeling ranges}
+To integrate knowledge about species ranging from the stand to the landscape, we present an application of hierarchical Bayesian methods for integrating the predictions of multiple models.
 Our approach is an alternative to other approaches to drawing inference from multiple models that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
-\ia{this is repetitive with the last paragraph of this section. \\
-delete everything up and and including "For instance..."}
-For instance, correlative models are often criticized because they assume that the selected explanatory variables limit (directly or indirectly) the species range, and, as mentioned above, that the relationship between the selected variables and the species occurrences will not change in the prediction context \citep{Araujo2006, Berteaux2006, Braunisch2013}.
-\aa{to here}
-However, additional information (in the form of both data and theory) is often available on the species of interest (e.g., growth rates in different climatic conditions, phenology, interactions with other species, sensitivity to disturbances) \citep{Holt2009, Thuiller2013}.
-The challenge is to account for all this available knowledge while producing a single prediction.
-\dm{change to: while producing predictions---I do have as bit of difficulty with this notion.
-Multiple predictions would seem to me to be quite reasonable given the range of understanding and issues with data.
-Reasonable people may differ if their understanding and beliefs hence multiple predictions that represent these may be desired.
-I am not sure my suggested wording in this sentence or the next is satisfactory.}
-Integrating multiple models to provide a single prediction would allow more knowledge to be used, potentially mitigating the limitations inherent in each modelling approach and contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
-
-One approach is to integrate multiple models in a single modelling framework \citep{Buckley2010}. 
+The challenge is to account for all this available knowledge while producing a single prediction, potentially mitigating the limitations inherent in each modeling approach and contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
 Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Catterall2012, Strigul2012, Stewart-Koster2013, Soranno2014}.
@@
 For instance, correlative models are used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while mechanistic approaches can include space and time dynamics and therefore account for dispersal processes \citep{Kearney2008}.
-This complementarity has been used to merge the two types of models into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Boulangeat2014}. 
-\mjf{add citation: Naujokaitis-Lewis et al., 2013}
-\mjf{Naujokaitis-Lewis, I.R., J.M.R.
-Curtis, L.
-Tischendorf, D.
-Badzinski, K.
-Lyndsay, M.-J.
-Fortin.
-2013.
-Uncertainties in coupled species distribution–metapopulation dynamics models for risk assessments under climate change.
-Diversity and Distributions, 19: 541-554.}
-However, the link between different sub-models is based on additional assumptions that are not easy to test \citep{Gallien2010}. 
-\aa{which assumptions? not clear}
-\ia{a bit vague here. Give an example?}
+This complementarity has been used to merge the two types of models into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Naujokaitis2013, Boulangeat2014}. 
+However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are not easy to test \citep{Gallien2010}. 
 Moreover, hybrid approaches are not suitable for merging models that are based on the same processes but with different underlying hypotheses.
-\ia{an example?}
 
 A second approach is to directly combine predictions.
-If models operate at the same scales and have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Araujo2007, Diniz-Filho2009}.
-\wt{add: Thuiller, W. (2004) Patterns and uncertainties of species' range shifts under climate change. Global Change Biology, 10, 2020-2027.\\
-remove: Diniz-Filho2009}
+If models operate at the same scales and have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}.
 If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Serra-Diaz2013}.
@@
 
-Here, we present a framework for modelling species range dynamics that aims to account for all the information available on a focal species, from a variety of sources and scales.
+Here, we present a framework for modeling species range dynamics that aims to account for all the information available on a focal species, from a variety of sources and scales.
 Contrary to hybrid and hierarchical models, the aim is not to link different sub-models into a single model, but to condition the predictions of a meta-model at the target scale (e.g., an entire species' range) with information from sub-models at a variety of spatial scales, allowing as much flexibility as possible regarding the type of information included.

### 2014-08-28 tex/manuscript.tex
First round of edits for the next draft

@@
 
-\title{Cross-scale integration of data and knowledge for predicting species ranges}
+\title{Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework}
 \date{}
@@
 
-%\listoftodos
 \maketitle
-\mjf{Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework}
-\dm{I like this suggestion for a title change}
-\aa{although you eliminated the sentence in which you defined knowledge as data+theory - which contradicted the title - if I play the role of a reviewer I would still ask you why you differentiate between data and knowledge and how would you define knowledge.
-Nothing too critical, though.}
 
 \begin{flushleft}
-\textbf{Keywords:} Uncertainty, scaling, range dynamics, trees, disturbances, patterns and processes, predict or understand, climate change, biotic interactions, spatial ecology, decision making, metamodelling, species distribution modelling
+\textbf{Keywords:} Uncertainty, scaling, range dynamics, trees, disturbances, patterns and processes, predict or understand, climate change, biotic interactions, spatial ecology, decision making, metamodeling, species distribution modeling
 \end{flushleft}
@@
 \begin{abstract}
-There is great interest in modelling species ranges, and such models are useful in guiding conservation and management decisions.
-There presently exists a great diversity of modelling approaches, and individual models often use only a small subset of the total available information about a species.
-\ia{Revision for first 2 sentences:\\
-Modelling species ranges, is of great interest in guiding conservation and management decisions.
-A wide diversity of modeling approaches are currently used, with individual models using only a small subset of the total available information about a species according to the scale and the specification of the model.
-}
-For example, a broad-scale correlative model might use climate variables to predict presence or absence, but ignore what is known about smaller-scale processes such as the effect of local climate on growth, fecundity, and dispersal rates.
-\ia{I would not put an example in the abstract}
-A central problem with the diversity of approaches is that multiple models produce differing predictions for the same organism, with no simple way to reconcile these predictions.
-\ia{This may not be seen as a problem as each model are focussing on a different aspect/have different objective/work at different scale. Maybe reword here?}
-We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
-\mjf{Multiple resolutions (same extent but different resolutions) and/or multiple scales (same resolution but different study extents).}
-The resulting meta-model produces probabilistic estimates of species presence with uncertainty that reflects error from all sub-models and data sources. 
-These predictions reflect all of the information used as input for the original sub-models. 
-We illustrate the approach through two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of some of the original data sources.
-Finally, we discuss the application of our method and its accessibility to conservation biologists and land managers.
-Although it requires extensive statistical programming experience, we anticipate the method will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.
-\ia{need rewording here}
-
-\mjf{There is great interest in modeling species ranges resulting in a great diversity of modeling approaches.
-Yet each model uses only a small subset of the total available information about a species.
+There is great interest in modeling species ranges, resulting in a great diversity of modeling approaches.
+However, in general, individual approaches include only a small subset of the total available information about a species. 
 For example, a broad-scale correlative model might use climate variables to predict presence or absence, but ignore what is known about smaller-scale processes such as the effect of local climate on growth, fecundity, and dispersal rates.
 Furthermore an inherent problem with the diversity of approaches is that different models often produce distinct predictions for the same species, with no simple way to reconcile these predictions.
-We present a flexible analytical framework for integrating models at multiple scales using hierarchical Bayesian methods.
-The resulting meta-model framework produces probabilistic estimates of species presence with uncertainty that reflects process-error from all sub-models as well as dataerror.
+We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
+The resulting meta-model framework produces probabilistic estimates of species presence with uncertainty that reflects error from data sources as well process error from all sub-models.
  These predictions reflect all of the information used as input for the original scale-specific sub-models.
-We illustrate the metamodeling approach using two forest modeling examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of some of the original data sources.
-We conclude by discussing the application of our metamodeling method and its accessibility to conservation biologists and land managers.
-Although this hierarchical Baysian method requires extensive statistical programming experience, we anticipate the results will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.}
-\dm{Again I like these edits; I think it provides a good tone. In the manuscript the tone at times felt a little too self confident ! I have made somme suggestions with this in mind.}
+We illustrate the metamodeling approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of some of the original data sources.
+We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
+Although this hierarchical Baysian method requires extensive statistical programming experience, we anticipate the results will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.
 \end{abstract}
@@
 %% ----------------------------------
-\dm{Attached are some thoughts as well.
-I think it is morphing into a nice paper and indeed a bit of review of sorts with the extent of literature cited.
-I did not have much to say (I liked Josee’s edits like the suggestion for the title change) but have made a few comments and editorial suggestions.
-One of my main thoughts (which perhaps is just a style suggestion) is that I think it needs to have a humble tone.
-Bayesian stuff is not as simple as I think you implying (for me anyway !!).
-Adoption will be challenging not just for researchers but also for decision-makers who might use the results.
-A bit more of a tip of the hat to its complexity is warranted (I acknowledge you do have a bit) is warranted in my opinion.
-At the end of the day you are the lead author so decisions like this will be your call.}
-\mjf{The introduction could be shorten to get to the point of the paper more quickly.
-While I like figure 6, I am not sure about the "adaptive management" part: is this figure really shows the links to management.}
+
 \input{introduction}
@@
 
-\input{brief_model_description}
-%---------------------------------------------
-%---------------------------------------------
-\input{examples}
+\input{modeling_framework}
 

### 2014-08-28 tex/modeling_framework.tex
First round of edits for the next draft

--- 2014-08-18 tex/examples.tex
@@
-\section*{Examples}
-\aa{Example 1: nothing to add, I think it is great as is. Figures 3 and 4 are very informative now. Great job! 
-Example 2: it is nicely exposed, although the results don't seem too promising. You said you got new, better results, so that’s good news. ;-)}
+\section*{Integrated modeling framework}
 
-Our modelling framework is best explained with concrete examples. 
-In this section we first present how experimental results could be introduced as prior knowledge of the fundamental niche in a species distribution model (\ac{SDM}), and in the following we apply our approach to integrating phenological information with occurrence data for a widespread North American tree.
-\ia{delete the above paragraph}
-\mt{I tend to agree}
+The key idea of our approach is to constrain the parameters of a correlative meta-model using predictions made by one or several mechanistic sub-models from other spatial scales.
+The role of the metamodel is to integrate data at the same ecological scale as predictions. 
+As an example, the metamodel could take the form of a simple correlative \ac{SDM}.
+Ideally, each sub-model incorporates different data types and hypotheses that specify underlying ecological mechanisms contributing to the model output.
+The output of sub-models should provide additional predictions at a comparable ecological scale at which the metamodel operates, although it is not necessary for the sub-models to actually operate at the same scale.
+For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, as we will show in the first example, these functions can be simple while still contributing useful information to the metamodel.
+All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Box 1, Fig. \ref{fig:diagram}).
+We present the complete framework in a hypothetical example where experimental results are  introduced as prior knowledge of the fundamental niche in a species distribution model (SDM).
+A more formal presentation of the modeling framework appears in Box 1.
+We follow with a second example where we apply our approach to integrate phenological information with occurrence data for a widespread North American tree.
+
+\begin{figure}
+\input{diagram}
+\caption{The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
+The metamodel ($\theta_M$) operates at a single ecological scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
+The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
+The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
+This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
+The final prediction can be obtained by applying the integrated model to the original explanatory variables (as shown) or, if projection is desired, on some new set of explanatory variables (e.g., future climate).
+This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).
+}
+\label{fig:diagram}
+\end{figure}
 
@@
 occurrence probability) of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
-The metamodel attempts to capture the realized distribution of a species, thereby encapsulating in a single correlative model the major physiological constraints and ecological processes constraining a species distribution.
-\mjf{How does it capture these constraints if it is a correlative model?}
+The metamodel attempts to capture the realized distribution of a species; as a correlative model, it implicitly captures the major physiological constraints and ecological processes constraining the distribution of the target species.
 However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the climatic envelope of where a species may be found in a natural setting.
@@
 For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature.
-\dm{May want to provide a bit more justification as to why this particular hypothetical example is useful?
-When I think of experimental data I think you are meaning lab-controlled data which is quite rare I think?}
 We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty. 
-Here, we focus only on the specification of the modelling framework; complete procedural details and code for generating the data sets and executing the model are provided as supplemental information.
+Here, we focus only on the specification of the modeling framework; complete procedural details and code for generating the data sets and executing the model are provided as supplemental information.
 
@@
 \end{equation}
-\dm{Just noting my initial comment that even absence data can be of course tricky eg structural zeros; a cryptic species; looked at the wrong time of day; some disturbance happened that year….etc….etc.
-I note your comment in the next sentence which is good}
 We further assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
@@
 We can fit this model in a Bayesian framework to allow for easy integration of models (as we will show later).
-In this context, the goal of modelling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, we recommend][for a concise introduction]{Link2010}:
-\dm{add citation: McCarthy M.A. 2007 Bayesian methods for Ecology. Cambridge University Press 296p}
-\dm{Mick McCarthy might be a good suggestion for a reviewer (I know him – an Aussie)}
+In this context, the goal of modeling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, concise introductions can be found in ][]{McCarthy2007, Link2010}:
 %-----
@@
 %-----
-where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modelling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
+where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
 
@@
 These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
-For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modelling process (due to the lack of additional mechanisms in the second model, as mentioned above).
+For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modeling process (due to the lack of additional mechanisms in the second model, as mentioned above).
 
@@
 %==================
-\dm{FIG 4: Again just a comment …one could discuss what does this probability of presence really mean?
-Presence with a self sustaining population.
-Someone could plant a specimen outside its historic natural range and expect it to live.
-I woder if rather than naïve the term "simple" would be better?
-I have substituted for that word in some places …Aren’t we all naïve ??!!}
 
@@
 The information from the sub-model will thus bring back the right amount of uncertainty in the relationship.
+Moreover, convergence in model predictions in overlapping domains (e.g., near the peak in the probability of presence in Fig. \ref{fig:ex1_precip}) can aid in detecting crucial ecological processes.
 
@@
 \defcitealias{IPCC2001}{IPCC, 2001} %% this is needed to make the organizational citation in the following paragraph work correctly
-For the metamodel, we used a dataset of 1013 sugar maple presence records at 0.5 degree resolution, with 498 corresponding pseudo-absence records sampled from North America.
-\mt{update numbers with final versions, and indicate that they are true absences}
-We used a binomial GLM to relate the suitability for sugar maple to 5 climate variables, scaled to the same resolution (see SI for a description of the climate variables used).
-\wt{We could also have used something more complex such as a GAM or boosted regression trees. }
-\mt{Perhaps note this; we selected this model for simplicity of explanation and implementation, but more flexible functions are possible}
-For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario (REF--Nakicenovic), and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
-\wt{citation:
-Nakicenovic, N. \& Swart, R. (ed.\^eds) (2000) Emissions Scenarios: A Special Report of Working Group III of the Intergovernmental Panel on Climate Change. Cambridge University Press, Cambridge.}
-\tf{Is it an optimistic or pessimistic or medium scenario? 
-Would it be interesting to compare two extreme scenarios? Would it change something to the demonstration of this manuscript?}
+For the meta-model, we used a dataset of 1013 presence and 13863 absence records, at 0.5 degree resolution, sampled from North America.
+We reserved one-third of the dataset for validation, and used the area under the receiver operating characters curve (AUC) for validation of contemporary predictions, and the coefficient of determination (\(R^2\)) as a measure of the fit between the future predictions of the metamodel and the Phenofit predictions.
+We used a binomial GLM to relate the suitability for sugar maple to 6 climate variables, scaled to the same resolution (see supplemental information for a descriptions of the climate variables and of the model selection process).
+We selected a GLM here for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
+For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario (Nakicenovic2000), and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
 Phenofit produces predictions at continental scales; thus the scaling function \(g\) was incorporated within Phenofit itself and we performed model integration directly on the outputs.
-To perform the integration, we used \ac{MCMC} to condition the parameters of the metamodel simultaneously on two likelihood functions.
-The first is the naive SDM, where the likelihood of a ``success'' in the presence/pseudo-absence dataset was a Bernoulli density with the (logit-transformed) probability expressed as a linear function of the climate predictors.
-The sub-model likelihood was the Beta density of the observed Phenofit prediction for the present climate given the predicted suitability under the naive model. 
-Thus, the posterior probability of the metamodel was:
+To perform the integration, we used \ac{MCMC} to condition the parameters of the metamodel  on two likelihood functions.
+The first is the naive SDM, where the likelihood of a ``success'' in the presence/absence dataset was a Bernoulli density with the (logit-transformed) probability expressed as a linear function of the climate predictors.
+For the sub-model, we considered that Phenofit, being a process-based model, may perform better when projecting into the future.
+Thus, we used the predicted future probability of presence as the basis for generating simulated future presence-absence datasets.
+This procedure inherently incorporates some uncertainty by randomly determining where the species is present (though uncertainty in the probabilities themselves could not be incorporated, as they were not available).
+Thus, the sub-model likelihood is simply the probability of observing a simulated dataset (that is randomly drawn from the phenofit predictions \(X_s\)),  given the predictions of the metamodel, resulting in the following posterior probability for the integrated metamodel:
 \begin{equation}
@@
 where \(\theta_M\) is the metamodel, 
-\(X_M\) is the vector of presences and pseudo-absences, 
+\(X_M\) is the vector of contemporary presences and absences, 
 \(D_M\) is the matrix of climate predictors,
-\(X_S\) is the prediction from Phenofit (in this case a single deterministic value, because we did not incorporate errors from the Phenofit modelling),
+\(X_S\) is the prediction from Phenofit,
 and \(\theta_S\) is the model relating the Phenofit predictions to the metamodel.
-Note that here, the integration is performed on the \emph{predictions} of the sub-model directly, rather than on compatible parameters as in Example 1.
+Note that the simulated datasets do not appear in this formulation, as they are only generated during estimation as a function of \(X_s\), and vary randomly between individual steps of the Markov chain.
+Furthermore, the integration is performed on the \emph{predictions} of the sub-model directly, rather than on compatible parameters as in Example 1. 
 Thus, the metamodel must appear in the sub-model likelihood.
 
-The naive model, despite being relatively simple (xx climate variables with xx total parameters), was highly fit to the data and consequently showed very high suitability scores near the core of the contemporary range with very little uncertainty (Fig \ref{fig:ex2}C, E).
+The naive model, despite being a relatively inflexible GLM based on 3 climatic variables
+relatively simple (xx climate variables with xx total parameters), was highly fit to the data and consequently showed very high suitability scores near the core of the contemporary range with very little uncertainty (Fig \ref{fig:ex2}C, E).
 Similar patterns were repeated for analogous climates under the future scenario (Fig \ref{fig:ex2}D, F).
@@
 %==================
-\tf{Fig 5: Is it realistic to make predictions as South as Mexico or In Cuba? 
-A question arises for me: to which spatial enveloppe should these models be used? Worldwide? Smaller enveloppe?
-This opens questions about species interactions.
-The predictions made are for sugar maple only without taking into account other speices.
-Sugr maple is present in EastAmrica.
-Is there another species which is only present in WesternAmerica ?
-Do we known how both species interact and who will “win” the competition game?
-I guess that the approach proposed is dedicated to answer these questions and that the figure are only here to present the distribution of sugar maple when only climate is changing.
-}
 
-For modelling the contemporary range of sugar maple, both Phenofit and the naive model perform very well relative to the integrated model, espeicially when capturing the range margin (Fig \ref{fig:ex2}A, C, G).
+For modeling the contemporary range of sugar maple, both Phenofit and the naive model perform very well relative to the integrated model, espeicially when capturing the range margin (Fig \ref{fig:ex2}A, C, G).
 Conventional SDMs are quite good at producing empirical models of species distributions under present climatic regimes, and the naive model here could be further improved with the addition of more parameters or the inclusion of ensemble forecasting (See SI for model-fitting details).
 However, Phenofit and the naive model differ substantially in their predictions of the future distribution of the species.
-The integrated model presents a reasonable compromise between the two predictions. 
+The integrated model presents a compromise between the two predictions. 
 The core range of the species under the integrated model generally reflects the regions of highest overlap between Phenofit and the naive model.

### 2014-08-29 tex/discussion.tex
Additional edits, as well as added first draft of proposal letter

@@
 
-\ia{suggest adding here:\\
-Model uncertainties is another key factor affecting applicability of model outputs (Addison et al., 2013). }
-Our approach also allows for a clear, transparent identification of uncertainties and how they are transmitted throughout the modeling process. 
-This is one of the main strengths of this approach, and offers several advantages. 
-For instance, transparency in uncertainty can be considered as a sort of sensitivity analysis, in which areas with large uncertainty can be detected and new experimental research or additional data collection can be designed (e.g., Example 1, Figs.
+Model uncertainty is another key factor affecting applicability of model outputs \citep{Addison2013}.
+One of the main strengths of our approach is that it allows for a clear, transparent identification of uncertainties and how they are transmitted throughout the modeling process. 
+Transparency in uncertainty can be considered as a sort of sensitivity analysis, in which areas with large uncertainty can be detected and new experimental research or additional data collection can be designed (e.g., Example 1, Figs.
 \ref{fig:ex1_sampling}, \ref{fig:ex1_precip}).
@@
 Moreover, our framework embraces change as a fundamental process and is able to adapt and respond to it. 
-The ease of incorporating new knowledge to the modeling framework (including new theory or the result of management and experimental efforts), will allow for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans.
-\fd{This framwork is in perfect accordance with the approach developed by Walters and Holling in Adaptive Management paradigm (Walters, C.
-J., \& Holling, C.
-S.
-(1990).
-Large-scale management experiments and learning by doing.
-Ecology, 71(6), 2060-2068.)\\
-Keith, D. A., Martin, T. G., McDonald-Madden, E., \& Walters, C. (2011). Uncertainty and adaptive management for biodiversity conservation. Biological Conservation, 144(4), 1175-1178.
-}
+The ease of incorporating new knowledge to the modeling framework (including new theory or the result of management and experimental efforts), will allow for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans \citep{Keith2010}.
 In an era of continuous and unprecedented change, adaptive approaches such as the one presented here are often highlighted as a pressing need in order to develop strategies to promote ecosystems that are both feasible and resilient \citep[Fig.
-\ref{fig:adaptive_management};][]{Seastedt2008, Mori2013}.
+\ref{fig:management};][]{Seastedt2008, Mori2013}.
 \wt{
@@
 }
-\fd{Moreover, overlapping scale domains of different sub-models showing divergent repsonse point out particular process/scale that require investigation for reconciling the sub-models or identify limits of use for each sub-models.}
-
-\dm{Could write out the word Objectives in the Figure}
-%==================
-% FIGURE
-
-\begin{figure}[tb]
-	\includegraphics{management.pdf}
-	\caption{Sample workflow for applying the models presented in the first example in a management context.
-	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
-	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the metamodel.}
-	\label{fig:adaptive_management}
-\end{figure}
-
-% ERUGIF
-%==================
-\aa{The box in the right could maybe just say: monitoring of management actions}
 
@@
 
-\tf{Revised this paragraph to the following}
-\tf{Finally, determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eqs.
-6, 7) is a critical point.
-In the context of integration models, the challenge is three fold: (i) which spatial and time scales, i.e.
-which processes, are to be considered, (ii) how to build scaling function in order to be consistent with the metamodel and (iii) how error and uncertainty are propagated from the finest sub-models scale to the largest metamodel scale.
-We argue that our proposed framework is able to easily deal with very different scale and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered.
-The building of scaling function deserve more investigation.
+Finally, determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eqs. \ref{eq:ex1_integrated}, \ref{eq:integrated2}) is a critical point.
+In the context of integrated models, the challenge is three fold: (i) which spatial and temporal scales (i.e.
+which processes), are to be considered, (ii) how to build a scaling function that is consistent with the metamodel and (iii) how error and uncertainty are propagated from the scales of the sub-models scale to that of the metamodel.
+Although we argue that our proposed framework is able to easily deal with very different scales and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered, the building of scaling function deserves further investigation.
 We have shown in our examples that even simple scaling functions can provide reasonable constraints on the metamodel. 
-However, with the objective to take into account all know processes and models, it is likely that the modeling process proposed may concern more than one scaling function and not necessarily from the same scale.
-Indeed, species distribution is a function of e.g.
-species growth rate (Sykes et al.
-1996, Guisan \& Zimmerman 2000) which involves processes at the individual (e.g.
+However, with the objective to take into account all know processes and models, it is likely that the modeling process may include multiple scaling functions, each at different scales. 
+Indeed, if species distributions are a function of, e.g.
+poopulation growth rate \citep{Sykes1996, Guisan2000}, they will involve processes at the individual (e.g.
 competition) or cellular (e.g.
 photosynthesis) scales.
-Such very large differences in spatial scales imply that simple functions will be inadequate requiring more traditional upscaling methods.
-Therefore, intermediate models and processes corresponding to additionnal data will be necessary: additionnal parameters inducing additionnal uncertainty have to be estimated before meaningful model integration will be possible.
-In addition, not only the spatial scale is to be considered, but also temporal scales.
-Most of time, this scale is taken into account in the spatial scale, as there is a strong concordance between them: the most rapid processes corresponded to the finest scale (e.g.
-photosynthesis is related to daylight) while longest processes are related to largest spatial scale (e.g.
-species niche).
-}
-\tf{Sykes, M.T., Prentice, I.C., Cramer, W., 1996.
-A Bioclimatic Model for the Potential Distributions of North European Tree Species Under Present and Future Climates.
-Journal of Biogeography 23, 203–233.}
-Finally, determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eqs.
-\ref{eq:ex1_integrated}, \ref{eq:integrated2}) will remain a significant challenge.
-We have shown in our examples that even simple scaling functions can provide reasonable constraints on the metamodel. 
-However, it is likely that with very large differences in scales (e.g., comparing cellular processes to landscapes), simple functions will be inadequate, requiring more traditional upscaling methods that combine explicit processes with additional data (and, necessarily, additional parameters) before meaningful model integration will be possible.
+Such very large differences in spatial scales imply that simple functions will be inadequate, requiring more traditional upscaling methods. 
+Therefore, intermediate models and processes will need to be introduced, and this will require additional parameters, assumptions, and data, and will add uncertainty to the final integrated model.
 
@@
 Incorporating information from multiple sources, particularly with respect to uncertainty,  fosters a connection between scientifically-generated knowledge and policy, and is therefore an important tool for adaptive management \citep[][Fig.
-\ref{fig:adaptive_management}; ]{Rehme2011}.
+\ref{fig:management}; ]{Rehme2011}.
 Such approaches are needed in formulating management plans for vulnerable species and ecosystems to avoid basing decisions on too-narrow subsets of the available information \citep{Dawson2011}.
@@
 Model workflow documentation becomes more crucial in the case of integrated modeling approaches, which incorporates information from various scales and resolutions within a number of aggregated sub-models.
-A transparent and well documented workflow describing the process of model integration ensures reproducibility and applicability (Fig. \ref{fig:adaptive_management}).
+A transparent and well documented workflow describing the process of model integration ensures reproducibility and applicability (Fig. \ref{fig:management}).
 
@@
 \mjf{Accessibility of what? Modeling? Data?}
+\mt{maybe remove the label and relocate to challenges}
 Results of models predicting species ranges are increasingly used by resource managers, conservation biologists, or forest ecologists for formulating or adjusting recommendations. 
@@
 In this context, practitioners may judge model usefulness on two criteria: (1) Does the model provide reliable predictions at the needed time and space scales, and (2) can the model be implemented given the available data and technical expertise?
-\tf{Practitioners may first want to validate models, no? This needs that practitioners are faithful to scientific results, which is not always the case.}
 Since most decision makers work at local space scales and follow specific time frames, modelers face the obvious challenge of providing detailed outputs while preserving implementation capacity.
@@
 However, developing the models requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming.
-\tf{This is needed to develop models. But not for using models?}
 In many cases, off-the-shelf software \citep[e.g.,][]{R, RJAGS} can adequately express the model likelihoods with minimal programming, but more complicated models will require the development of custom software.
-Thus, this framework will require significant investment in developing customized code for the samplers in order to actually estimate parameters.
-No doubt, the implementation of our proposed method is currently too technically challenging for many practitioners. 
-\mt{This seems a bit negative to me... maybe nicen it up a bit}
-\dm{There is a difference between practioners and Decision-makers.
-You might want to work something in like: Adoption will be no doubt influenced by the percieved complexity and and usefulness to both practitioners (species distribution modelers) and decision-makers (those who would use to results to support operational or policy-related decisions).
-}
+Thus, this framework will require significant investment in developing customized code for the samplers in order to actually estimate parameters, which may be technically challenging for some practitioners and may limit the adoption of these methods by both practitioners and decision-makers.
 However, the same was true of all statistical techniques or modeling approaches when they were initially developed. 
 In addition, some of the computation costs that were associated with many techniques have now vanished, and even the conceptual challenges associated Bayesian statistics are being reduced as they gain recognition in the scientific literature.
-\tf{I find this argument poor. This is not because a method is popular that it is a good one. Maybe recognition is better.}
 We therefore argue that our proposed approach as a strong potential for direct use in our real world where climate is quickly changing and conservation practices must be adjusted accordingly.
-\ia{This would be possible with a strengthen collaboration between modelers and practitioners}
+We further argue that forging stronger collaborations between modelers, decision-makers, and practitioners will improve the incorporation of this and other new methods into applied ecology and conservation.
 

### 2014-08-29 tex/figures.tex
Additional edits, as well as added first draft of proposal letter

--- 2014-08-28 tex/modeling_framework.tex
@@
-\section*{Integrated modeling framework}
+\newpage
+\section*{Figure Captions}
+\mt{captions are limited to 150 words}
 
-The key idea of our approach is to constrain the parameters of a correlative meta-model using predictions made by one or several mechanistic sub-models from other spatial scales.
-The role of the metamodel is to integrate data at the same ecological scale as predictions. 
-As an example, the metamodel could take the form of a simple correlative \ac{SDM}.
-Ideally, each sub-model incorporates different data types and hypotheses that specify underlying ecological mechanisms contributing to the model output.
-The output of sub-models should provide additional predictions at a comparable ecological scale at which the metamodel operates, although it is not necessary for the sub-models to actually operate at the same scale.
-For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, as we will show in the first example, these functions can be simple while still contributing useful information to the metamodel.
-All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Box 1, Fig. \ref{fig:diagram}).
-We present the complete framework in a hypothetical example where experimental results are  introduced as prior knowledge of the fundamental niche in a species distribution model (SDM).
-A more formal presentation of the modeling framework appears in Box 1.
-We follow with a second example where we apply our approach to integrate phenological information with occurrence data for a widespread North American tree.
+\begin{flushleft}
 
-\begin{figure}
-\input{diagram}
-\caption{The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
+%==================
+% FIGURE 1
+%% 175 words
+\textbf{Figure \ref{fig:diagram}:}
+	The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
 	The metamodel ($\theta_M$) operates at a single ecological scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
@@
 	This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).
-}
-\label{fig:diagram}
-\end{figure}
-
-%---------------------------------------------
-\subsection*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
-
-In this hypothetical example, we build a metamodel relating the distribution (i.e.
-occurrence probability) of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
-The metamodel attempts to capture the realized distribution of a species; as a correlative model, it implicitly captures the major physiological constraints and ecological processes constraining the distribution of the target species.
-However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the climatic envelope of where a species may be found in a natural setting.
-Prior information of the physiological constraints affecting species distribution is sometimes available, but usually too incomplete to perform a reasonable comparison with the realized distribution.
-For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature.
-We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty. 
-Here, we focus only on the specification of the modeling framework; complete procedural details and code for generating the data sets and executing the model are provided as supplemental information.
-
-We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation. 
-We will consider \(X_M\) (see Box 1, Fig. \ref{fig:diagram}) to be a vector of $n$ observations that takes the value of one to indicate presence and zero for absence:
-\begin{equation}
-X_M = \{x_{M,1}, x_{M,2}, \ldots, x_{M,n}\}
-\end{equation}
-We further assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
-The model is evaluated by relating these observations to environmental data \((D_M)\), which for the sake of the example consists of mean annual temperature $T_M$ and annual precipitation $P_M$. 
-For simplicity, we use a simple logistic model for the metamodel \((\theta_M)\) with a second order effect of temperature and precipitation. 
-This naive model (i.e., the metamodel with no constraints from sub-models) predicts  occurrence probability (\(\psi_N\)) as a function of the environment:
-%-----
-\begin{equation}
-\begin{aligned}
-	\psi_N &= f\left(\theta_M, D_M \right) \\
-	&= p \left (X_M = 1 \mid \theta_M, T_M, P_M \right) \\
-	&=\text{logit}^{-1}\left( \mathbf{\Theta_M} \mathbf{D_M} \right)
-\end{aligned}
-\end{equation}
-%-----
-where \(\mathbf{\Theta_M}\) is the parameter vector of the model \(\theta_M\), \(\mathbf{D_M} \) is the covariate matrix (i.e., \(T_M, P_M\), with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
-We can fit this model in a Bayesian framework to allow for easy integration of models (as we will show later).
-In this context, the goal of modeling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, concise introductions can be found in ][]{McCarthy2007, Link2010}:
-%-----
-\begin{equation}
-\label{eq:ex1_bayes}
-	p\left (\theta_M \mid X_M,T_M,P_M \right ) \propto 
-	p \left(X_M \mid \theta_M, T_M, P_M \right)
-	p \left(\theta_M \right)
-\end{equation}
-%-----
-where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
 
 %==================
-% FIGURE
-
-\begin{figure}[tb]
-	\includegraphics{ex1_sampling.pdf}
-	\caption{Two simulated data sets used to illustrate the model integration framework.
+% FIGURE 2
+%% 88 words
+\textbf{Figure \ref{fig:ex1_sampling}:}
+	Two simulated data sets used to illustrate the model integration framework.
 	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
@@
 	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
-	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.}
-	\label{fig:ex1_sampling}
-\end{figure}
-
-% ERUGIF
-%==================
-
-Thus far, we have considered only a single source of information to fit this model, and therefore the prior distribution \(p\left(\theta_M \right)\) from Eq. \ref{eq:ex1_bayes} is uninformative.
-As a secondary source of information, we will consider an experiment relating the population growth rate of the plant to manipulations to the precipitation regime, with results (but no raw data) available from the literature (Fig.
-\ref{fig:ex1_sampling}b).
-Furthermore, no information is available regarding the temperature regime for the experiment.
-Transplant experiments that evaluate performance beyond the range of a species are common, and represent a plausible scenario for model integration \citep{Hargreaves2014}.
-According to niche theory \citep{Holt2009}, the fundamental niche corresponds to the set of environmental conditions where the per capita intrinsic growth rate $r$ is positive.
-This concept gives us a reasonable model to fit the scaling function $g$ for our sub-model (Fig. \ref{fig:diagram}).
-If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, then for observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
-\begin{equation}
-	\psi_{S,i} = \int_0^\infty N \left(X_{S,i}, \sigma_{S,i} \right)
-\end{equation}
-where \(N\) is the Normal density function.
-We can then estimate the posterior distribution for the sub-model by fitting the relationship between \(\psi_S\) and precipitation \( \left( P_S \right) \) using Bayesian beta regression \citep{Ferrari2004}:
-%-----
-\begin{equation}
-\label{eq:ex1_thetas}
-	p\left (\theta_S \mid \psi_S,P_S \right ) \propto 
-	p \left(\psi_S \mid \theta_S, P_S \right)
-	p \left(\theta_S \right) \\
-\end{equation}
-%-----
-
-Although the two data sets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation).
-This scaling is quite simplistic, and would never be used to predict a species' range using this mechanistic sub-model in isolation.
-Despite the strong theoretical foundation (i.e., niche theory) linking the observations to presence-absence at large scales, the upscaled sub-model (i.e., the transformation of \(X_S\) and \(\sigma_S\) to \(\psi_S\)) is built on the strong hypothesis that the species will be present when $r>0$, neglecting other population dynamics constraints such as Allee effects and metapopulation dynamics.
-The fundamental niche is also incomplete, as we lack environmental variables such as temperature, as well as all other ecological processes responsible for the realized distribution.
-As such, it would be unwise to expect predictions from this model alone to resemble the actual distribution of the species; as a mechanistic model, it is simply too incomplete to predict distribution.
-As we will show, however, the information from this sub-model, when applied as a constraint on the metamodel, can result in improved predictions that incorporate the information within each model.
-
-We accomplish model integration by treating the posterior of \(\theta_S\) as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
-%-----
-\begin{equation}
-	\label{eq:ex1_integrated}
-	\overbrace{p(\theta_M \mid X_M, T_M, P_M, \theta_S, \psi_S, P_S)}^\text{integrated posterior}
-	\propto
-	\overbrace{p\left (\psi_S \mid \theta_S,P_S \right )}^{\substack{\text{new information} \\ \text{from sub-model}}}
-	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) P \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
-	\overbrace{p \left(\theta_S \right)}^{\substack{\text{prior for} \\ \text{sub-model}}}	
-\end{equation}
-%-----
-As before, the metamodel \(\theta_M\) can be used to predict the species distribution \((\psi_I)\).
-However, these predictions will not reflect only the presence-absence samples in \(X_M\), but will also reflect the information from \(\theta_S\), including all of the data sources used to produce this sub-model.
-In other words, with this evaluation, the set of parameters $\theta_M$ is most likely when it agrees with both the original data and predictions from the sub-model. 
-Note that in this particular example, because the sub-model is only contingent on precipitation, there is no integration on temperature. 
-Finally, we note the presence of marginal distributions for both models (i.e., \(p(\theta_M)\) and \(p(\theta_S)\)).
-These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
-For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modeling process (due to the lack of additional mechanisms in the second model, as mentioned above).
-
-We performed this procedure with our hypothetical data sets.
-When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data.
-Unsurprisingly given the quality of the data, the sub-model was highly precise, providing a fairly strong constraint when producing the integrated model.
-The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
-At the scale of the metamodel, considering both temperature and precipitation, we observed similar results, with reduced uncertainty in the predictions over the domain not covered by the presence-absence data (Fig.
-\ref{fig:ex1_map}).
+	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.
 
 %==================
-% FIGURE
-
-\begin{figure}[tb]
-	\includegraphics{ex1_precip.pdf}
-	\caption{Results of integration showing the probability of presence (\(\psi\)) when considering only precipitation.
+% FIGURE 3
+%%  120 words
+\textbf{Figure \ref{fig:ex1_precip}:}
+	Results of integration showing the probability of presence (\(\psi\)) when considering only precipitation.
 	(a) Naive model, using only presence-absence data.
@@
 	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
-	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.}
-	\label{fig:ex1_precip}
-\end{figure}
-% ERUGIF
-%==================
+	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.
 
 %==================
-% FIGURE
-
-\begin{figure}[tb]
-	\includegraphics[width=5.25in]{ex1_map.pdf}
-	\caption{Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
+% FIGURE 4
+%% 75 words
+\textbf{Figure \ref{fig:ex1_map}:}
+	Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
 	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
 	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
-	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.}
-	\label{fig:ex1_map}
-\end{figure}
-%
-% ERUGIF
-%==================
-
-Once the technical challenges of evaluating the parameters are overcome, the integration has several advantages over the both the naive parameterization of the metamodel and the mechanistic sub-model.
-In the situation where both models strongly agree on the relationship between water availability and occurrence probability, then we should expect significantly reduced uncertainty in that domain (Figs.
-\ref{fig:ex1_precip}, \ref{fig:ex1_map}).
-Because the metamodel includes both temperature and precipitation, any reduction in the uncertainty with respect to precipitation might also influence the evaluation of the relationship with temperature (reducing bias in the case where there is a correlation between T and P in the data).
-However, uncertainty could increase in the case of model disagreement.
-For instance, the relationship between precipitation and presence in the naive metamodel could be caused by a spurious relationship (e.g.
-spatial autocorrelation caused by historical contingencies), leading to a false confidence in the species distribution over precipitation gradients.
-The information from the sub-model will thus bring back the right amount of uncertainty in the relationship.
-Moreover, convergence in model predictions in overlapping domains (e.g., near the peak in the probability of presence in Fig. \ref{fig:ex1_precip}) can aid in detecting crucial ecological processes.
-
-\subsection*{Example 2: Constraining an SDM using phenological information}
-For the second example, we consider the problem of predicting the future distribution of a species following climate change.
-There is considerable interest in comparing correlative and mechanistic projections with respect to climate change \citep{Morin2009}, and correctly characterizing uncertainty is a critical aspect of this problem \citep{Cheaib2012}.
-Despite being a relatively common application of \ac{SDM}s \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios is problematic for a number of reasons previously discussed \citep{Araujo2006, Austin2011}.
-We used our metamodeling framework to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
-We use data from \citet{Morin2009} for sugar maple (\emph{Acer saccharum}), an economically and ecologically important species occurring in eastern North America.
-Here we describe briefly the dataset, methods, and the results of the analysis.
-Complete details, including all data and scripts to reproduce the analysis, are included as supplemental information.
-
-\defcitealias{IPCC2001}{IPCC, 2001} %% this is needed to make the organizational citation in the following paragraph work correctly
-For the meta-model, we used a dataset of 1013 presence and 13863 absence records, at 0.5 degree resolution, sampled from North America.
-We reserved one-third of the dataset for validation, and used the area under the receiver operating characters curve (AUC) for validation of contemporary predictions, and the coefficient of determination (\(R^2\)) as a measure of the fit between the future predictions of the metamodel and the Phenofit predictions.
-We used a binomial GLM to relate the suitability for sugar maple to 6 climate variables, scaled to the same resolution (see supplemental information for a descriptions of the climate variables and of the model selection process).
-We selected a GLM here for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
-For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario (Nakicenovic2000), and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
-Phenofit produces predictions at continental scales; thus the scaling function \(g\) was incorporated within Phenofit itself and we performed model integration directly on the outputs.
-To perform the integration, we used \ac{MCMC} to condition the parameters of the metamodel  on two likelihood functions.
-The first is the naive SDM, where the likelihood of a ``success'' in the presence/absence dataset was a Bernoulli density with the (logit-transformed) probability expressed as a linear function of the climate predictors.
-For the sub-model, we considered that Phenofit, being a process-based model, may perform better when projecting into the future.
-Thus, we used the predicted future probability of presence as the basis for generating simulated future presence-absence datasets.
-This procedure inherently incorporates some uncertainty by randomly determining where the species is present (though uncertainty in the probabilities themselves could not be incorporated, as they were not available).
-Thus, the sub-model likelihood is simply the probability of observing a simulated dataset (that is randomly drawn from the phenofit predictions \(X_s\)),  given the predictions of the metamodel, resulting in the following posterior probability for the integrated metamodel:
-\begin{equation}
-\label{eq:integrated2}
-	p( \theta_M \mid X_M, D_M, X_S, \theta_S )
-	\propto 
-	p( X_S \mid \theta_M, \theta_S )
-	p( Y_M \mid \theta_M, D_M ) 
-	p( \theta_M )
-	p( \theta_S )
-\end{equation}
-where \(\theta_M\) is the metamodel, 
-\(X_M\) is the vector of contemporary presences and absences, 
-\(D_M\) is the matrix of climate predictors,
-\(X_S\) is the prediction from Phenofit,
-and \(\theta_S\) is the model relating the Phenofit predictions to the metamodel.
-Note that the simulated datasets do not appear in this formulation, as they are only generated during estimation as a function of \(X_s\), and vary randomly between individual steps of the Markov chain.
-Furthermore, the integration is performed on the \emph{predictions} of the sub-model directly, rather than on compatible parameters as in Example 1. 
-Thus, the metamodel must appear in the sub-model likelihood.
-
-The naive model, despite being a relatively inflexible GLM based on 3 climatic variables
-relatively simple (xx climate variables with xx total parameters), was highly fit to the data and consequently showed very high suitability scores near the core of the contemporary range with very little uncertainty (Fig \ref{fig:ex2}C, E).
-Similar patterns were repeated for analogous climates under the future scenario (Fig \ref{fig:ex2}D, F).
-However, standard errors were extreme for the range margins where suitability scores were intermediate.
-In comparison, Phenofit suitability scores were more moderate, and Phenofit predicted less northward movement of sugar maple under the future scenario \citep[Fig. \ref{fig:ex2}A, B;][]{Morin2009}.
-The integrated model demonstrated generally lower suitability scores throughout both the present and projected ranges, and thus represents lower confidence in the presence of sugar maple at any given location in its future projected range (Fig \ref{fig:ex2}G, H).
-Standard errors at range margins were significantly lower (as these are regions where both the naive model and phenofit were in agreement about intermediate suitability scores), whereas standard errors were slightly higher at the core range (where Phenofit predicted lower suitability than the naive model) (Fig \ref{fig:ex2}I, J).
+	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.
 
 %==================
-% FIGURE
-
-\begin{figure}[tb]
-	\includegraphics[width=6in]{ex2.pdf}
-	\caption{Results of model integration for sugar maple, \emph{Acer saccharum}.
+% FIGURE 5
+%% 133 words
+\textbf{Figure \ref{fig:ex2}:}
+	Results of model integration for sugar maple, \emph{Acer saccharum}.
 	The mechanistic sub-model Phenofit predicts present (A) and future (B) suitability using phenological information, but does not provide estimates of uncertainty.
@@
 	Integrated model uncertainty for both present (I) and future (J) climates is reduced near range margins but slightly increased in areas where uncertainty was low in the naive model.
-	}
+
+%==================
+% FIGURE 6
+%% 65 words
+\textbf{Figure \ref{fig:management}:}
+	Sample workflow for applying the models presented in the first example in a management context.
+	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
+	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the metamodel.
+
+\end{flushleft}
+
+%==================
+% FIGURE 1
+
+\newpage
+\begin{figure}[p!]
+\input{diagram}
+\caption{}
+\label{fig:diagram}
+\end{figure}
+
+%==================
+% FIGURE 2
+\newpage
+\begin{figure}[p!]
+	\includegraphics{ex1_sampling.pdf}
+	\caption{}
+	\label{fig:ex1_sampling}
+\end{figure}
+
+%==================
+% FIGURE 3
+\newpage
+\begin{figure}[p!]
+	\includegraphics{ex1_precip.pdf}
+	\caption{}
+	\label{fig:ex1_precip}
+\end{figure}
+
+%==================
+% FIGURE 4
+\newpage
+\begin{figure}[p!]
+	\includegraphics[width=5.25in]{ex1_map.pdf}
+	\caption{}
+	\label{fig:ex1_map}
+\end{figure}
+
+%==================
+% FIGURE 5
+\newpage
+\begin{figure}[p!]
+	\includegraphics[width=6in]{ex2.pdf}
+	\caption{}
 	\label{fig:ex2}
 \end{figure}
-%
-% ERUGIF
+
 %==================
-
-For modeling the contemporary range of sugar maple, both Phenofit and the naive model perform very well relative to the integrated model, espeicially when capturing the range margin (Fig \ref{fig:ex2}A, C, G).
-Conventional SDMs are quite good at producing empirical models of species distributions under present climatic regimes, and the naive model here could be further improved with the addition of more parameters or the inclusion of ensemble forecasting (See SI for model-fitting details).
-However, Phenofit and the naive model differ substantially in their predictions of the future distribution of the species.
-The integrated model presents a compromise between the two predictions. 
-The core range of the species under the integrated model generally reflects the regions of highest overlap between Phenofit and the naive model.
-Regions surrounding the core range generally show low suitabilities, but higher than those under either sub-model, demonstrating reduced certainty about where the species will be absent in the future.
+% FIGURE 6
+\newpage
+\dm{Could write out the word Objectives in the Figure}
+\aa{The box in the right could maybe just say: monitoring of management actions}
+\begin{figure}[p!]
+	\includegraphics{management.pdf}
+	\caption{}
+	\label{fig:management}
+\end{figure}

### 2014-08-29 tex/manuscript.tex
Additional edits, as well as added first draft of proposal letter

@@
 
+\usepackage{setspace}
 \usepackage{natbib}
@@
 \usepackage{authblk} % adds some nice options for displaying the author list
+\renewcommand\Authsep{\protect\\}
+\renewcommand\Authands{\protect\\}
 
@@
 \newcommand{\fd}[1]{\todo[color=orange!30]{\textsuperscript{Frédérik D}#1}}
-%\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
+\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
 \newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
@@
 \newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
-%\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
+\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
 \newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
@@
 % \renewcommand{\ab}[1]{\relax}%
+% \renewcommand{\fd}[1]{\relax}%
 % \renewcommand{\rd}[1]{\relax}%
@@
 \date{}
-\author[1,2]{Matthew V. Talluto}
+\author[1,2,13]{Matthew V. Talluto (mtalluto@gmail.com)}
 \author[1,2]{Isabelle Boulangeat}
@@
 \author[1,2]{Alyssa Butler}
-\author[6,7]{Frédérik Doyon} %??
+\author[6,7]{Frédérik Doyon}
 \author[8]{C. Ronnie Drever}
@@
 \affil[12]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
+\affil[13]{Author for correspondance. Address: Departament de Biologie, chimie, et geographie, 300, Allée des Ursulines, Rimouski, Quebec G5L 3A1, Canada}
 
@@
 \begin{document}
-
+\doublespacing
 %% ----------------------------------
@@
 
+\begin{titlingpage}
 	\maketitle
 
+	\mt{need to include email addresses of all authors}
+
 	\begin{flushleft}
-\textbf{Keywords:} Uncertainty, scaling, range dynamics, trees, disturbances, patterns and processes, predict or understand, climate change, biotic interactions, spatial ecology, decision making, metamodeling, species distribution modeling
+	\textbf{Short title: } \mt{less than 45 chars including spaces}
+
+	\textbf{Type of article: } Ideas and Perspectives
+
+	\textbf{Word counts: } \mt{main text excludes abstract, acknowledgements, references, table and figure legends; limited to 7500 words}
+	\begin{itemize}
+		\item[] Abstract: 
+		\item[] Main Text: 
+		\item[] Boxes:
+	\end{itemize}
+
+	\textbf{Number of references: } \mt{limited to 80}
+
+	\textbf{Number of figures: } 6
+
+	\textbf{Number of tables: } 0
+
+	\textbf{Number of text boxes: } 1
+
+	\textbf{Keywords:} Climate change, decision making, metamodeling, patterns and processes, range dynamics, scaling, spatial ecology, species distribution modeling, uncertainty
 	\end{flushleft}
+\end{titlingpage}
 
 \begin{abstract}
+\mt{limited to 200 words, currently at 235}
 There is great interest in modeling species ranges, resulting in a great diversity of modeling approaches.
@@
 
-\input{model_description}
-
-%---------------------------------------------
-%---------------------------------------------
-
 \section*{Acknowledgements}
@@
 
+\input{figures}
+
+%% ----------------------------------
+%
+%     TEXT BOX
+%
+%% ----------------------------------
+
+\input{model_description}
+
+%---------------------------------------------
+%---------------------------------------------
+
 \end{document}

### 2014-08-29 tex/modeling_framework.tex
Additional edits, as well as added first draft of proposal letter

@@
 We follow with a second example where we apply our approach to integrate phenological information with occurrence data for a widespread North American tree.
-
-\begin{figure}
-\input{diagram}
-\caption{The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
-The metamodel ($\theta_M$) operates at a single ecological scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
-The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
-The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
-This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
-The final prediction can be obtained by applying the integrated model to the original explanatory variables (as shown) or, if projection is desired, on some new set of explanatory variables (e.g., future climate).
-This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).
-}
-\label{fig:diagram}
-\end{figure}
 
@@
 where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
-
-%==================
-% FIGURE
-
-\begin{figure}[tb]
-	\includegraphics{ex1_sampling.pdf}
-	\caption{Two simulated data sets used to illustrate the model integration framework.
-	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
-	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., \(-\)1--1).
-	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
-	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.}
-	\label{fig:ex1_sampling}
-\end{figure}
-
-% ERUGIF
-%==================
 
@@
 \ref{fig:ex1_map}).
-
-%==================
-% FIGURE
-
-\begin{figure}[tb]
-	\includegraphics{ex1_precip.pdf}
-	\caption{Results of integration showing the probability of presence (\(\psi\)) when considering only precipitation.
-	(a) Naive model, using only presence-absence data.
-Uncertainty increases dramatically when attempting to project beyond the scope of the source data (represented by the horizontal line above the bottom axis).
-	(b) Mechanistic model, using observations of an experiment to infer probability of presence.
-Predictions are highly precise due to high quality source data, but are likely to be overly optimistic given the simplicity in assuming only precipitation determines probability of presence (see text).
-	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
-	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.}
-	\label{fig:ex1_precip}
-\end{figure}
-% ERUGIF
-%==================
-
-%==================
-% FIGURE
-
-\begin{figure}[tb]
-	\includegraphics[width=5.25in]{ex1_map.pdf}
-	\caption{Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
-	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
-	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
-	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.}
-	\label{fig:ex1_map}
-\end{figure}
-%
-% ERUGIF
-%==================
 
@@
 
-%==================
-% FIGURE
-
-\begin{figure}[tb]
-	\includegraphics[width=6in]{ex2.pdf}
-	\caption{Results of model integration for sugar maple, \emph{Acer saccharum}.
-	The mechanistic sub-model Phenofit predicts present (A) and future (B) suitability using phenological information, but does not provide estimates of uncertainty.
-	Predictions from a naive metal-model for the present (C) were quite similar to the sub-model, but future predictions (D) were somewhat different, and uncertainty was extreme at the predicted range margins for both present (E) and future (F).
-	Applying the integration results in reduced suitability overall, possibly due to overfitting regions where the models disagree, but patterns in relative suitability largely reflected information from both models for both present (G) and future (H) climate.
-	Integrated model uncertainty for both present (I) and future (J) climates is reduced near range margins but slightly increased in areas where uncertainty was low in the naive model.
-	}
-	\label{fig:ex2}
-\end{figure}
-%
-% ERUGIF
-%==================
-
 For modeling the contemporary range of sugar maple, both Phenofit and the naive model perform very well relative to the integrated model, espeicially when capturing the range margin (Fig \ref{fig:ex2}A, C, G).

### 2014-08-29 tex/proposal_letter.tex
Additional edits, as well as added first draft of proposal letter

\documentclass[11pt]{letter}
\usepackage[margin=1in]{geometry}

\makeatletter
\renewcommand{\closing}[1]{\par\nobreak\vspace{\parskip}%
  \stopbreaks
  \noindent
  \ifx\@empty\fromaddress\else
  \hspace*{\longindentation}\fi
  \parbox{\indentedwidth}{\raggedright
       \ignorespaces #1\\[1\medskipamount]%
       \ifx\@empty\fromsig
           \fromname
       \else \fromsig \fi\strut}%
   \par}
\makeatother

\usepackage{fontspec} 
\defaultfontfeatures{Mapping=tex-text} % converts LaTeX specials (``quotes'' --- dashes etc.) to unicode
\newopentypefeature{Contextuals}{NoAlternate}{-calt} % enable turning off the crazy Q in Garamond Premier Pro
\usepackage{xunicode}
\setmainfont{Garamond Premier Pro}
\setsansfont[Scale=0.9]{Optima Regular}

\signature{Matthew V. Talluto, Isabelle Boulangeat, \& Dominique Gravel \\
	Université du Québec à Rimouski \\ 
	mtalluto@gmail.com}
\address{Matthew V. Talluto, Ph.D. \\
	Départament du Biologie, Chimie, et Géographie \\
	300, Allée des Ursulines \\
	Rimouski, Québec G5L 3A1, Canada \\
	mtalluto@gmail.com}

\begin{document}
\addfontfeatures{Contextuals=NoAlternate}   % turn off the crazy Q in Garamond Premier Pro

\begin{letter}{Marcel Holyoak \\ 
	Editor-in-chief, \emph{Ecology Letters} \\ 
	Department of Environmental Science and Policy \\
	University of California, Davis \\
	One Shields Avenue \\
	Davis, CA 95616, USA }

\opening{Dear Dr. Holyoak,}

We are pleased to submit a proposal for a manuscript titled ``Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework''. 
We would appreciate your consideration of our manuscript for the \emph{Ideas and Perspectives} section of \emph{Ecology Letters}.

%%% Proposals should be no more than 300 words long, describe the nature and novelty of the work, the contribution of the proposed article to the discipline, and the qualifications of the author(s) who will write the manuscript. Proposals should be sent to the Editorial Office

% Make this a bit nicer, maybe add some text from the intro including some refs if there is room
% Mention qualifications
% Mention all authors?
There is great interest in modeling species ranges, resulting in a great diversity of modeling approaches.
However, in general, individual approaches include only a small subset of the total available information about a species. 
For example, a broad-scale correlative model might use climate variables to predict presence or absence, but ignore what is known about smaller-scale processes such as the effect of local climate on growth, fecundity, and dispersal rates.
Furthermore an inherent problem with the diversity of approaches is that different models often produce distinct predictions for the same species, with no simple way to reconcile these predictions.
We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
The resulting meta-model framework produces probabilistic estimates of species presence with uncertainty that reflects error from data sources as well process error from all sub-models.
 These predictions reflect all of the information used as input for the original scale-specific sub-models.
We illustrate the metamodeling approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of some of the original data sources.
We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
Although this hierarchical Baysian method requires extensive statistical programming experience, we anticipate the results will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.
We believe that our manuscript represents an important step toward a more holistic approach for forecasting future species ranges.
Due to the importance of range shifts for both theoretical and applied ecology, the topic is of broad interest, and we hope you will find it of interest for the \emph{Ideas and Perspectives} section.
We will be ready to submit the manuscript within one month.

Thank you for considering out submission. We look forward to your reply.

\closing{Sincerely,}

\end{letter}
\end{document}

### 2014-09-01 tex/proposal_letter.tex
Edits to proposal letter

@@
 % Mention all authors?
+We present a novel application of hierarchical Bayesian methods to integrating the predictions of multiple models into a single metamodel of a species distribution.
 There is great interest in modeling species ranges, resulting in a great diversity of modeling approaches.
@@
 For example, a broad-scale correlative model might use climate variables to predict presence or absence, but ignore what is known about smaller-scale processes such as the effect of local climate on growth, fecundity, and dispersal rates.
-Furthermore an inherent problem with the diversity of approaches is that different models often produce distinct predictions for the same species, with no simple way to reconcile these predictions.
-We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
-The resulting meta-model framework produces probabilistic estimates of species presence with uncertainty that reflects error from data sources as well process error from all sub-models.
- These predictions reflect all of the information used as input for the original scale-specific sub-models.
+Furthermore, individual models often produce distinct predictions for the same species, with no simple way to reconcile these predictions.
+Our framework is highly flexible and can integrate information from sub-models operating a multiple spatial and temporal scales.
+The resulting meta-model framework produces probabilistic estimates of species presence that includes uncertainty from data sources as well as all sub-models.
+ These predictions reflect all of the information used as input for the original sub-models.
 We illustrate the metamodeling approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of some of the original data sources.
 We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
-Although this hierarchical Baysian method requires extensive statistical programming experience, we anticipate the results will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.
 We believe that our manuscript represents an important step toward a more holistic approach for forecasting future species ranges.

### 2014-09-17 tex/discussion.tex
Added some text for ex2 results

@@
 \subsection*{Future directions}
-\wt{Honestly, I do not really think you need this last part (future work) given space limitation (watch out the total number of words and references).}
-\mjf{Will work on this section}
-\begin{itemize}
-	\item incorporation of spatially explicit models into the framework? \citep[e.g.,][]{Fortin2012}
-	\item more sophisticated metamodels (e.g., state-transition models)
-	\item Applications to demonstrate relevance; Use for productivity and abundance modeling rather than range delineation; Data development is also important
-	\item development of a post-hoc analytical procedure in multileveleld meta-models integration allowing to understand source of uncertainties and to test multiple intergration schemes.
-\end{itemize}
+Hierarchical Bayesian modeling requires data at every scale to fit the parameters required to constrain one level with another.
+Moreover, beyond the needs for additional data, more knowledge of the important processes driving species range limits is needed to allow for extrapolation beyond the scope of the original datasets to address future changes due to climate and land-use changes.
+This knowledge can be obtained through a series of experimental manipulations and by substituting space for time data along environmental gradients.
+When experiments are possible, gaps in knowledge can be inferred using surrogate variables, theoretical models and spatially explicit models to explore the range of outcomes possible \citep{Fortin2012}.
+Another way to gain knowledge is to simplify species complexity into a series of states and to use state-transition models (e.g., Markov chain, semi-Markov chain, matrix projection model, integral projection model) which require more theoretical or coarser knowledge of species responses to biotic and abiotic changes.
+Although fewer states allow us to better capture the essence of species responses to environmental conditions, knowledge is still required to model feedback effects across processes and spatio-temporal scales.
+Hence we need to determine which processes are the most important at a given scale and how to weight their effects while scaling-up.
+Our meta-modeling approach can act as a springboard for incorporating imperfect knowledge of how multi-scale processes influence species ranges.
+By incorporating this knowledge into species distribution models, we can hopefully both reduce bias and more accurately assess model uncertainty when forecasting future changes to species ranges.

### 2014-09-17 tex/introduction.tex
Additional edits, as well as added first draft of proposal letter
Added some text for ex2 results

@@
 
-An important constraint on modeling ecological processes is having the appropriate ecological theory needed to link data to modeling objectives, and further in maintaining coherence between the spatial and temporal scales of data and theory.
+Two important constraints on modeling ecological processes prevent models from making unbiased predictions with acceptable uncertainty: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
 Another constraint is a paucity of data over wide ranges of ecological conditions and scales.
@@
 Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
+In cases where a selected explanatory variable indirectly accounts for a process (e.g., where the strength of biotic interactions are correlated with an environmental variable), extrapolation will produce biased predictions in temporal and spatial domains where the process and the environmental variable are decoupled.
 Furthermore, correlative models often do not explicitly incorporate ecological processes into their formulations, and when applied at coarse scales, they may miss important fine-scale processes, making a multi-scale approach desirable when accurate forecasting is required \citep{Austin2011, Thuiller2013}.
@@
 This complementarity has been used to merge the two types of models into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Naujokaitis2013, Boulangeat2014}. 
-However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are not easy to test \citep{Gallien2010}. 
+However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are not easy to test \citep{Gallien2010}, and uncertainties are very difficult to quantify and identify from its many different sources.
 Moreover, hybrid approaches are not suitable for merging models that are based on the same processes but with different underlying hypotheses.
@@
 In a second example, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}) to climate with the predictions of a phenological model, Phenofit, with the goal of improving uncertainty estimation when projecting model predictions to future climate.
-
-\fd{figure 1 \\
-Shouldn't be:\\
-
- "g(0s,Ds)" ther than g(0s, Xs, Ds) here?}

### 2014-09-17 tex/manuscript.tex
Added some text for ex2 results

@@
 
-%% most recent draft - nothing yet
-% IB
-% Dominique B
-% Ronnie D
-% MJF
-% Dan M
-% JL/NS
-% KS
-% Fred D
-
-% discussion - nothing yet
-% Ronnie D
-% MJF
-% Dan M?
-% Fred D
-% Kevin S
-
 % uncomment to disable individual authors
@@
 \author[1,2,13]{Matthew V. Talluto (mtalluto@gmail.com)}
-\author[1,2]{Isabelle Boulangeat}
-\author[3]{Aitor Ameztegui}
-\author[4]{Isabelle Aubin}
-\author[1,2,5]{Dominique Berteaux}
-\author[1,2]{Alyssa Butler}
-\author[6,7]{Frédérik Doyon}
-\author[8]{C. Ronnie Drever}
-\author[9]{Marie-Josée Fortin}
-\author[1]{Tony Franceschini}
-\author[10]{Jean Liénard}
-\author[4]{Dan McKenney}
-\author[2,3]{Kevin A. Solarik}
-\author[10]{Nickolay Strigul}
-\author[11, 12]{Wilfried Thuiller}
-\author[1,2]{Dominique Gravel}
+\author[1,2]{Isabelle Boulangeat (isabelle.boulangeat@gmail.com)}
+\author[3]{Aitor Ameztegui (ameztegui@gmail.com)}
+\author[4]{Isabelle Aubin (Isabelle.Aubin@NRCan-RNCan.gc.ca)}
+\author[1,2,5]{Dominique Berteaux (Dominique\_Berteaux@uqar.ca)}
+\author[1,2]{Alyssa Butler (ca.butler10@gmail.com)}
+\author[6,7]{Frédérik Doyon (Frederik.Doyon@uqo.ca)}
+\author[8]{C. Ronnie Drever (cdrever@TNC.ORG)}
+\author[9]{Marie-Josée Fortin (fortinmj@gmail.com)}
+\author[1]{Tony Franceschini (Tony.Franceschini@uqar.ca)}
+\author[10]{Jean Liénard (jean.lienard@gmail.com)}
+\author[4]{Dan McKenney (Dan.McKenney@nrcan-rncan.gc.ca)}
+\author[2,3]{Kevin A. Solarik (kevinsolarik@hotmail.com)}
+\author[10]{Nickolay Strigul (nick.strigul@vancouver.wsu.edu)}
+\author[11, 12]{Wilfried Thuiller (wilfried.thuiller@ujf-grenoble.fr)}
+\author[1,2]{Dominique Gravel (dominique\_gravel@uqar.ca)}
 \affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
@@
 	\maketitle
-
-	\mt{need to include email addresses of all authors}
 

### 2014-09-17 tex/modeling_framework.tex
Added some text for ex2 results

@@
 We selected a GLM here for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
-For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario (Nakicenovic2000), and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
+For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
 Phenofit produces predictions at continental scales; thus the scaling function \(g\) was incorporated within Phenofit itself and we performed model integration directly on the outputs.
@@
 This procedure inherently incorporates some uncertainty by randomly determining where the species is present (though uncertainty in the probabilities themselves could not be incorporated, as they were not available).
-Thus, the sub-model likelihood is simply the probability of observing a simulated dataset (that is randomly drawn from the phenofit predictions \(X_s\)),  given the predictions of the metamodel, resulting in the following posterior probability for the integrated metamodel:
+Thus, the sub-model likelihood is simply the probability, given the predictions of the metamodel, of observing a simulated dataset \(X_s\) that is randomly drawn from the phenofit predictions \(\psi_s\); i.e., \(X_s = f(\psi_s)\):
 \begin{equation}
 \label{eq:integrated2}
-	p( \theta_M \mid X_M, D_M, X_S, \theta_S )
+	p( \theta_M \mid X_M, D_M, \psi_s, \theta_S )
 	\propto 
-	p( X_S \mid \theta_M, \theta_S )
+	p( X_S = f(\psi_s)\mid \theta_M, \theta_S )
 	p( Y_M \mid \theta_M, D_M ) 
@@
 \(D_M\) is the matrix of climate predictors,
-\(X_S\) is the prediction from Phenofit,
 and \(\theta_S\) is the model relating the Phenofit predictions to the metamodel.
-Note that the simulated datasets do not appear in this formulation, as they are only generated during estimation as a function of \(X_s\), and vary randomly between individual steps of the Markov chain.
 Furthermore, the integration is performed on the \emph{predictions} of the sub-model directly, rather than on compatible parameters as in Example 1. 
@@
 
-The naive model, despite being a relatively inflexible GLM based on 3 climatic variables
-relatively simple (xx climate variables with xx total parameters), was highly fit to the data and consequently showed very high suitability scores near the core of the contemporary range with very little uncertainty (Fig \ref{fig:ex2}C, E).
-Similar patterns were repeated for analogous climates under the future scenario (Fig \ref{fig:ex2}D, F).
-However, standard errors were extreme for the range margins where suitability scores were intermediate.
-In comparison, Phenofit suitability scores were more moderate, and Phenofit predicted less northward movement of sugar maple under the future scenario \citep[Fig. \ref{fig:ex2}A, B;][]{Morin2009}.
-The integrated model demonstrated generally lower suitability scores throughout both the present and projected ranges, and thus represents lower confidence in the presence of sugar maple at any given location in its future projected range (Fig \ref{fig:ex2}G, H).
-Standard errors at range margins were significantly lower (as these are regions where both the naive model and phenofit were in agreement about intermediate suitability scores), whereas standard errors were slightly higher at the core range (where Phenofit predicted lower suitability than the naive model) (Fig \ref{fig:ex2}I, J).
-
-For modeling the contemporary range of sugar maple, both Phenofit and the naive model perform very well relative to the integrated model, espeicially when capturing the range margin (Fig \ref{fig:ex2}A, C, G).
-Conventional SDMs are quite good at producing empirical models of species distributions under present climatic regimes, and the naive model here could be further improved with the addition of more parameters or the inclusion of ensemble forecasting (See SI for model-fitting details).
-However, Phenofit and the naive model differ substantially in their predictions of the future distribution of the species.
-The integrated model presents a compromise between the two predictions. 
-The core range of the species under the integrated model generally reflects the regions of highest overlap between Phenofit and the naive model.
-Regions surrounding the core range generally show low suitabilities, but higher than those under either sub-model, demonstrating reduced certainty about where the species will be absent in the future.
+For present climates, the predictions of Phenofit (Fig \ref{fig:ex2}A) were largely similar to those of the naive SDM (Fig \ref{fig:ex2}C). 
+Furthermore, variance in the predictions of the naive model was relatively low throughout the spatial domain (Fig \ref{fig:ex2}E).
+Similar patterns were repeated for analogous climates under the future scenario (Fig \ref{fig:ex2}D); however these predictions differed considerably from the predictions of the mechanistic sub-model (Fig \ref{fig:ex2}B), and uncertainty for the projection was relatively high, particularly for eastern Canada (Fig \ref{fig:ex2}F).
+In general, Phenofit was predicted considerably less northward migration by 2080.
+The integrated model, which included both the presence-absence data as well as the Phenofit predictions for the future, made intermediate predictions in terms of northward range shifts for sugar maple (Fig \ref{fig:ex2}H).
+Globally, the integrated model expressed somewhat more uncertainty than the naive model (Fig \ref{fig:ex2}J), reflecting the divergent predictions of Phenofit and the naive model.
+However, in areas where the naive model performed worst (e.g., eastern Canada), the level of uncertainty in the integrated model was similar to the uncertainty in other locations.
+Unsurprisingly, integration resulted in worse predictions for the present distribution of the species (Fig \ref{fig:ex2}G) compared with either the naive model or Phenofit in isolation.
+Conventional SDMs are quite good at producing empirical models of species distributions under present climatic regimes, and the naive model here could be further improved with the addition of more parameters or the inclusion of ensemble forecasting.
+This is due to the integration focusing on future predictions; if integration was desired for the present range of the species (perhaps as a way to more accurately quantify uncertainty), a better approach would have been to use the present predictions from Phenofit.

### 2014-09-18 tex/manuscript.tex
Small edits

@@
 	\begin{flushleft}
-	\textbf{Short title: } \mt{less than 45 chars including spaces}
+
+	% less than 45 chars including spaces
+	\textbf{Short title: Integrated models of species ranges}
 
@@
 	\begin{itemize}
-		\item[] Abstract: 
+		\item[] Abstract: 200 % limited to 200 words
 		\item[] Main Text: 
-		\item[] Boxes:
+		% current rough estimates:
+		% intro: 1187
+		% modeling framework: 2459
+		% discussion: 2734
+		% total: 6380
+		\item[] Boxes: 517 % limited to 750 words per box
 	\end{itemize}
@@
 \begin{abstract}
-\mt{limited to 200 words, currently at 235}
+% limited to 200 words
 There is great interest in modeling species ranges, resulting in a great diversity of modeling approaches.
 However, in general, individual approaches include only a small subset of the total available information about a species. 
-For example, a broad-scale correlative model might use climate variables to predict presence or absence, but ignore what is known about smaller-scale processes such as the effect of local climate on growth, fecundity, and dispersal rates.
-Furthermore an inherent problem with the diversity of approaches is that different models often produce distinct predictions for the same species, with no simple way to reconcile these predictions.
+Correlative models often predict presence or absence as a function of climate, but ignore  smaller-scale processes such as growth, fecundity, and dispersal.
+Furthermore, the diversity of approaches often results in divergent predictions for the same species, with no simple way to reconcile these predictions.
 We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
-The resulting meta-model framework produces probabilistic estimates of species presence with uncertainty that reflects error from data sources as well process error from all sub-models.
- These predictions reflect all of the information used as input for the original scale-specific sub-models.
-We illustrate the metamodeling approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of some of the original data sources.
+The resulting metamodel produces probabilistic estimates of species presence, reports uncertainty from both data and process error from all sub-models, and incorporates all of the information used as input for the original scale-specific sub-models.
+We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
 We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
-Although this hierarchical Baysian method requires extensive statistical programming experience, we anticipate the results will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.
+Although implementation of our method can be technically challenging, we anticipate the results will be of wide interest, and we encourage collaboration between modelers and practitioners in applying our framework.
 \end{abstract}

### 2014-09-18 tex/model_description.tex
Additional edits, as well as added first draft of proposal letter
Small edits

--- 2014-08-18 tex/model_description.tex
@@
 
+\newpage
 \section*{Box 1: Model description}
-
+% Boxes are limited to 750 words
 We begin with a simple correlative model (i.e. calibrated with data provided at the same ecological scale as predictions), characterized in a Bayesian framework to allow further integration. 
@@
 
-For integration to be successful, it must be possible to compute the likelihood of this prediction given the metamodel, \(p \left( \psi_S \mid \theta_M ) = \theta_S \mid \theta_M, X_S, D_S \right)\), thus, we refer to $g$ as the \emph{scaling function}.
+For integration to be successful, it must be possible to compute the likelihood of this prediction given the metamodel, \(p \left( \psi_S \mid \theta_M \right) = p \left(\theta_S \mid \theta_M, X_S, D_S \right)\); the function \(g\) serves to transform the parameters of the sub-model to the scale of the metamodel.
+Thus, we refer to $g$ as the \emph{scaling function}.
 In many cases, computing this likelihood will be challenging.
@@
 We simply use the posterior distribution of $\theta_M$ from the previous step (Eq.
-\ref{eq:sdm4}) as the new prior probability of $\theta_M$, and evaluate the likelihood of $\theta_M$ in light of the new data and the prior knowledge, as before:
+\ref{eq:sdm4}) as the new prior probability of $\theta_M$, and evaluate the posterior probability of $\theta_M$ in light of the new data (i.e., information from the submodel \(\theta_S\)) and the prior knowledge, as before:
+
 %-----

### 2014-09-18 tex/discussion.tex
Changes to discussion and to proposal letter

@@
 Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005, Efendiev2005a}, plant physiology \citep{Ogle2008, Ogle2009}, and climate and atmospheric modeling \citep{Zimmerman2005, Mcmillan2010, Kang2012}.
-\wt{Why not being more precise and exploratory here.
-You may actually run a model averaging around your approach, may not you?
-Indeed, if we think to Ex 1 you could have several native metamodel posterior distributions, right?
-Or even a direct model averaging from them.
-As said, you could actually run something like biomod2 to get the distribution of values for each pixel given the models/algorithms and data and use them directly as a posterior into the integration.
-.
-The integration could also be done in the Ex 2 as far as I could see it. 
-It might nicely link your approach to the on-going move of SDM people towards ensemble modeling and forecasting. It also lonk to Bayesian model averaging. 
-If I was meant, I would say that a simple use of binomial GLM is a 10-years way back for most people where more 
-}
+Although our method as presented is distinct from Bayesian model averaging, it is fully compatible with these methods.
+Multiple metamodels could be averaged in this way if a consensus prediction from multiple metamodels (containing, e.g., different sub-model sets or different metamodel formulations).
+Furthermore, the metamodel can easily be incorporated into a broader multimodel inference scheme.
+For example, the metamodel can be used as one input in an ensemble forecasting scheme that also incorporates other common SDM formulations (e.g., Random Forest, MaxEnt).
 

### 2014-09-18 tex/proposal_letter.tex
Changes to discussion and to proposal letter

@@
 
+%%% will need to shorten this quite a bit
+
 % Make this a bit nicer, maybe add some text from the intro including some refs if there is room
@@
 % Mention all authors?
+
+%%%% Comments from dom:
+
+%- First, it gives the impression that it is only a methodological contribution while we go one step ahead.
+The conceptual contribution is overlooked in this paragraph and the first sentence should point at it.
+%
+%- Second, I wonder if it does not over emphasize the Bayesian approach. again, I think we go beyond Bayes stats, it's just the tool we use to accomplish our needs. 
+%
+%- Isa is right, the paragraph does not underline clearly how this paper differs from previous ones: how does it differ from model averaging?
+From reviews and Ideas papers by Clark on Bayes stats in the same journal?
+%
+%May be you can start with an issue that is more fundamental, the one of mechanistic vs empirical/correlative models.
+The problem of working at the right scale and of complex systems could also be a good entry point.
+
+%% From Isa:
+% Yes. in the argument, you may insist on that the proposition solve the problem of uncertainty propagation.
+
 We present a novel application of hierarchical Bayesian methods to integrating the predictions of multiple models into a single metamodel of a species distribution.
@@
 Due to the importance of range shifts for both theoretical and applied ecology, the topic is of broad interest, and we hope you will find it of interest for the \emph{Ideas and Perspectives} section.
-We will be ready to submit the manuscript within one month.
 
-Thank you for considering out submission. We look forward to your reply.
+% Suggestion from Isa for a novelty statement:
+Our contribution is the first to propose a conceptual and analytical framework that goes further than aggregating a posteriori different predictions from different models. 
+Based on Bayesian statistics, it provides a way to integrate all available knowledge gathered for a focal species at different ecological scale in order to provide more robust predictions of its future distributions.
+Moreover it is able not only to quantify uncertainties and their variablity in space, but also to propagate them from each piece of information we add to the framework.
+This is of considerable importance in order to identify the sources of uncertainties and the knowledge gaps, as well as to guide futur research.
+Finally, the framework allows integrating future knowledge step by step in order to refine the predictions, which makes its potential application for adaptive management very appealing.
+
+Thank you for considering our submission. We will be ready to submit the manuscript within one month. We look forward to your reply.
 

### 2014-09-19 tex/manuscript.tex
Mostly final changes

@@
 
-%\usepackage[margin=1in]{geometry}
+\usepackage[margin=1in]{geometry}
 
@@
 \usepackage{pdflscape}   % for figures in landscape orientation
+\usepackage[labelsep=none]{caption}
 
@@
 %% ----------------------------------
-\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
-
-\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
+%\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
+%
+%\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
 
@@
 %
-\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
-\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
-\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
-\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
-\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
-\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
-\newcommand{\fd}[1]{\todo[color=orange!30]{\textsuperscript{Frédérik D}#1}}
-\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
-\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
-\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
-\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
-\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
-\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
-\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
-\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}
+%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
+%\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
+%\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
+%\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
+%\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
+%\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
+%\newcommand{\fd}[1]{\todo[color=orange!30]{\textsuperscript{Frédérik D}#1}}
+%\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
+%\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
+%\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
+%\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
+%\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
+%\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
+%\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
+%\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}
 
@@
 
+%TC:ignore
 \begin{titlingpage}
@@
 
-	\textbf{Word counts: } \mt{main text excludes abstract, acknowledgements, references, table and figure legends; limited to 7500 words}
-	\begin{itemize}
-		\item[] Abstract: 200 % limited to 200 words
-		\item[] Main Text: 
-		% current rough estimates:
-		% intro: 1187
-		% modeling framework: 2459
-		% discussion: 2734
-		% total: 6380
-		\item[] Boxes: 517 % limited to 750 words per box
-	\end{itemize}
+	% abstract limited to 200 words
+	% main text excludes abstract, acknowledgements, references, table and figure legends; limited to 7500 words
+	% boxes limited to 750 words each
+	% figure captions limited to 150 words each
+	\textbf{Word counts: } Abstract: 200; Main Text: 6091; Boxes: 517
 
-	\textbf{Number of references: } \mt{limited to 80}
+	\textbf{Number of references: } % limited to 80
 
@@
 
+	\textbf{Authorship:} All authors contributed substantially to an initial outline of the paper and ideas.
+MT, IB, and DG built the model and wrote the first draft of the introduction and model description.
+All authors contributed to the complete first draft and all subsequent revisions to the manuscript.
+
 	\textbf{Keywords:} Climate change, decision making, metamodeling, patterns and processes, range dynamics, scaling, spatial ecology, species distribution modeling, uncertainty
@@
 \end{titlingpage}
+%TC:endignore
 
+%TC:break abstract
 \begin{abstract}
@@
 
+%TC:break manuscript
 %% ----------------------------------
 %
-%     REST OF DOCUMENT
+%     INTRODUCTION
 %
@@
 
-\input{introduction}
+\section*{Introduction}
 
+Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
+Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}.
+Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution \citep{Jay2012}, hybrid zone dynamics \citep{Engler2013}, and the impacts of climate change on species distributions \citep{Thuiller2011, Blois2013, Thuiller2014}.
+However, despite having recognized potential to support decision making, the contributions of these models to conservation and applied ecology remain limited \citep{Guisan2013}.
+
+Two important constraints on modeling ecological processes prevent models from making unbiased predictions with acceptable uncertainty: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
+Another constraint is a paucity of data over wide ranges of ecological conditions and scales.
+Poor data or incoherent models will result in predictions with unacceptably high uncertainty or precise but biased predictions.
+In recent decades, however, modeling techniques have proliferated to take advantage of what is available in terms of both data and modeling platforms.
+Model diversification also arises due to the diverse processes generating species ranges \citep{Soberon2007}.
+In the context of forecasting, which is the focus here, mechanistic and correlative approaches represent quite different modes of inference, but both have strong limitations. 
+Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, they may perform poorly when applied at the scale of species ranges.
+For instance, biotic interactions (including trophic interactions) are usually not modeled mechanistically at the scale of species distributions because they are not well-known, or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Soberon2007, Pigot2013}.
+On the other hand, coarse-scale correlative models statistically relating species occurrences to associated distributions of other variables at the same scale have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
+However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}.
+Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
+In cases where a selected explanatory variable indirectly accounts for a process (e.g., where the strength of biotic interactions are correlated with an environmental variable), extrapolation will produce biased predictions in temporal and spatial domains where the process and the environmental variable are decoupled.
+Furthermore, correlative models often do not explicitly incorporate ecological processes into their formulations, and when applied at coarse scales, they may miss important fine-scale processes, making a multi-scale approach desirable when accurate forecasting is required \citep{Thuiller2013}.
+On the other hand process-based models often require a variety of assumptions which may be poorly understood across space and time.
+
+%% ----------------------------------
+%% ----------------------------------
+%% ----------------------------------
+
+\subsection*{Toward an integrated approach for modeling ranges}
+To integrate knowledge about species ranging from the stand to the landscape, we present an application of hierarchical Bayesian methods for integrating the predictions of multiple models.
+Our approach is an alternative to other approaches to drawing inference from multiple models that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
+The challenge is to account for all this available knowledge while producing a single prediction, potentially mitigating the limitations inherent in each modeling approach and contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
+Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Catterall2012, Strigul2012, Soranno2014}. 
+Recent literature has also focused on the development of hybrid models, which allow for flexible combinations between mechanistic and phenomenological sub-models \citep{Gallien2010, Thuiller2013}. 
+For instance, correlative models are used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while mechanistic approaches can include space and time dynamics and therefore account for dispersal processes \citep{Kearney2008}.
+This complementarity has been used to merge the two types of models into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Naujokaitis2013, Boulangeat2014}. 
+However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are not easy to test \citep{Gallien2010}, and uncertainties are very difficult to quantify and identify from its many different sources.
+Moreover, hybrid approaches are not suitable for merging models that are based on the same processes but with different underlying hypotheses.
+
+A second approach is to directly combine predictions.
+If models operate at the same scales and have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}.
+If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Serra-Diaz2013}.
+However, the applicability of ensemble forecasts are limited; for example, it is not currently possible to evaluate the effects of convergent predictions on total uncertainty of outcomes.
+One of the greatest challenges inherent in both approaches above is estimating uncertainty, which is of utmost importance in a prediction context.
+A more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying parameters that make large contributions to the total uncertainty in the model predictions \citep{McMahon2011}.
+In models that encompass multiple scales or levels of organization, it is particularly challenging to evaluate all sources of uncertainty, especially those related to model specification (\citealt{Calder2003} but see \citealt{Conlisk2013}).
+Clearly, an approach that integrates predictions and uncertainty from multiple model types is needed \citep{Beck2012, Thuiller2013}. 
+Such an approach would incorporate multiple modes of inference and integrate data from multiple sources and scales \citep{Levin1992, Peters2004, Thuiller2013}.
+
+Here, we present a framework for modeling species range dynamics that aims to account for all the information available on a focal species, from a variety of sources and scales.
+Contrary to hybrid and hierarchical models, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from sub-models at a variety of spatial scales, allowing as much flexibility as possible regarding the type of information included.
+We use the power and generality of a hierarchical Bayesian framework, which allows us to include multiple data sources and modes of inference \citep{Clark2005, VanOijen2005, Clark2006, Hobbs2011, Hartig2012}.
+Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to interpret and reflects uncertainty at all levels of organization \citep{Clark2005, Cressie2009, Hobbs2011}. 
+We illustrate our approach with two examples.
+First, we use simulated data to present a complete example of the application of the framework to multiple information sources that are relevant at very different scales.
+In a second example, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}) to climate with the predictions of a phenological model, Phenofit, with the goal of improving uncertainty estimation when projecting model predictions to future climate.
+
+%% ----------------------------------
+%
+%     MODELING FRAMEWORK
+%
+%% ----------------------------------
+
+\section*{Integrated modeling framework}
+
+The key idea of our approach is to constrain the parameters of a correlative metamodel using predictions made by one or several mechanistic sub-models from other spatial scales.
+The role of the metamodel is to integrate data at the same ecological scale as predictions. 
+As an example, the metamodel could take the form of a simple correlative \ac{SDM}.
+Ideally, each sub-model incorporates different data types and hypotheses that specify underlying ecological mechanisms contributing to the model output.
+The output of sub-models should provide additional predictions at a comparable ecological scale at which the metamodel operates, although it is not necessary for the sub-models to actually operate at the same scale.
+For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, as we will show in the first example, these functions can be simple while still contributing useful information to the metamodel.
+All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Box 1, Fig. \ref{fig:diagram}).
+We present the complete framework in a hypothetical example where experimental results are  introduced as prior knowledge of the fundamental niche in a species distribution model (SDM).
+A more formal presentation of the modeling framework appears in Box 1.
+We follow with a second example where we apply our approach to integrate phenological information with occurrence data for a widespread North American tree.
+
+%% ----------------------------------
+%
+%     EXAMPLE 1
+%
+%% ----------------------------------
+\subsection*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
+
+In this hypothetical example, we build a metamodel relating the distribution (i.e.
+occurrence probability) of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
+The metamodel attempts to capture the realized distribution of a species; as a correlative model, it implicitly captures the major physiological constraints and ecological processes constraining the distribution of the target species.
+However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the climatic envelope of where a species may be found in a natural setting.
+Prior information of the physiological constraints affecting species distribution is sometimes available, but usually too incomplete to perform a reasonable comparison with the realized distribution.
+For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature.
+We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty. 
+Here, we focus only on the specification of the modeling framework; complete procedural details and code for generating the data sets and executing the model are provided as supplemental information.
+
+We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation. 
+We will consider \(X_M\) (see Box 1, Fig. \ref{fig:diagram}) to be a vector of $n$ observations that takes the value of one to indicate presence and zero for absence:
+\begin{equation}
+X_M = \{x_{M,1}, x_{M,2}, \ldots, x_{M,n}\}
+\end{equation}
+We further assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
+The model is evaluated by relating these observations to environmental data \((D_M)\), which for the sake of the example consists of mean annual temperature $T_M$ and annual precipitation $P_M$. 
+For simplicity, we use a simple logistic model for the metamodel \((\theta_M)\) with a second order effect of temperature and precipitation. 
+This naive model (i.e., the metamodel with no constraints from sub-models) predicts  occurrence probability (\(\psi_N\)) as a function of the environment:
+%-----
+\begin{equation}
+\begin{aligned}
+	\psi_N &= f\left(\theta_M, D_M \right) \\
+	&= p \left (X_M = 1 \mid \theta_M, T_M, P_M \right) \\
+	&=\text{logit}^{-1}\left( \mathbf{\Theta_M} \mathbf{D_M} \right)
+\end{aligned}
+\end{equation}
+%-----
+where \(\mathbf{\Theta_M}\) is the parameter vector of the model \(\theta_M\), \(\mathbf{D_M} \) is the covariate matrix (i.e., \(T_M, P_M\), with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
+We can fit this model in a Bayesian framework to allow for easy integration of models (as we will show later).
+In this context, the goal of modeling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, concise introductions can be found in ][]{McCarthy2007, Link2010}:
+%-----
+\begin{equation}
+\label{eq:ex1_bayes}
+	p\left (\theta_M \mid X_M,T_M,P_M \right ) \propto 
+	p \left(X_M \mid \theta_M, T_M, P_M \right)
+	p \left(\theta_M \right)
+\end{equation}
+%-----
+where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
+
+Thus far, we have considered only a single source of information to fit this model, and therefore the prior distribution \(p\left(\theta_M \right)\) from Eq. \ref{eq:ex1_bayes} is uninformative.
+As a secondary source of information, we will consider an experiment relating the population growth rate of the plant to manipulations to the precipitation regime, with results (but no raw data) available from the literature (Fig.
+\ref{fig:ex1_sampling}b).
+Furthermore, no information is available regarding the temperature regime for the experiment.
+Transplant experiments that evaluate performance beyond the range of a species are common, and represent a plausible scenario for model integration \citep{Hargreaves2014}.
+According to niche theory \citep{Holt2009}, the fundamental niche corresponds to the set of environmental conditions where the per capita intrinsic growth rate $r$ is positive.
+This concept gives us a reasonable model to fit the scaling function $g$ for our sub-model (Fig. \ref{fig:diagram}).
+If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, then for observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
+\begin{equation}
+	\psi_{S,i} = \int_0^\infty N \left(X_{S,i}, \sigma_{S,i} \right)
+\end{equation}
+where \(N\) is the Normal density function.
+We can then estimate the posterior distribution for the sub-model by fitting the relationship between \(\psi_S\) and precipitation \( \left( P_S \right) \) using Bayesian beta regression \citep{Ferrari2004}:
+%-----
+\begin{equation}
+\label{eq:ex1_thetas}
+	p\left (\theta_S \mid \psi_S,P_S \right ) \propto 
+	p \left(\psi_S \mid \theta_S, P_S \right)
+	p \left(\theta_S \right) \\
+\end{equation}
+%-----
+
+Although the two data sets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation).
+This scaling is quite simplistic, and would never be used to predict a species' range using this mechanistic sub-model in isolation.
+Despite the strong theoretical foundation (i.e., niche theory) linking the observations to presence-absence at large scales, the upscaled sub-model (i.e., the transformation of \(X_S\) and \(\sigma_S\) to \(\psi_S\)) is built on the strong hypothesis that the species will be present when $r>0$, neglecting other population dynamics constraints such as Allee effects and metapopulation dynamics.
+The fundamental niche is also incomplete, as we lack environmental variables such as temperature, as well as all other ecological processes responsible for the realized distribution.
+As such, it would be unwise to expect predictions from this model alone to resemble the actual distribution of the species; as a mechanistic model, it is simply too incomplete to predict distribution.
+As we will show, however, the information from this sub-model, when applied as a constraint on the metamodel, can result in improved predictions that incorporate the information within each model.
+
+We accomplish model integration by treating the posterior of \(\theta_S\) as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
+%-----
+\begin{equation}
+	\label{eq:ex1_integrated}
+	\overbrace{p(\theta_M \mid X_M, T_M, P_M, \theta_S, \psi_S, P_S)}^\text{integrated posterior}
+	\propto
+	\overbrace{p\left (\psi_S \mid \theta_S,P_S \right )}^{\substack{\text{new information} \\ \text{from sub-model}}}
+	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) P \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
+	\overbrace{p \left(\theta_S \right)}^{\substack{\text{prior for} \\ \text{sub-model}}}	
+\end{equation}
+%-----
+As before, the metamodel \(\theta_M\) can be used to predict the species distribution \((\psi_I)\).
+However, these predictions will not reflect only the presence-absence samples in \(X_M\), but will also reflect the information from \(\theta_S\), including all of the data sources used to produce this sub-model.
+In other words, with this evaluation, the set of parameters $\theta_M$ is most likely when it agrees with both the original data and predictions from the sub-model. 
+Note that in this particular example, because the sub-model is only contingent on precipitation, there is no integration on temperature. 
+Finally, we note the presence of marginal distributions for both models (i.e., \(p(\theta_M)\) and \(p(\theta_S)\)).
+These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
+For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modeling process (due to the lack of additional mechanisms in the second model, as mentioned above).
+
+We performed this procedure with our hypothetical data sets.
+When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data.
+Unsurprisingly given the quality of the data, the sub-model was highly precise, providing a fairly strong constraint when producing the integrated model.
+The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
+At the scale of the metamodel, considering both temperature and precipitation, we observed similar results, with reduced uncertainty in the predictions over the domain not covered by the presence-absence data (Fig.
+\ref{fig:ex1_map}).
+
+Once the technical challenges of evaluating the parameters are overcome, the integration has several advantages over the both the naive parameterization of the metamodel and the mechanistic sub-model.
+In the situation where both models strongly agree on the relationship between water availability and occurrence probability, then we should expect significantly reduced uncertainty in that domain (Figs.
+\ref{fig:ex1_precip}, \ref{fig:ex1_map}).
+Because the metamodel includes both temperature and precipitation, any reduction in the uncertainty with respect to precipitation might also influence the evaluation of the relationship with temperature (reducing bias in the case where there is a correlation between T and P in the data).
+However, uncertainty could increase in the case of model disagreement.
+For instance, the relationship between precipitation and presence in the naive metamodel could be caused by a spurious relationship (e.g.
+spatial autocorrelation caused by historical contingencies), leading to a false confidence in the species distribution over precipitation gradients.
+The information from the sub-model will thus bring back the right amount of uncertainty in the relationship.
+Moreover, convergence in model predictions in overlapping domains (e.g., near the peak in the probability of presence in Fig. \ref{fig:ex1_precip}) can aid in detecting crucial ecological processes.
+
+%% ----------------------------------
+%
+%     EXAMPLE 2
+%
+%% ----------------------------------
+
+\subsection*{Example 2: Constraining an SDM using phenological information}
+For the second example, we consider the problem of predicting the future distribution of a species following climate change.
+There is considerable interest in comparing correlative and mechanistic projections with respect to climate change \citep{Morin2009}, and correctly characterizing uncertainty is a critical aspect of this problem \citep{Cheaib2012}.
+Despite being a relatively common application of \ac{SDM}s \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios is problematic for a number of reasons previously discussed \citep{Araujo2006}.
+We used our metamodeling framework to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
+We use data from \citet{Morin2009} for sugar maple (\emph{Acer saccharum}), an economically and ecologically important species occurring in eastern North America.
+Here we describe briefly the dataset, methods, and the results of the analysis.
+Complete details, including all data and scripts to reproduce the analysis, are included as supplemental information.
+
+\defcitealias{IPCC2001}{IPCC, 2001} % this is needed to make the organizational citation in the following paragraph work correctly
+For the metamodel, we used a dataset of 1013 presence and 13863 absence records, at 0.5 degree resolution, sampled from North America.
+We reserved one-third of the dataset for validation, and used the area under the receiver operating characters curve (AUC) for validation of contemporary predictions.
+We used a binomial GLM to relate the suitability for sugar maple to 6 climate variables, scaled to the same resolution (see supplemental information for a descriptions of the climate variables and of the model selection process).
+We selected a GLM here for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
+For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
+Phenofit produces predictions at continental scales; thus the scaling function \(g\) was incorporated within Phenofit itself and we performed model integration directly on the outputs.
+To perform the integration, we used \ac{MCMC} to condition the parameters of the metamodel  on two likelihood functions.
+The first is the naive SDM, where the likelihood of a ``success'' in the presence/absence dataset was a Bernoulli density with the (logit-transformed) probability expressed as a linear function of the climate predictors.
+For the sub-model, we considered that Phenofit, being a process-based model, may perform better when projecting into the future.
+Thus, we used the predicted future probability of presence as the basis for generating simulated future presence-absence datasets.
+This procedure inherently incorporates some uncertainty by randomly determining where the species is present (though uncertainty in the probabilities themselves could not be incorporated, as they were not available).
+Thus, the sub-model likelihood is simply the probability, given the predictions of the metamodel, of observing a simulated dataset \(X_s\) that is randomly drawn from the phenofit predictions \(\psi_s\); i.e., \(X_s = f(\psi_s)\):
+\begin{equation}
+\label{eq:integrated2}
+	p( \theta_M \mid X_M, D_M, \psi_s, \theta_S )
+	\propto 
+	p( X_S = f(\psi_s)\mid \theta_M, \theta_S )
+	p( Y_M \mid \theta_M, D_M ) 
+	p( \theta_M )
+	p( \theta_S )
+\end{equation}
+where \(\theta_M\) is the metamodel, 
+\(X_M\) is the vector of contemporary presences and absences, 
+\(D_M\) is the matrix of climate predictors,
+and \(\theta_S\) is the model relating the Phenofit predictions to the metamodel.
+Furthermore, the integration is performed on the \emph{predictions} of the sub-model directly, rather than on compatible parameters as in Example 1. 
+Thus, the metamodel must appear in the sub-model likelihood.
+
+For present climates, the predictions of Phenofit (Fig \ref{fig:ex2}A) were largely similar to those of the naive SDM (Fig \ref{fig:ex2}C). 
+Furthermore, variance in the predictions of the naive model was relatively low throughout the spatial domain (Fig \ref{fig:ex2}E).
+Similar patterns were repeated for analogous climates under the future scenario (Fig \ref{fig:ex2}D); however these predictions differed considerably from the predictions of the mechanistic sub-model (Fig \ref{fig:ex2}B), and uncertainty for the projection was relatively high, particularly for eastern Canada (Fig \ref{fig:ex2}F).
+In general, Phenofit was predicted considerably less northward migration by 2080.
+The integrated model, which included both the presence-absence data as well as the Phenofit predictions for the future, made intermediate predictions in terms of northward range shifts for sugar maple (Fig \ref{fig:ex2}H).
+Globally, the integrated model expressed somewhat more uncertainty than the naive model (Fig \ref{fig:ex2}J), reflecting the divergent predictions of Phenofit and the naive model.
+However, in areas where the naive model performed worst (e.g., eastern Canada), the level of uncertainty in the integrated model was similar to the uncertainty in other locations.
+Unsurprisingly, integration resulted in worse predictions for the present distribution of the species (Fig \ref{fig:ex2}G) compared with either the naive model or Phenofit in isolation.
+Conventional SDMs are quite good at producing empirical models of species distributions under present climatic regimes, and the naive model here could be further improved with the addition of more parameters or the inclusion of ensemble forecasting.
+This is due to the integration focusing on future predictions; if integration was desired for the present range of the species (perhaps as a way to more accurately quantify uncertainty), a better approach would have been to use the present predictions from Phenofit.
+Unfortunately, validation is difficult for projections, as we lack data to evaluate model performance.
+For the desired outcome of this model (future projections), validations against present data may not be useful; the predictions of naive model were superior (AUC = 0.99) to those of the integrated model (AUC = 0.96), yet the future predictions of the integrated model may be more realistic than those of the naive model.
+
+%% ----------------------------------
+%
+%     DISCUSSION
+%
+%% ----------------------------------
+
+\section*{Discussion}
+
+\subsection*{Comparison with other methods}
+The methods provided here expand upon the motivation of hybrid models to develop a more robust and unifying approach for ecological models in a predictive context while striving to overcome some of the limitations characteristic of other integrated approaches.
+Attempting to incorporate underlying ecological processes so that SDMs are not oversimplified by combining different model parameters or scales can be challenging using these hybrid approaches. 
+In particular, it can often be difficult to identify parameters that can be used to connect different modeling frameworks and what parameters can produce meaningful responses \citep{Thuiller2013}. 
+The methods presented in this study aim to overcome some of these limitations through the use of Bayesian statistics. 
+Bayesian methods produce results as an entire distribution instead of a single point, allowing for a comprehensive understanding of the uncertainty of estimates \citep{Link2010}. 
+The estimation of model parameters (and predictions) as distributions can also help decrease overparameterization that can be a concern for hybrid models. 
+Bayesian frameworks also provide a natural framework for the incorporation of prior information. 
+The inability or difficulty to include information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Thuiller2008, Smolik2010}. 
+Even when a link can be formed to include this information, it often is oversimplified to simplify computation \citep{Gallien2010}. 
+Finally, Bayesian methods inherently allow for feedbacks or interactions between sub models, which may portray a more realistic response to many environmental dynamics where factors may simultaneously influence the change of one another.
+
+Our approach is an original application of Bayesian multilevel modeling, and can be considered a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales while using several models simultaneously.
+In particular, this development has certain similarities with Bayesian model averaging, Bayesian calibration of process-based models and hierarchical Bayesian modeling. 
+Bayesian model averaging aims to combine several alternative models to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.][]{Wintle2003, Link2006}.
+Bayesian model averaging considers models operating at the same scale and the posterior distribution is usually determined with Gibbs sampling. 
+Bayesian calibration of process-based model focuses on uncertainty of the parameter values in the process-based models, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, VanOijen2011, Hartig2012}.
+In contrast with Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of a generally more correlative metamodel.
+Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005, Efendiev2005a}, plant physiology \citep{Ogle2008, Ogle2009}, and climate and atmospheric modeling \citep{Zimmerman2005, Mcmillan2010, Kang2012}.
+Although our method as presented is distinct from Bayesian model averaging, it is fully compatible with these methods.
+Multiple metamodels could be averaged in this way if a consensus prediction from multiple metamodels (containing, e.g., different sub-model sets or different metamodel formulations).
+Furthermore, the metamodel can easily be incorporated into a broader multimodel inference scheme.
+For example, the metamodel can be used as one input in an ensemble forecasting scheme that also incorporates other common SDM formulations (e.g., Random Forest, MaxEnt).
+
+%\fd{FD will work on this part: \\
+%It would be good to compare this approach with the kind of one develop by Moorcroft et al.
+2001 : Moorcroft, P.
+R., Hurtt, G.
+C., \& Pacala, S.
+W.
+(2001).
+A method for scaling vegetation dynamics: the ecosystem demography model (ED).
+Ecological monographs, 71(4), 557-586.\\
+%%Medvigy, D., Wofsy, S.
+C., Munger, J.
+W., Hollinger, D.
+Y., \& Moorcroft, P.
+R.
+(2009).
+Mechanistic scaling of ecosystem function and dynamics in space and time: Ecosystem Demography model version 2.
+Journal of Geophysical Research: Biogeosciences (2005–2012), 114(G1).\\
+%Fisher, R., McDowell, N., Purves, D., Moorcroft, P., Sitch, S., Cox, P., \& Ian Woodward, F.
+(2010).
+Assessing uncertainties in a second-generation dynamic vegetation model caused by ecological scale limitations.
+New Phytologist, 187(3), 666-681.
+%\\
+%These models are multilevel and use scaling functions using statisitical moments series.
+%}
+
+%----------------------------
+%----------------------------
+%----------------------------
+
+\subsection*{Advantages of Model Integration}
+\defcitealias{TERN2013}{TERN, 2013} %% this is needed to make the organizational citation in the following paragraph work correctly
+Models are important tools that are increasingly used by land managers to face challenges associated to decision-making \citep{Guisan2013}
+A growing number of models are available, but they can provide diverging answers to very similar questions due to the differences in their assumptions and methodology. 
+This can create confusion and even some mistrust towards models, and as a consequence some managers may be discouraged to incorporate model results in their management plans. 
+Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organizations around the world \citep[e.g.][\citetalias{TERN2013}]{Bernier2013}.
+Incorporating information from multiple sources, particularly with respect to uncertainty,  fosters a connection between scientifically-generated knowledge and policy, and is therefore an important tool for adaptive management \citep[][Fig.
+\ref{fig:management}; ]{Rehme2011}.
+Such approaches are needed in formulating management plans for vulnerable species and ecosystems to avoid basing decisions on too-narrow subsets of the available information \citep{Dawson2011}.
+The flexibility of our approach may also represent an advantage for management by easing the integration of specific decision making criteria (e.g.
+desired grain and extent of outputs) into model development.
+Successful use of an integrated modeling approach will always remain dependent on an intimate understanding of the decision-making process by modelers, emphasizing the importance of close collaboration between with practitioners at all stages of model development \citep{Guisan2013}.
+
+Model uncertainty is another key factor affecting applicability of model outputs \citep{Addison2013}.
+One of the main strengths of our approach is that it allows for a clear, transparent identification of uncertainties and how they are transmitted throughout the modeling process. 
+Transparency in uncertainty can be considered as a sort of sensitivity analysis, in which areas with large uncertainty can be detected and new experimental research or additional data collection can be designed (e.g., Example 1, Figs.
+\ref{fig:ex1_sampling}, \ref{fig:ex1_precip}).
+The new knowledge resulting from this research can be readily incorporated into the metamodel, allowing for an iterative learning process that will undoubtedly contribute to reduce uncertainty in predictions for a wide range of environmental conditions.
+Moreover, our framework embraces change as a fundamental process and is able to adapt and respond to it. 
+The ease of incorporating new knowledge to the modeling framework (including new theory or the result of management and experimental efforts) will allow for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans \citep{Keith2011}, and the integration of sub-models allows for clear specification of desired model outputs (via the specification of the metamodel) while easily retaining important ecological objectives (via careful specification of sub-models).
+Transparency in the model building process must be accompanied by a clearly documented modeling workflow; the transparency of our approach (and the utility of specifying specific objectives accompanied by appropriate sub-models) will assist in the development of transparent and well documented workflows describing the process of model integration and will ensure reproducibility and applicability (Fig.
+\ref{fig:management}).
+In an era of continuous and unprecedented change, adaptive approaches such as the one presented here are often highlighted as a pressing need in order to develop strategies to promote ecosystems that are both feasible and resilient \citep{Seastedt2008, Mori2013}.
+
+In many cases, both the data and theory needed to apply our method already exist, and all that is needed is the development of sub-models and the links between the sub-models and a metamodel.
+For example, climatic gradients may mediate competitive interactions \citep{Kunstler2011}, which means that simple correlative models that fail to account for competition may be wrong if the climate-competition association changes in the future.
+In North America, the US Forest Service maintains a long term Forest Inventory Analysis database that could be utilized to parameterize a competition model.
+Such a model need not explicitly predict occurrence limits; rather it could be integrated with a larger-scale model to include information about competition in a distribution model.
+The phenological information needed to parameterize the sub-model for Example 2 is similarly available for a wide range of species \citep{Morin2009}.
+There are also a number of networks collecting high-quality data with good temporal and spatial coverage (e.g., the National Ecological Observatory Network; NEON, and Long-Term Ecological Research sites; LTER).
+There is great potential for these kinds of data to be used in fitting sub-models of the kind used in Example 1.
+In other cases, qualitative model comparison efforts have already been made \citep{Morin2009, Cheaib2012}.
+Our framework could be used \emph{post-hoc} on the outputs of these models to quantify uncertainty resulting from model disagreement.
+Our goal in this paper is to provide a framework for including much of this existing knowledge about species' responses to the environment into species distribution models.
+
+%----------------------------
+%----------------------------
+%----------------------------
+
+\subsection*{Challenges} 
+Although our approach is highly flexible and can be applied in a number of situations, there are some challenges to successfully using the framework.
+As with any modeling effort, good model specification with strong links to theory are essential \citep{Austin2007}.
+Poorly specified models will produce outputs that are uninformative or misleading, and model integration is not a cure for these problems.
+In the worst case, constraining a metamodel with a poor sub-model can result in outcomes that are worse in terms of bias and uncertainty than those produced by a naive model.
+We expect model selection will play an important role in applications of our framework, and such schemes can be incorporated relatively easily \citep{Madigan1995, Wasserman2000, Tenan2014}.
+
+In a similar fashion, the quality and availability of data impose a significant constraint on the number and type of models that can be implemented.
+The capacity to implement a model is very low if data requirements are prohibitive. 
+Adequate coverage of explanatory variables is a significant obstacle, and exploratory analyses can be a significant aid in understanding how data coverage impacts resulting predictions \citep{Mckenney2002}.
+Integration can ameliorate these issues to some extent by using supplemental information (and conceptual advances) in additional sub-models where coverage is weak (e.g., Example 1, Figs.
+\ref{fig:ex1_precip}, \ref{fig:ex1_map}).
+For example, \citet{Freckleton2009} estimated that data are too scant to successfully develop highly mechanistic models predicting weed population numbers. 
+A strong asset of our approach is that it can be used without the full suite of data that would be required to run a fully mechanistic model. 
+Given that the metamodel is correlative, it can be effectively implemented with, e.g., only presence-absence data, or, in the case where true absences are difficult to obtain, with presences and pseudo absences (provided sufficient care is used in interpreting the results of such a model).
+Then any additional mechanistic data that are available will enhance predictions by constraining outputs of the metamodel.
+
+Finally, determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eqs. \ref{eq:ex1_integrated}, \ref{eq:integrated2}) is a critical point.
+In the context of integrated models, the challenge is three fold: (i) which spatial and temporal scales (i.e.
+which processes), are to be considered, (ii) how to build a scaling function that is consistent with the metamodel and (iii) how error and uncertainty are propagated from the scales of the sub-models scale to that of the metamodel.
+Although we argue that our proposed framework is able to easily deal with very different scales and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered, the building of scaling function deserves further investigation.
+We have shown in our examples that even simple scaling functions can provide reasonable constraints on the metamodel. 
+However, with the objective to take into account all know processes and models, it is likely that the modeling process may include multiple scaling functions, each at different scales. 
+Indeed, if species distributions are a function of, e.g.
+poopulation growth rate \citep{Sykes1996, Guisan2000}, they will involve processes at the individual (e.g.
+competition) or cellular (e.g.
+photosynthesis) scales.
+Such very large differences in spatial scales imply that simple functions will be inadequate, requiring more traditional upscaling methods. 
+Therefore, intermediate models and processes will need to be introduced, and this will require additional parameters, assumptions, and data, and will add uncertainty to the final integrated model.
+
+In an applied context, the results of models predicting species ranges are increasingly used by resource managers, conservation biologists, or forest ecologists for formulating or adjusting recommendations.
+The utility of such models depends on their ability to help evaluate events beyond the bounds of the available data, including in future situations constrained by climate change. 
+Practitioners may judge model usefulness on two criteria: (1) Does the model provide reliable predictions at the needed time and space scales, and (2) can the model be implemented given the available data and technical expertise?
+Since most decision makers work at local space scales and follow specific time frames, modelers face the obvious challenge of providing detailed outputs while facilitating implementation.
+In terms of model outputs, our framework is transparent in terms of uncertainty by virtue of providing easily interpretable probability distributions for model parameters and predictions.
+However, developing the models requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming.
+In many cases, off-the-shelf software can adequately express the model likelihoods with minimal programming, but more complicated models will require the development of custom software.
+Thus, this framework will require significant investment in developing customized code for the samplers in order to actually estimate parameters, which may be technically challenging for some practitioners and may limit the adoption of these methods by both practitioners and decision-makers.
+However, the same was true of all statistical techniques or modeling approaches when they were initially developed, and many of the conceptual challenges associated with Bayesian statistics are being reduced as they gain recognition in the scientific literature.
+We therefore argue that our proposed approach as a strong potential for direct use in our real world where climate is quickly changing and conservation practices must be adjusted accordingly.
+We further argue that forging stronger collaborations between modelers, decision-makers, and practitioners will improve the incorporation of this and other new methods into applied ecology and conservation.
+
+%----------------------------
+%----------------------------
+%----------------------------
+
+\subsection*{Future directions}
+Hierarchical Bayesian modeling requires data at every scale to fit the parameters required to constrain one level with another.
+Moreover, beyond the needs for additional data, more knowledge of the important processes driving species range limits is needed to allow for extrapolation beyond the scope of the original datasets to address future changes due to climate and land-use changes.
+This knowledge can be obtained through a series of experimental manipulations and by substituting space for time data along environmental gradients.
+When experiments are possible, gaps in knowledge can be inferred using surrogate variables, theoretical models and spatially explicit models to explore the range of outcomes possible \citep{Fortin2012}.
+Another way to gain knowledge is to simplify species complexity into a series of states and to use state-transition models (e.g., Markov chain, semi-Markov chain, matrix projection model, integral projection model) which require more theoretical or coarser knowledge of species responses to biotic and abiotic changes.
+Although fewer states allow us to better capture the essence of species responses to environmental conditions, knowledge is still required to model feedback effects across processes and spatio-temporal scales.
+Hence we need to determine which processes are the most important at a given scale and how to weight their effects while scaling-up.
+Our metamodeling approach can act as a springboard for incorporating imperfect knowledge of how multi-scale processes influence species ranges.
+By incorporating this knowledge into species distribution models, we can hopefully both reduce bias and more accurately assess model uncertainty when forecasting future changes to species ranges.
 %---------------------------------------------
@@
 
-\input{modeling_framework}
-
-%---------------------------------------------
-%---------------------------------------------
-
-\input{discussion}
-
-%---------------------------------------------
-%---------------------------------------------
-
+%TC:ignore
 \section*{Acknowledgements}
@@
 Comments from Charles Canham and William Godsoe improved a previous version of this manuscript.
+%TC:endignore
 
+%TC:break references
 %% ----------------------------------
@@
 
-\input{figures}
+\newpage
+\section*{Figure Captions}
+
+\begin{flushleft}
+
+%==================
+% FIGURE 1
+%% 149 words
+%TC:break fig 1
+\textbf{Figure \ref{fig:diagram}:}
+	The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
+	The metamodel ($\theta_M$) operates at a single scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
+	The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
+	The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
+	This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
+	This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).
+
+%==================
+% FIGURE 2
+%% 88 words
+%TC:break fig 2
+\textbf{Figure \ref{fig:ex1_sampling}:}
+	Two simulated data sets used to illustrate the model integration framework.
+	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
+	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., \(-\)1--1).
+	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
+	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.
+
+%==================
+% FIGURE 3
+%%  120 words
+%TC:break fig 3
+\textbf{Figure \ref{fig:ex1_precip}:}
+	Results of integration showing the probability of presence (\(\psi\)) when considering only precipitation.
+	(a) Naive model, using only presence-absence data.
+Uncertainty increases dramatically when attempting to project beyond the scope of the source data (represented by the horizontal line above the bottom axis).
+	(b) Mechanistic model, using observations of an experiment to infer probability of presence.
+Predictions are highly precise due to high quality source data, but are likely to be overly optimistic given the simplicity in assuming only precipitation determines probability of presence (see text).
+	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
+	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.
+
+%==================
+% FIGURE 4
+%% 75 words
+%TC:break fig 4
+\textbf{Figure \ref{fig:ex1_map}:}
+	Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
+	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
+	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
+	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.
+
+%==================
+% FIGURE 5
+%% 140 words
+%TC:break fig 5
+\textbf{Figure \ref{fig:ex2}:}
+	Results of model integration for sugar maple, \emph{Acer saccharum}.
+	The mechanistic sub-model Phenofit predicts present (A) and future (B) suitability using phenological information, but does not provide estimates of uncertainty.
+	Predictions from a naive metamodel for the present (C) were quite similar to the sub-model, but future predictions  differed (D).
+	Uncertainty in future projections was quite high in the eastern portion of the range (F) relative to uncertainty in predictions for the present (E).
+	Model integration resulted in poor predicitons for present range (G), with future predictions that were intermediate between the two models (H).
+	Uncertainty in the integrated model (I, J) was generally greater than in the naive model, reflecting the disagreement between the naive model and the sub-model, but there were fewer extremes in model uncertainty when projecting the integrated model (J) compared with the naive model (F).
+
+%==================
+% FIGURE 6
+%% 65 words
+%TC:break fig 6
+\textbf{Figure \ref{fig:management}:}
+	Sample workflow for applying the models presented in the first example in a management context.
+	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
+	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the metamodel.
+
+\end{flushleft}
+
+%==================
+% FIGURE 1
+
+\newpage
+\begin{figure}[h!]
+\input{diagram}
+\caption{}
+\label{fig:diagram}
+\end{figure}
+
+%==================
+% FIGURE 2
+\newpage
+\begin{figure}[h!]
+	\includegraphics{ex1_sampling.pdf}
+	\caption{}
+	\label{fig:ex1_sampling}
+\end{figure}
+
+%==================
+% FIGURE 3
+\newpage
+\begin{figure}[h!]
+	\includegraphics{ex1_precip.pdf}
+	\caption{}
+	\label{fig:ex1_precip}
+\end{figure}
+
+%==================
+% FIGURE 4
+\newpage
+\begin{figure}[h!]
+	\includegraphics[width=5.25in]{ex1_map.pdf}
+	\caption{}
+	\label{fig:ex1_map}
+\end{figure}
+
+%==================
+% FIGURE 5
+\newpage
+\begin{figure}[h!]
+	\includegraphics[width=6in]{ex2.pdf}
+	\caption{}
+	\label{fig:ex2}
+\end{figure}
+
+%==================
+% FIGURE 6
+\newpage
+\begin{figure}[h!]
+	\includegraphics{management.pdf}
+	\caption{}
+	\label{fig:management}
+\end{figure}
 
@@
 
-\input{model_description}
+%TC:break box 1
+
+\newpage
+\section*{Box 1: Model description}
+% Boxes are limited to 750 words
+We begin with a simple correlative model (i.e. calibrated with data provided at the same ecological scale as predictions), characterized in a Bayesian framework to allow further integration. 
+We refer to this model (with associated parameters) as the metamodel, \(\theta_M\), and it is parameterized with information on the species' presence, \(X_M\), along with associated covariates (e.g., climate, presence/absence of interacting species, etc.), collectively referred to as \(D_M\).
+The naive (i.e., not integrated with information from sub-models) probability that a species is present, \(\psi_N\), is a deterministic function of the metamodel and its covariates::
+%-----
+\begin{equation}
+\label{eq:sdm1}
+	\psi_N = f(\theta_M, D_M)
+\end{equation}
+%-----
+The goal of model fitting is to estimate the posterior probability distribution for the parameters of \(\theta_M\), given the observations (\(X_M\)) and covariates (\(D_M\)):
+%-----
+\begin{equation}
+\label{eq:sdm2}
+	p( \theta_M \mid X_M, D_M)
+\end{equation}
+%-----
+
+Using Bayes' theorem, we find the posterior probability of \(\theta_M\) as follows:
+%-----
+\begin{equation}
+\label{eq:sdm3}
+	p( \theta_M \mid X_M, D_M ) = \frac{ p( X_M \mid \theta_M, D_M ) p( \theta_M ) } { p(X_M, D_M) }
+\end{equation}
+%-----
+where \(p( X_M \mid \theta_M, D_M )\) is referred to as the likelihood of the observations, \(p(\theta_M )\) is the prior probability of the model, and \(p(X_M, D_M)\) is the normalization constant.
+The normalization constant involves computing integrals that are often impossible to solve analytically. 
+In practice, simulation techniques such as \ac{MCMC} can be used to sample directly from the posterior distribution, making such computations unnecessary.
+Thus, we use the proportional form of Bayes' Theorem:
+%-----
+\begin{equation}
+\label{eq:sdm4}
+	p( \theta_M \mid X_M, D_M ) \propto p( X_M \mid \theta_M, D_M ) p( \theta_M ) 
+\end{equation}
+%-----
+
+The role of the metamodel is to integrate data at the same ecological scale of predictions.
+Additional sub-models require outputs that should be comparable to this given ecological scale (e.g. constraining presence or absence on the landscape).
+Formally, we will add an additional model, \(\theta_S\), that is based on a different set of hypotheses and that makes predictions \((\psi_S)\) based on an additional dataset \((X_S, D_S)\):
+%-----
+\begin{equation}
+\label{eq:model2-1}
+	\psi_S = g(\theta_S, X_S, D_S)
+\end{equation}
+%-----
+
+For integration to be successful, it must be possible to compute the likelihood of this prediction given the metamodel, \(p \left( \psi_S \mid \theta_M \right) = p \left(\theta_S \mid \theta_M, X_S, D_S \right)\); the function \(g\) serves to transform the parameters of the sub-model to the scale of the metamodel.
+Thus, we refer to $g$ as the \emph{scaling function}.
+In many cases, computing this likelihood will be challenging.
+The simplest case is when the parameters of \(\theta_S\) are directly comparable to those in \(\theta_M\) (see Example 1), or when the scaling is performed directly within the sub-model (see Example 2), but other solutions are possible.
+This new likelihood, which integrates the information from the second model into the first, can be treated as new ``data'' to evaluate the parameters of \(\theta_M\).
+We simply use the posterior distribution of $\theta_M$ from the previous step (Eq.
+\ref{eq:sdm4}) as the new prior probability of $\theta_M$, and evaluate the posterior probability of $\theta_M$ in light of the new data (i.e., information from the submodel \(\theta_S\)) and the prior knowledge, as before:
+
+%-----
+\begin{equation}
+\label{eq:integrated2}
+	\overbrace{p( \theta_M \mid X_M, D_M, \theta_S, X_S, D_S )}^\text{integrated posterior}
+	\propto 
+	\overbrace{p( \theta_S \mid \theta_M, X_S, D_S )}^\text{likelihood}
+	\overbrace{p( X_M \mid D_M, \theta_M ) P( \theta_M )}^{\text{prior for } \theta_M}
+	\overbrace{p(\theta_S)}^{\text{prior for } \theta_S}
+\end{equation}
+%-----
+
+This procedure may be repeated for an arbitrary number of models (limited only by available data and computational power). 
+The models may be implemented simultaneously, when multiple data sets are available and can be evaluated under different underlying models, or they may be run sequentially, updating the posterior distribution of $\theta_M$ as new information becomes available.
+Note that $\theta_S$ itself need not be implemented in a Bayesian framework and the output need not be identical in form to the output of \(\theta_M\); it is enough to produce an output that can be evaluated as a likelihood under the first model and to have some assessment of the confidence in the parameter estimates of $\theta_S$ to use as a prior.
 

### 2014-09-19 tex/proposal_letter.tex
Mostly final changes

@@
 Our framework is highly flexible and can integrate information from sub-models operating a multiple spatial and temporal scales.
-The resulting meta-model framework produces probabilistic estimates of species presence that includes uncertainty from data sources as well as all sub-models.
+The resulting metamodel framework produces probabilistic estimates of species presence that includes uncertainty from data sources as well as all sub-models.
  These predictions reflect all of the information used as input for the original sub-models.

### 2014-09-22 tex/manuscript.tex
Finished latest draft

@@
 	% less than 45 chars including spaces
-	\textbf{Short title: Integrated models of species ranges}
+	\textbf{Short title:} Integrated models of species ranges
 
-	\textbf{Type of article: } Ideas and Perspectives
+	\textbf{Type of article:} Ideas and Perspectives
 
@@
 	% figure captions limited to 150 words each
-	\textbf{Word counts: } Abstract: 200; Main Text: 6091; Boxes: 517
+	\textbf{Word counts: } Abstract: 200; Main Text: 6086; Boxes: 517
 
-	\textbf{Number of references: } % limited to 80
+	\textbf{Number of references:} 79 % limited to 80
 
-	\textbf{Number of figures: } 6
+	\textbf{Number of figures:} 6
 
-	\textbf{Number of tables: } 0
+	\textbf{Number of tables:} 0
 
-	\textbf{Number of text boxes: } 1
+	\textbf{Number of text boxes:} 1
 
@@
 For instance, correlative models are used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while mechanistic approaches can include space and time dynamics and therefore account for dispersal processes \citep{Kearney2008}.
-This complementarity has been used to merge the two types of models into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Naujokaitis2013, Boulangeat2014}. 
+This complementarity has been used to merge the two types of models into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Boulangeat2014}. 
 However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are not easy to test \citep{Gallien2010}, and uncertainties are very difficult to quantify and identify from its many different sources.
@@
 Contrary to hybrid and hierarchical models, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from sub-models at a variety of spatial scales, allowing as much flexibility as possible regarding the type of information included.
-We use the power and generality of a hierarchical Bayesian framework, which allows us to include multiple data sources and modes of inference \citep{Clark2005, VanOijen2005, Clark2006, Hobbs2011, Hartig2012}.
-Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to interpret and reflects uncertainty at all levels of organization \citep{Clark2005, Cressie2009, Hobbs2011}. 
+We use the power and generality of a hierarchical Bayesian framework, which allows us to include multiple data sources and modes of inference \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}. 
+Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to interpret and reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}. 
 We illustrate our approach with two examples.
@@
 We can fit this model in a Bayesian framework to allow for easy integration of models (as we will show later).
-In this context, the goal of modeling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, concise introductions can be found in ][]{McCarthy2007, Link2010}:
+In this context, the goal of modeling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, a concise introduction can be found in ][]{Link2010}:
 %-----
@@
 In particular, this development has certain similarities with Bayesian model averaging, Bayesian calibration of process-based models and hierarchical Bayesian modeling. 
-Bayesian model averaging aims to combine several alternative models to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.][]{Wintle2003, Link2006}.
+Bayesian model averaging aims to combine several alternative models to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.,][]{Wintle2003, Link2006}.
 Bayesian model averaging considers models operating at the same scale and the posterior distribution is usually determined with Gibbs sampling. 
-Bayesian calibration of process-based model focuses on uncertainty of the parameter values in the process-based models, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, VanOijen2011, Hartig2012}.
+Bayesian calibration of process-based model focuses on uncertainty of the parameter values in the process-based models, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, Hartig2012}.
 In contrast with Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of a generally more correlative metamodel.
-Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005, Efendiev2005a}, plant physiology \citep{Ogle2008, Ogle2009}, and climate and atmospheric modeling \citep{Zimmerman2005, Mcmillan2010, Kang2012}.
+Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005}, and climate and atmospheric modeling \citep{Mcmillan2010, Kang2012}.
 Although our method as presented is distinct from Bayesian model averaging, it is fully compatible with these methods.
@@
 \subsection*{Advantages of Model Integration}
-\defcitealias{TERN2013}{TERN, 2013} %% this is needed to make the organizational citation in the following paragraph work correctly
 Models are important tools that are increasingly used by land managers to face challenges associated to decision-making \citep{Guisan2013}
@@
 This can create confusion and even some mistrust towards models, and as a consequence some managers may be discouraged to incorporate model results in their management plans. 
-Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organizations around the world \citep[e.g.][\citetalias{TERN2013}]{Bernier2013}.
+Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organizations around the world \citep[e.g.,][]{Bernier2013}.
 Incorporating information from multiple sources, particularly with respect to uncertainty,  fosters a connection between scientifically-generated knowledge and policy, and is therefore an important tool for adaptive management \citep[][Fig.
-\ref{fig:management}; ]{Rehme2011}.
+\ref{fig:management}]{Rehme2011}.
 Such approaches are needed in formulating management plans for vulnerable species and ecosystems to avoid basing decisions on too-narrow subsets of the available information \citep{Dawson2011}.

### 2014-09-22 tex/proposal_letter.tex
Finished latest draft

--- 2014-09-01 tex/proposal_letter.tex
@@
 \documentclass[11pt]{letter}
-\usepackage[margin=1in]{geometry}
+\usepackage[margin=1.5in]{geometry}
 
@@
 
-\signature{Matthew V. Talluto, Isabelle Boulangeat, \& Dominique Gravel \\
+\usepackage{wallpaper}
+
+\signature{Matthew V. Talluto, \\
+	Isabelle Boulangeat, \& \\
+	Dominique Gravel \\
 	Université du Québec à Rimouski \\ 
@@
 \addfontfeatures{Contextuals=NoAlternate}   % turn off the crazy Q in Garamond Premier Pro
+\ThisULCornerWallPaper{0.22}{uqar-logo2}
 
@@
 
+%TC:break opening
 We are pleased to submit a proposal for a manuscript titled ``Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework''. 
@@
 
-%%% Proposals should be no more than 300 words long, describe the nature and novelty of the work, the contribution of the proposed article to the discipline, and the qualifications of the author(s) who will write the manuscript.
-Proposals should be sent to the Editorial Office
+%TC:break body
+We present a novel framework for integrating small-scale, more mechanistic models into large-scale correlative species distribution models.
+The processes generating species ranges are complex and operate at multiple scales; thus, a wealth of theory (e.g., metapopulation theory, niche theory, and complexity theory) underlies models of species ranges.
+Furthermore, there is great interest in modeling ranges, resulting in a diversity of approaches.
+In practice, however, most models do not include all available information; for example, correlative models predicting presence/absence as a function of climate often ignore smaller-scale processes such as growth, fecundity, and dispersal.
+Furthermore, individual models often produce distinct predictions for the same species, with no simple way to reconcile these predictions.
+Our contribution is the first to propose a conceptual and analytical framework that goes further than \emph{a posteriori} aggregation of predictions from different models. 
+It draws on the flexibility of hierarchical Bayesian methods to integrate knowledge of a focal species at varying ecological scales in order to provide more robust predictions of its future distribution.
+Moreover, our method quantifies uncertainty and its variablity in space while also propagating uncertainty from each information source.
+This is of crucial for identifying knowledge gaps, as thus as a to guide future research.
+Finally, under our framework future information can be added as it becomes available, which makes its potential application for adaptive management very appealing.
+We illustrate our approach using two examples, and demonstrate that the framework can reduce both bias and uncertainty when projecting beyond the range of the original data.
+We believe that our manuscript represents an important step toward a more holistic approach for forecasting species ranges.
+This is of broad interest due to the importance of range shifts for theoretical and applied ecology, and we hope you will consider it for the \emph{Ideas and Perspectives} section.
 
-% Make this a bit nicer, maybe add some text from the intro including some refs if there is room
-% Mention qualifications
-% Mention all authors?
-We present a novel application of hierarchical Bayesian methods to integrating the predictions of multiple models into a single metamodel of a species distribution.
-There is great interest in modeling species ranges, resulting in a great diversity of modeling approaches.
-However, in general, individual approaches include only a small subset of the total available information about a species. 
-For example, a broad-scale correlative model might use climate variables to predict presence or absence, but ignore what is known about smaller-scale processes such as the effect of local climate on growth, fecundity, and dispersal rates.
-Furthermore, individual models often produce distinct predictions for the same species, with no simple way to reconcile these predictions.
-Our framework is highly flexible and can integrate information from sub-models operating a multiple spatial and temporal scales.
-The resulting meta-model framework produces probabilistic estimates of species presence that includes uncertainty from data sources as well as all sub-models.
- These predictions reflect all of the information used as input for the original sub-models.
-We illustrate the metamodeling approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of some of the original data sources.
-We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
-We believe that our manuscript represents an important step toward a more holistic approach for forecasting future species ranges.
-Due to the importance of range shifts for both theoretical and applied ecology, the topic is of broad interest, and we hope you will find it of interest for the \emph{Ideas and Perspectives} section.
-We will be ready to submit the manuscript within one month.
+% Suggestion from Isa for a novelty statement:
 
-Thank you for considering out submission. We look forward to your reply.
+%TC:break closing
+Thank you for considering our submission. We will be ready to submit the manuscript within one month. We look forward to your reply.
 

### 2014-09-22 tex/bibprep.pl

@@ -1,0 +1,30 @@
+#!/usr/bin/perl
+
+use strict;
+use warnings;
+
+my $inFilename = $ARGV[0];
+my @fileContents = [];
+my $i = 1;
+my $newItem = 0;
+my $name = "";
+
+open(inFile, "<", $inFilename) or die "Cannot open file $inFilename\n";
+while(<inFile>) {
+	chomp;
+	if ( (not $newItem) && m/\\bibitem\[\{(\w+)/) {
+		$name = $1;
+		$newItem = 1;
+	} elsif($newItem && s/^$name/$i\. \\newline $name/) {
+		$i++;
+		$newItem = 0;
+		$name = ""
+	}
+	push (@fileContents, $_);
+}
+close(inFile);
+
+open(inFile, ">", $inFilename) or die "Cannot open file $inFilename\n";
+foreach (@fileContents) {
+	print inFile "$_\n";
+}

### 2014-09-22 tex/manuscript.tex
Changes to wording for SI

@@
 For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature.
-We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty. 
-Here, we focus only on the specification of the modeling framework; complete procedural details and code for generating the data sets and executing the model are provided as supplemental information.
+We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty (see Appendix S1 in Supporting Information for procedural details and scripts for executing the model).
 
@@
 We use data from \citet{Morin2009} for sugar maple (\emph{Acer saccharum}), an economically and ecologically important species occurring in eastern North America.
-Here we describe briefly the dataset, methods, and the results of the analysis.
-Complete details, including all data and scripts to reproduce the analysis, are included as supplemental information.
+Here we describe briefly the dataset, methods, and the results of the analysis (see Appendix S2 for data and scripts to reproduce the analysis).
 
@@
 We reserved one-third of the dataset for validation, and used the area under the receiver operating characters curve (AUC) for validation of contemporary predictions.
-We used a binomial GLM to relate the suitability for sugar maple to 6 climate variables, scaled to the same resolution (see supplemental information for a descriptions of the climate variables and of the model selection process).
+We used a binomial GLM to relate the suitability for sugar maple to 6 climate variables, scaled to the same resolution (see Appendix S2 for details on the climate variables).
 We selected a GLM here for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.

### 2014-09-26 tex/proposal_letter.tex
Finished latest draft and finalized proposal letter

@@
 \documentclass[11pt]{letter}
-\usepackage[margin=1.5in]{geometry}
+\usepackage[margin=1in]{geometry}
 
@@
 
-\begin{letter}{Marcel Holyoak \\ 
-	Editor-in-chief, \emph{Ecology Letters} \\ 
-	Department of Environmental Science and Policy \\
-	University of California, Davis \\
-	One Shields Avenue \\
-	Davis, CA 95616, USA }
-
-\opening{Dear Dr. Holyoak,}
+\begin{letter}{Jordi Bascompte \\ 
+	Ideas \& Perspectives Editor \\
+	\emph{Ecology Letters}
+}
+\opening{Dear Dr. Bascompte,}
 
@@
 We are pleased to submit a proposal for a manuscript titled ``Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework''. 
-We would appreciate your consideration of our manuscript for the \emph{Ideas and Perspectives} section of \emph{Ecology Letters}.
+We would appreciate your consideration of our manuscript for the \emph{Ideas and Perspectives} section of \emph{Ecology Letters}. Please find our proposal text below.
 
 %TC:break body
-We present a novel framework for integrating small-scale, more mechanistic models into large-scale correlative species distribution models.
-The processes generating species ranges are complex and operate at multiple scales; thus, a wealth of theory (e.g., metapopulation theory, niche theory, and complexity theory) underlies models of species ranges.
-Furthermore, there is great interest in modeling ranges, resulting in a diversity of approaches.
+We present a novel framework for integrating small-scale, more mechanistic models into large-scale correlative species distribution models in order to improve the forecasts of these models.
+The processes generating species ranges are complex and operate at multiple scales.
+Consequently, a wealth of theory (e.g., metapopulation theory, niche theory, and complexity theory) underlies models of species ranges, resulting in a diversity of approaches.
 In practice, however, most models do not include all available information; for example, correlative models predicting presence/absence as a function of climate often ignore smaller-scale processes such as growth, fecundity, and dispersal.
@@
 Our contribution is the first to propose a conceptual and analytical framework that goes further than \emph{a posteriori} aggregation of predictions from different models. 
-It draws on the flexibility of hierarchical Bayesian methods to integrate knowledge of a focal species at varying ecological scales in order to provide more robust predictions of its future distribution.
-Moreover, our method quantifies uncertainty and its variablity in space while also propagating uncertainty from each information source.
+We use hierarchical Bayesian methods to integrate knowledge at varying ecological scales in order to provide more robust predictions future species distributions.
+Moreover, our method explicitly quantifies uncertainty and propagates it from all information sources.
 This is of crucial for identifying knowledge gaps, as thus as a to guide future research.
-Finally, under our framework future information can be added as it becomes available, which makes its potential application for adaptive management very appealing.
-We illustrate our approach using two examples, and demonstrate that the framework can reduce both bias and uncertainty when projecting beyond the range of the original data.
-We believe that our manuscript represents an important step toward a more holistic approach for forecasting species ranges.
-This is of broad interest due to the importance of range shifts for theoretical and applied ecology, and we hope you will consider it for the \emph{Ideas and Perspectives} section.
-
-% Suggestion from Isa for a novelty statement:
+We illustrate our approach using two examples, and demonstrate that the framework can reduce bias and uncertainty when projecting beyond the range of the original data.
+In order to provide a platform for continuing to develop the framework, we include all code and data for the examples as supporting information with the manuscript.
+Understanding range shifts are of broad interest for both theoretical and applied ecologists, and we believe that our manuscript represents an important step toward a more holistic approach for forecasting species ranges.
+We hope you will consider it for the \emph{Ideas and Perspectives} section. 
+We will be ready to submit the manuscript within one month.
 
 %TC:break closing
-Thank you for considering our submission. We will be ready to submit the manuscript within one month. We look forward to your reply.
-
 \closing{Sincerely,}

### 2014-09-26 tex/bibprep.pl

@@ -4,7 +4,7 @@ use strict;
 use warnings;
 
 my $inFilename = $ARGV[0];
-my @fileContents = [];
+my @fileContents;
 my $i = 1;
 my $newItem = 0;
 my $name = "";
@@ -26,5 +26,7 @@ close(inFile);
 
 open(inFile, ">", $inFilename) or die "Cannot open file $inFilename\n";
 foreach (@fileContents) {
+	print "$_\n";
 	print inFile "$_\n";
 }
+

### 2014-10-10 tex/manuscript.tex
Finished latest draft and finalized proposal letter
Added comments from various coauthors

@@
 
+\usepackage[pagewise]{lineno}
+
 %% ----------------------------------
@@
 %% ----------------------------------
-%\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
-%
-%\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
+\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
 
-% available colors:
-% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
+\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
+
+%%% available colors:
+%%% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
 %
@@
 %\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
-%\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
-%\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
+\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
+\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
 %\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
@@
 %\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
-%\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
+\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
 %\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
-%\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
-%\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
+\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
+\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
 %\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
-%\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
-%\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}
+\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
+\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}
 
@@
 %TC:break abstract
+\linenumbers
 \begin{abstract}
+\ia{Some section of the manuscript are a bit long and may be shorterned. I have made some suggestion but I think it should be shorten more.}
+\ia{The abstract would need a bit of work in the formulation and editing. I have made some suggestions.}
 % limited to 200 words
 There is great interest in modeling species ranges, resulting in a great diversity of modeling approaches.
+\dm{There is great interest in modeling species ranges – both potential and actual, resulting in a great diversity of modeling approaches.}
 However, in general, individual approaches include only a small subset of the total available information about a species. 
+\ia{delete first two sentences, replace with: \\
+Current worldwide preoccupations with species range shifts have resulted in a great diversity of approaches in species range modelling.
+Each approach typically includes only a small subset of the total available information about a species; correlative models customarily predict...}
 Correlative models often predict presence or absence as a function of climate, but ignore  smaller-scale processes such as growth, fecundity, and dispersal.
+\ia{see suggestion for a rewrite of the remainder:\\
+What’s more, these approaches frequently result in divergent predictions, with no straightforward method to reconcile them.
+We present a flexible framework allowing the integration of models across multiple scales.
+Using hierarchical Bayesian methods, we produce ametamodel which, by incorporating all of the information used as input for the original scale-specific sub-models, produce probabilistic estimates of species presence andreports uncertainty from both data and process error.
+We provide two contrasting examples to illustrate our approach, and demonstrate that it can substantially reduce uncertainty when projecting beyond the range of the original data.
+Our method could be an invaluable tool for conservation biologists and land managers, despite the technical challenge of its actual implementation.
+We strongly encourage collaboration between modelers and practitioners in overcoming these hurdles and achieving greater integration of knowledge, a goal which we are confident will be of wide interest.
+}
 Furthermore, the diversity of approaches often results in divergent predictions for the same species, with no simple way to reconcile these predictions.
@@
 The resulting metamodel produces probabilistic estimates of species presence, reports uncertainty from both data and process error from all sub-models, and incorporates all of the information used as input for the original scale-specific sub-models.
+\ks{Change above to:\\
+The resulting metamodel produces probabilistic estimates of species presence, reports uncertainty from both data and process error from all sub-models used, while also  synthesizing all of the information as an input for the original scale-specific sub-models.}
 We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
 We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
+\dg{That’s the smartest comment of the whole paper!!!}
 Although implementation of our method can be technically challenging, we anticipate the results will be of wide interest, and we encourage collaboration between modelers and practitioners in applying our framework.
@@
 Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
+\ia{This sentence is repetitive and somewhat contradictory with the last one of the paragraph. Suggest deletion}
 Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}.
-Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution \citep{Jay2012}, hybrid zone dynamics \citep{Engler2013}, and the impacts of climate change on species distributions \citep{Thuiller2011, Blois2013, Thuiller2014}.
+Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution \citep{Jay2012}, hybrid zone dynamics \citep{Engler2013}, and forecasting the impacts of climate change on species distributions \citep{Thuiller2011, Blois2013, Thuiller2014}.
 However, despite having recognized potential to support decision making, the contributions of these models to conservation and applied ecology remain limited \citep{Guisan2013}.
 
-Two important constraints on modeling ecological processes prevent models from making unbiased predictions with acceptable uncertainty: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
+\ia{Miss one sentence here. To make a link with the sentence above, you should briefly state why models contribution remains limited. }
+Two important constraints on modeling ecological processes can prevent models from making unbiased predictions with acceptable uncertainty: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
 Another constraint is a paucity of data over wide ranges of ecological conditions and scales.
+\dg{change to:\\
+The paucity of data over theses ecological conditions and scales is a severe constraint. }
 Poor data or incoherent models will result in predictions with unacceptably high uncertainty or precise but biased predictions.
@@
 Model diversification also arises due to the diverse processes generating species ranges \citep{Soberon2007}.
+\dg{Just to encourage self citation, this paper is more recent and supported by data, not only discussion: \\
+
+ Boulangeat, I., Gravel, D.
+and Thuiller, W.
+2012.
+Disentangling the underlying mechanisms of species abundance distribution using a comprehensive and nested modeling framework.
+Ecology Letters 15 : 584-593.
+}
 In the context of forecasting, which is the focus here, mechanistic and correlative approaches represent quite different modes of inference, but both have strong limitations. 
+\ia{delete "which is the focus here"}
 Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, they may perform poorly when applied at the scale of species ranges.
-For instance, biotic interactions (including trophic interactions) are usually not modeled mechanistically at the scale of species distributions because they are not well-known, or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Soberon2007, Pigot2013}.
-On the other hand, coarse-scale correlative models statistically relating species occurrences to associated distributions of other variables at the same scale have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
+For instance, biotic interactions (including trophic interactions) are usually not modeled mechanistically at the scale of species distributions because they are poorly known, or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Soberon2007, Pigot2013}.
+On the other hand, coarse-scale correlative models statistically relating species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
 However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}.
@@
 Furthermore, correlative models often do not explicitly incorporate ecological processes into their formulations, and when applied at coarse scales, they may miss important fine-scale processes, making a multi-scale approach desirable when accurate forecasting is required \citep{Thuiller2013}.
+\ia{replace: \\
+ desirable when accurate forecasting is required \\
+with:
+ more suitable}
 On the other hand process-based models often require a variety of assumptions which may be poorly understood across space and time.
-
+\ia{Finish abruptely here. Feels like one sentence is missing}
+\ks{Feels like it is getting snuck in here…almost an after thought… I would recommend either removing it or developing it a tad further. }
 %% ----------------------------------
@@
 \subsection*{Toward an integrated approach for modeling ranges}
-To integrate knowledge about species ranging from the stand to the landscape, we present an application of hierarchical Bayesian methods for integrating the predictions of multiple models.
-Our approach is an alternative to other approaches to drawing inference from multiple models that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
+To include knowledge of species from multiple sources and scales into distribution models, we present an application of hierarchical Bayesian methods that uses outputs from multiple model outputs to inform the results of the final model.
+Our approach is an alternative to other approaches to multimodel inference that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
 The challenge is to account for all this available knowledge while producing a single prediction, potentially mitigating the limitations inherent in each modeling approach and contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
-Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Catterall2012, Strigul2012, Soranno2014}. 
+\ia{replace above sentence with: \\
+``By producing a single prediction from multiple models, we aim at mitigating the limitations inherent in each modeling approach and contributing to more robust predictions ''}
+\ks{change previous 2 sentences to:\\
+Our approach succeeds in providing a much more throrough implementation of data from multiple models, while incorportating multiple scales and transparency in our estimation of uncertainty.
+A number of challenge(s) arised with this approach: (1) accounting for all the available data, (2) producing a single congruent prediction from all the models implemented, and (3) mitigating the potential limitations within each modeling approach }
+Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Boulangeat2012, Catterall2012, Strigul2012, Soranno2014}.
 Recent literature has also focused on the development of hybrid models, which allow for flexible combinations between mechanistic and phenomenological sub-models \citep{Gallien2010, Thuiller2013}. 
 For instance, correlative models are used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while mechanistic approaches can include space and time dynamics and therefore account for dispersal processes \citep{Kearney2008}.
-This complementarity has been used to merge the two types of models into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Boulangeat2014}. 
-However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are not easy to test \citep{Gallien2010}, and uncertainties are very difficult to quantify and identify from its many different sources.
+This complementarity has been used to merge the two model types into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Boulangeat2014}. 
+\wt{Dormann et al. suggested to separate hybrid and integrated. 
+Dormann, C.F., Schymanski, S.J., Cabral, J.S., Chuine, I., Graham, C.H., Hartig, F., Kearney, M., Morin, X., Römermann, C., Schröder, B.
+\& Singer, A.
+(2012) Correlation and process in species distribution models: bridging a dichotomy.
+Journal of Biogeography,
+}
+However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are not easy to test \citep{Gallien2010},
+\wt{and hardly known}
+and uncertainties are very difficult to quantify and identify from its many different sources. 
 Moreover, hybrid approaches are not suitable for merging models that are based on the same processes but with different underlying hypotheses.
@@
 A second approach is to directly combine predictions.
-If models operate at the same scales and have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}.
+\ia{A second approach recently explored by modelers is to directly combine predictions.}
+If models operate at the same scale and have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}.
 If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Serra-Diaz2013}.
-However, the applicability of ensemble forecasts are limited; for example, it is not currently possible to evaluate the effects of convergent predictions on total uncertainty of outcomes.
-One of the greatest challenges inherent in both approaches above is estimating uncertainty, which is of utmost importance in a prediction context.
+\wt{Cheaib, A., Badeau, V., Boe, J., Chuine, I., Delire, C., Dufrêne, E., François, C., Grittid, E.S., Legay, M., Pagé, C., Thuiller, W., Viovy, N.
+\& Leadley, P.
+(2012) Climate change impacts on tree ranges: model inter-comparison facilitates understanding and quantification of uncertainty.
+Ecology Letters, 15, 533-544.}
+However, the applicability of ensemble forecasts is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes.
+\ia{paragraph break here}
+One of the greatest challenges inherent in both approaches is estimating uncertainty, which is of utmost importance in a prediction context.
 A more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying parameters that make large contributions to the total uncertainty in the model predictions \citep{McMahon2011}.
 In models that encompass multiple scales or levels of organization, it is particularly challenging to evaluate all sources of uncertainty, especially those related to model specification (\citealt{Calder2003} but see \citealt{Conlisk2013}).
+\dg{‘uncertainty’ is used 6 times over 5 sentences. Would help some rewording to limit redundancy}
 Clearly, an approach that integrates predictions and uncertainty from multiple model types is needed \citep{Beck2012, Thuiller2013}. 
@@
 Here, we present a framework for modeling species range dynamics that aims to account for all the information available on a focal species, from a variety of sources and scales.
-Contrary to hybrid and hierarchical models, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from sub-models at a variety of spatial scales, allowing as much flexibility as possible regarding the type of information included.
+Contrary to hybrid and hierarchical models, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for as much flexibility as possible regarding the type of information included.
 We use the power and generality of a hierarchical Bayesian framework, which allows us to include multiple data sources and modes of inference \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}. 
@@
 First, we use simulated data to present a complete example of the application of the framework to multiple information sources that are relevant at very different scales.
-In a second example, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}) to climate with the predictions of a phenological model, Phenofit, with the goal of improving uncertainty estimation when projecting model predictions to future climate.
+In a second example, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}), a widespread foundation tree species from eastern North America, to climate with the predictions of a phenological model, Phenofit, with the goal of improving uncertainty estimation when projecting model predictions to future climate.
+\dg{ref missing for phenofit}
 
@@
 The key idea of our approach is to constrain the parameters of a correlative metamodel using predictions made by one or several mechanistic sub-models from other spatial scales.
+\wt{does not really have to, right? (wrt different scales)}
 The role of the metamodel is to integrate data at the same ecological scale as predictions. 
@@
 The output of sub-models should provide additional predictions at a comparable ecological scale at which the metamodel operates, although it is not necessary for the sub-models to actually operate at the same scale.
-For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, as we will show in the first example, these functions can be simple while still contributing useful information to the metamodel.
+For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, these functions can be simple while still contributing useful information to the metamodel (see Example 1).
 All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Box 1, Fig. \ref{fig:diagram}).
-We present the complete framework in a hypothetical example where experimental results are  introduced as prior knowledge of the fundamental niche in a species distribution model (SDM).
-A more formal presentation of the modeling framework appears in Box 1.
-We follow with a second example where we apply our approach to integrate phenological information with occurrence data for a widespread North American tree.
+We present and apply the framework here with two examples, and follow with a more formal presentation in Box 1.
 
@@
 \subsection*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
+\ia{To keep the ms short, the 2 examples can be summarized a bit more and the details put in the Appendix}
 
@@
 \end{equation}
-We further assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
-The model is evaluated by relating these observations to environmental data \((D_M)\), which for the sake of the example consists of mean annual temperature $T_M$ and annual precipitation $P_M$. 
-For simplicity, we use a simple logistic model for the metamodel \((\theta_M)\) with a second order effect of temperature and precipitation. 
+We assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
+The model is evaluated by relating these observations to environmental data \((D_M)\), which, for the sake of the example, consists of mean annual temperature $T_M$ and annual precipitation $P_M$. 
+For simplicity, we formulate the metamodel \((\theta_M)\) as a simple logistic regression with a second order effect of temperature and precipitation. 
 This naive model (i.e., the metamodel with no constraints from sub-models) predicts  occurrence probability (\(\psi_N\)) as a function of the environment:
@@
 %-----
-where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
+where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), 
+\wt{model parameters?}
+\(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
 
@@
 
-Although the two data sets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation).
+Although the two datasets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation).
 This scaling is quite simplistic, and would never be used to predict a species' range using this mechanistic sub-model in isolation.
@@
 
-We performed this procedure with our hypothetical data sets.
+We performed this procedure with our hypothetical datasets.
 When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data.
-Unsurprisingly given the quality of the data, the sub-model was highly precise, providing a fairly strong constraint when producing the integrated model.
+\ks{add ref to figure}
+Unsurprisingly given the quality of the data, the sub-model was highly precise with respect to precipitation and not useful for temperature, thus providing a fairly strong constraint when producing the integrated model.
 The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
@@
 
-Once the technical challenges of evaluating the parameters are overcome, the integration has several advantages over the both the naive parameterization of the metamodel and the mechanistic sub-model.
+Once the technical challenges of evaluating the parameters are overcome, integration has several advantages over both the naive parameterization of the metamodel and the mechanistic sub-model. 
 In the situation where both models strongly agree on the relationship between water availability and occurrence probability, then we should expect significantly reduced uncertainty in that domain (Figs.
@@
 \subsection*{Example 2: Constraining an SDM using phenological information}
-For the second example, we consider the problem of predicting the future distribution of a species following climate change.
+For the second example, we consider the problem of predicting the future distribution 
+\wt{forecasting the potential distribution of a species}
+of a species following climate change.
 There is considerable interest in comparing correlative and mechanistic projections with respect to climate change \citep{Morin2009}, and correctly characterizing uncertainty is a critical aspect of this problem \citep{Cheaib2012}.
-Despite being a relatively common application of \ac{SDM}s \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios is problematic for a number of reasons previously discussed \citep{Araujo2006}.
+Despite being a relatively common application of \ac{SDM}s \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios remains problematic \citep{Araujo2006}.
 We used our metamodeling framework to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
@@
 Furthermore, variance in the predictions of the naive model was relatively low throughout the spatial domain (Fig \ref{fig:ex2}E).
+\ks{provide specific figures for variance of the different models}
 Similar patterns were repeated for analogous climates under the future scenario (Fig \ref{fig:ex2}D); however these predictions differed considerably from the predictions of the mechanistic sub-model (Fig \ref{fig:ex2}B), and uncertainty for the projection was relatively high, particularly for eastern Canada (Fig \ref{fig:ex2}F).
-In general, Phenofit was predicted considerably less northward migration by 2080.
+In general, Phenofit predicted considerably less northward migration by 2080.
 The integrated model, which included both the presence-absence data as well as the Phenofit predictions for the future, made intermediate predictions in terms of northward range shifts for sugar maple (Fig \ref{fig:ex2}H).
-Globally, the integrated model expressed somewhat more uncertainty than the naive model (Fig \ref{fig:ex2}J), reflecting the divergent predictions of Phenofit and the naive model.
+Globally, the integrated model expressed somewhat more uncertainty 
+\wt{for the future}
+than the naive model (Fig \ref{fig:ex2}J), reflecting the divergent predictions of Phenofit and the naive model.
+\wt{Are you sure? From the map, I would have said the reverse.}
 However, in areas where the naive model performed worst (e.g., eastern Canada), the level of uncertainty in the integrated model was similar to the uncertainty in other locations.
@@
 The methods provided here expand upon the motivation of hybrid models to develop a more robust and unifying approach for ecological models in a predictive context while striving to overcome some of the limitations characteristic of other integrated approaches.
-Attempting to incorporate underlying ecological processes so that SDMs are not oversimplified by combining different model parameters or scales can be challenging using these hybrid approaches. 
-In particular, it can often be difficult to identify parameters that can be used to connect different modeling frameworks and what parameters can produce meaningful responses \citep{Thuiller2013}. 
+Attempting to incorporate underlying ecological processes so that SDMs are not oversimplified by combining different model parameters or scales can be challenging using hybrid approaches. 
+In particular, it can often be difficult to identify parameters that can be used to connect different modeling frameworks and which parameters can produce meaningful responses \citep{Thuiller2013}. 
 The methods presented in this study aim to overcome some of these limitations through the use of Bayesian statistics. 
@@
 The estimation of model parameters (and predictions) as distributions can also help decrease overparameterization that can be a concern for hybrid models. 
-Bayesian frameworks also provide a natural framework for the incorporation of prior information. 
+Bayesian methods also provide a natural framework for the incorporation of multiple sources of information. 
 The inability or difficulty to include information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Thuiller2008, Smolik2010}. 
-Even when a link can be formed to include this information, it often is oversimplified to simplify computation \citep{Gallien2010}. 
-Finally, Bayesian methods inherently allow for feedbacks or interactions between sub models, which may portray a more realistic response to many environmental dynamics where factors may simultaneously influence the change of one another.
+\wt{Thuiller, W., Münkemüller, T., Schiffers, K.H., Georges, D., Dullinger, S., Eckhart, V.M., Edwards, T.C., Gravel, D., Kunstler, G., Merow, C., Moore, K., Piedallu, C., Vissault, S., Zimmermann, N.E., Zurell, D.
+\& Schurr, F.M.
+(2014) Does probability of occurrence relate to population dynamics?
+Ecography,}
+Even when a link can be formed to include this information, it often is oversimplified to facilitate computation \citep{Gallien2010}. 
+Finally, Bayesian methods inherently allow for feedbacks or interactions between sub models, which may be a more realistic representation of ecological dynamics where many factors may simultaneously influence the system.
 
 Our approach is an original application of Bayesian multilevel modeling, and can be considered a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales while using several models simultaneously.
-In particular, this development has certain similarities with Bayesian model averaging, Bayesian calibration of process-based models and hierarchical Bayesian modeling. 
+In particular, this development has certain similarities with Bayesian model averaging, Bayesian calibration of process-based models, and hierarchical Bayesian modeling. 
 Bayesian model averaging aims to combine several alternative models to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.,][]{Wintle2003, Link2006}.
 Bayesian model averaging considers models operating at the same scale and the posterior distribution is usually determined with Gibbs sampling. 
-Bayesian calibration of process-based model focuses on uncertainty of the parameter values in the process-based models, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, Hartig2012}.
+Bayesian calibration of process-based model focuses on uncertainty of the parameter values, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, Hartig2012}.
 In contrast with Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of a generally more correlative metamodel.
@@
 For example, the metamodel can be used as one input in an ensemble forecasting scheme that also incorporates other common SDM formulations (e.g., Random Forest, MaxEnt).
+\wt{That would be  a bit weird, is not it? I do not see the point of incapsulating the metamodel into a more basic model. I would actually drop that last sentence}
 
@@
 \subsection*{Advantages of Model Integration}
-Models are important tools that are increasingly used by land managers to face challenges associated to decision-making \citep{Guisan2013}
-A growing number of models are available, but they can provide diverging answers to very similar questions due to the differences in their assumptions and methodology. 
-This can create confusion and even some mistrust towards models, and as a consequence some managers may be discouraged to incorporate model results in their management plans. 
+Species distribution models are important tools that are increasingly used by land managers to face challenges associated to decision-making \citep{Guisan2013}.
+A growing number of approaches are available, but they can provide diverging answers to very similar questions due to the differences in their assumptions and methodology. 
+This can create confusion and even some mistrust towards models, and some managers may be discouraged from incorporating model results in their management plans. 
 Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organizations around the world \citep[e.g.,][]{Bernier2013}.
@@
 Successful use of an integrated modeling approach will always remain dependent on an intimate understanding of the decision-making process by modelers, emphasizing the importance of close collaboration between with practitioners at all stages of model development \citep{Guisan2013}.
+\dm{add\\
+Notably the connection between modelers and decision-makers who would use these types of modelsis also critical and the complexities of this relationship should not be trivialized.}
 
@@
 One of the main strengths of our approach is that it allows for a clear, transparent identification of uncertainties and how they are transmitted throughout the modeling process. 
-Transparency in uncertainty can be considered as a sort of sensitivity analysis, in which areas with large uncertainty can be detected and new experimental research or additional data collection can be designed (e.g., Example 1, Figs.
+Transparency in uncertainty can be considered as a sort of sensitivity analysis, in which areas with large uncertainty can be detected and new experimental research or additional data collection can be designed accordingly (e.g., Example 1, Figs.
 \ref{fig:ex1_sampling}, \ref{fig:ex1_precip}).
@@
 Our framework could be used \emph{post-hoc} on the outputs of these models to quantify uncertainty resulting from model disagreement.
-Our goal in this paper is to provide a framework for including much of this existing knowledge about species' responses to the environment into species distribution models.
 
@@
 
+\dg{but we should also see our approach as a mean to circumvent cases where this happens, since additional information will still be better than anything else}
 In a similar fashion, the quality and availability of data impose a significant constraint on the number and type of models that can be implemented.
 The capacity to implement a model is very low if data requirements are prohibitive. 
-Adequate coverage of explanatory variables is a significant obstacle, and exploratory analyses can be a significant aid in understanding how data coverage impacts resulting predictions \citep{Mckenney2002}.
+Adequate and unbiased coverage of explanatory variables is a significant obstacle, and exploratory analyses can be a significant aid in understanding how data coverage impacts resulting predictions \citep{Mckenney2002}.
 Integration can ameliorate these issues to some extent by using supplemental information (and conceptual advances) in additional sub-models where coverage is weak (e.g., Example 1, Figs.
 \ref{fig:ex1_precip}, \ref{fig:ex1_map}).
-For example, \citet{Freckleton2009} estimated that data are too scant to successfully develop highly mechanistic models predicting weed population numbers. 
+For example, \citet{Freckleton2009} estimated that data are too limited to successfully develop highly mechanistic models predicting weed population numbers. 
 A strong asset of our approach is that it can be used without the full suite of data that would be required to run a fully mechanistic model. 
@@
 In the context of integrated models, the challenge is three fold: (i) which spatial and temporal scales (i.e.
-which processes), are to be considered, (ii) how to build a scaling function that is consistent with the metamodel and (iii) how error and uncertainty are propagated from the scales of the sub-models scale to that of the metamodel.
-Although we argue that our proposed framework is able to easily deal with very different scales and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered, the building of scaling function deserves further investigation.
+which processes), are to be considered, (ii) how to build a scaling function that is consistent with the metamodel and (iii) how error and uncertainty propagate from the sub-models to the metamodel.
+Although we argue that our proposed framework is able to easily deal with very different scales and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered, the building of scaling functions deserves further investigation.
 We have shown in our examples that even simple scaling functions can provide reasonable constraints on the metamodel. 
-However, with the objective to take into account all know processes and models, it is likely that the modeling process may include multiple scaling functions, each at different scales. 
+However, it is likely that the modeling process may include multiple scaling functions operating at different scales when taking all known processes and models into account. 
 Indeed, if species distributions are a function of, e.g.
@@
 
-In an applied context, the results of models predicting species ranges are increasingly used by resource managers, conservation biologists, or forest ecologists for formulating or adjusting recommendations.
+In an applied context, the results of models predicting species ranges are increasingly used by resource managers, conservation biologists, and forest ecologists for formulating and adjusting recommendations.
 The utility of such models depends on their ability to help evaluate events beyond the bounds of the available data, including in future situations constrained by climate change. 
@@
 However, the same was true of all statistical techniques or modeling approaches when they were initially developed, and many of the conceptual challenges associated with Bayesian statistics are being reduced as they gain recognition in the scientific literature.
-We therefore argue that our proposed approach as a strong potential for direct use in our real world where climate is quickly changing and conservation practices must be adjusted accordingly.
+We therefore argue that our proposed approach has a strong potential for direct use in our real world where climate is quickly changing and conservation practices must be adjusted accordingly.
 We further argue that forging stronger collaborations between modelers, decision-makers, and practitioners will improve the incorporation of this and other new methods into applied ecology and conservation.
@@
 %TC:ignore
+\nolinenumbers
 \section*{Acknowledgements}
@@
 Comments from Charles Canham and William Godsoe improved a previous version of this manuscript.
+WT received support funding from the European Research Council under the European Community's Seven Framework Programme FP7/2007-2013 Grant Agreement no.
+281422 (TEEMBIO).
 %TC:endignore
@@
 \textbf{Figure \ref{fig:ex1_sampling}:}
-	Two simulated data sets used to illustrate the model integration framework.
+	Two simulated datasets used to illustrate the model integration framework.
 	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
@@
 	Uncertainty in future projections was quite high in the eastern portion of the range (F) relative to uncertainty in predictions for the present (E).
-	Model integration resulted in poor predicitons for present range (G), with future predictions that were intermediate between the two models (H).
+	Model integration resulted in poor predictions for present range (G), with future predictions that were intermediate between the two models (H).
 	Uncertainty in the integrated model (I, J) was generally greater than in the naive model, reflecting the disagreement between the naive model and the sub-model, but there were fewer extremes in model uncertainty when projecting the integrated model (J) compared with the naive model (F).
@@
 \newpage
+\linenumbers
 \section*{Box 1: Model description}
@@
 
-This procedure may be repeated for an arbitrary number of models (limited only by available data and computational power). 
-The models may be implemented simultaneously, when multiple data sets are available and can be evaluated under different underlying models, or they may be run sequentially, updating the posterior distribution of $\theta_M$ as new information becomes available.
+This procedure may be repeated 
+\wt{replace by applied or used. 
+Here looks like you have to repeat this to add a new model for instance. }
+for an arbitrary number of models (limited only by available data and computational power). 
+The models may be implemented simultaneously, when multiple datasets are available and can be evaluated under different underlying models, or they may be run sequentially, updating the posterior distribution of $\theta_M$ as new information becomes available.
 Note that $\theta_S$ itself need not be implemented in a Bayesian framework and the output need not be identical in form to the output of \(\theta_M\); it is enough to produce an output that can be evaluated as a likelihood under the first model and to have some assessment of the confidence in the parameter estimates of $\theta_S$ to use as a prior.

### 2014-10-13 tex/manuscript.tex
First round of edits to the manuscript

@@
 %
-%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
+\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
 %\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
-%\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
+\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
 \newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
@@
 \begin{abstract}
-\ia{Some section of the manuscript are a bit long and may be shorterned. I have made some suggestion but I think it should be shorten more.}
-\ia{The abstract would need a bit of work in the formulation and editing. I have made some suggestions.}
 % limited to 200 words
-There is great interest in modeling species ranges, resulting in a great diversity of modeling approaches.
-\dm{There is great interest in modeling species ranges – both potential and actual, resulting in a great diversity of modeling approaches.}
-However, in general, individual approaches include only a small subset of the total available information about a species. 
-\ia{delete first two sentences, replace with: \\
-Current worldwide preoccupations with species range shifts have resulted in a great diversity of approaches in species range modelling.
-Each approach typically includes only a small subset of the total available information about a species; correlative models customarily predict...}
+Current interest in modeling potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
+However, in general, individual approaches include only a small subset of the available information about a species. 
 Correlative models often predict presence or absence as a function of climate, but ignore  smaller-scale processes such as growth, fecundity, and dispersal.
-\ia{see suggestion for a rewrite of the remainder:\\
-What’s more, these approaches frequently result in divergent predictions, with no straightforward method to reconcile them.
-We present a flexible framework allowing the integration of models across multiple scales.
-Using hierarchical Bayesian methods, we produce ametamodel which, by incorporating all of the information used as input for the original scale-specific sub-models, produce probabilistic estimates of species presence andreports uncertainty from both data and process error.
-We provide two contrasting examples to illustrate our approach, and demonstrate that it can substantially reduce uncertainty when projecting beyond the range of the original data.
-Our method could be an invaluable tool for conservation biologists and land managers, despite the technical challenge of its actual implementation.
-We strongly encourage collaboration between modelers and practitioners in overcoming these hurdles and achieving greater integration of knowledge, a goal which we are confident will be of wide interest.
-}
-Furthermore, the diversity of approaches often results in divergent predictions for the same species, with no simple way to reconcile these predictions.
+Furthermore, different approaches often produce divergent predictions, with no simple method to reconcile them.
 We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
-The resulting metamodel produces probabilistic estimates of species presence, reports uncertainty from both data and process error from all sub-models, and incorporates all of the information used as input for the original scale-specific sub-models.
-\ks{Change above to:\\
-The resulting metamodel produces probabilistic estimates of species presence, reports uncertainty from both data and process error from all sub-models used, while also  synthesizing all of the information as an input for the original scale-specific sub-models.}
+Our method results in a metamodel that produces probabilistic estimates of species presence, reports uncertainty from both data and process error from all sub-models, and incorporates all of the information used as input for the original scale-specific sub-models.
 We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
 We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
-\dg{That’s the smartest comment of the whole paper!!!}
 Although implementation of our method can be technically challenging, we anticipate the results will be of wide interest, and we encourage collaboration between modelers and practitioners in applying our framework.
@@
 Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
-\ia{This sentence is repetitive and somewhat contradictory with the last one of the paragraph. Suggest deletion}
 Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}.
 Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution \citep{Jay2012}, hybrid zone dynamics \citep{Engler2013}, and forecasting the impacts of climate change on species distributions \citep{Thuiller2011, Blois2013, Thuiller2014}.
-However, despite having recognized potential to support decision making, the contributions of these models to conservation and applied ecology remain limited \citep{Guisan2013}.
 
-\ia{Miss one sentence here. To make a link with the sentence above, you should briefly state why models contribution remains limited. }
-Two important constraints on modeling ecological processes can prevent models from making unbiased predictions with acceptable uncertainty: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
-Another constraint is a paucity of data over wide ranges of ecological conditions and scales.
-\dg{change to:\\
-The paucity of data over theses ecological conditions and scales is a severe constraint. }
+Despite the recognized potential of these models, difficulties in producing predictions with acceptable precision and low bias has limited their contributions \citep{Guisan2013}.
+Two important constraints on modeling ecological processes can prevent models from making unbiased predictions with acceptable uncertainty: 
+(1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
 Poor data or incoherent models will result in predictions with unacceptably high uncertainty or precise but biased predictions.
 In recent decades, however, modeling techniques have proliferated to take advantage of what is available in terms of both data and modeling platforms.
-Model diversification also arises due to the diverse processes generating species ranges \citep{Soberon2007}.
-\dg{Just to encourage self citation, this paper is more recent and supported by data, not only discussion: \\
-
- Boulangeat, I., Gravel, D.
-and Thuiller, W.
-2012.
-Disentangling the underlying mechanisms of species abundance distribution using a comprehensive and nested modeling framework.
-Ecology Letters 15 : 584-593.
-}
-In the context of forecasting, which is the focus here, mechanistic and correlative approaches represent quite different modes of inference, but both have strong limitations. 
-\ia{delete "which is the focus here"}
+Model diversification also arises due to the diversity of processes generating species ranges \citep{Boulangeat2012}.
+In the context of forecasting, mechanistic and correlative approaches represent quite different modes of inference, but both have strong limitations. 
 Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, they may perform poorly when applied at the scale of species ranges.
-For instance, biotic interactions (including trophic interactions) are usually not modeled mechanistically at the scale of species distributions because they are poorly known, or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Soberon2007, Pigot2013}.
+For instance, biotic interactions (including trophic interactions) are usually not modeled mechanistically at the scale of species distributions because they are poorly known, or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Pigot2013}.
 On the other hand, coarse-scale correlative models statistically relating species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
@@
 Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
-In cases where a selected explanatory variable indirectly accounts for a process (e.g., where the strength of biotic interactions are correlated with an environmental variable), extrapolation will produce biased predictions in temporal and spatial domains where the process and the environmental variable are decoupled.
-Furthermore, correlative models often do not explicitly incorporate ecological processes into their formulations, and when applied at coarse scales, they may miss important fine-scale processes, making a multi-scale approach desirable when accurate forecasting is required \citep{Thuiller2013}.
-\ia{replace: \\
- desirable when accurate forecasting is required \\
-with:
- more suitable}
+Furthermore, correlative models often do not explicitly incorporate ecological processes into their formulations, and when applied at coarse scales, they may miss important fine-scale processes, making a multi-scale approach more suitable when accurate forecasting is required \citep{Thuiller2013}.
 On the other hand process-based models often require a variety of assumptions which may be poorly understood across space and time.
+\mt{This last sentence could go, and the sentence before is actually a bit repetitive. Suggest reorganizing a bit to still end on this point, but reduce repetition}
+\aa{you mentioned fine-scale mechanistic models and, on the other hand, coarse-scale correlative models.
+To talk now about process-based models sounds a bit bizarre and may make the reader wonder why are you mentioning them now.}
 \ia{Finish abruptely here. Feels like one sentence is missing}
@@
 Our approach is an alternative to other approaches to multimodel inference that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
-The challenge is to account for all this available knowledge while producing a single prediction, potentially mitigating the limitations inherent in each modeling approach and contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
-\ia{replace above sentence with: \\
-``By producing a single prediction from multiple models, we aim at mitigating the limitations inherent in each modeling approach and contributing to more robust predictions ''}
-\ks{change previous 2 sentences to:\\
-Our approach succeeds in providing a much more throrough implementation of data from multiple models, while incorportating multiple scales and transparency in our estimation of uncertainty.
-A number of challenge(s) arised with this approach: (1) accounting for all the available data, (2) producing a single congruent prediction from all the models implemented, and (3) mitigating the potential limitations within each modeling approach }
+By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
 Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Boulangeat2012, Catterall2012, Strigul2012, Soranno2014}.
@@
 For instance, correlative models are used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while mechanistic approaches can include space and time dynamics and therefore account for dispersal processes \citep{Kearney2008}.
-This complementarity has been used to merge the two model types into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Boulangeat2014}. 
-\wt{Dormann et al. suggested to separate hybrid and integrated. 
-Dormann, C.F., Schymanski, S.J., Cabral, J.S., Chuine, I., Graham, C.H., Hartig, F., Kearney, M., Morin, X., Römermann, C., Schröder, B.
-\& Singer, A.
-(2012) Correlation and process in species distribution models: bridging a dichotomy.
-Journal of Biogeography,
-}
-However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are not easy to test \citep{Gallien2010},
-\wt{and hardly known}
-and uncertainties are very difficult to quantify and identify from its many different sources. 
+This complementarity has been used to merge the two model types into hybrid models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Dormann2012, Boulangeat2014}. 
+However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and the sources and magnitudes of uncertainties are difficult to quantify.
 Moreover, hybrid approaches are not suitable for merging models that are based on the same processes but with different underlying hypotheses.
 
-A second approach is to directly combine predictions.
-\ia{A second approach recently explored by modelers is to directly combine predictions.}
+A commonly-employed alternative to hybrid models is the direct combination of predictions.
 If models operate at the same scale and have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}.
-If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Serra-Diaz2013}.
-\wt{Cheaib, A., Badeau, V., Boe, J., Chuine, I., Delire, C., Dufrêne, E., François, C., Grittid, E.S., Legay, M., Pagé, C., Thuiller, W., Viovy, N.
-\& Leadley, P.
-(2012) Climate change impacts on tree ranges: model inter-comparison facilitates understanding and quantification of uncertainty.
-Ecology Letters, 15, 533-544.}
-However, the applicability of ensemble forecasts is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes.
-\ia{paragraph break here}
-One of the greatest challenges inherent in both approaches is estimating uncertainty, which is of utmost importance in a prediction context.
-A more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying parameters that make large contributions to the total uncertainty in the model predictions \citep{McMahon2011}.
-In models that encompass multiple scales or levels of organization, it is particularly challenging to evaluate all sources of uncertainty, especially those related to model specification (\citealt{Calder2003} but see \citealt{Conlisk2013}).
-\dg{‘uncertainty’ is used 6 times over 5 sentences. Would help some rewording to limit redundancy}
+If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Cheaib2012, Serra-Diaz2013}.
+However, the applicability of ensemble forecasts is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
+A more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying the parameters that contribute most to variance in the model predictions \citep{McMahon2011}.
 Clearly, an approach that integrates predictions and uncertainty from multiple model types is needed \citep{Beck2012, Thuiller2013}. 
@@
 First, we use simulated data to present a complete example of the application of the framework to multiple information sources that are relevant at very different scales.
-In a second example, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}), a widespread foundation tree species from eastern North America, to climate with the predictions of a phenological model, Phenofit, with the goal of improving uncertainty estimation when projecting model predictions to future climate.
-\dg{ref missing for phenofit}
+In a second example, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}), a widespread foundation tree species from eastern North America, to climate with the predictions of a phenological model, Phenofit \citep{Chuine2001}, with the goal of improving uncertainty estimation when projecting model predictions to future climate.
 
@@
 
-The key idea of our approach is to constrain the parameters of a correlative metamodel using predictions made by one or several mechanistic sub-models from other spatial scales.
-\wt{does not really have to, right? (wrt different scales)}
-The role of the metamodel is to integrate data at the same ecological scale as predictions. 
-As an example, the metamodel could take the form of a simple correlative \ac{SDM}.
+The key idea of our approach is formulate a metamodel that integrates data at the same ecological scale as the desired predictions.
+The parameters of this model can be constrained using the output of one or more sub-models that can operate at different scales than the metamodel.
+As an example, the metamodel could take the form of a simple correlative \ac{SDM}, with one or more fine-scaled process-based sub-models.
 Ideally, each sub-model incorporates different data types and hypotheses that specify underlying ecological mechanisms contributing to the model output.
@@
 \subsection*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
-\ia{To keep the ms short, the 2 examples can be summarized a bit more and the details put in the Appendix}
 
@@
 %-----
-where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), 
-\wt{model parameters?}
-\(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
+where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
 
@@
 We performed this procedure with our hypothetical datasets.
-When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data.
-\ks{add ref to figure}
-Unsurprisingly given the quality of the data, the sub-model was highly precise with respect to precipitation and not useful for temperature, thus providing a fairly strong constraint when producing the integrated model.
+When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data (Figs.
+\ref{fig:ex1_precip}a, \ref{fig:ex1_map}a, \ref{fig:ex1_map}b)
+Unsurprisingly given the quality of the data, the sub-model was highly precise with respect to precipitation and not useful for temperature, thus providing a fairly strong constraint when producing the integrated model (Fig \ref{fig:ex1_precip}b).
 The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
@@
 \subsection*{Example 2: Constraining an SDM using phenological information}
-For the second example, we consider the problem of predicting the future distribution 
-\wt{forecasting the potential distribution of a species}
-of a species following climate change.
+For the second example, we consider the problem of forecasting the potential distribution of a species following climate change.
 There is considerable interest in comparing correlative and mechanistic projections with respect to climate change \citep{Morin2009}, and correctly characterizing uncertainty is a critical aspect of this problem \citep{Cheaib2012}.
@@
 We used a binomial GLM to relate the suitability for sugar maple to 6 climate variables, scaled to the same resolution (see Appendix S2 for details on the climate variables).
+\aa{I would cite here the climatic variables used (just cite them, without further description)}
 We selected a GLM here for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
@@
 
+\mt{Comments about this from the group discussion.
+The integrated present predictions are confusing, because that panel is not addressing the main modeling question.
+It would be a good idea to emphasize this; that the modeling approach reflects the questions being asked, and it is important to carefully identify these questions.}
 For present climates, the predictions of Phenofit (Fig \ref{fig:ex2}A) were largely similar to those of the naive SDM (Fig \ref{fig:ex2}C). 
@@
 Similar patterns were repeated for analogous climates under the future scenario (Fig \ref{fig:ex2}D); however these predictions differed considerably from the predictions of the mechanistic sub-model (Fig \ref{fig:ex2}B), and uncertainty for the projection was relatively high, particularly for eastern Canada (Fig \ref{fig:ex2}F).
+\aa{ instead of calling it 'mechanistic sub-model' I suggest saying Phenofit, for consistency.}
 In general, Phenofit predicted considerably less northward migration by 2080.
@@
 Unsurprisingly, integration resulted in worse predictions for the present distribution of the species (Fig \ref{fig:ex2}G) compared with either the naive model or Phenofit in isolation.
+\aa{it may not be unsurprising that integration yields worse predictions than the other 2 models, but it is striking to see that it predicts such high suitabilities in southeastern US.
+In my opinion, it is not sufficiently explained why this should be considered as expected.
+I may have missed the point, but this is likely to happen to other readers also.}
+\mt{This ties in with the group discussion in Grenoble}
 Conventional SDMs are quite good at producing empirical models of species distributions under present climatic regimes, and the naive model here could be further improved with the addition of more parameters or the inclusion of ensemble forecasting.
@@
 Bayesian methods also provide a natural framework for the incorporation of multiple sources of information. 
-The inability or difficulty to include information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Thuiller2008, Smolik2010}. 
-\wt{Thuiller, W., Münkemüller, T., Schiffers, K.H., Georges, D., Dullinger, S., Eckhart, V.M., Edwards, T.C., Gravel, D., Kunstler, G., Merow, C., Moore, K., Piedallu, C., Vissault, S., Zimmermann, N.E., Zurell, D.
-\& Schurr, F.M.
-(2014) Does probability of occurrence relate to population dynamics?
-Ecography,}
+The inability or difficulty to include information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Thuiller2008, Smolik2010, Thuiller2014a}.
 Even when a link can be formed to include this information, it often is oversimplified to facilitate computation \citep{Gallien2010}. 
@@
 Bayesian model averaging considers models operating at the same scale and the posterior distribution is usually determined with Gibbs sampling. 
-Bayesian calibration of process-based model focuses on uncertainty of the parameter values, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, Hartig2012}.
-In contrast with Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of a generally more correlative metamodel.
+Bayesian calibration of process-based models focuses on uncertainty of the parameter values, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, Hartig2012}.
+In contrast with Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of the metamodel.
 Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005}, and climate and atmospheric modeling \citep{Mcmillan2010, Kang2012}.
@@
 Furthermore, the metamodel can easily be incorporated into a broader multimodel inference scheme.
-For example, the metamodel can be used as one input in an ensemble forecasting scheme that also incorporates other common SDM formulations (e.g., Random Forest, MaxEnt).
-\wt{That would be  a bit weird, is not it? I do not see the point of incapsulating the metamodel into a more basic model. I would actually drop that last sentence}
 
@@
 desired grain and extent of outputs) into model development.
-Successful use of an integrated modeling approach will always remain dependent on an intimate understanding of the decision-making process by modelers, emphasizing the importance of close collaboration between with practitioners at all stages of model development \citep{Guisan2013}.
-\dm{add\\
-Notably the connection between modelers and decision-makers who would use these types of modelsis also critical and the complexities of this relationship should not be trivialized.}
+Successful use of an integrated modeling approach will always remain dependent on an intimate understanding of the decision-making process by modelers, emphasizing the importance of close collaboration between modelers and practitioners at all stages of model development \citep{Guisan2013}.
 
@@
 
-\dg{but we should also see our approach as a mean to circumvent cases where this happens, since additional information will still be better than anything else}
 In a similar fashion, the quality and availability of data impose a significant constraint on the number and type of models that can be implemented.
@@
 Indeed, if species distributions are a function of, e.g.
-poopulation growth rate \citep{Sykes1996, Guisan2000}, they will involve processes at the individual (e.g.
+population growth rate \citep{Sykes1996, Guisan2000}, they will involve processes at the individual (e.g.
 competition) or cellular (e.g.
@@
 Since most decision makers work at local space scales and follow specific time frames, modelers face the obvious challenge of providing detailed outputs while facilitating implementation.
-In terms of model outputs, our framework is transparent in terms of uncertainty by virtue of providing easily interpretable probability distributions for model parameters and predictions.
+Our framework is transparent in reporting uncertainty of model output by virtue of providing easily interpretable probability distributions for model parameters and predictions.
 However, developing the models requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming.
@@
 However, the same was true of all statistical techniques or modeling approaches when they were initially developed, and many of the conceptual challenges associated with Bayesian statistics are being reduced as they gain recognition in the scientific literature.
-We therefore argue that our proposed approach has a strong potential for direct use in our real world where climate is quickly changing and conservation practices must be adjusted accordingly.
+We therefore argue that our approach has a strong potential for direct use in our real world where climate is quickly changing and conservation practices must be adjusted accordingly.
 We further argue that forging stronger collaborations between modelers, decision-makers, and practitioners will improve the incorporation of this and other new methods into applied ecology and conservation.
@@
 %TC:break fig 5
+\aa{Figure 5E and 5F: the legend should say 'naive SDM', as in 5C and 5D.}
+\aa{Another comment about this figure: have you considered showing the actual current distribution of the species?
+I am not saying it is needed to understand anything, but if I was a reviewer I think I would ask for it, just to compare with predictions from both the SDM and Phenofit, and especially to assess how worse are predictions of the integrated model as compared to SDM and Phenofit.}
 \textbf{Figure \ref{fig:ex2}:}
@@
 
-This procedure may be repeated 
-\wt{replace by applied or used. 
-Here looks like you have to repeat this to add a new model for instance. }
-for an arbitrary number of models (limited only by available data and computational power). 
+This procedure may be applied to an arbitrary number of models, limited only by available data and computational power. 
 The models may be implemented simultaneously, when multiple datasets are available and can be evaluated under different underlying models, or they may be run sequentially, updating the posterior distribution of $\theta_M$ as new information becomes available.

### 2014-10-13 tex/cover_letter.tex
Added most of the final revisions, started cover letter

--- 2014-09-26 tex/proposal_letter.tex
@@
+%%%%
+%
+%	cover letter for initial submission the week of Oct 13, 2014
+%
+%%%%
+
 \documentclass[11pt]{letter}
@@
 \signature{Matthew V. Talluto, \\
-	Isabelle Boulangeat, \& \\
+	Isabelle Boulangeat, \\
 	Dominique Gravel \\
+	Départament du Biologie \\
 	Université du Québec à Rimouski \\ 
@@
 \address{Matthew V. Talluto, Ph.D. \\
-	Départament du Biologie, Chimie, et Géographie \\
+	Départament du Biologie \\
 	300, Allée des Ursulines \\
@@
 
-\begin{letter}{Jordi Bascompte \\ 
-	Ideas \& Perspectives Editor \\
+\begin{letter}{Marcel Holyoak \\ 
+	Editor-in-Chief \\
 	\emph{Ecology Letters}
 }
-\opening{Dear Dr. Bascompte,}
+\opening{Dear Dr. Holyoak,}
 
-%TC:break opening
-We are pleased to submit a proposal for a manuscript titled ``Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework''. 
-We would appreciate your consideration of our manuscript for the \emph{Ideas and Perspectives} section of \emph{Ecology Letters}. Please find our proposal text below.
+We are pleased to submit a manuscript titled ``Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework''. 
+A proposal for the manuscript was reviewed and accepted by Dr. Bascompte on 2 October, 2014. 
+We would appreciate your consideration of our manuscript for the \emph{Ideas and Perspectives} section of \emph{Ecology Letters}.
 
-%TC:break body
 We present a novel framework for integrating small-scale, more mechanistic models into large-scale correlative species distribution models in order to improve the forecasts of these models.
@@
 Understanding range shifts are of broad interest for both theoretical and applied ecologists, and we believe that our manuscript represents an important step toward a more holistic approach for forecasting species ranges.
-We hope you will consider it for the \emph{Ideas and Perspectives} section. 
-We will be ready to submit the manuscript within one month.
 
-%TC:break closing
+This manuscript represents original unpublished work and is not under consideration for publication elsewhere. 
+No portion of it is published, in press, submitted, or in preparation for submission elsewhere; however, we use previously published data (with permission and appropriate citation) in our second example
+The authors are aware of no conflicts of interest regarding publication of this manuscript. 
+All authors have reviewed this work and approved it for submission. 
+Thank you for considering our manuscript for the \emph{Ideas and Perspectives} section, and we look forward to your response.
+
 \closing{Sincerely,}

### 2014-10-13 tex/manuscript.tex
Added most of the final revisions, started cover letter

@@
 %
-\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
+%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
 %\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
-\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
-\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
-\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
+%\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
+%\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
+%\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
 %\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
@@
 %\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
-\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
+%\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
 %\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
-\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
-\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
+%\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
+%\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
 %\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
-\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
-\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}
+%\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
+%\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}
 
@@
 Despite the recognized potential of these models, difficulties in producing predictions with acceptable precision and low bias has limited their contributions \citep{Guisan2013}.
-Two important constraints on modeling ecological processes can prevent models from making unbiased predictions with acceptable uncertainty: 
+For mechanistic models, two important constraints on modeling ecological processes can prevent models from making unbiased predictions with acceptable uncertainty: 
 (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
-Poor data or incoherent models will result in predictions with unacceptably high uncertainty or precise but biased predictions.
+Poor data or incoherent models will result in predictions with high uncertainty or precise but biased predictions.
 In recent decades, however, modeling techniques have proliferated to take advantage of what is available in terms of both data and modeling platforms.
 Model diversification also arises due to the diversity of processes generating species ranges \citep{Boulangeat2012}.
-In the context of forecasting, mechanistic and correlative approaches represent quite different modes of inference, but both have strong limitations. 
 Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, they may perform poorly when applied at the scale of species ranges.
 For instance, biotic interactions (including trophic interactions) are usually not modeled mechanistically at the scale of species distributions because they are poorly known, or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Pigot2013}.
-On the other hand, coarse-scale correlative models statistically relating species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
+In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
 However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}.
 Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
-Furthermore, correlative models often do not explicitly incorporate ecological processes into their formulations, and when applied at coarse scales, they may miss important fine-scale processes, making a multi-scale approach more suitable when accurate forecasting is required \citep{Thuiller2013}.
-On the other hand process-based models often require a variety of assumptions which may be poorly understood across space and time.
-\mt{This last sentence could go, and the sentence before is actually a bit repetitive. Suggest reorganizing a bit to still end on this point, but reduce repetition}
-\aa{you mentioned fine-scale mechanistic models and, on the other hand, coarse-scale correlative models.
-To talk now about process-based models sounds a bit bizarre and may make the reader wonder why are you mentioning them now.}
-\ia{Finish abruptely here. Feels like one sentence is missing}
-\ks{Feels like it is getting snuck in here…almost an after thought… I would recommend either removing it or developing it a tad further. }
 %% ----------------------------------
@@
 We used our metamodeling framework to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
-We use data from \citet{Morin2009} for sugar maple (\emph{Acer saccharum}), an economically and ecologically important species occurring in eastern North America.
+We used data from \citet{Morin2009} for sugar maple (\emph{Acer saccharum}), an economically and ecologically important species occurring in eastern North America.
 Here we describe briefly the dataset, methods, and the results of the analysis (see Appendix S2 for data and scripts to reproduce the analysis).
@@
 \defcitealias{IPCC2001}{IPCC, 2001} % this is needed to make the organizational citation in the following paragraph work correctly
-For the metamodel, we used a dataset of 1013 presence and 13863 absence records, at 0.5 degree resolution, sampled from North America.
-We reserved one-third of the dataset for validation, and used the area under the receiver operating characters curve (AUC) for validation of contemporary predictions.
-We used a binomial GLM to relate the suitability for sugar maple to 6 climate variables, scaled to the same resolution (see Appendix S2 for details on the climate variables).
-\aa{I would cite here the climatic variables used (just cite them, without further description)}
-We selected a GLM here for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
+For the metamodel, we used a dataset of 1013 presence and 13863 absence records at 0.5 degree resolution.
+We used a binomial GLM to relate the suitability for sugar maple to 3 climate variables: the number of degree days, mean summer precipitation, and the ratio of mean annual precipitation to potential evapotranspiration.
+These variables were selected from the initial set of 6 variables from \citet{Morin2009}; see Appendix S2 for details on model construction and selection.
+We selected a GLM for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
 For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
 Phenofit produces predictions at continental scales; thus the scaling function \(g\) was incorporated within Phenofit itself and we performed model integration directly on the outputs.
-To perform the integration, we used \ac{MCMC} to condition the parameters of the metamodel  on two likelihood functions.
-The first is the naive SDM, where the likelihood of a ``success'' in the presence/absence dataset was a Bernoulli density with the (logit-transformed) probability expressed as a linear function of the climate predictors.
-For the sub-model, we considered that Phenofit, being a process-based model, may perform better when projecting into the future.
+To perform the integration, we used a Metropolis-Hastings scheme within \ac{MCMC} to condition the parameters of the metamodel on two likelihood functions.
+The first was the naive SDM, where the likelihood of a ``success'' in the presence/absence dataset was a Bernoulli density with the (logit-transformed) probability expressed as a linear function of the climate predictors.
+For the sub-model, we considered that Phenofit, being a process-based model, may perform better when projecting into the future, particularly if suitability is primarily driven by the phenological match to the environment.
 Thus, we used the predicted future probability of presence as the basis for generating simulated future presence-absence datasets.
-This procedure inherently incorporates some uncertainty by randomly determining where the species is present (though uncertainty in the probabilities themselves could not be incorporated, as they were not available).
+This procedure inherently incorporates some uncertainty by randomly determining where the species is present (though uncertainty in the probabilities themselves could not be incorporated, as it is not possible to obtain them from Phenofit).
 Thus, the sub-model likelihood is simply the probability, given the predictions of the metamodel, of observing a simulated dataset \(X_s\) that is randomly drawn from the phenofit predictions \(\psi_s\); i.e., \(X_s = f(\psi_s)\):
@@
 
-\mt{Comments about this from the group discussion.
-The integrated present predictions are confusing, because that panel is not addressing the main modeling question.
-It would be a good idea to emphasize this; that the modeling approach reflects the questions being asked, and it is important to carefully identify these questions.}
-For present climates, the predictions of Phenofit (Fig \ref{fig:ex2}A) were largely similar to those of the naive SDM (Fig \ref{fig:ex2}C). 
-Furthermore, variance in the predictions of the naive model was relatively low throughout the spatial domain (Fig \ref{fig:ex2}E).
-\ks{provide specific figures for variance of the different models}
-Similar patterns were repeated for analogous climates under the future scenario (Fig \ref{fig:ex2}D); however these predictions differed considerably from the predictions of the mechanistic sub-model (Fig \ref{fig:ex2}B), and uncertainty for the projection was relatively high, particularly for eastern Canada (Fig \ref{fig:ex2}F).
-\aa{ instead of calling it 'mechanistic sub-model' I suggest saying Phenofit, for consistency.}
-In general, Phenofit predicted considerably less northward migration by 2080.
-The integrated model, which included both the presence-absence data as well as the Phenofit predictions for the future, made intermediate predictions in terms of northward range shifts for sugar maple (Fig \ref{fig:ex2}H).
-Globally, the integrated model expressed somewhat more uncertainty 
-\wt{for the future}
-than the naive model (Fig \ref{fig:ex2}J), reflecting the divergent predictions of Phenofit and the naive model.
-\wt{Are you sure? From the map, I would have said the reverse.}
-However, in areas where the naive model performed worst (e.g., eastern Canada), the level of uncertainty in the integrated model was similar to the uncertainty in other locations.
-Unsurprisingly, integration resulted in worse predictions for the present distribution of the species (Fig \ref{fig:ex2}G) compared with either the naive model or Phenofit in isolation.
-\aa{it may not be unsurprising that integration yields worse predictions than the other 2 models, but it is striking to see that it predicts such high suitabilities in southeastern US.
-In my opinion, it is not sufficiently explained why this should be considered as expected.
-I may have missed the point, but this is likely to happen to other readers also.}
-\mt{This ties in with the group discussion in Grenoble}
-Conventional SDMs are quite good at producing empirical models of species distributions under present climatic regimes, and the naive model here could be further improved with the addition of more parameters or the inclusion of ensemble forecasting.
-This is due to the integration focusing on future predictions; if integration was desired for the present range of the species (perhaps as a way to more accurately quantify uncertainty), a better approach would have been to use the present predictions from Phenofit.
-Unfortunately, validation is difficult for projections, as we lack data to evaluate model performance.
-For the desired outcome of this model (future projections), validations against present data may not be useful; the predictions of naive model were superior (AUC = 0.99) to those of the integrated model (AUC = 0.96), yet the future predictions of the integrated model may be more realistic than those of the naive model.
+Considering the models individually, the predictions of Phenofit were much more conservative than those of the naive model (Fig \ref{fig:ex2}).
+Although both models predicted a northward range shift, the change under Phenofit was limited to approximately 200 km north of the present range limit of the species, compared with more than 900 km for the naive model.
+Phenofit also predicted little change in the southern range limit of the species, while the naive model projected a lack of suitability over the southern half of the present distribution.
+Model integration produced predictions that were intermediate between the extremes set by the individual models (Fig \ref{fig:ex2}C).
+The northern limit of the species was forecast to move by approximately 400--500 km, with reduced suitabilities but little extirpation throughout the southern portion of the present range.
+Uncertainty in the naive model was relatively low, particularly for the southern portion of the range where the model predicted extirpation (Fig \ref{fig:ex2}D).
+However, uncertainty in the predictions was locally extreme, particularly in portions of eastern Canada.
+Prediction uncertainty was generally increased for the integrated model relative to the naive model (Fig \ref{fig:ex2}E), likely due to the large amount of disagreement in the predictions of the two models.
+Unfortunately, validation of future projections is not possible due to a lack of data with which to compare model predictions.
+However, model integration in this case provides a greater degree of humility in interpreting the predictions; for example, the naive model predicted with a high degree of certainty that much of the southern portion of the present range of sugar maple will be unsuitable by 2080.
+The integrated model suggests that we should not be so confident in this interpretation.
+Furthermore, the relatively high variance in predictions in the south suggests that further exploration of limiting factors in that region may be useful in providing further understanding and more precise predictions for changes to the distribution of this species.
 
@@
 % FIGURE 5
-%% 140 words
+%% 125 words
 %TC:break fig 5
-\aa{Figure 5E and 5F: the legend should say 'naive SDM', as in 5C and 5D.}
-\aa{Another comment about this figure: have you considered showing the actual current distribution of the species?
-I am not saying it is needed to understand anything, but if I was a reviewer I think I would ask for it, just to compare with predictions from both the SDM and Phenofit, and especially to assess how worse are predictions of the integrated model as compared to SDM and Phenofit.}
 \textbf{Figure \ref{fig:ex2}:}
 	Results of model integration for sugar maple, \emph{Acer saccharum}.
-	The mechanistic sub-model Phenofit predicts present (A) and future (B) suitability using phenological information, but does not provide estimates of uncertainty.
-	Predictions from a naive metamodel for the present (C) were quite similar to the sub-model, but future predictions  differed (D).
-	Uncertainty in future projections was quite high in the eastern portion of the range (F) relative to uncertainty in predictions for the present (E).
-	Model integration resulted in poor predictions for present range (G), with future predictions that were intermediate between the two models (H).
-	Uncertainty in the integrated model (I, J) was generally greater than in the naive model, reflecting the disagreement between the naive model and the sub-model, but there were fewer extremes in model uncertainty when projecting the integrated model (J) compared with the naive model (F).
+	All panels show information using predicted climate from 2080 under the HadA2 scenario (see main text).
+	For reference, the present range of the species is outlined in red \citep{Little1971}.
+	The mechanistic sub-model Phenofit (A) predicted fairly small shifts sugar maple range.
+	In contrast, the naive model (B) projected large northward shifts in suitable habitat.
+	Model integration (C) produced predictions that were intermediate between the two models.
+	Uncertainty in the naive model (D) was fairly low, except for some portions of the northeast where uncertainty was extreme.
+	Model integration produced uncertainty (E) that was globally somewhat higher reflecting the disagreement between the naive model (B) and the sub-model (A), but with fewer extremes in prediction uncertainty.
 

### 2014-10-14 tex/manuscript.tex
Finished first draft for submission

@@
 \author[2,3]{Kevin A. Solarik (kevinsolarik@hotmail.com)}
-\author[10]{Nickolay Strigul (nick.strigul@vancouver.wsu.edu)}
+\author[10]{Nikolay Strigul (nick.strigul@vancouver.wsu.edu)}
 \author[11, 12]{Wilfried Thuiller (wilfried.thuiller@ujf-grenoble.fr)}
@@
 \affil[9]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
-\affil[10]{Department of Mathematics, Washington State University Vancouver, Washington, USA}
+\affil[10]{Department of Mathematics, Washington State University, Vancouver, Washington, USA}
 \affil[11]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
@@
 	% figure captions limited to 150 words each
-	\textbf{Word counts: } Abstract: 200; Main Text: 6086; Boxes: 517
+	\textbf{Word counts: } Abstract: 200; Main Text: 5953; Boxes: 516
 
-	\textbf{Number of references:} 79 % limited to 80
+	\textbf{Number of references:} 80 % limited to 80
 
@@
 For the metamodel, we used a dataset of 1013 presence and 13863 absence records at 0.5 degree resolution.
-We used a binomial GLM to relate the suitability for sugar maple to 3 climate variables: the number of degree days, mean summer precipitation, and the ratio of mean annual precipitation to potential evapotranspiration.
+We used a binomial GLM to relate the suitability for sugar maple to 3 climate variables: the number of degree days, mean summer precipitation, and the ratio of annual precipitation to potential evapotranspiration.
 These variables were selected from the initial set of 6 variables from \citet{Morin2009}; see Appendix S2 for details on model construction and selection.
@@
 Prediction uncertainty was generally increased for the integrated model relative to the naive model (Fig \ref{fig:ex2}E), likely due to the large amount of disagreement in the predictions of the two models.
-Unfortunately, validation of future projections is not possible due to a lack of data with which to compare model predictions.
+
+Although validation is an important tool for analyzing the performance of SDMs, it is not possible to validation predictions for the future due to a lack of data with which to compare model predictions.
 However, model integration in this case provides a greater degree of humility in interpreting the predictions; for example, the naive model predicted with a high degree of certainty that much of the southern portion of the present range of sugar maple will be unsuitable by 2080.
 The integrated model suggests that we should not be so confident in this interpretation.
-Furthermore, the relatively high variance in predictions in the south suggests that further exploration of limiting factors in that region may be useful in providing further understanding and more precise predictions for changes to the distribution of this species.
+The naive model predicts suitability based on the values of selected climate variables (in this case, number of degree days, mean summer precipitation, and the ratio of annual precipitation to potential evapotranspiration).
+These variables were selected because they provided the best fit to the present distribution; however, this does not mean these variables are the most important climatic drivers of the processes generating the distribution, nor that they will continue to be important in the future.
+In contrast, Phenofit predicts fitness based on how climate affects phenological timings, frost injury, reproduction, and survival \citep{Chuine2001, Morin2009}.
+Thus, climatic factors that ultimately limit species distribution might be quite different under Phenofit, as illustrated by the substantial differences in the potential future distributions predicted by Phenofit and the naive model.
+In the absence of information about which model is ``best,'' the integrated model provides a consensus view of these two very different models by fitting parameters that maximize the joint likelihood of the presence-absence dataset and the Phenofit predictions.
+Furthermore, the integrated model provides robust estimates of uncertainty; in the southern portion of the range, the relatively high variance in predictions reflects the disagreement in the models and suggests that further exploration of limiting factors in that region may be useful in improving understanding and providing more precise predictions for changes to the distribution of sugar maple.
 
@@
 We acknowledge funding from the Quebec Centre for Biodiversity Science as well as NSERC strategic grant 430393-12.
-Comments from Charles Canham and William Godsoe improved a previous version of this manuscript.
+Comments from Charles Canham, William Godsoe, and Tamara Münkemüller improved a previous version of this manuscript.
 WT received support funding from the European Research Council under the European Community's Seven Framework Programme FP7/2007-2013 Grant Agreement no.
@@
 \bibliographystyle{model_integration}
+
+\section*{Supporting Information}
+Additional Supporting Information may be downloaded via the online version of this article at Wiley Online Library (www.ecologyletters.com).
+As a service to our authors and readers, this journal provides supporting information supplied by the authors.
+Such materials are peer-reviewed and may be reorganized for online delivery, but are not copy-edited or typeset.
+Technical support issues arising from supporting information (other than missing files) should be addressed to the authors.
 

### 2014-12-19 tex/manuscript.tex
Finished MEE submission

@@
 
-%% ----------------------------------
-%
-%     Packages to be used in the final manuscript
-%
-%% ----------------------------------
-
-\usepackage[margin=1in]{geometry}
+\usepackage[margin=1in, left=1.2in]{geometry}
 
@@
 \usepackage{natbib} 
+\bibpunct{(}{)}{;}{a}{}{,}
 
@@
 
-\usepackage{acronym} % set up some acronyms
-\acrodef{SDM}{species distribution model}
-\acrodef{MCMC}{Markov Chain Monte Carlo}
-
 %% graphics packages
 \usepackage{graphicx}
-\usepackage{pdflscape}   % for figures in landscape orientation
-\usepackage[labelsep=none]{caption}
 
@@
 
-\usepackage[pagewise]{lineno}
-
-%% ----------------------------------
-%
-%     Packages to be used for drafts and editing
-%     Remove from the final manuscript
-%
-%% ----------------------------------
-%\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
-%
-%\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
-
-%%% available colors:
-%%% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
-%
-%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
-%\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
-%\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
-%\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
-%\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
-%\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
-%\newcommand{\fd}[1]{\todo[color=orange!30]{\textsuperscript{Frédérik D}#1}}
-%\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
-%\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
-%\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
-%\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
-%\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
-%\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
-%\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
-%\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}
-
-% uncomment to disable individual authors
-% \renewcommand{\mt}[1]{\relax}%
-% \renewcommand{\ib}[1]{\relax}%
-% \renewcommand{\aa}[1]{\relax}%
-% \renewcommand{\ia}[1]{\relax}%
-% \renewcommand{\db}[1]{\relax}%
-% \renewcommand{\ab}[1]{\relax}%
-% \renewcommand{\fd}[1]{\relax}%
-% \renewcommand{\rd}[1]{\relax}%
-% \renewcommand{\mjf}[1]{\relax}%
-% \renewcommand{\tf}[1]{\relax}%
-% \renewcommand{\dm}[1]{\relax}%
-% \renewcommand{\ks}[1]{\relax}%
-% \renewcommand{\ns}[1]{\relax}%
-% \renewcommand{\wt}[1]{\relax}%
-% \renewcommand{\dg}[1]{\relax}%
+\usepackage{lineno}
 
@@
 
-	% less than 45 chars including spaces
 	\textbf{Short title:} Integrated models of species ranges
 
-	\textbf{Type of article:} Ideas and Perspectives
-
-	% abstract limited to 200 words
-	% main text excludes abstract, acknowledgements, references, table and figure legends; limited to 7500 words
-	% boxes limited to 750 words each
-	% figure captions limited to 150 words each
-	\textbf{Word counts: } Abstract: 200; Main Text: 5953; Boxes: 516
-
-	\textbf{Number of references:} 80 % limited to 80
-
-	\textbf{Number of figures:} 6
-
-	\textbf{Number of tables:} 0
-
-	\textbf{Number of text boxes:} 1
+	\textbf{Word counts: } 7100
 
@@
 \begin{abstract}
-% limited to 200 words
-Current interest in modeling potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
+\noindent \begin{enumerate}
+
+\item Current interest in modeling potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
 However, in general, individual approaches include only a small subset of the available information about a species. 
@@
 We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods.
-Our method results in a metamodel that produces probabilistic estimates of species presence, reports uncertainty from both data and process error from all sub-models, and incorporates all of the information used as input for the original scale-specific sub-models.
-We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
-We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
+
+\item Our framework uses a metamodel, estimated in a hierarchical Bayesian framework, that is constrained by both data directly applied to the metamodel and by the results of one or more sub-models.
+The output of the metamodel is in the form of probabilistic estimates of species presence.
+Metamodel outputs also incorporate uncertainty from both data and process error from all sub-models, and include all of the information used as input for the original scale-specific sub-models.
+\item We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
+\item We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
+The framework has wide potential for use in species distribution modelling, and can drive better integration of multi-scale data into ecological decision-making.
 Although implementation of our method can be technically challenging, we anticipate the results will be of wide interest, and we encourage collaboration between modelers and practitioners in applying our framework.
+
+\end{enumerate}
 \end{abstract}
@@
 We illustrate our approach with two examples.
-First, we use simulated data to present a complete example of the application of the framework to multiple information sources that are relevant at very different scales.
-In a second example, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}), a widespread foundation tree species from eastern North America, to climate with the predictions of a phenological model, Phenofit \citep{Chuine2001}, with the goal of improving uncertainty estimation when projecting model predictions to future climate.
+We begin with a hypothetical example, using simulated data, that presents a simplified application of our framework to multiple sources of information from different scales (Example 1).
+In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread foundation tree species from eastern North America, in response to climate change.
+Both examples are presented with complete code and data (in Supporting Information).
+Finally, we provide a more formal mathematical presentation in Box 1.
 
@@
 The parameters of this model can be constrained using the output of one or more sub-models that can operate at different scales than the metamodel.
-As an example, the metamodel could take the form of a simple correlative \ac{SDM}, with one or more fine-scaled process-based sub-models.
+As an example, the metamodel could take the form of a simple correlative species distribution model (SDM) operating at very large spatial scales, with one or more fine-scaled process-based sub-models.
 Ideally, each sub-model incorporates different data types and hypotheses that specify underlying ecological mechanisms contributing to the model output.
@@
 For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature.
-We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty (see Appendix S1 in Supporting Information for procedural details and scripts for executing the model).
+We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty (see Appendix 1 in Supporting Information for procedural details and scripts for executing the model).
 
@@
 There is considerable interest in comparing correlative and mechanistic projections with respect to climate change \citep{Morin2009}, and correctly characterizing uncertainty is a critical aspect of this problem \citep{Cheaib2012}.
-Despite being a relatively common application of \ac{SDM}s \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios remains problematic \citep{Araujo2006}.
+Despite being a relatively common application of SDMs \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios remains problematic \citep{Araujo2006}.
 We used our metamodeling framework to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
-We used data from \citet{Morin2009} for sugar maple (\emph{Acer saccharum}), an economically and ecologically important species occurring in eastern North America.
-Here we describe briefly the dataset, methods, and the results of the analysis (see Appendix S2 for data and scripts to reproduce the analysis).
+We used data from \citet{Morin2009} for sugar maple, an economically and ecologically important species occurring in eastern North America.
+Here we describe briefly the dataset, methods, and the results of the analysis (see Appendix 2 for data and scripts to reproduce the analysis).
 
@@
 We used a binomial GLM to relate the suitability for sugar maple to 3 climate variables: the number of degree days, mean summer precipitation, and the ratio of annual precipitation to potential evapotranspiration.
-These variables were selected from the initial set of 6 variables from \citet{Morin2009}; see Appendix S2 for details on model construction and selection.
+These variables were selected from the initial set of 6 variables from \citet{Morin2009}; see Appendix 2 for details on model construction and selection.
 We selected a GLM for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
@@
 Phenofit produces predictions at continental scales; thus the scaling function \(g\) was incorporated within Phenofit itself and we performed model integration directly on the outputs.
-To perform the integration, we used a Metropolis-Hastings scheme within \ac{MCMC} to condition the parameters of the metamodel on two likelihood functions.
+To perform the integration, we used a Metropolis-Hastings scheme within Markov Chain Monte Carlo (MCMC) sampling to condition the parameters of the metamodel on two likelihood functions.
 The first was the naive SDM, where the likelihood of a ``success'' in the presence/absence dataset was a Bernoulli density with the (logit-transformed) probability expressed as a linear function of the climate predictors.
@@
 \subsection*{Comparison with other methods}
-The methods provided here expand upon the motivation of hybrid models to develop a more robust and unifying approach for ecological models in a predictive context while striving to overcome some of the limitations characteristic of other integrated approaches.
-Attempting to incorporate underlying ecological processes so that SDMs are not oversimplified by combining different model parameters or scales can be challenging using hybrid approaches. 
-In particular, it can often be difficult to identify parameters that can be used to connect different modeling frameworks and which parameters can produce meaningful responses \citep{Thuiller2013}. 
+The methods provided here expand upon the motivation of hybrid models to develop a more robust and unifying approach for ecological models in a predictive context while overcoming some limitation characteristic of other integrated approaches.
+In particular, it is often difficult in a hybrid model to identify parameters that can be used to connect different modeling frameworks and produce a meaningful response \citep{Thuiller2013}. 
+The difficulty of including information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Thuiller2008, Smolik2010, Thuiller2014a}. 
+Even when a link can be formed to include this information, it often is oversimplified to facilitate computation \citep{Gallien2010}. 
 The methods presented in this study aim to overcome some of these limitations through the use of Bayesian statistics.
-Bayesian methods produce results as an entire distribution instead of a single point, allowing for a comprehensive understanding of the uncertainty of estimates \citep{Link2010}. 
+Bayesian methods provide a natural framework for the incorporation of multiple sources of information.
+Because the metamodel is expressed as a series of conditional probabilities (see Box 1), this information can be included simply provided the probability of the metamodel, conditional on information from the sub-models, can be expressed mathematically.
+Furthermore, Bayesian methods produce results as an entire distribution instead of a single point, allowing for a comprehensive understanding of the uncertainty of estimates \citep{Link2010}. 
 The estimation of model parameters (and predictions) as distributions can also help decrease overparameterization that can be a concern for hybrid models. 
-Bayesian methods also provide a natural framework for the incorporation of multiple sources of information. 
-The inability or difficulty to include information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Thuiller2008, Smolik2010, Thuiller2014a}.
-Even when a link can be formed to include this information, it often is oversimplified to facilitate computation \citep{Gallien2010}. 
 Finally, Bayesian methods inherently allow for feedbacks or interactions between sub models, which may be a more realistic representation of ecological dynamics where many factors may simultaneously influence the system.
 
-Our approach is an original application of Bayesian multilevel modeling, and can be considered a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales while using several models simultaneously.
-In particular, this development has certain similarities with Bayesian model averaging, Bayesian calibration of process-based models, and hierarchical Bayesian modeling. 
+Our approach is an original application of hierarchical Bayesian modeling, and can be considered a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales while using several models simultaneously.
+In particular, this development has certain similarities with Bayesian model averaging and Bayesian calibration of process-based models. 
 Bayesian model averaging aims to combine several alternative models to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.,][]{Wintle2003, Link2006}.
@@
 %TC:ignore
+\section*{Data Accessibility}
+All data, as well as all code required to repeat the analyses, have been uploaded as online supporting information.
+
 \nolinenumbers
@@
 
-\section*{Supporting Information}
-Additional Supporting Information may be downloaded via the online version of this article at Wiley Online Library (www.ecologyletters.com).
-As a service to our authors and readers, this journal provides supporting information supplied by the authors.
-Such materials are peer-reviewed and may be reorganized for online delivery, but are not copy-edited or typeset.
-Technical support issues arising from supporting information (other than missing files) should be addressed to the authors.
-
 %% ----------------------------------
@@
 
-\newpage
-\section*{Figure Captions}
-
-\begin{flushleft}
-
 %==================
 % FIGURE 1
-%% 149 words
 %TC:break fig 1
-\textbf{Figure \ref{fig:diagram}:}
-	The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
+\newpage
+\begin{figure}[h!]
+\input{diagram}
+\caption{The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
 	The metamodel ($\theta_M$) operates at a single scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
@@
 	This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
-	This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).
+	This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).}
+\label{fig:diagram}
+\end{figure}
 
@@
 % FIGURE 2
-%% 88 words
 %TC:break fig 2
-\textbf{Figure \ref{fig:ex1_sampling}:}
-	Two simulated datasets used to illustrate the model integration framework.
+\newpage
+\begin{figure}[h!]
+	\includegraphics{ex1_sampling.pdf}
+	\caption{Two simulated datasets used to illustrate the model integration framework.
 	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
@@
 	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
-	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.
+	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.}
+	\label{fig:ex1_sampling}
+\end{figure}
 
@@
 % FIGURE 3
-%%  120 words
 %TC:break fig 3
-\textbf{Figure \ref{fig:ex1_precip}:}
-	Results of integration showing the probability of presence (\(\psi\)) when considering only precipitation.
+\newpage
+\begin{figure}[h!]
+	\includegraphics{ex1_precip.pdf}
+	\caption{Results of integration showing the probability of presence (\(\psi\)) when considering only precipitation.
 	(a) Naive model, using only presence-absence data.
@@
 	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
-	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.
+	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.}
+	\label{fig:ex1_precip}
+\end{figure}
 
@@
 % FIGURE 4
-%% 75 words
 %TC:break fig 4
-\textbf{Figure \ref{fig:ex1_map}:}
-	Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
+\newpage
+\begin{figure}[h!]
+	\includegraphics[width=5.25in]{ex1_map.pdf}
+	\caption{Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
 	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
 	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
-	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.
+	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.}
+	\label{fig:ex1_map}
+\end{figure}
 
@@
 % FIGURE 5
-%% 125 words
 %TC:break fig 5
-\textbf{Figure \ref{fig:ex2}:}
-	Results of model integration for sugar maple, \emph{Acer saccharum}.
+\newpage
+\begin{figure}[h!]
+	\includegraphics[width=6in]{ex2.pdf}
+	\caption{Results of model integration for sugar maple.
 	All panels show information using predicted climate from 2080 under the HadA2 scenario (see main text).
@@
 	Uncertainty in the naive model (D) was fairly low, except for some portions of the northeast where uncertainty was extreme.
-	Model integration produced uncertainty (E) that was globally somewhat higher reflecting the disagreement between the naive model (B) and the sub-model (A), but with fewer extremes in prediction uncertainty.
-
-%==================
-% FIGURE 6
-%% 65 words
-%TC:break fig 6
-\textbf{Figure \ref{fig:management}:}
-	Sample workflow for applying the models presented in the first example in a management context.
-	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
-	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the metamodel.
-
-\end{flushleft}
-
-%==================
-% FIGURE 1
-
-\newpage
-\begin{figure}[h!]
-\input{diagram}
-\caption{}
-\label{fig:diagram}
-\end{figure}
-
-%==================
-% FIGURE 2
-\newpage
-\begin{figure}[h!]
-	\includegraphics{ex1_sampling.pdf}
-	\caption{}
-	\label{fig:ex1_sampling}
-\end{figure}
-
-%==================
-% FIGURE 3
-\newpage
-\begin{figure}[h!]
-	\includegraphics{ex1_precip.pdf}
-	\caption{}
-	\label{fig:ex1_precip}
-\end{figure}
-
-%==================
-% FIGURE 4
-\newpage
-\begin{figure}[h!]
-	\includegraphics[width=5.25in]{ex1_map.pdf}
-	\caption{}
-	\label{fig:ex1_map}
-\end{figure}
-
-%==================
-% FIGURE 5
-\newpage
-\begin{figure}[h!]
-	\includegraphics[width=6in]{ex2.pdf}
-	\caption{}
+	Model integration produced uncertainty (E) that was globally somewhat higher reflecting the disagreement between the naive model (B) and the sub-model (A), but with fewer extremes in prediction uncertainty.}
 	\label{fig:ex2}
@@
 % FIGURE 6
+%TC:break fig 6
 \newpage
@@
 	\includegraphics{management.pdf}
-	\caption{}
+	\caption{Sample workflow for applying the models presented in the first example in a management context.
+	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
+	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the metamodel.}
 	\label{fig:management}
@@
 The normalization constant involves computing integrals that are often impossible to solve analytically. 
-In practice, simulation techniques such as \ac{MCMC} can be used to sample directly from the posterior distribution, making such computations unnecessary.
+In practice, simulation techniques such as MCMC can be used to sample directly from the posterior distribution, making such computations unnecessary.
 Thus, we use the proportional form of Bayes' Theorem:

### 2014-12-19 tex/bibprep.pl

@@ -1,32 +1,0 @@
-#!/usr/bin/perl
-
-use strict;
-use warnings;
-
-my $inFilename = $ARGV[0];
-my @fileContents;
-my $i = 1;
-my $newItem = 0;
-my $name = "";
-
-open(inFile, "<", $inFilename) or die "Cannot open file $inFilename\n";
-while(<inFile>) {
-	chomp;
-	if ( (not $newItem) && m/\\bibitem\[\{(\w+)/) {
-		$name = $1;
-		$newItem = 1;
-	} elsif($newItem && s/^$name/$i\. \\newline $name/) {
-		$i++;
-		$newItem = 0;
-		$name = ""
-	}
-	push (@fileContents, $_);
-}
-close(inFile);
-
-open(inFile, ">", $inFilename) or die "Cannot open file $inFilename\n";
-foreach (@fileContents) {
-	print "$_\n";
-	print inFile "$_\n";
-}
-

### 2015-01-14 tex/manuscript.tex
Fix a small typo
Prep for GEB submission

@@
 \bibpunct{(}{)}{;}{a}{}{,}
+
+\usepackage{caption}
+\captionsetup{labelformat=simple, labelsep=none}
 
@@
 
-	\textbf{Word counts: } 7100
+	\textbf{Abstract word count: } 258
+
+	\textbf{Main text word count: } 7121
 
@@
 \begin{abstract}
-\noindent \begin{enumerate}
+\noindent
 
-\item Current interest in modeling potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
+Current interest in modeling potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
 However, in general, individual approaches include only a small subset of the available information about a species. 
@@
 We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
-
-\item Our framework uses a metamodel, estimated in a hierarchical Bayesian framework, that is constrained by both data directly applied to the metamodel and by the results of one or more sub-models.
+Our framework uses a metamodel, estimated in a hierarchical Bayesian framework, that is constrained by both data directly applied to the metamodel and by the results of one or more sub-models.
 The output of the metamodel is in the form of probabilistic estimates of species presence.
 Metamodel outputs also incorporate uncertainty from both data and process error from all sub-models, and include all of the information used as input for the original scale-specific sub-models.
-\item We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
-\item We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
+We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
+We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
 The framework has wide potential for use in species distribution modelling, and can drive better integration of multi-scale data into ecological decision-making.
@@
 
-\end{enumerate}
  \end{abstract}
@@
 For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature.
-We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty (see Appendix 1 in Supporting Information for procedural details and scripts for executing the model).
+We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty (see Appendix S1 in Supporting Information for procedural details and scripts for executing the model).
 
@@
 	\overbrace{p\left (\psi_S \mid \theta_S,P_S \right )}^{\substack{\text{new information} \\ \text{from sub-model}}}
-	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) P \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
+	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) p \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
 	\overbrace{p \left(\theta_S \right)}^{\substack{\text{prior for} \\ \text{sub-model}}}	
@@
 We used data from \citet{Morin2009} for sugar maple, an economically and ecologically important species occurring in eastern North America.
-Here we describe briefly the dataset, methods, and the results of the analysis (see Appendix 2 for data and scripts to reproduce the analysis).
+Here we describe briefly the dataset, methods, and the results of the analysis (see Appendices S2 and S3 for data and scripts to reproduce the analysis).
 
@@
 We used a binomial GLM to relate the suitability for sugar maple to 3 climate variables: the number of degree days, mean summer precipitation, and the ratio of annual precipitation to potential evapotranspiration.
-These variables were selected from the initial set of 6 variables from \citet{Morin2009}; see Appendix 2 for details on model construction and selection.
+These variables were selected from the initial set of 6 variables from \citet{Morin2009}; see Appendix S2 for details on model construction and selection.
 We selected a GLM for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
@@
 	p( X_S = f(\psi_s)\mid \theta_M, \theta_S )
-	p( Y_M \mid \theta_M, D_M ) 
+	p( X_M \mid \theta_M, D_M ) 
 	p( \theta_M )
@@
 \section*{Data Accessibility}
-All data, as well as all code required to repeat the analyses, have been uploaded as online supporting information.
+All data, as well as all code required to repeat the analyses, have been uploaded as online supporting information in Appendix S3.
 
-\nolinenumbers
+%% ----------------------------------
+%
+%     Supporting information
+%
+%% ----------------------------------
+\section*{Supporting information}
+Appendix S1: Supporting information on using the code and data for example 1. \\
+Appendix S2: Supporting information on using the code and data for example 2. \\
+Appendix S3: Scripts and data required to reproduce the analyses.
+
 \section*{Acknowledgements}
@@
 %TC:endignore
+
+\nolinenumbers
 
@@
 %
+%     BIOSKETCH (GEB)
+%
+%% ----------------------------------
+
+\section*{Biosketch}
+%TC:break biosketch
+The Theoretical Ecosystem Ecology lab group at the Université du Québec à Rimouski focuses on interactions between species distributions, community structure and ecosystem functioning.
+We apply principles of spatial ecology to a variety of organisms and systems, from bacteria to entire forests.
+We also use theoretical and simulation models to develop hypotheses and extend our work beyond the technical limitations of empirical studies.
+More information can be found at: http://chaire-eec.uqar.ca/
+
+%% ----------------------------------
+%
 %     FIGURES
@@
 %% ----------------------------------
+
+\newpage
+%TC:break fig legends
+\section*{Figure Legends}
+\begin{flushleft}
+
+\textbf{Figure 1:} The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
+	The metamodel ($\theta_M$) operates at a single scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
+	The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
+	The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
+	This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
+	This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).
+
+\textbf{Figure 2:} Two simulated datasets used to illustrate the model integration framework.
+	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
+	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., \(-\)1--1).
+	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
+	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.
+
+\textbf{Figure 3:} Results of integration showing the probability of presence (\(\psi\)) when considering only precipitation.
+	(a) Naive model, using only presence-absence data.
+Uncertainty increases dramatically when attempting to project beyond the scope of the source data (represented by the horizontal line above the bottom axis).
+	(b) Mechanistic model, using observations of an experiment to infer probability of presence.
+Predictions are highly precise due to high quality source data, but are likely to be overly optimistic given the simplicity in assuming only precipitation determines probability of presence (see text).
+	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
+	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.
+
+\textbf{Figure 4:} Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
+	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
+	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
+	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.
+
+\textbf{Figure 5:} Results of model integration for sugar maple.
+	All panels show information using predicted climate from 2080 under the HadA2 scenario (see main text).
+	For reference, the present range of the species is outlined in red \citep{Little1971}.
+	The mechanistic sub-model Phenofit (A) predicted fairly small shifts sugar maple range.
+	In contrast, the naive model (B) projected large northward shifts in suitable habitat.
+	Model integration (C) produced predictions that were intermediate between the two models.
+	Uncertainty in the naive model (D) was fairly low, except for some portions of the northeast where uncertainty was extreme.
+	Model integration produced uncertainty (E) that was globally somewhat higher reflecting the disagreement between the naive model (B) and the sub-model (A), but with fewer extremes in prediction uncertainty.
+
+\textbf{Figure 6:} Sample workflow for applying the models presented in the first example in a management context.
+	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
+	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the metamodel.
+
+\end{flushleft}
 
@@
 \input{diagram}
-\caption{The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
-	The metamodel ($\theta_M$) operates at a single scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
-	The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
-	The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
-	This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
-	This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).}
+\caption{}
 \label{fig:diagram}
@@
 	\includegraphics{ex1_sampling.pdf}
-	\caption{Two simulated datasets used to illustrate the model integration framework.
-	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
-	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., \(-\)1--1).
-	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
-	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.}
+	\caption{}
 	\label{fig:ex1_sampling}
@@
 	\includegraphics{ex1_precip.pdf}
-	\caption{Results of integration showing the probability of presence (\(\psi\)) when considering only precipitation.
-	(a) Naive model, using only presence-absence data.
-Uncertainty increases dramatically when attempting to project beyond the scope of the source data (represented by the horizontal line above the bottom axis).
-	(b) Mechanistic model, using observations of an experiment to infer probability of presence.
-Predictions are highly precise due to high quality source data, but are likely to be overly optimistic given the simplicity in assuming only precipitation determines probability of presence (see text).
-	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
-	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.}
+	\caption{}
 	\label{fig:ex1_precip}
@@
 	\includegraphics[width=5.25in]{ex1_map.pdf}
-	\caption{Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
-	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
-	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
-	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.}
+	\caption{}
 	\label{fig:ex1_map}
@@
 	\includegraphics[width=6in]{ex2.pdf}
-	\caption{Results of model integration for sugar maple.
-	All panels show information using predicted climate from 2080 under the HadA2 scenario (see main text).
-	For reference, the present range of the species is outlined in red \citep{Little1971}.
-	The mechanistic sub-model Phenofit (A) predicted fairly small shifts sugar maple range.
-	In contrast, the naive model (B) projected large northward shifts in suitable habitat.
-	Model integration (C) produced predictions that were intermediate between the two models.
-	Uncertainty in the naive model (D) was fairly low, except for some portions of the northeast where uncertainty was extreme.
-	Model integration produced uncertainty (E) that was globally somewhat higher reflecting the disagreement between the naive model (B) and the sub-model (A), but with fewer extremes in prediction uncertainty.}
+	\caption{}
 	\label{fig:ex2}
@@
 	\includegraphics{management.pdf}
-	\caption{Sample workflow for applying the models presented in the first example in a management context.
-	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
-	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the metamodel.}
+	\caption{}
 	\label{fig:management}
@@
 \section*{Box 1: Model description}
-% Boxes are limited to 750 words
 We begin with a simple correlative model (i.e. calibrated with data provided at the same ecological scale as predictions), characterized in a Bayesian framework to allow further integration. 

### 2015-02-11 tex/appendix_s1.tex
Moved box 1 to an appendix

--- 2014-12-19 tex/manuscript.tex
@@
-\documentclass[11pt]{article}
-
-\usepackage[margin=1in, left=1.2in]{geometry}
-
-% use proper unicode fonts
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-
-\usepackage{amsmath} % for better display of equations
-
-\usepackage{setspace}
-\usepackage{natbib} 
-\bibpunct{(}{)}{;}{a}{}{,}
-
-\usepackage{titling} % controls the way the title information is displayed
-\pretitle{\begin{flushleft}\Large}
-\posttitle{\end{flushleft}}
-\predate{}
-\postdate{}
-\preauthor{\begin{flushleft}}
-\postauthor{\end{flushleft}}
-\setlength{\droptitle}{-3em}
-
-\usepackage{authblk} % adds some nice options for displaying the author list
-\renewcommand\Authsep{\protect\\}
-\renewcommand\Authands{\protect\\}
-
-%% graphics packages
-\usepackage{graphicx}
-
-% Tikz libraries for building the diagram
-\usepackage{tikz}
-\usepackage{tikz-cd}
-\usetikzlibrary{calc, shapes}
-\usetikzlibrary{shapes.geometric,shapes.arrows,decorations.pathmorphing}
-\usetikzlibrary{matrix,chains,scopes,positioning,arrows,fit}
-
-\usepackage{lineno}
-
-%% ----------------------------------
-%
-%     Title and authorship information
-%
-%% ----------------------------------
-
-\title{Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework}
-\date{}
-\author[1,2,13]{Matthew V. Talluto (mtalluto@gmail.com)}
-\author[1,2]{Isabelle Boulangeat (isabelle.boulangeat@gmail.com)}
-\author[3]{Aitor Ameztegui (ameztegui@gmail.com)}
-\author[4]{Isabelle Aubin (Isabelle.Aubin@NRCan-RNCan.gc.ca)}
-\author[1,2,5]{Dominique Berteaux (Dominique\_Berteaux@uqar.ca)}
-\author[1,2]{Alyssa Butler (ca.butler10@gmail.com)}
-\author[6,7]{Frédérik Doyon (Frederik.Doyon@uqo.ca)}
-\author[8]{C. Ronnie Drever (cdrever@TNC.ORG)}
-\author[9]{Marie-Josée Fortin (fortinmj@gmail.com)}
-\author[1]{Tony Franceschini (Tony.Franceschini@uqar.ca)}
-\author[10]{Jean Liénard (jean.lienard@gmail.com)}
-\author[4]{Dan McKenney (Dan.McKenney@nrcan-rncan.gc.ca)}
-\author[2,3]{Kevin A. Solarik (kevinsolarik@hotmail.com)}
-\author[10]{Nikolay Strigul (nick.strigul@vancouver.wsu.edu)}
-\author[11, 12]{Wilfried Thuiller (wilfried.thuiller@ujf-grenoble.fr)}
-\author[1,2]{Dominique Gravel (dominique\_gravel@uqar.ca)}
-\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
-\affil[2]{Quebec Centre for Biodiversity Science, Montreal, Quebec, Canada}
-\affil[3]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
-\affil[4]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
-\affil[5]{Centre for Northern Studies, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
-\affil[6]{Université du Québec en Outaouais, Gatineau, Quebec, Canada}
-\affil[7]{Institut des Sciences de la Forêt Tempérée (ISFORT), Ripon, Quebec, Canada}
-\affil[8]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
-\affil[9]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
-\affil[10]{Department of Mathematics, Washington State University, Vancouver, Washington, USA}
-\affil[11]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
-\affil[12]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
-\affil[13]{Author for correspondance. Address: Departament de Biologie, chimie, et geographie, 300, Allée des Ursulines, Rimouski, Quebec G5L 3A1, Canada}
-
-%% ----------------------------------
-%
-%     END PREAMBLE
-%
-%% ----------------------------------
-
-\begin{document}
-\doublespacing
-%% ----------------------------------
-%
-%     TITLE PAGE
-%
-%% ----------------------------------
-
-%TC:ignore
-\begin{titlingpage}
-	\maketitle
-
-	\begin{flushleft}
-
-	\textbf{Short title:} Integrated models of species ranges
-
-	\textbf{Word counts: } 7100
-
-	\textbf{Authorship:} All authors contributed substantially to an initial outline of the paper and ideas.
-MT, IB, and DG built the model and wrote the first draft of the introduction and model description.
-All authors contributed to the complete first draft and all subsequent revisions to the manuscript.
-
-	\textbf{Keywords:} Climate change, decision making, metamodeling, patterns and processes, range dynamics, scaling, spatial ecology, species distribution modeling, uncertainty
-	\end{flushleft}
-\end{titlingpage}
-%TC:endignore
-
-%TC:break abstract
-\linenumbers
-\begin{abstract}
-\noindent \begin{enumerate}
-
-\item Current interest in modeling potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
-However, in general, individual approaches include only a small subset of the available information about a species. 
-Correlative models often predict presence or absence as a function of climate, but ignore  smaller-scale processes such as growth, fecundity, and dispersal.
-Furthermore, different approaches often produce divergent predictions, with no simple method to reconcile them.
-We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods.
-
-\item Our framework uses a metamodel, estimated in a hierarchical Bayesian framework, that is constrained by both data directly applied to the metamodel and by the results of one or more sub-models.
-The output of the metamodel is in the form of probabilistic estimates of species presence.
-Metamodel outputs also incorporate uncertainty from both data and process error from all sub-models, and include all of the information used as input for the original scale-specific sub-models.
-\item We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
-\item We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
-The framework has wide potential for use in species distribution modelling, and can drive better integration of multi-scale data into ecological decision-making.
-Although implementation of our method can be technically challenging, we anticipate the results will be of wide interest, and we encourage collaboration between modelers and practitioners in applying our framework.
-
-\end{enumerate}
-\end{abstract}
-
-%TC:break manuscript
-%% ----------------------------------
-%
-%     INTRODUCTION
-%
-%% ----------------------------------
-
-\section*{Introduction}
-
-Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
-Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}.
-Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution \citep{Jay2012}, hybrid zone dynamics \citep{Engler2013}, and forecasting the impacts of climate change on species distributions \citep{Thuiller2011, Blois2013, Thuiller2014}.
-
-Despite the recognized potential of these models, difficulties in producing predictions with acceptable precision and low bias has limited their contributions \citep{Guisan2013}.
-For mechanistic models, two important constraints on modeling ecological processes can prevent models from making unbiased predictions with acceptable uncertainty: 
-(1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
-Poor data or incoherent models will result in predictions with high uncertainty or precise but biased predictions.
-In recent decades, however, modeling techniques have proliferated to take advantage of what is available in terms of both data and modeling platforms.
-Model diversification also arises due to the diversity of processes generating species ranges \citep{Boulangeat2012}.
-Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, they may perform poorly when applied at the scale of species ranges.
-For instance, biotic interactions (including trophic interactions) are usually not modeled mechanistically at the scale of species distributions because they are poorly known, or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Pigot2013}.
-In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
-However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}.
-Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
-%% ----------------------------------
-%% ----------------------------------
-%% ----------------------------------
-
-\subsection*{Toward an integrated approach for modeling ranges}
-To include knowledge of species from multiple sources and scales into distribution models, we present an application of hierarchical Bayesian methods that uses outputs from multiple model outputs to inform the results of the final model.
-Our approach is an alternative to other approaches to multimodel inference that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
-By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
-Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Boulangeat2012, Catterall2012, Strigul2012, Soranno2014}.
-Recent literature has also focused on the development of hybrid models, which allow for flexible combinations between mechanistic and phenomenological sub-models \citep{Gallien2010, Thuiller2013}. 
-For instance, correlative models are used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while mechanistic approaches can include space and time dynamics and therefore account for dispersal processes \citep{Kearney2008}.
-This complementarity has been used to merge the two model types into hybrid models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Dormann2012, Boulangeat2014}. 
-However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and the sources and magnitudes of uncertainties are difficult to quantify.
-Moreover, hybrid approaches are not suitable for merging models that are based on the same processes but with different underlying hypotheses.
-
-A commonly-employed alternative to hybrid models is the direct combination of predictions.
-If models operate at the same scale and have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}.
-If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Cheaib2012, Serra-Diaz2013}.
-However, the applicability of ensemble forecasts is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
-A more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying the parameters that contribute most to variance in the model predictions \citep{McMahon2011}.
-Clearly, an approach that integrates predictions and uncertainty from multiple model types is needed \citep{Beck2012, Thuiller2013}. 
-Such an approach would incorporate multiple modes of inference and integrate data from multiple sources and scales \citep{Levin1992, Peters2004, Thuiller2013}.
-
-Here, we present a framework for modeling species range dynamics that aims to account for all the information available on a focal species, from a variety of sources and scales.
-Contrary to hybrid and hierarchical models, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for as much flexibility as possible regarding the type of information included.
-We use the power and generality of a hierarchical Bayesian framework, which allows us to include multiple data sources and modes of inference \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}. 
-Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to interpret and reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}. 
-We illustrate our approach with two examples.
-We begin with a hypothetical example, using simulated data, that presents a simplified application of our framework to multiple sources of information from different scales (Example 1).
-In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread foundation tree species from eastern North America, in response to climate change.
-Both examples are presented with complete code and data (in Supporting Information).
-Finally, we provide a more formal mathematical presentation in Box 1.
-
-%% ----------------------------------
-%
-%     MODELING FRAMEWORK
-%
-%% ----------------------------------
-
-\section*{Integrated modeling framework}
-
-The key idea of our approach is formulate a metamodel that integrates data at the same ecological scale as the desired predictions.
-The parameters of this model can be constrained using the output of one or more sub-models that can operate at different scales than the metamodel.
-As an example, the metamodel could take the form of a simple correlative species distribution model (SDM) operating at very large spatial scales, with one or more fine-scaled process-based sub-models.
-Ideally, each sub-model incorporates different data types and hypotheses that specify underlying ecological mechanisms contributing to the model output.
-The output of sub-models should provide additional predictions at a comparable ecological scale at which the metamodel operates, although it is not necessary for the sub-models to actually operate at the same scale.
-For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, these functions can be simple while still contributing useful information to the metamodel (see Example 1).
-All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Box 1, Fig. \ref{fig:diagram}).
-We present and apply the framework here with two examples, and follow with a more formal presentation in Box 1.
-
-%% ----------------------------------
-%
-%     EXAMPLE 1
-%
-%% ----------------------------------
-\subsection*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
-
-In this hypothetical example, we build a metamodel relating the distribution (i.e.
-occurrence probability) of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
-The metamodel attempts to capture the realized distribution of a species; as a correlative model, it implicitly captures the major physiological constraints and ecological processes constraining the distribution of the target species.
-However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the climatic envelope of where a species may be found in a natural setting.
-Prior information of the physiological constraints affecting species distribution is sometimes available, but usually too incomplete to perform a reasonable comparison with the realized distribution.
-For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature.
-We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty (see Appendix 1 in Supporting Information for procedural details and scripts for executing the model).
-
-We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation. 
-We will consider \(X_M\) (see Box 1, Fig. \ref{fig:diagram}) to be a vector of $n$ observations that takes the value of one to indicate presence and zero for absence:
-\begin{equation}
-X_M = \{x_{M,1}, x_{M,2}, \ldots, x_{M,n}\}
-\end{equation}
-We assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
-The model is evaluated by relating these observations to environmental data \((D_M)\), which, for the sake of the example, consists of mean annual temperature $T_M$ and annual precipitation $P_M$. 
-For simplicity, we formulate the metamodel \((\theta_M)\) as a simple logistic regression with a second order effect of temperature and precipitation. 
-This naive model (i.e., the metamodel with no constraints from sub-models) predicts  occurrence probability (\(\psi_N\)) as a function of the environment:
-%-----
-\begin{equation}
-\begin{aligned}
-	\psi_N &= f\left(\theta_M, D_M \right) \\
-	&= p \left (X_M = 1 \mid \theta_M, T_M, P_M \right) \\
-	&=\text{logit}^{-1}\left( \mathbf{\Theta_M} \mathbf{D_M} \right)
-\end{aligned}
-\end{equation}
-%-----
-where \(\mathbf{\Theta_M}\) is the parameter vector of the model \(\theta_M\), \(\mathbf{D_M} \) is the covariate matrix (i.e., \(T_M, P_M\), with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
-We can fit this model in a Bayesian framework to allow for easy integration of models (as we will show later).
-In this context, the goal of modeling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, a concise introduction can be found in ][]{Link2010}:
-%-----
-\begin{equation}
-\label{eq:ex1_bayes}
-	p\left (\theta_M \mid X_M,T_M,P_M \right ) \propto 
-	p \left(X_M \mid \theta_M, T_M, P_M \right)
-	p \left(\theta_M \right)
-\end{equation}
-%-----
-where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.
-
-Thus far, we have considered only a single source of information to fit this model, and therefore the prior distribution \(p\left(\theta_M \right)\) from Eq. \ref{eq:ex1_bayes} is uninformative.
-As a secondary source of information, we will consider an experiment relating the population growth rate of the plant to manipulations to the precipitation regime, with results (but no raw data) available from the literature (Fig.
-\ref{fig:ex1_sampling}b).
-Furthermore, no information is available regarding the temperature regime for the experiment.
-Transplant experiments that evaluate performance beyond the range of a species are common, and represent a plausible scenario for model integration \citep{Hargreaves2014}.
-According to niche theory \citep{Holt2009}, the fundamental niche corresponds to the set of environmental conditions where the per capita intrinsic growth rate $r$ is positive.
-This concept gives us a reasonable model to fit the scaling function $g$ for our sub-model (Fig. \ref{fig:diagram}).
-If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, then for observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
-\begin{equation}
-	\psi_{S,i} = \int_0^\infty N \left(X_{S,i}, \sigma_{S,i} \right)
-\end{equation}
-where \(N\) is the Normal density function.
-We can then estimate the posterior distribution for the sub-model by fitting the relationship between \(\psi_S\) and precipitation \( \left( P_S \right) \) using Bayesian beta regression \citep{Ferrari2004}:
-%-----
-\begin{equation}
-\label{eq:ex1_thetas}
-	p\left (\theta_S \mid \psi_S,P_S \right ) \propto 
-	p \left(\psi_S \mid \theta_S, P_S \right)
-	p \left(\theta_S \right) \\
-\end{equation}
-%-----
-
-Although the two datasets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation).
-This scaling is quite simplistic, and would never be used to predict a species' range using this mechanistic sub-model in isolation.
-Despite the strong theoretical foundation (i.e., niche theory) linking the observations to presence-absence at large scales, the upscaled sub-model (i.e., the transformation of \(X_S\) and \(\sigma_S\) to \(\psi_S\)) is built on the strong hypothesis that the species will be present when $r>0$, neglecting other population dynamics constraints such as Allee effects and metapopulation dynamics.
-The fundamental niche is also incomplete, as we lack environmental variables such as temperature, as well as all other ecological processes responsible for the realized distribution.
-As such, it would be unwise to expect predictions from this model alone to resemble the actual distribution of the species; as a mechanistic model, it is simply too incomplete to predict distribution.
-As we will show, however, the information from this sub-model, when applied as a constraint on the metamodel, can result in improved predictions that incorporate the information within each model.
-
-We accomplish model integration by treating the posterior of \(\theta_S\) as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
-%-----
-\begin{equation}
-	\label{eq:ex1_integrated}
-	\overbrace{p(\theta_M \mid X_M, T_M, P_M, \theta_S, \psi_S, P_S)}^\text{integrated posterior}
-	\propto
-	\overbrace{p\left (\psi_S \mid \theta_S,P_S \right )}^{\substack{\text{new information} \\ \text{from sub-model}}}
-	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) P \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
-	\overbrace{p \left(\theta_S \right)}^{\substack{\text{prior for} \\ \text{sub-model}}}	
-\end{equation}
-%-----
-As before, the metamodel \(\theta_M\) can be used to predict the species distribution \((\psi_I)\).
-However, these predictions will not reflect only the presence-absence samples in \(X_M\), but will also reflect the information from \(\theta_S\), including all of the data sources used to produce this sub-model.
-In other words, with this evaluation, the set of parameters $\theta_M$ is most likely when it agrees with both the original data and predictions from the sub-model. 
-Note that in this particular example, because the sub-model is only contingent on precipitation, there is no integration on temperature. 
-Finally, we note the presence of marginal distributions for both models (i.e., \(p(\theta_M)\) and \(p(\theta_S)\)).
-These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
-For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modeling process (due to the lack of additional mechanisms in the second model, as mentioned above).
-
-We performed this procedure with our hypothetical datasets.
-When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data (Figs.
-\ref{fig:ex1_precip}a, \ref{fig:ex1_map}a, \ref{fig:ex1_map}b)
-Unsurprisingly given the quality of the data, the sub-model was highly precise with respect to precipitation and not useful for temperature, thus providing a fairly strong constraint when producing the integrated model (Fig \ref{fig:ex1_precip}b).
-The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
-At the scale of the metamodel, considering both temperature and precipitation, we observed similar results, with reduced uncertainty in the predictions over the domain not covered by the presence-absence data (Fig.
-\ref{fig:ex1_map}).
-
-Once the technical challenges of evaluating the parameters are overcome, integration has several advantages over both the naive parameterization of the metamodel and the mechanistic sub-model. 
-In the situation where both models strongly agree on the relationship between water availability and occurrence probability, then we should expect significantly reduced uncertainty in that domain (Figs.
-\ref{fig:ex1_precip}, \ref{fig:ex1_map}).
-Because the metamodel includes both temperature and precipitation, any reduction in the uncertainty with respect to precipitation might also influence the evaluation of the relationship with temperature (reducing bias in the case where there is a correlation between T and P in the data).
-However, uncertainty could increase in the case of model disagreement.
-For instance, the relationship between precipitation and presence in the naive metamodel could be caused by a spurious relationship (e.g.
-spatial autocorrelation caused by historical contingencies), leading to a false confidence in the species distribution over precipitation gradients.
-The information from the sub-model will thus bring back the right amount of uncertainty in the relationship.
-Moreover, convergence in model predictions in overlapping domains (e.g., near the peak in the probability of presence in Fig. \ref{fig:ex1_precip}) can aid in detecting crucial ecological processes.
-
-%% ----------------------------------
-%
-%     EXAMPLE 2
-%
-%% ----------------------------------
-
-\subsection*{Example 2: Constraining an SDM using phenological information}
-For the second example, we consider the problem of forecasting the potential distribution of a species following climate change.
-There is considerable interest in comparing correlative and mechanistic projections with respect to climate change \citep{Morin2009}, and correctly characterizing uncertainty is a critical aspect of this problem \citep{Cheaib2012}.
-Despite being a relatively common application of SDMs \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios remains problematic \citep{Araujo2006}.
-We used our metamodeling framework to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
-We used data from \citet{Morin2009} for sugar maple, an economically and ecologically important species occurring in eastern North America.
-Here we describe briefly the dataset, methods, and the results of the analysis (see Appendix 2 for data and scripts to reproduce the analysis).
-
-\defcitealias{IPCC2001}{IPCC, 2001} % this is needed to make the organizational citation in the following paragraph work correctly
-For the metamodel, we used a dataset of 1013 presence and 13863 absence records at 0.5 degree resolution.
-We used a binomial GLM to relate the suitability for sugar maple to 3 climate variables: the number of degree days, mean summer precipitation, and the ratio of annual precipitation to potential evapotranspiration.
-These variables were selected from the initial set of 6 variables from \citet{Morin2009}; see Appendix 2 for details on model construction and selection.
-We selected a GLM for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
-For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
-Phenofit produces predictions at continental scales; thus the scaling function \(g\) was incorporated within Phenofit itself and we performed model integration directly on the outputs.
-To perform the integration, we used a Metropolis-Hastings scheme within Markov Chain Monte Carlo (MCMC) sampling to condition the parameters of the metamodel on two likelihood functions.
-The first was the naive SDM, where the likelihood of a ``success'' in the presence/absence dataset was a Bernoulli density with the (logit-transformed) probability expressed as a linear function of the climate predictors.
-For the sub-model, we considered that Phenofit, being a process-based model, may perform better when projecting into the future, particularly if suitability is primarily driven by the phenological match to the environment.
-Thus, we used the predicted future probability of presence as the basis for generating simulated future presence-absence datasets.
-This procedure inherently incorporates some uncertainty by randomly determining where the species is present (though uncertainty in the probabilities themselves could not be incorporated, as it is not possible to obtain them from Phenofit).
-Thus, the sub-model likelihood is simply the probability, given the predictions of the metamodel, of observing a simulated dataset \(X_s\) that is randomly drawn from the phenofit predictions \(\psi_s\); i.e., \(X_s = f(\psi_s)\):
-\begin{equation}
-\label{eq:integrated2}
-	p( \theta_M \mid X_M, D_M, \psi_s, \theta_S )
-	\propto 
-	p( X_S = f(\psi_s)\mid \theta_M, \theta_S )
-	p( Y_M \mid \theta_M, D_M ) 
-	p( \theta_M )
-	p( \theta_S )
-\end{equation}
-where \(\theta_M\) is the metamodel, 
-\(X_M\) is the vector of contemporary presences and absences, 
-\(D_M\) is the matrix of climate predictors,
-and \(\theta_S\) is the model relating the Phenofit predictions to the metamodel.
-Furthermore, the integration is performed on the \emph{predictions} of the sub-model directly, rather than on compatible parameters as in Example 1. 
-Thus, the metamodel must appear in the sub-model likelihood.
-
-Considering the models individually, the predictions of Phenofit were much more conservative than those of the naive model (Fig \ref{fig:ex2}).
-Although both models predicted a northward range shift, the change under Phenofit was limited to approximately 200 km north of the present range limit of the species, compared with more than 900 km for the naive model.
-Phenofit also predicted little change in the southern range limit of the species, while the naive model projected a lack of suitability over the southern half of the present distribution.
-Model integration produced predictions that were intermediate between the extremes set by the individual models (Fig \ref{fig:ex2}C).
-The northern limit of the species was forecast to move by approximately 400--500 km, with reduced suitabilities but little extirpation throughout the southern portion of the present range.
-Uncertainty in the naive model was relatively low, particularly for the southern portion of the range where the model predicted extirpation (Fig \ref{fig:ex2}D).
-However, uncertainty in the predictions was locally extreme, particularly in portions of eastern Canada.
-Prediction uncertainty was generally increased for the integrated model relative to the naive model (Fig \ref{fig:ex2}E), likely due to the large amount of disagreement in the predictions of the two models.
-
-Although validation is an important tool for analyzing the performance of SDMs, it is not possible to validation predictions for the future due to a lack of data with which to compare model predictions.
-However, model integration in this case provides a greater degree of humility in interpreting the predictions; for example, the naive model predicted with a high degree of certainty that much of the southern portion of the present range of sugar maple will be unsuitable by 2080.
-The integrated model suggests that we should not be so confident in this interpretation.
-The naive model predicts suitability based on the values of selected climate variables (in this case, number of degree days, mean summer precipitation, and the ratio of annual precipitation to potential evapotranspiration).
-These variables were selected because they provided the best fit to the present distribution; however, this does not mean these variables are the most important climatic drivers of the processes generating the distribution, nor that they will continue to be important in the future.
-In contrast, Phenofit predicts fitness based on how climate affects phenological timings, frost injury, reproduction, and survival \citep{Chuine2001, Morin2009}.
-Thus, climatic factors that ultimately limit species distribution might be quite different under Phenofit, as illustrated by the substantial differences in the potential future distributions predicted by Phenofit and the naive model.
-In the absence of information about which model is ``best,'' the integrated model provides a consensus view of these two very different models by fitting parameters that maximize the joint likelihood of the presence-absence dataset and the Phenofit predictions.
-Furthermore, the integrated model provides robust estimates of uncertainty; in the southern portion of the range, the relatively high variance in predictions reflects the disagreement in the models and suggests that further exploration of limiting factors in that region may be useful in improving understanding and providing more precise predictions for changes to the distribution of sugar maple.
-
-%% ----------------------------------
-%
-%     DISCUSSION
-%
-%% ----------------------------------
-
-\section*{Discussion}
-
-\subsection*{Comparison with other methods}
-The methods provided here expand upon the motivation of hybrid models to develop a more robust and unifying approach for ecological models in a predictive context while overcoming some limitation characteristic of other integrated approaches.
-In particular, it is often difficult in a hybrid model to identify parameters that can be used to connect different modeling frameworks and produce a meaningful response \citep{Thuiller2013}. 
-The difficulty of including information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Thuiller2008, Smolik2010, Thuiller2014a}. 
-Even when a link can be formed to include this information, it often is oversimplified to facilitate computation \citep{Gallien2010}. 
-The methods presented in this study aim to overcome some of these limitations through the use of Bayesian statistics.
-Bayesian methods provide a natural framework for the incorporation of multiple sources of information.
-Because the metamodel is expressed as a series of conditional probabilities (see Box 1), this information can be included simply provided the probability of the metamodel, conditional on information from the sub-models, can be expressed mathematically.
-Furthermore, Bayesian methods produce results as an entire distribution instead of a single point, allowing for a comprehensive understanding of the uncertainty of estimates \citep{Link2010}. 
-The estimation of model parameters (and predictions) as distributions can also help decrease overparameterization that can be a concern for hybrid models. 
-Finally, Bayesian methods inherently allow for feedbacks or interactions between sub models, which may be a more realistic representation of ecological dynamics where many factors may simultaneously influence the system.
-
-Our approach is an original application of hierarchical Bayesian modeling, and can be considered a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales while using several models simultaneously.
-In particular, this development has certain similarities with Bayesian model averaging and Bayesian calibration of process-based models. 
-Bayesian model averaging aims to combine several alternative models to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.,][]{Wintle2003, Link2006}.
-Bayesian model averaging considers models operating at the same scale and the posterior distribution is usually determined with Gibbs sampling. 
-Bayesian calibration of process-based models focuses on uncertainty of the parameter values, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, Hartig2012}.
-In contrast with Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of the metamodel.
-Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005}, and climate and atmospheric modeling \citep{Mcmillan2010, Kang2012}.
-Although our method as presented is distinct from Bayesian model averaging, it is fully compatible with these methods.
-Multiple metamodels could be averaged in this way if a consensus prediction from multiple metamodels (containing, e.g., different sub-model sets or different metamodel formulations).
-Furthermore, the metamodel can easily be incorporated into a broader multimodel inference scheme.
-
-%\fd{FD will work on this part: \\
-%It would be good to compare this approach with the kind of one develop by Moorcroft et al.
-2001 : Moorcroft, P.
-R., Hurtt, G.
-C., \& Pacala, S.
-W.
-(2001).
-A method for scaling vegetation dynamics: the ecosystem demography model (ED).
-Ecological monographs, 71(4), 557-586.\\
-%%Medvigy, D., Wofsy, S.
-C., Munger, J.
-W., Hollinger, D.
-Y., \& Moorcroft, P.
-R.
-(2009).
-Mechanistic scaling of ecosystem function and dynamics in space and time: Ecosystem Demography model version 2.
-Journal of Geophysical Research: Biogeosciences (2005–2012), 114(G1).\\
-%Fisher, R., McDowell, N., Purves, D., Moorcroft, P., Sitch, S., Cox, P., \& Ian Woodward, F.
-(2010).
-Assessing uncertainties in a second-generation dynamic vegetation model caused by ecological scale limitations.
-New Phytologist, 187(3), 666-681.
-%\\
-%These models are multilevel and use scaling functions using statisitical moments series.
-%}
-
-%----------------------------
-%----------------------------
-%----------------------------
-
-\subsection*{Advantages of Model Integration}
-Species distribution models are important tools that are increasingly used by land managers to face challenges associated to decision-making \citep{Guisan2013}.
-A growing number of approaches are available, but they can provide diverging answers to very similar questions due to the differences in their assumptions and methodology. 
-This can create confusion and even some mistrust towards models, and some managers may be discouraged from incorporating model results in their management plans. 
-Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organizations around the world \citep[e.g.,][]{Bernier2013}.
-Incorporating information from multiple sources, particularly with respect to uncertainty,  fosters a connection between scientifically-generated knowledge and policy, and is therefore an important tool for adaptive management \citep[][Fig.
-\ref{fig:management}]{Rehme2011}.
-Such approaches are needed in formulating management plans for vulnerable species and ecosystems to avoid basing decisions on too-narrow subsets of the available information \citep{Dawson2011}.
-The flexibility of our approach may also represent an advantage for management by easing the integration of specific decision making criteria (e.g.
-desired grain and extent of outputs) into model development.
-Successful use of an integrated modeling approach will always remain dependent on an intimate understanding of the decision-making process by modelers, emphasizing the importance of close collaboration between modelers and practitioners at all stages of model development \citep{Guisan2013}.
-
-Model uncertainty is another key factor affecting applicability of model outputs \citep{Addison2013}.
-One of the main strengths of our approach is that it allows for a clear, transparent identification of uncertainties and how they are transmitted throughout the modeling process. 
-Transparency in uncertainty can be considered as a sort of sensitivity analysis, in which areas with large uncertainty can be detected and new experimental research or additional data collection can be designed accordingly (e.g., Example 1, Figs.
-\ref{fig:ex1_sampling}, \ref{fig:ex1_precip}).
-The new knowledge resulting from this research can be readily incorporated into the metamodel, allowing for an iterative learning process that will undoubtedly contribute to reduce uncertainty in predictions for a wide range of environmental conditions.
-Moreover, our framework embraces change as a fundamental process and is able to adapt and respond to it. 
-The ease of incorporating new knowledge to the modeling framework (including new theory or the result of management and experimental efforts) will allow for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans \citep{Keith2011}, and the integration of sub-models allows for clear specification of desired model outputs (via the specification of the metamodel) while easily retaining important ecological objectives (via careful specification of sub-models).
-Transparency in the model building process must be accompanied by a clearly documented modeling workflow; the transparency of our approach (and the utility of specifying specific objectives accompanied by appropriate sub-models) will assist in the development of transparent and well documented workflows describing the process of model integration and will ensure reproducibility and applicability (Fig.
-\ref{fig:management}).
-In an era of continuous and unprecedented change, adaptive approaches such as the one presented here are often highlighted as a pressing need in order to develop strategies to promote ecosystems that are both feasible and resilient \citep{Seastedt2008, Mori2013}.
-
-In many cases, both the data and theory needed to apply our method already exist, and all that is needed is the development of sub-models and the links between the sub-models and a metamodel.
-For example, climatic gradients may mediate competitive interactions \citep{Kunstler2011}, which means that simple correlative models that fail to account for competition may be wrong if the climate-competition association changes in the future.
-In North America, the US Forest Service maintains a long term Forest Inventory Analysis database that could be utilized to parameterize a competition model.
-Such a model need not explicitly predict occurrence limits; rather it could be integrated with a larger-scale model to include information about competition in a distribution model.
-The phenological information needed to parameterize the sub-model for Example 2 is similarly available for a wide range of species \citep{Morin2009}.
-There are also a number of networks collecting high-quality data with good temporal and spatial coverage (e.g., the National Ecological Observatory Network; NEON, and Long-Term Ecological Research sites; LTER).
-There is great potential for these kinds of data to be used in fitting sub-models of the kind used in Example 1.
-In other cases, qualitative model comparison efforts have already been made \citep{Morin2009, Cheaib2012}.
-Our framework could be used \emph{post-hoc} on the outputs of these models to quantify uncertainty resulting from model disagreement.
-
-%----------------------------
-%----------------------------
-%----------------------------
-
-\subsection*{Challenges} 
-Although our approach is highly flexible and can be applied in a number of situations, there are some challenges to successfully using the framework.
-As with any modeling effort, good model specification with strong links to theory are essential \citep{Austin2007}.
-Poorly specified models will produce outputs that are uninformative or misleading, and model integration is not a cure for these problems.
-In the worst case, constraining a metamodel with a poor sub-model can result in outcomes that are worse in terms of bias and uncertainty than those produced by a naive model.
-We expect model selection will play an important role in applications of our framework, and such schemes can be incorporated relatively easily \citep{Madigan1995, Wasserman2000, Tenan2014}.
-
-In a similar fashion, the quality and availability of data impose a significant constraint on the number and type of models that can be implemented.
-The capacity to implement a model is very low if data requirements are prohibitive. 
-Adequate and unbiased coverage of explanatory variables is a significant obstacle, and exploratory analyses can be a significant aid in understanding how data coverage impacts resulting predictions \citep{Mckenney2002}.
-Integration can ameliorate these issues to some extent by using supplemental information (and conceptual advances) in additional sub-models where coverage is weak (e.g., Example 1, Figs.
-\ref{fig:ex1_precip}, \ref{fig:ex1_map}).
-For example, \citet{Freckleton2009} estimated that data are too limited to successfully develop highly mechanistic models predicting weed population numbers. 
-A strong asset of our approach is that it can be used without the full suite of data that would be required to run a fully mechanistic model. 
-Given that the metamodel is correlative, it can be effectively implemented with, e.g., only presence-absence data, or, in the case where true absences are difficult to obtain, with presences and pseudo absences (provided sufficient care is used in interpreting the results of such a model).
-Then any additional mechanistic data that are available will enhance predictions by constraining outputs of the metamodel.
-
-Finally, determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eqs. \ref{eq:ex1_integrated}, \ref{eq:integrated2}) is a critical point.
-In the context of integrated models, the challenge is three fold: (i) which spatial and temporal scales (i.e.
-which processes), are to be considered, (ii) how to build a scaling function that is consistent with the metamodel and (iii) how error and uncertainty propagate from the sub-models to the metamodel.
-Although we argue that our proposed framework is able to easily deal with very different scales and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered, the building of scaling functions deserves further investigation.
-We have shown in our examples that even simple scaling functions can provide reasonable constraints on the metamodel. 
-However, it is likely that the modeling process may include multiple scaling functions operating at different scales when taking all known processes and models into account. 
-Indeed, if species distributions are a function of, e.g.
-population growth rate \citep{Sykes1996, Guisan2000}, they will involve processes at the individual (e.g.
-competition) or cellular (e.g.
-photosynthesis) scales.
-Such very large differences in spatial scales imply that simple functions will be inadequate, requiring more traditional upscaling methods. 
-Therefore, intermediate models and processes will need to be introduced, and this will require additional parameters, assumptions, and data, and will add uncertainty to the final integrated model.
-
-In an applied context, the results of models predicting species ranges are increasingly used by resource managers, conservation biologists, and forest ecologists for formulating and adjusting recommendations.
-The utility of such models depends on their ability to help evaluate events beyond the bounds of the available data, including in future situations constrained by climate change. 
-Practitioners may judge model usefulness on two criteria: (1) Does the model provide reliable predictions at the needed time and space scales, and (2) can the model be implemented given the available data and technical expertise?
-Since most decision makers work at local space scales and follow specific time frames, modelers face the obvious challenge of providing detailed outputs while facilitating implementation.
-Our framework is transparent in reporting uncertainty of model output by virtue of providing easily interpretable probability distributions for model parameters and predictions.
-However, developing the models requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming.
-In many cases, off-the-shelf software can adequately express the model likelihoods with minimal programming, but more complicated models will require the development of custom software.
-Thus, this framework will require significant investment in developing customized code for the samplers in order to actually estimate parameters, which may be technically challenging for some practitioners and may limit the adoption of these methods by both practitioners and decision-makers.
-However, the same was true of all statistical techniques or modeling approaches when they were initially developed, and many of the conceptual challenges associated with Bayesian statistics are being reduced as they gain recognition in the scientific literature.
-We therefore argue that our approach has a strong potential for direct use in our real world where climate is quickly changing and conservation practices must be adjusted accordingly.
-We further argue that forging stronger collaborations between modelers, decision-makers, and practitioners will improve the incorporation of this and other new methods into applied ecology and conservation.
-
-%----------------------------
-%----------------------------
-%----------------------------
-
-\subsection*{Future directions}
-Hierarchical Bayesian modeling requires data at every scale to fit the parameters required to constrain one level with another.
-Moreover, beyond the needs for additional data, more knowledge of the important processes driving species range limits is needed to allow for extrapolation beyond the scope of the original datasets to address future changes due to climate and land-use changes.
-This knowledge can be obtained through a series of experimental manipulations and by substituting space for time data along environmental gradients.
-When experiments are possible, gaps in knowledge can be inferred using surrogate variables, theoretical models and spatially explicit models to explore the range of outcomes possible \citep{Fortin2012}.
-Another way to gain knowledge is to simplify species complexity into a series of states and to use state-transition models (e.g., Markov chain, semi-Markov chain, matrix projection model, integral projection model) which require more theoretical or coarser knowledge of species responses to biotic and abiotic changes.
-Although fewer states allow us to better capture the essence of species responses to environmental conditions, knowledge is still required to model feedback effects across processes and spatio-temporal scales.
-Hence we need to determine which processes are the most important at a given scale and how to weight their effects while scaling-up.
-Our metamodeling approach can act as a springboard for incorporating imperfect knowledge of how multi-scale processes influence species ranges.
-By incorporating this knowledge into species distribution models, we can hopefully both reduce bias and more accurately assess model uncertainty when forecasting future changes to species ranges.
-%---------------------------------------------
-%---------------------------------------------
-
-%TC:ignore
-\section*{Data Accessibility}
-All data, as well as all code required to repeat the analyses, have been uploaded as online supporting information.
-
-\nolinenumbers
-\section*{Acknowledgements}
-We acknowledge funding from the Quebec Centre for Biodiversity Science as well as NSERC strategic grant 430393-12.
-Comments from Charles Canham, William Godsoe, and Tamara Münkemüller improved a previous version of this manuscript.
-WT received support funding from the European Research Council under the European Community's Seven Framework Programme FP7/2007-2013 Grant Agreement no.
-281422 (TEEMBIO).
-%TC:endignore
-
-%TC:break references
-%% ----------------------------------
-%
-%     REFERENCES
-%
-%% ----------------------------------
-
-\renewcommand\refname{Literature Cited}
-\bibliography{model_integration}{}
-\bibliographystyle{model_integration}
-
-%% ----------------------------------
-%
-%     FIGURES
-%
-%% ----------------------------------
-
-%==================
-% FIGURE 1
-%TC:break fig 1
-\newpage
-\begin{figure}[h!]
-\input{diagram}
-\caption{The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
-	The metamodel ($\theta_M$) operates at a single scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
-	The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
-	The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
-	This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
-	This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).}
-\label{fig:diagram}
-\end{figure}
-
-%==================
-% FIGURE 2
-%TC:break fig 2
-\newpage
-\begin{figure}[h!]
-	\includegraphics{ex1_sampling.pdf}
-	\caption{Two simulated datasets used to illustrate the model integration framework.
-	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
-	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., \(-\)1--1).
-	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
-	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.}
-	\label{fig:ex1_sampling}
-\end{figure}
-
-%==================
-% FIGURE 3
-%TC:break fig 3
-\newpage
-\begin{figure}[h!]
-	\includegraphics{ex1_precip.pdf}
-	\caption{Results of integration showing the probability of presence (\(\psi\)) when considering only precipitation.
-	(a) Naive model, using only presence-absence data.
-Uncertainty increases dramatically when attempting to project beyond the scope of the source data (represented by the horizontal line above the bottom axis).
-	(b) Mechanistic model, using observations of an experiment to infer probability of presence.
-Predictions are highly precise due to high quality source data, but are likely to be overly optimistic given the simplicity in assuming only precipitation determines probability of presence (see text).
-	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
-	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.}
-	\label{fig:ex1_precip}
-\end{figure}
-
-%==================
-% FIGURE 4
-%TC:break fig 4
-\newpage
-\begin{figure}[h!]
-	\includegraphics[width=5.25in]{ex1_map.pdf}
-	\caption{Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
-	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
-	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
-	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.}
-	\label{fig:ex1_map}
-\end{figure}
-
-%==================
-% FIGURE 5
-%TC:break fig 5
-\newpage
-\begin{figure}[h!]
-	\includegraphics[width=6in]{ex2.pdf}
-	\caption{Results of model integration for sugar maple.
-	All panels show information using predicted climate from 2080 under the HadA2 scenario (see main text).
-	For reference, the present range of the species is outlined in red \citep{Little1971}.
-	The mechanistic sub-model Phenofit (A) predicted fairly small shifts sugar maple range.
-	In contrast, the naive model (B) projected large northward shifts in suitable habitat.
-	Model integration (C) produced predictions that were intermediate between the two models.
-	Uncertainty in the naive model (D) was fairly low, except for some portions of the northeast where uncertainty was extreme.
-	Model integration produced uncertainty (E) that was globally somewhat higher reflecting the disagreement between the naive model (B) and the sub-model (A), but with fewer extremes in prediction uncertainty.}
-	\label{fig:ex2}
-\end{figure}
-
-%==================
-% FIGURE 6
-%TC:break fig 6
-\newpage
-\begin{figure}[h!]
-	\includegraphics{management.pdf}
-	\caption{Sample workflow for applying the models presented in the first example in a management context.
-	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
-	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the metamodel.}
-	\label{fig:management}
-\end{figure}
-
 %% ----------------------------------
@@
 
-%TC:break box 1
-
-\newpage
-\linenumbers
-\section*{Box 1: Model description}
-% Boxes are limited to 750 words
+\section*{Appendix S1: Model description}
 We begin with a simple correlative model (i.e. calibrated with data provided at the same ecological scale as predictions), characterized in a Bayesian framework to allow further integration. 
@@
 For integration to be successful, it must be possible to compute the likelihood of this prediction given the metamodel, \(p \left( \psi_S \mid \theta_M \right) = p \left(\theta_S \mid \theta_M, X_S, D_S \right)\); the function \(g\) serves to transform the parameters of the sub-model to the scale of the metamodel.
-Thus, we refer to $g$ as the \emph{scaling function}.
+Thus, we refer to \(g\) as the \emph{scaling function}.
 In many cases, computing this likelihood will be challenging.
@@
 This new likelihood, which integrates the information from the second model into the first, can be treated as new ``data'' to evaluate the parameters of \(\theta_M\).
-We simply use the posterior distribution of $\theta_M$ from the previous step (Eq.
-\ref{eq:sdm4}) as the new prior probability of $\theta_M$, and evaluate the posterior probability of $\theta_M$ in light of the new data (i.e., information from the submodel \(\theta_S\)) and the prior knowledge, as before:
+We simply use the posterior distribution of \(\theta_M\) from the previous step (Eq.
+\ref{eq:sdm4}) as the new prior probability of \(\theta_M\), and evaluate the posterior probability of \(\theta_M\) in light of the new data (i.e., information from the submodel \(\theta_S\)) and the prior knowledge, as before:
 
@@
 This procedure may be applied to an arbitrary number of models, limited only by available data and computational power. 
-The models may be implemented simultaneously, when multiple datasets are available and can be evaluated under different underlying models, or they may be run sequentially, updating the posterior distribution of $\theta_M$ as new information becomes available.
-Note that $\theta_S$ itself need not be implemented in a Bayesian framework and the output need not be identical in form to the output of \(\theta_M\); it is enough to produce an output that can be evaluated as a likelihood under the first model and to have some assessment of the confidence in the parameter estimates of $\theta_S$ to use as a prior.
+The models may be implemented simultaneously, when multiple datasets are available and can be evaluated under different underlying models, or they may be run sequentially, updating the posterior distribution of \(\theta_M\) as new information becomes available.
+Note that \(\theta_S\) itself need not be implemented in a Bayesian framework and the output need not be identical in form to the output of \(\theta_M\); it is enough to produce an output that can be evaluated as a likelihood under the first model and to have some assessment of the confidence in the parameter estimates of \(\theta_S\) to use as a prior.
 
@@
 %---------------------------------------------
-
-\end{document}

### 2015-02-11 tex/manuscript.tex
Moved box 1 to an appendix

@@
 \usepackage{natbib} 
-\bibpunct{(}{)}{;}{a}{}{,}
+%\bibpunct{(}{)}{;}{a}{}{,} -- uncomment this for the final version
 
@@
 Both examples are presented with complete code and data (in Supporting Information).
-Finally, we provide a more formal mathematical presentation in Box 1.
+Finally, we provide a more formal mathematical presentation in Appendix S1.
 
@@
 For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, these functions can be simple while still contributing useful information to the metamodel (see Example 1).
-All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Box 1, Fig. \ref{fig:diagram}).
-We present and apply the framework here with two examples, and follow with a more formal presentation in Box 1.
+All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Appendix S1, Fig. \ref{fig:diagram}).
+We present and apply the framework here with two examples, and follow with a more formal presentation in Appendix S1.
 
@@
 For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature.
-We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty (see Appendix S1 in Supporting Information for procedural details and scripts for executing the model).
+We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty (see Appendix S2 in Supporting Information for procedural details and scripts for executing the model).
 
 We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation. 
-We will consider \(X_M\) (see Box 1, Fig. \ref{fig:diagram}) to be a vector of $n$ observations that takes the value of one to indicate presence and zero for absence:
+We will consider \(X_M\) (see Appendix S1, Fig. \ref{fig:diagram}) to be a vector of $n$ observations that takes the value of one to indicate presence and zero for absence:
 \begin{equation}
@@
 We used a binomial GLM to relate the suitability for sugar maple to 3 climate variables: the number of degree days, mean summer precipitation, and the ratio of annual precipitation to potential evapotranspiration.
-These variables were selected from the initial set of 6 variables from \citet{Morin2009}; see Appendix S2 for details on model construction and selection.
+These variables were selected from the initial set of 6 variables from \citet{Morin2009}; see Appendix S3 for details on model construction and selection.
 We selected a GLM for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
@@
 Bayesian methods provide a natural framework for the incorporation of multiple sources of information.
-Because the metamodel is expressed as a series of conditional probabilities (see Box 1), this information can be included simply provided the probability of the metamodel, conditional on information from the sub-models, can be expressed mathematically.
+Because the metamodel is expressed as a series of conditional probabilities (see Appendix S1), this information can be included simply provided the probability of the metamodel, conditional on information from the sub-models, can be expressed mathematically.
 Furthermore, Bayesian methods produce results as an entire distribution instead of a single point, allowing for a comprehensive understanding of the uncertainty of estimates \citep{Link2010}. 
@@
 \section*{Data Accessibility}
-All data, as well as all code required to repeat the analyses, have been uploaded as online supporting information in Appendix S3.
+All data, as well as all code required to repeat the analyses, have been uploaded as online supporting information in Appendix S4.
 
@@
 \section*{Supporting information}
-Appendix S1: Supporting information on using the code and data for example 1. \\
-Appendix S2: Supporting information on using the code and data for example 2. \\
-Appendix S3: Scripts and data required to reproduce the analyses.
+Appendix S1: Complete model description. \\
+Appendix S2: Supporting information on using the code and data for example 1. \\
+Appendix S3: Supporting information on using the code and data for example 2. \\
+Appendix S4: Scripts and data required to reproduce the analyses.
 
@@
 
-%% ----------------------------------
-%
-%     TEXT BOX
-%
-%% ----------------------------------
-
-%TC:break box 1
-
-\newpage
-\linenumbers
-\section*{Box 1: Model description}
-We begin with a simple correlative model (i.e. calibrated with data provided at the same ecological scale as predictions), characterized in a Bayesian framework to allow further integration. 
-We refer to this model (with associated parameters) as the metamodel, \(\theta_M\), and it is parameterized with information on the species' presence, \(X_M\), along with associated covariates (e.g., climate, presence/absence of interacting species, etc.), collectively referred to as \(D_M\).
-The naive (i.e., not integrated with information from sub-models) probability that a species is present, \(\psi_N\), is a deterministic function of the metamodel and its covariates::
-%-----
-\begin{equation}
-\label{eq:sdm1}
-	\psi_N = f(\theta_M, D_M)
-\end{equation}
-%-----
-The goal of model fitting is to estimate the posterior probability distribution for the parameters of \(\theta_M\), given the observations (\(X_M\)) and covariates (\(D_M\)):
-%-----
-\begin{equation}
-\label{eq:sdm2}
-	p( \theta_M \mid X_M, D_M)
-\end{equation}
-%-----
-
-Using Bayes' theorem, we find the posterior probability of \(\theta_M\) as follows:
-%-----
-\begin{equation}
-\label{eq:sdm3}
-	p( \theta_M \mid X_M, D_M ) = \frac{ p( X_M \mid \theta_M, D_M ) p( \theta_M ) } { p(X_M, D_M) }
-\end{equation}
-%-----
-where \(p( X_M \mid \theta_M, D_M )\) is referred to as the likelihood of the observations, \(p(\theta_M )\) is the prior probability of the model, and \(p(X_M, D_M)\) is the normalization constant.
-The normalization constant involves computing integrals that are often impossible to solve analytically. 
-In practice, simulation techniques such as MCMC can be used to sample directly from the posterior distribution, making such computations unnecessary.
-Thus, we use the proportional form of Bayes' Theorem:
-%-----
-\begin{equation}
-\label{eq:sdm4}
-	p( \theta_M \mid X_M, D_M ) \propto p( X_M \mid \theta_M, D_M ) p( \theta_M ) 
-\end{equation}
-%-----
-
-The role of the metamodel is to integrate data at the same ecological scale of predictions.
-Additional sub-models require outputs that should be comparable to this given ecological scale (e.g. constraining presence or absence on the landscape).
-Formally, we will add an additional model, \(\theta_S\), that is based on a different set of hypotheses and that makes predictions \((\psi_S)\) based on an additional dataset \((X_S, D_S)\):
-%-----
-\begin{equation}
-\label{eq:model2-1}
-	\psi_S = g(\theta_S, X_S, D_S)
-\end{equation}
-%-----
-
-For integration to be successful, it must be possible to compute the likelihood of this prediction given the metamodel, \(p \left( \psi_S \mid \theta_M \right) = p \left(\theta_S \mid \theta_M, X_S, D_S \right)\); the function \(g\) serves to transform the parameters of the sub-model to the scale of the metamodel.
-Thus, we refer to $g$ as the \emph{scaling function}.
-In many cases, computing this likelihood will be challenging.
-The simplest case is when the parameters of \(\theta_S\) are directly comparable to those in \(\theta_M\) (see Example 1), or when the scaling is performed directly within the sub-model (see Example 2), but other solutions are possible.
-This new likelihood, which integrates the information from the second model into the first, can be treated as new ``data'' to evaluate the parameters of \(\theta_M\).
-We simply use the posterior distribution of $\theta_M$ from the previous step (Eq.
-\ref{eq:sdm4}) as the new prior probability of $\theta_M$, and evaluate the posterior probability of $\theta_M$ in light of the new data (i.e., information from the submodel \(\theta_S\)) and the prior knowledge, as before:
-
-%-----
-\begin{equation}
-\label{eq:integrated2}
-	\overbrace{p( \theta_M \mid X_M, D_M, \theta_S, X_S, D_S )}^\text{integrated posterior}
-	\propto 
-	\overbrace{p( \theta_S \mid \theta_M, X_S, D_S )}^\text{likelihood}
-	\overbrace{p( X_M \mid D_M, \theta_M ) P( \theta_M )}^{\text{prior for } \theta_M}
-	\overbrace{p(\theta_S)}^{\text{prior for } \theta_S}
-\end{equation}
-%-----
-
-This procedure may be applied to an arbitrary number of models, limited only by available data and computational power. 
-The models may be implemented simultaneously, when multiple datasets are available and can be evaluated under different underlying models, or they may be run sequentially, updating the posterior distribution of $\theta_M$ as new information becomes available.
-Note that $\theta_S$ itself need not be implemented in a Bayesian framework and the output need not be identical in form to the output of \(\theta_M\); it is enough to produce an output that can be evaluated as a likelihood under the first model and to have some assessment of the confidence in the parameter estimates of $\theta_S$ to use as a prior.
-
-%---------------------------------------------
-%---------------------------------------------
-
 \end{document}

### 2015-02-13 tex/appendix_s1.tex
Revisions to intro and automatic word counting

@@
+\documentclass[11pt]{article}
+\usepackage[margin=1in, left=1.2in]{geometry}
+
+% use proper unicode fonts
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
+
+\usepackage{amsmath} % for better display of equations
+
 %% ----------------------------------
@@
 %% ----------------------------------
+\begin{document}
 
@@
 In many cases, computing this likelihood will be challenging.
-The simplest case is when the parameters of \(\theta_S\) are directly comparable to those in \(\theta_M\) (see Example 1), or when the scaling is performed directly within the sub-model (see Example 2), but other solutions are possible.
+The simplest case is when the parameters of \(\theta_S\) are directly comparable to those in \(\theta_M\) (see Main Text, Example 1), or when the scaling is performed directly within the sub-model (see Main Text, Example 2), but other solutions are possible.
 This new likelihood, which integrates the information from the second model into the first, can be treated as new ``data'' to evaluate the parameters of \(\theta_M\).
@@
 %---------------------------------------------
+\end{document}

### 2015-02-13 tex/manuscript.tex
Revisions to intro and automatic word counting

@@
 
-\usepackage[margin=1in, left=1.2in]{geometry}
+%\usepackage[margin=1in, left=1.2in]{geometry}
 
@@
 \usepackage{natbib}
-%\bibpunct{(}{)}{;}{a}{}{,} -- uncomment this for the final version
 
@@
 \usepackage{lineno}
+
+%% ----------------------------------
+%
+%     Packages to be used for drafts and editing
+%     Remove from the final manuscript
+%
+%% ----------------------------------
+\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
+\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
+\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
 
@@
 
-	\textbf{Main text word count: } 7121
+	\textbf{Main text word count: } 6197 %WORD_COUNT %Target 5000
 
@@
 
-Despite the recognized potential of these models, difficulties in producing predictions with acceptable precision and low bias has limited their contributions \citep{Guisan2013}.
-For mechanistic models, two important constraints on modeling ecological processes can prevent models from making unbiased predictions with acceptable uncertainty: 
-(1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
-Poor data or incoherent models will result in predictions with high uncertainty or precise but biased predictions.
-In recent decades, however, modeling techniques have proliferated to take advantage of what is available in terms of both data and modeling platforms.
-Model diversification also arises due to the diversity of processes generating species ranges \citep{Boulangeat2012}.
-Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, they may perform poorly when applied at the scale of species ranges.
-For instance, biotic interactions (including trophic interactions) are usually not modeled mechanistically at the scale of species distributions because they are poorly known, or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Pigot2013}.
+Despite the recognized potential of these models, it can be difficult to produce predictions with acceptable levels of precision and bias \citep{Guisan2013}.
+For mechanistic models, two important constraints can present difficulties in reducing uncertainty and bias: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
+In recent decades, however, modeling techniques have proliferated to take advantage of increased availability of data. 
+A growing body of theory, reflecting the diversity of processes generating species ranges, also has contributed to model diversification \citep{Boulangeat2012}.
+Of these model types, fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, perform poorly when applied at the scale of species ranges.
+For instance, biotic interactions are usually not modeled mechanistically at the scale of species distributions because they are poorly known or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Pigot2013}.
 In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
@@
 \subsection*{Toward an integrated approach for modeling ranges}
+\mt{this transition is fairly abrupt. it would be nice to have more of an intro explaining the motivation. perhaps it comes later in the intro; if so, it should be moved here}
+\mt{a structure to consider:\\
+1. intro sentence (gee it would be great to combine data sources) \\
+2. "Some researchers... hierarchical nature..." (but make it less awful) \\
+3. Hybrid models (and rewrite -- it is a bit redundant) \\
+4. Ensemble forecasting
+5. Our approach
+-- could break this into different paragraphs
+}
 To include knowledge of species from multiple sources and scales into distribution models, we present an application of hierarchical Bayesian methods that uses outputs from multiple model outputs to inform the results of the final model.
@@
 By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
+\mt{Blech. This transition is also terrible}
 Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Boulangeat2012, Catterall2012, Strigul2012, Soranno2014}.
@@
 A commonly-employed alternative to hybrid models is the direct combination of predictions.
-If models operate at the same scale and have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}.
-If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Cheaib2012, Serra-Diaz2013}.
-However, the applicability of ensemble forecasts is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
+Predictions from models operating at the same spatio-temporal scales and having compatible parameterizations can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}.
+However, the applicability of such techniques is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
 A more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying the parameters that contribute most to variance in the model predictions \citep{McMahon2011}.
-Clearly, an approach that integrates predictions and uncertainty from multiple model types is needed \citep{Beck2012, Thuiller2013}. 
-Such an approach would incorporate multiple modes of inference and integrate data from multiple sources and scales \citep{Levin1992, Peters2004, Thuiller2013}.
+Clearly, we require an approach that can both (1) flexibly incorporate multiple modes of inference (i.e., mechanistic, correlative models), (2) encompasses data from multiple sources and scales, and (3) explicitly includes uncertainty from all sources \citep{Levin1992, Peters2004, Beck2012, Thuiller2013}.
 
@@
 We present and apply the framework here with two examples, and follow with a more formal presentation in Appendix S1.
+\mt{There is still some repetition here -- two refs to Appendix S1 in 2 sentences}
 
@@
 We will consider \(X_M\) (see Appendix S1, Fig. \ref{fig:diagram}) to be a vector of $n$ observations that takes the value of one to indicate presence and zero for absence:
+\mt{will need to edit this, since \(X_M\) is no longer defined clearly in the text (has been moved to appendix S1)}
 \begin{equation}
@@
 
-%TC:ignore
 \section*{Data Accessibility}
@@
 281422 (TEEMBIO).
-%TC:endignore
 
@@
 
-%TC:break references
 %% ----------------------------------
@@
 
+%TC:break biosketch
 \section*{Biosketch}
-%TC:break biosketch
 The Theoretical Ecosystem Ecology lab group at the Université du Québec à Rimouski focuses on interactions between species distributions, community structure and ecosystem functioning.
@@
 
+%TC:break figures
 \newpage
-%TC:break fig legends
 \section*{Figure Legends}

### 2015-02-13 tex/wordcount.sh

@@ -1,0 +1,2 @@
+word=`texcount -sub=none manuscript.tex | sed -n -E 's/[[:space:]]+([0-9]+).+manuscript/\1/p'`
+sed -i backup -E 's/[0-9]+([[:space:]]+%WORD_COUNT)/'"${word}"'\1/g' manuscript.tex 

### 2015-03-24 tex/appendix_s1.tex
Major revisions in prep for GEB submission

@@
 \usepackage{amsmath} % for better display of equations
+\usepackage{natbib}
 
-%% ----------------------------------
-%
-%     TEXT BOX
-%
-%% ----------------------------------
+\usepackage{graphicx}
+%\usepackage[nomarkers,nolists,heads]{endfloat}
+\usepackage{caption}
+\DeclareCaptionLabelFormat{appendix}{#1 S#2}
+\captionsetup{labelformat=appendix}
+
+% Tikz libraries for building the diagram
+\usepackage{tikz}
+\usepackage{tikz-cd}
+\usetikzlibrary{calc, shapes}
+\usetikzlibrary{shapes.geometric,shapes.arrows,decorations.pathmorphing}
+\usetikzlibrary{matrix,chains,scopes,positioning,arrows,fit}
+
 \begin{document}
 
-\section*{Appendix S1: Model description}
+\section*{Appendix S1: Model description and methodology}
+\subsection*{Model description}
+
 We begin with a simple correlative model (i.e. calibrated with data provided at the same ecological scale as predictions), characterized in a Bayesian framework to allow further integration. 
-We refer to this model (with associated parameters) as the metamodel, \(\theta_M\), and it is parameterized with information on the species' presence, \(X_M\), along with associated covariates (e.g., climate, presence/absence of interacting species, etc.), collectively referred to as \(D_M\).
-The naive (i.e., not integrated with information from sub-models) probability that a species is present, \(\psi_N\), is a deterministic function of the metamodel and its covariates::
+We refer to this model (with associated parameters) as the metamodel, \(\theta_M\), and it is parameterized with information on the species' presence, \(X_M\), along with associated covariates (e.g., climate, presence/absence of interacting species, etc.), collectively referred to as \(D_M\) (Fig.
+S\ref{fig:diagram}).
+The naive (i.e., not integrated with information from sub-models) probability that a species is present, \(\psi_N\), is a deterministic function of the metamodel and its covariates:
 %-----
@@
 
+%==================
+% FIGURE 1
+\begin{figure}[t]
+\input{diagram}
+\caption{The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
+	The metamodel ($\theta_M$) operates at a single scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
+	The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
+	The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
+	This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
+	This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).}
+\label{fig:diagram}
+\end{figure}
+%==================
+
 Using Bayes' theorem, we find the posterior probability of \(\theta_M\) as follows:
@@
 The normalization constant involves computing integrals that are often impossible to solve analytically. 
-In practice, simulation techniques such as MCMC can be used to sample directly from the posterior distribution, making such computations unnecessary.
+In practice, simulation techniques such as MCMC can be used to sample directly from the posterior distribution, making such computations unnecessary \citep{Link2006}.
 Thus, we use the proportional form of Bayes' Theorem:
@@
 In many cases, computing this likelihood will be challenging.
-The simplest case is when the parameters of \(\theta_S\) are directly comparable to those in \(\theta_M\) (see Main Text, Example 1), or when the scaling is performed directly within the sub-model (see Main Text, Example 2), but other solutions are possible.
+The simplest case is when the parameters of \(\theta_S\) are directly comparable to those in \(\theta_M\), or when the scaling is performed directly within the sub-model, but other solutions are possible.
 This new likelihood, which integrates the information from the second model into the first, can be treated as new ``data'' to evaluate the parameters of \(\theta_M\).
@@
 %---------------------------------------------
+
+\section*{Example 1}
+To produce the simulated dataset used in the first example, we first simulated 100 presence-absence points at randomized locations in ecological space (i.e., temperature and precipitation) using the spsample function from package sp in R \citep{Bivand2013, R}.
+This function generates points with a specified amount of spatial clustering (ranging from 0 to 1, where 1 represents complete spatial independence); we selected a value of 0.2 for our analysis.
+Presence or absence at each point was determined by randomly sampling from a Bernoulli distribution, where the probability was selected using a pre-determined function of temperature and precipitation.
+We then fit the naive model in JAGS as a logistic regression, with both linear and quadratic terms for both temperature and precipitation and using uninformative priors (Normal with \(\mu = 0\) and \(\sigma = 10000\)).
+We discarded the first 5000 MCMC samples as burnin, and then collected an additional 2000 samples for analysis.
+The final sample size was selected to provide for relatively rapid computational time; the addition of longer final samples or extended burnin periods had no effect on the results.
+
+For the mechanistic submodel, we computed the predicted probability of presence as a function of precipitation, using the experimental data and the theoretical prediction that the species will be present when the population growth rate is greater than 0.
+For each of the five experimental treatments, we computed the probability of presence as the integral of the normal density from 0 to infinity, where the mean of the normal distribution was equal to the average population growth rate at each precipitation treatment and the standard deviation was the corresponding standard error for each treatment.
+We then fit these data to the same model, priors, and sample sizes as those from the previous step.
+Fitting the metamodel is simply a matter of repeating the exact procedure for fitting the naive model, but using informative priors generated from the naive model step.
+Because the sub-model considered only precipitation, and because the response was highly correlated to precipitation, we expected the submodel to produce highly precise estimates of the parameters for the effect of precipitation on the probability of presence.
+However, the sub-model was quite simplistic, and thus likely over-estimates its precision when considering the species range wholistically (as is done with the metamodel).
+Thus, we applied a prior weight of 0.05 to the sub-model.
+This allowed the sub-model to inform the integrated model without dominating the results.
+
+All scripts were prepared in R and tested under R version 3.0.3.
+The MCMC analyses were performed using the JAGS (Just Another Gibbs Sampler) software package, version 3.4.0 \citep{RJAGS}.
+Additionally, the rjags, coda, and sp packages are required to complete the analysis.
+
+\section*{Example 2}
+We formulated a metamodel for the second example that allowed for the full use of three sources of data: occurrences from forestry plots (see below) as well as coarser-scale presence-absence information and Phenofit projections \citep[both from][]{Morin2009}.
+Plot data originated from four forest permanent plot databases and included samples widely distributed in Eastern North America, from Florida to north Canada (Fig. S\ref{fig:plotmap}).
+In the US, we included 86,000 plots standardized since 1990 and monitored until 2013 with up to 4 remeasurements by forest plot \citep{OConnell2013}.
+Quebec data were provided from the Ministère des Forêts de la Faune et des Parcs with 12,409 permanent plots, and DOMTAR, a forest company in paper production with 1,741 plots, and surveyed from 1960--2011 with up to 10 remeasurements \citep{MFFP2013}.
+Ontario and New-Brunswick included 1,038 and 2,748 plots respectively \citep{Porter1999, Ontario2014}.
+Ontario monitored forest plots from 1992--2006 with up to 3 remeasurement, and New-Brunswick from 1985--2010 with up to 7 remeasurements.
+
+\begin{figure}[t]
+\includegraphics{figs/plotmap.pdf}
+\caption{Map of permanent forest plot locations}
+\label{fig:plotmap}
+\end{figure}
+
+We divided the dataset into calibration and validation subsets, with 1/3 of the data reserved for validation and the remaining 2/3 used for calibration.
+As in the previous example, we modeled the probability of presence as a function of climate.
+We considered seven climate variables: the number of degree days (ddeg), minimum temperature (min\_temp), potential evapotranspiration (PET), annual precipitation (an\_prcp), summer precipitation (sum\_prcp), winter precipitation (win\_prcp), and the ratio of annual precipitation to potential evapotranspiration (pToPET).
+See \citet{Morin2009} for a complete description of these variables.
+For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the models to forecast suitability into the future.
+Exploratory analysis revealed that there were collinearities within the predictor variables.
+Thus, we only included three variables in the model: ddeg, an\_prcp, and pToPET.
+To select the form of the naive model (and thus the metamodel), we used stepwise regression with BIC as the evaluation criteria.
+The search scope included all models with linear, quadratic, and cubic terms for all three variables.
+Because Phenofit provided predictions in the form of a probability of presence (rather than directly in terms of the metamodel parameters as in the previous example), it was necessary to develop a method for relating these predictions to the metamodel.
+We opted to do this via simulation because it is a natural method to use in an MCMC scheme.
+For each MCMC iteration, we generated a simulated dataset by drawing \(N_x\) samples with probability \(p_x\) from a Binomial distribution, where \(N_x\) is the number of observations in the occurrence dataset for a given grid cell \(x\), and \(p_x\) is the predicted probability of presence from Phenofit for that cell.
+The metamodel was then fit to these simulated data (with prior conditioning on the naive model).
+This procedure, by generating a random dataset with each iteration, incorporated the variance in the Phenofit predictions.
+Convergence for all models was assessed by comparing the results of multiple chains with overdispersed starting points.
+Based on the results of these initial runs, we thinned all posterior samples by recording only every 50th sample, and we obtained 25000 posterior samples for each model following a burn-in period of 20000 samples.
+
+\renewcommand\refname{Literature Cited}
+\bibliography{model_integration}{}
+\bibliographystyle{model_integration}
+
 \end{document}

### 2015-03-24 tex/manuscript.tex
Major revisions in prep for GEB submission

@@
 
-%\usepackage[margin=1in, left=1.2in]{geometry}
+\usepackage[margin=1in]{geometry}
 
@@
 \usepackage{natbib}
-
-\usepackage{caption}
-\captionsetup{labelformat=simple, labelsep=none}
 
@@
 \usepackage{graphicx}
-
-% Tikz libraries for building the diagram
-\usepackage{tikz}
-\usepackage{tikz-cd}
-\usetikzlibrary{calc, shapes}
-\usetikzlibrary{shapes.geometric,shapes.arrows,decorations.pathmorphing}
-\usetikzlibrary{matrix,chains,scopes,positioning,arrows,fit}
+%\usepackage[nomarkers,nolists,heads]{endfloat}
+\usepackage{caption}
+\captionsetup{labelformat=simple}
+\usepackage{booktabs}
+\usepackage{multirow}
 
@@
 %% ----------------------------------
-\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
-\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
-\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
+%\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
+%\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
+%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
 
@@
 
-	\textbf{Abstract word count: } 258
+	\textbf{Abstract word count: } 258 %ABS_COUNT
 
-	\textbf{Main text word count: } 6197 %WORD_COUNT %Target 5000
+	\textbf{Main text word count: } 4641 %WORD_COUNT %Target 5000
 
@@
 Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}.
-Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution \citep{Jay2012}, hybrid zone dynamics \citep{Engler2013}, and forecasting the impacts of climate change on species distributions \citep{Thuiller2011, Blois2013, Thuiller2014}.
+Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, hybrid zone dynamics \citep{Engler2013}, and forecasting the impacts of climate change on species distributions \citep{Blois2013, Thuiller2014}.
 
-Despite the recognized potential of these models, it can be difficult to produce predictions with acceptable levels of precision and bias \citep{Guisan2013}.
-For mechanistic models, two important constraints can present difficulties in reducing uncertainty and bias: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
-In recent decades, however, modeling techniques have proliferated to take advantage of increased availability of data. 
+Despite the recognized potential of these models, it can be difficult to produce species distribution models (SDMs) with acceptable levels of precision and bias \citep{Guisan2013}.
+For mechanistic models, two important constraints can present difficulties: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
+In recent decades, however, modeling techniques have proliferated to take advantage of increased data availability. 
 A growing body of theory, reflecting the diversity of processes generating species ranges, also has contributed to model diversification \citep{Boulangeat2012}.
 Of these model types, fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, perform poorly when applied at the scale of species ranges.
-For instance, biotic interactions are usually not modeled mechanistically at the scale of species distributions because they are poorly known or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Pigot2013}.
+For instance, biotic interactions are usually not modeled mechanistically at the scale of species distributions because they are poorly known or have not been recorded, despite being considered a key determinant of range limits \citep{Pigot2013}.
 In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
@@
 
-\subsection*{Toward an integrated approach for modeling ranges}
-\mt{this transition is fairly abrupt. it would be nice to have more of an intro explaining the motivation. perhaps it comes later in the intro; if so, it should be moved here}
-\mt{a structure to consider:\\
-1. intro sentence (gee it would be great to combine data sources) \\
-2. "Some researchers... hierarchical nature..." (but make it less awful) \\
-3. Hybrid models (and rewrite -- it is a bit redundant) \\
-4. Ensemble forecasting
-5. Our approach
--- could break this into different paragraphs
-}
-To include knowledge of species from multiple sources and scales into distribution models, we present an application of hierarchical Bayesian methods that uses outputs from multiple model outputs to inform the results of the final model.
-Our approach is an alternative to other approaches to multimodel inference that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
-By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
-\mt{Blech. This transition is also terrible}
-Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Boulangeat2012, Catterall2012, Strigul2012, Soranno2014}.
-Recent literature has also focused on the development of hybrid models, which allow for flexible combinations between mechanistic and phenomenological sub-models \citep{Gallien2010, Thuiller2013}. 
-For instance, correlative models are used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while mechanistic approaches can include space and time dynamics and therefore account for dispersal processes \citep{Kearney2008}.
-This complementarity has been used to merge the two model types into hybrid models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Dormann2012, Boulangeat2014}. 
+Clearly, an approach is needed that unifies the strength of the diversity of modeling approaches and can incorporate multiple data sources and modeling paradigms.
+Toward this end, we present here an application of hierarchical Bayesian methods that uses outputs from multiple model outputs to inform the results of the final model.
+Techniques for multimodel inference have proliferated in recent years.
+For example, hybrid models that allow for flexible combinations between mechanistic and phenomenological sub-models are commonly employed in SDMs \citep{Gallien2010, Thuiller2013, Boulangeat2014}.
+Within the hybrid framework, a correlative model might be used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while a more mechanistic approach could include space and time dynamics and therefore account for dispersal \citep{Smolik2010}.
 However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and the sources and magnitudes of uncertainties are difficult to quantify.
-Moreover, hybrid approaches are not suitable for merging models that are based on the same processes but with different underlying hypotheses.
+A commonly-employed alternative to hybrid models is the direct combination of predictions, where predictions from models operating at the same spatio-temporal scales can be combined \citep[e.g., model averaging, ensemble forecasting;][]{Araujo2007}.
+However, the applicability of such techniques is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
 
-A commonly-employed alternative to hybrid models is the direct combination of predictions.
-Predictions from models operating at the same spatio-temporal scales and having compatible parameterizations can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}.
-However, the applicability of such techniques is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
-A more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying the parameters that contribute most to variance in the model predictions \citep{McMahon2011}.
-Clearly, we require an approach that can both (1) flexibly incorporate multiple modes of inference (i.e., mechanistic, correlative models), (2) encompasses data from multiple sources and scales, and (3) explicitly includes uncertainty from all sources \citep{Levin1992, Peters2004, Beck2012, Thuiller2013}.
+We propose an alternative to these other approaches to multimodel inference.
+We use the power and generality of a hierarchical Bayesian framework, which provides a number of desirable advantages, including (1) the ability to incorporate multiple modes of inference (e.g., mechanistic, correlative models) \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}, (2) an easy mechanism to include multiple data sources at various scales \citep{Levin1992, Peters2004}, and (3) intuitive and comprehensive reporting of uncertainty in model predictions that  reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}.
+Contrary to hybrid and ensemble methods, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for as much flexibility as possible regarding the type of information included.
+By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Guisan2005, Araujo2006}.
+Moreover, a more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying the parameters that contribute most to variance in the model predictions \citep{McMahon2011}.
 
-Here, we present a framework for modeling species range dynamics that aims to account for all the information available on a focal species, from a variety of sources and scales.
-Contrary to hybrid and hierarchical models, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for as much flexibility as possible regarding the type of information included.
-We use the power and generality of a hierarchical Bayesian framework, which allows us to include multiple data sources and modes of inference \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}. 
-Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to interpret and reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}. 
 We illustrate our approach with two examples.
-We begin with a hypothetical example, using simulated data, that presents a simplified application of our framework to multiple sources of information from different scales (Example 1).
+We begin with a hypothetical example, using simulated data, where we define the framework and demonstrate its application to multiple sources of information from different scales.
 In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread foundation tree species from eastern North America, in response to climate change.
@@
 %
-%     MODELING FRAMEWORK
-%
-%% ----------------------------------
-
-\section*{Integrated modeling framework}
-
-The key idea of our approach is formulate a metamodel that integrates data at the same ecological scale as the desired predictions.
-The parameters of this model can be constrained using the output of one or more sub-models that can operate at different scales than the metamodel.
-As an example, the metamodel could take the form of a simple correlative species distribution model (SDM) operating at very large spatial scales, with one or more fine-scaled process-based sub-models.
-Ideally, each sub-model incorporates different data types and hypotheses that specify underlying ecological mechanisms contributing to the model output.
-The output of sub-models should provide additional predictions at a comparable ecological scale at which the metamodel operates, although it is not necessary for the sub-models to actually operate at the same scale.
-For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, these functions can be simple while still contributing useful information to the metamodel (see Example 1).
-All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Appendix S1, Fig. \ref{fig:diagram}).
-We present and apply the framework here with two examples, and follow with a more formal presentation in Appendix S1.
-\mt{There is still some repetition here -- two refs to Appendix S1 in 2 sentences}
-
-%% ----------------------------------
-%
 %     EXAMPLE 1
@@
 %% ----------------------------------
-\subsection*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
-
+\section*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
+The key idea of our approach is formulate a metamodel that integrates data at the same ecological scale as the desired predictions, and to constrain the parameters of this model using the output of one or more sub-models
 In this hypothetical example, we build a metamodel relating the distribution (i.e.
@@
 However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the climatic envelope of where a species may be found in a natural setting.
-Prior information of the physiological constraints affecting species distribution is sometimes available, but usually too incomplete to perform a reasonable comparison with the realized distribution.
-For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature.
-We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty (see Appendix S2 in Supporting Information for procedural details and scripts for executing the model).
+Thus, we use additional information on the physiological constraints affecting species distribution.
+Because these data are often collected at a finer scale than that of rangewide occurence data, we apply a simple scaling function, drawing on ecological theory, to compute the likelihood of a set of metamodel parameters given both the occurrence and the physiological data.
+See Appendix S1 in Supporting Information for procedural details and scripts for executing the model.
+
+%==================
+%TC:ignore
+\begin{figure}[t]
+	\includegraphics{figs/ex1_sampling.pdf}
+	\caption{Two simulated datasets used to illustrate the model integration framework.
+	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
+	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., \(-\)1--1).
+	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
+	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.}
+	\label{fig:ex1_sampling}
+\end{figure}
+%TC:endignore
+%==================
 
 We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation.
-We will consider \(X_M\) (see Appendix S1, Fig. \ref{fig:diagram}) to be a vector of $n$ observations that takes the value of one to indicate presence and zero for absence:
-\mt{will need to edit this, since \(X_M\) is no longer defined clearly in the text (has been moved to appendix S1)}
-\begin{equation}
-X_M = \{x_{M,1}, x_{M,2}, \ldots, x_{M,n}\}
-\end{equation}
-We assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
-The model is evaluated by relating these observations to environmental data \((D_M)\), which, for the sake of the example, consists of mean annual temperature $T_M$ and annual precipitation $P_M$. 
-For simplicity, we formulate the metamodel \((\theta_M)\) as a simple logistic regression with a second order effect of temperature and precipitation. 
-This naive model (i.e., the metamodel with no constraints from sub-models) predicts  occurrence probability (\(\psi_N\)) as a function of the environment:
+We assume a presence-absence dataset, which we term \(X_M\), where the subscripted \(M\) indicates that the data were collected at the same scale as the metamodel.
+We simulated a relatively high-quality dataset that provides coverage of a variety of ecological conditions (Fig. \ref{fig:ex1_sampling}).
+We desire to model the species' distribution as a function of temperature, \(T_M\), and precipitation, \(P_M\).
+An initial version of the metamodel \((\theta_M)\) that has no constraints from other datasets will be referred to as the naive model.
+This naive model uses a simple Bayesian logistic regression to estimate the naive probability of occurrence \((\psi_N)\) as a function of second order effects of temperature and precipitation:
 %-----
@@
 \begin{aligned}
-	\psi_N &= f\left(\theta_M, D_M \right) \\
+	\psi_N &= f\left(\theta_M, T_M, P_M \right) \\
 	&= p \left (X_M = 1 \mid \theta_M, T_M, P_M \right) \\
-	&=\text{logit}^{-1}\left( \mathbf{\Theta_M} \mathbf{D_M} \right)
+	&=\text{logit}^{-1}\left( \theta_M D_M \right)
 \end{aligned}
@@
 %-----
-where \(\mathbf{\Theta_M}\) is the parameter vector of the model \(\theta_M\), \(\mathbf{D_M} \) is the covariate matrix (i.e., \(T_M, P_M\), with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
-We can fit this model in a Bayesian framework to allow for easy integration of models (as we will show later).
-In this context, the goal of modeling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, a concise introduction can be found in ][]{Link2010}:
+where \(theta_M\) is the parameter vector of the model, \(D_M\) is the covariate matrix (i.e., \(T_M, P_M\), and quadratic terms, with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
+We estimate parameters using a Markov Chain Monte Carlo (MCMC) method using the proportional form of Bayes' theorem:
 %-----
@@
 According to niche theory \citep{Holt2009}, the fundamental niche corresponds to the set of environmental conditions where the per capita intrinsic growth rate $r$ is positive.
-This concept gives us a reasonable model to fit the scaling function $g$ for our sub-model (Fig. \ref{fig:diagram}).
-If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, then for observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
+This concept gives us a reasonable model to fit a scaling function for our sub-model (Appendix S1).
+If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, then for an observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
 \begin{equation}
@@
 Although the two datasets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation).
-This scaling is quite simplistic, and would never be used to predict a species' range using this mechanistic sub-model in isolation.
-Despite the strong theoretical foundation (i.e., niche theory) linking the observations to presence-absence at large scales, the upscaled sub-model (i.e., the transformation of \(X_S\) and \(\sigma_S\) to \(\psi_S\)) is built on the strong hypothesis that the species will be present when $r>0$, neglecting other population dynamics constraints such as Allee effects and metapopulation dynamics.
-The fundamental niche is also incomplete, as we lack environmental variables such as temperature, as well as all other ecological processes responsible for the realized distribution.
+The scaling is quite simplistic; it treats the fundamental niche as the only driver of species distribution, and only considers a single dimension of the niche.
 As such, it would be unwise to expect predictions from this model alone to resemble the actual distribution of the species; as a mechanistic model, it is simply too incomplete to predict distribution.
@@
 %-----
-As before, the metamodel \(\theta_M\) can be used to predict the species distribution \((\psi_I)\).
-However, these predictions will not reflect only the presence-absence samples in \(X_M\), but will also reflect the information from \(\theta_S\), including all of the data sources used to produce this sub-model.
+As before, the metamodel \(\theta_M\) can be used to predict probability of occurrence \((\psi_I)\).
+However, these predictions now reflect the presence-absence data \(X_M\) as well as the information from \(\theta_S\), including all of the data sources used to produce this sub-model.
 In other words, with this evaluation, the set of parameters $\theta_M$ is most likely when it agrees with both the original data and predictions from the sub-model. 
-Note that in this particular example, because the sub-model is only contingent on precipitation, there is no integration on temperature. 
 Finally, we note the presence of marginal distributions for both models (i.e., \(p(\theta_M)\) and \(p(\theta_S)\)).
@@
 
-We performed this procedure with our hypothetical datasets.
+%==================
+\begin{figure}[bt]
+	\includegraphics{figs/ex1_precip.pdf}
+	\caption{Comparison of the naive model, mechanistic sub-model, and the integrated model showing the probability of presence (\(\psi\)) as a function of precipitation.
+	(a) Naive model, using only presence-absence data. Uncertainty increases dramatically when attempting to project beyond the scope of the source data.
+	(b) Mechanistic model, using observations of an experiment to infer probability of presence.
+	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
+	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals. The shaded region shows the calibration range for the naive model.}
+	\label{fig:ex1_precip}
+\end{figure}
+%==================
+
+%==================
+\begin{figure}[tb]
+	\includegraphics[width=5.25in]{figs/ex1_map.pdf}
+	\caption{Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
+	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
+	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
+	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.
+}
+	\label{fig:ex1_map}
+\end{figure}
+%==================
+
 When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data (Figs.
-\ref{fig:ex1_precip}a, \ref{fig:ex1_map}a, \ref{fig:ex1_map}b)
-Unsurprisingly given the quality of the data, the sub-model was highly precise with respect to precipitation and not useful for temperature, thus providing a fairly strong constraint when producing the integrated model (Fig \ref{fig:ex1_precip}b).
+\ref{fig:ex1_precip}a, \ref{fig:ex1_map}a, \ref{fig:ex1_map}b).
+The sub-model was highly precise with respect to precipitation, thus providing a fairly strong constraint when producing the integrated model (Fig \ref{fig:ex1_precip}b).
 The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
@@
 \ref{fig:ex1_map}).
-
-Once the technical challenges of evaluating the parameters are overcome, integration has several advantages over both the naive parameterization of the metamodel and the mechanistic sub-model. 
-In the situation where both models strongly agree on the relationship between water availability and occurrence probability, then we should expect significantly reduced uncertainty in that domain (Figs.
-\ref{fig:ex1_precip}, \ref{fig:ex1_map}).
-Because the metamodel includes both temperature and precipitation, any reduction in the uncertainty with respect to precipitation might also influence the evaluation of the relationship with temperature (reducing bias in the case where there is a correlation between T and P in the data).
-However, uncertainty could increase in the case of model disagreement.
-For instance, the relationship between precipitation and presence in the naive metamodel could be caused by a spurious relationship (e.g.
-spatial autocorrelation caused by historical contingencies), leading to a false confidence in the species distribution over precipitation gradients.
-The information from the sub-model will thus bring back the right amount of uncertainty in the relationship.
-Moreover, convergence in model predictions in overlapping domains (e.g., near the peak in the probability of presence in Fig. \ref{fig:ex1_precip}) can aid in detecting crucial ecological processes.
 
@@
 Despite being a relatively common application of SDMs \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios remains problematic \citep{Araujo2006}.
-We used our metamodeling framework to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
-We used data from \citet{Morin2009} for sugar maple, an economically and ecologically important species occurring in eastern North America.
-Here we describe briefly the dataset, methods, and the results of the analysis (see Appendices S2 and S3 for data and scripts to reproduce the analysis).
+We used our framework to constrain a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
+Here we describe briefly the dataset, methods, and the results of the analysis (see Appendix S1 for data and scripts to reproduce the analysis).
 
-\defcitealias{IPCC2001}{IPCC, 2001} % this is needed to make the organizational citation in the following paragraph work correctly
-For the metamodel, we used a dataset of 1013 presence and 13863 absence records at 0.5 degree resolution.
-We used a binomial GLM to relate the suitability for sugar maple to 3 climate variables: the number of degree days, mean summer precipitation, and the ratio of annual precipitation to potential evapotranspiration.
-These variables were selected from the initial set of 6 variables from \citet{Morin2009}; see Appendix S3 for details on model construction and selection.
-We selected a GLM for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., GAM) are compatible with the framework.
-For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the GLM to forecast suitability into the future (i.e., \(\psi_N\)).
-Phenofit produces predictions at continental scales; thus the scaling function \(g\) was incorporated within Phenofit itself and we performed model integration directly on the outputs.
-To perform the integration, we used a Metropolis-Hastings scheme within Markov Chain Monte Carlo (MCMC) sampling to condition the parameters of the metamodel on two likelihood functions.
-The first was the naive SDM, where the likelihood of a ``success'' in the presence/absence dataset was a Bernoulli density with the (logit-transformed) probability expressed as a linear function of the climate predictors.
-For the sub-model, we considered that Phenofit, being a process-based model, may perform better when projecting into the future, particularly if suitability is primarily driven by the phenological match to the environment.
-Thus, we used the predicted future probability of presence as the basis for generating simulated future presence-absence datasets.
-This procedure inherently incorporates some uncertainty by randomly determining where the species is present (though uncertainty in the probabilities themselves could not be incorporated, as it is not possible to obtain them from Phenofit).
-Thus, the sub-model likelihood is simply the probability, given the predictions of the metamodel, of observing a simulated dataset \(X_s\) that is randomly drawn from the phenofit predictions \(\psi_s\); i.e., \(X_s = f(\psi_s)\):
-\begin{equation}
-\label{eq:integrated2}
-	p( \theta_M \mid X_M, D_M, \psi_s, \theta_S )
-	\propto 
-	p( X_S = f(\psi_s)\mid \theta_M, \theta_S )
-	p( X_M \mid \theta_M, D_M ) 
-	p( \theta_M )
-	p( \theta_S )
-\end{equation}
-where \(\theta_M\) is the metamodel, 
-\(X_M\) is the vector of contemporary presences and absences, 
-\(D_M\) is the matrix of climate predictors,
-and \(\theta_S\) is the model relating the Phenofit predictions to the metamodel.
-Furthermore, the integration is performed on the \emph{predictions} of the sub-model directly, rather than on compatible parameters as in Example 1. 
-Thus, the metamodel must appear in the sub-model likelihood.
+We obtained climate variables, occurrence data, and Phenofit projections at 0.5-degree resolution for both the present and for 2080 for sugar maple, an economically and ecologically important species occurring in eastern North America \citep{Morin2009}.
+These data defined the metamodel scale.
+To these data, we added 4903 recorded presences and 21701 absences derived from permanent forest sample plots located in the United States and Canada.
+We reserved 1/3 of this dataset for validation, and used the remaining records to calibrate the models.
+We constructed the naive model by using a binomial generalized linear model (GLM) to relate the presence-absence dataset to three climate variables: the number of degree days (ddeg), mean annual precipitation (an\_prcp), and the ratio of annual precipitation to potential evapotranspiration (pToPET).
+These variables were selected from an initial set of 6 variables \citep[see Appendix S1 and ][for details on the climate variables]{Morin2009}.
+We selected a GLM for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., generalized additive models) are compatible with the framework.
+The parameters for the model were estimated using a Metropolis-Hastings algorithm within Markov Chain Monte Carlo (MCMC).
 
-Considering the models individually, the predictions of Phenofit were much more conservative than those of the naive model (Fig \ref{fig:ex2}).
+To perform the integration, we constrained the estimates of the naive model with the additional information from Phenofit while considering two different potential modeling goals.
+The first is improving our model of the present range of the species.
+Using both datasets to develop a range model for the species has a number of advantages.
+Assuming we have chosen climate variables that well represent the constraints on the species, including Phenofit in our model is likely to reduce bias in our estimate of the bioclimatic niche.
+Second, the posterior predictions of the model (that is, the probability of presence projected into geographic space) will incorporate uncertainty from all sources, including the modeling process, data, parameterization, and the ecological processes included.
+This can provide a much more accurate estimate of the uncertainty of our predictions.
+Thus, for our first integration we combined the naive model with the Phenofit predictions for the present; we refer to this model as ``Integrated-Present'' hereafter.
+We also validated this model and compared it against the naive model by computing the area under the receiver operating curve (AUC), a commonly-used metric that measures the rate of correct classification and varies from 0 to 1 (where a value of 1 indicates perfect classification) \citep{Swets1988}.
+The second modeling goal is projecting changes to the range of sugar maple following climate change.
+Process-based models and correlative models often differ substantially when projecting beyond the range of the original data (see Example 1).
+Thus, we obtained predictions for 2100 from Phenofit \citep{Morin2009} and used the predictions to condition the metamodel given the predicitons of the naive model under future climate.
+This procedure provided a consensus view of the future range of the species; we refer to this model as ``Integrated-Future.''
+
+\begin{table}[t]
+	\footnotesize
+	\caption{Standardized logit-scale parameter estimates and 95\% credible intervals for the three models.}
+	\begin{tabular}{lcccccccc}
+	\toprule
+	& Naive & Integrated-present & Integrated-future \\
+	\midrule
+	\multirow{2}{*}{intercept} & -0.886 & -0.103 & -1.037 \\
+	& (-1.12, -0.66) & (-0.23, 0.026) & (-1.16,-0.92) \\
+	& & & \\
+
+	\multirow{2}{*}{ddeg} & 2.904 & 3.701 & 6.431 \\
+	& (2.34, 3.45) & (3.37, 4.03) & (6.07, 6.79) \\
+	& & & \\
+
+	\multirow{2}{*}{ddeg\textsuperscript{2}} & -6.697 & -6.216 & -5.241 \\
+	& (-7.04, 6.35) & (-6.43,-6.00) & (-5.43,-5.05) \\
+	& & & \\
+
+	\multirow{2}{*}{ddeg\textsuperscript{3}} & 1.669 & 1.454 & 0.893 \\
+	& (1.52, 1.79) & (1.37, 1.53) & (0.85, 0.94) \\
+	& & & \\
+
+	\multirow{2}{*}{an\_prcp} & 0.358 & 0.612 & 1.412 \\
+	& (-0.26, 1.02) & (0.28, 0.94) & (1.11, 1.71) \\
+	& & & \\
+
+	\multirow{2}{*}{an\_prcp\textsuperscript{2}} & -0.571 & -0.848 & -0.975 \\
+	& (-0.76,-0.40) & (-0.95,-0.75) & (-1.06,-0.90) \\
+	& & & \\
+
+	\multirow{2}{*}{pToPET} & 2.960 & 2.637 & 2.093 \\
+	& (2.29, 3.61) & (2.28, 2.99) & (1.70, 2.48) \\
+	& & & \\
+
+	\multirow{2}{*}{pToPET\textsuperscript{2}} & -0.557 & -0.093 & 0.0064 \\
+	& (-0.74,-0.37) & (-0.15,-0.03) & (-0.070,-0.073) \\
+	\bottomrule
+	\end{tabular}
+	\label{tab:ex2_params}
+\end{table}
+
+%==================
+\begin{figure}[t]
+	\includegraphics[width=6.5in]{figs/ex2_response.pdf}
+	\caption{Response curves for each environmental variable for the three models.
+	Predictions are broadly similar for the three models, with an increase the optimal temperature regime predicted by the Integrated-Future model (left panel).
+	Integration reduced prediction uncertainty for all three variables, particularly for domains outside the naive calibration range.
+	Single-variable predictions were computed with the other variables set to their medians.}
+	\label{fig:ex2_response}
+\end{figure}
+%==================
+
+%==================
+\begin{figure}[t]
+\includegraphics[width=6.5in]{figs/ex2_pres_map.pdf}
+\caption{Range predictions for present climate for Phenofit (A), the naive model (B), the Integrated-Present model (C), and the difference between posterior prediction standard errors of the Integrated-Present and Naive models.
+Model predictions were quite similar, with reduced uncertainty where the models were in strongest agreement, and increased uncertainty near the range boundary where the models disagreed.
+For reference, the present range of the species is outlined in red \citep{Little1971}}.
+\label{fig:ex2_map_pres}
+\end{figure}
+
+%==================
+
+% results
+%% compare parameters
+Model integration resulted substantial reduction in posterior uncertainty in the parameters, and, for the Integrated-Future model, a relatively large revision in the the estimate for the response of sugar maple to temperature (i.e., variable ddeg) (Table \ref{tab:ex2_params}, Fig.
+\ref{fig:ex2_response}).
+When projecting beyond the calibration range for the naive model, the greater coverage provided by the integrated model produced substantial reductions in uncertainty (Fig. \ref{fig:ex2_response}).
+When considering the species distribution, the naive and Integrated-Present models made very similar predictions (Fig. \ref{fig:ex2_map_pres}).
+Furthermore, both models performed well when validating against reserved data, with AUC values of 0.72 and 0.73 for the naive and integrated models, respectively.
+The small difference between the models indicates that both models effectively predict the probability of presence of sugar maple.
+The major advantage to integration is the improved understanding of uncertainty in the predictions, with greater uncertainty in southern portions of the range (Fig. \ref{fig:ex2_map_pres}).
+It is important to note that the naive model was the basis for the integrated model; thus the increased uncertainty present in the integrated model is not the result of a ``worse'' model, but rather should be viewed as a correction to overfitting in the naive model.
+The uncertainty in the integrated model includes uncertainty arising from the processes that we have incorporated into the metamodel.
+Phenofit predicts fitness based on how climate affects phenological timings, frost injury, reproduction, and survival \citep{Chuine2001, Morin2009}.
+Thus, climatic factors that ultimately limit species distribution might be quite different under Phenofit, as illustrated by the substantial differences in the potential future distributions predicted by Phenofit and the naive model.
+
+\begin{figure}[t]
+\includegraphics[width=6.5in]{figs/ex2_fut_map.pdf}
+\caption{Range predictions for future climate for Phenofit (A), the naive model (B), the Integrated-Future model (C), and the difference between posterior prediction standard errors of the Integrated-Future and Naive models.
+The mechanistic sub-model Phenofit (A) predicted small shifts sugar maple range.
+In contrast, the naive model (B) projected a large northward change in suitable habitat.
+Model integration (C) produced predictions that were intermediate between the two models.
+Uncertainty decreased for the northern portion of the present range (red outline) where the models were in agreement, while it increased in the southern portion of the range where the models were in strong disagreement.}
+\label{fig:ex2_map_fut}
+\end{figure}
+
+The Integrated-Future model presents a different interpretation of the response of sugar maple to warmer temperatures. 
+The naive model predicted a substantial northward migration; in other words, the expectation under the naive model is that the present occurrence data is an accurate estimation of the niche, and thus the species should track its niche northward as the climate warms.
+The Integrated-Future model, in contrast, predicted substantially more tolerance to warmer temperatures, reflecting similar predictions from Phenofit (Figs.
+\ref{fig:ex2_response}, \ref{fig:ex2_map_fut}).
 Although both models predicted a northward range shift, the change under Phenofit was limited to approximately 200 km north of the present range limit of the species, compared with more than 900 km for the naive model.
-Phenofit also predicted little change in the southern range limit of the species, while the naive model projected a lack of suitability over the southern half of the present distribution.
-Model integration produced predictions that were intermediate between the extremes set by the individual models (Fig \ref{fig:ex2}C).
-The northern limit of the species was forecast to move by approximately 400--500 km, with reduced suitabilities but little extirpation throughout the southern portion of the present range.
-Uncertainty in the naive model was relatively low, particularly for the southern portion of the range where the model predicted extirpation (Fig \ref{fig:ex2}D).
-However, uncertainty in the predictions was locally extreme, particularly in portions of eastern Canada.
-Prediction uncertainty was generally increased for the integrated model relative to the naive model (Fig \ref{fig:ex2}E), likely due to the large amount of disagreement in the predictions of the two models.
-
-Although validation is an important tool for analyzing the performance of SDMs, it is not possible to validation predictions for the future due to a lack of data with which to compare model predictions.
-However, model integration in this case provides a greater degree of humility in interpreting the predictions; for example, the naive model predicted with a high degree of certainty that much of the southern portion of the present range of sugar maple will be unsuitable by 2080.
-The integrated model suggests that we should not be so confident in this interpretation.
-The naive model predicts suitability based on the values of selected climate variables (in this case, number of degree days, mean summer precipitation, and the ratio of annual precipitation to potential evapotranspiration).
-These variables were selected because they provided the best fit to the present distribution; however, this does not mean these variables are the most important climatic drivers of the processes generating the distribution, nor that they will continue to be important in the future.
-In contrast, Phenofit predicts fitness based on how climate affects phenological timings, frost injury, reproduction, and survival \citep{Chuine2001, Morin2009}.
-Thus, climatic factors that ultimately limit species distribution might be quite different under Phenofit, as illustrated by the substantial differences in the potential future distributions predicted by Phenofit and the naive model.
-In the absence of information about which model is ``best,'' the integrated model provides a consensus view of these two very different models by fitting parameters that maximize the joint likelihood of the presence-absence dataset and the Phenofit predictions.
-Furthermore, the integrated model provides robust estimates of uncertainty; in the southern portion of the range, the relatively high variance in predictions reflects the disagreement in the models and suggests that further exploration of limiting factors in that region may be useful in improving understanding and providing more precise predictions for changes to the distribution of sugar maple.
+Phenofit also predicted little change in the southern range limit of the species, while the naive model projected loss of the species over much of the southern portion of the range (Fig.
+\ref{fig:ex2_map_fut}).
 
@@
 \subsection*{Comparison with other methods}
-The methods provided here expand upon the motivation of hybrid models to develop a more robust and unifying approach for ecological models in a predictive context while overcoming some limitation characteristic of other integrated approaches.
+The methods provided here expand upon the motivation of hybrid models to develop a more robust approach for using ecological models for predictoin while overcoming some limitations characteristic of other integrated approaches.
 In particular, it is often difficult in a hybrid model to identify parameters that can be used to connect different modeling frameworks and produce a meaningful response \citep{Thuiller2013}. 
-The difficulty of including information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Thuiller2008, Smolik2010, Thuiller2014a}. 
-Even when a link can be formed to include this information, it often is oversimplified to facilitate computation \citep{Gallien2010}. 
-The methods presented in this study aim to overcome some of these limitations through the use of Bayesian statistics.
-Bayesian methods provide a natural framework for the incorporation of multiple sources of information.
-Because the metamodel is expressed as a series of conditional probabilities (see Appendix S1), this information can be included simply provided the probability of the metamodel, conditional on information from the sub-models, can be expressed mathematically.
-Furthermore, Bayesian methods produce results as an entire distribution instead of a single point, allowing for a comprehensive understanding of the uncertainty of estimates \citep{Link2010}. 
-The estimation of model parameters (and predictions) as distributions can also help decrease overparameterization that can be a concern for hybrid models. 
+The difficulty of including information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Smolik2010, Thuiller2014a}. 
+Bayesian methods such as ours are able to overcome some of these limitations, as they provide a natural framework for the incorporation of multiple sources of information.
+Because the metamodel is expressed as a series of conditional probabilities (see Appendix S1), this information can be included simply as long as the probability of the metamodel can be expressed mathematically.
+Furthermore, Bayesian methods produce posterior distributions of parameters and predictions, rather than point estimates, allowing for a comprehensive understanding of the uncertainty. 
 Finally, Bayesian methods inherently allow for feedbacks or interactions between sub models, which may be a more realistic representation of ecological dynamics where many factors may simultaneously influence the system.
 
-Our approach is an original application of hierarchical Bayesian modeling, and can be considered a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales while using several models simultaneously.
-In particular, this development has certain similarities with Bayesian model averaging and Bayesian calibration of process-based models. 
-Bayesian model averaging aims to combine several alternative models to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.,][]{Wintle2003, Link2006}.
-Bayesian model averaging considers models operating at the same scale and the posterior distribution is usually determined with Gibbs sampling. 
+Our approach is a logical extension of other Bayesian approaches developed to deal with processes that occur at multiple scales while using several models simultaneously. 
+In particular, it has certain similarities with Bayesian model averaging and Bayesian calibration of process-based models. 
+Bayesian model averaging aims to combine several alternative models that operate at the same scale to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.,][]{Wintle2003, Link2006}.
 Bayesian calibration of process-based models focuses on uncertainty of the parameter values, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, Hartig2012}.
-In contrast with Bayesian model averaging and calibration techniques, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of the metamodel.
-Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009, Efendiev2005}, and climate and atmospheric modeling \citep{Mcmillan2010, Kang2012}.
-Although our method as presented is distinct from Bayesian model averaging, it is fully compatible with these methods.
-Multiple metamodels could be averaged in this way if a consensus prediction from multiple metamodels (containing, e.g., different sub-model sets or different metamodel formulations).
-Furthermore, the metamodel can easily be incorporated into a broader multimodel inference scheme.
-
-%\fd{FD will work on this part: \\
-%It would be good to compare this approach with the kind of one develop by Moorcroft et al.
-2001 : Moorcroft, P.
-R., Hurtt, G.
-C., \& Pacala, S.
-W.
-(2001).
-A method for scaling vegetation dynamics: the ecosystem demography model (ED).
-Ecological monographs, 71(4), 557-586.\\
-%%Medvigy, D., Wofsy, S.
-C., Munger, J.
-W., Hollinger, D.
-Y., \& Moorcroft, P.
-R.
-(2009).
-Mechanistic scaling of ecosystem function and dynamics in space and time: Ecosystem Demography model version 2.
-Journal of Geophysical Research: Biogeosciences (2005–2012), 114(G1).\\
-%Fisher, R., McDowell, N., Purves, D., Moorcroft, P., Sitch, S., Cox, P., \& Ian Woodward, F.
-(2010).
-Assessing uncertainties in a second-generation dynamic vegetation model caused by ecological scale limitations.
-New Phytologist, 187(3), 666-681.
-%\\
-%These models are multilevel and use scaling functions using statisitical moments series.
-%}
+In contrast with these methods, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of the metamodel.
+%Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009}, and climate and atmospheric modeling \citep{Kang2012}.
+%Although our method as presented is distinct from Bayesian model averaging, it is fully compatible with these methods.
+%Multiple metamodels could be averaged in this way if a consensus prediction from multiple metamodels (containing, e.g., different sub-model sets or different metamodel formulations).
+%Furthermore, the metamodel can easily be incorporated into a broader multimodel inference scheme.
 
@@
 Species distribution models are important tools that are increasingly used by land managers to face challenges associated to decision-making \citep{Guisan2013}.
-A growing number of approaches are available, but they can provide diverging answers to very similar questions due to the differences in their assumptions and methodology. 
-This can create confusion and even some mistrust towards models, and some managers may be discouraged from incorporating model results in their management plans. 
+However, the potential for diverse approaches to provide diverging answers as a result of different assumptions and methodologies can create confusion and mistrust towards models, and some managers may be discouraged from incorporating model results in their management plans.
 Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organizations around the world \citep[e.g.,][]{Bernier2013}.
@@
 Such approaches are needed in formulating management plans for vulnerable species and ecosystems to avoid basing decisions on too-narrow subsets of the available information \citep{Dawson2011}.
-The flexibility of our approach may also represent an advantage for management by easing the integration of specific decision making criteria (e.g.
-desired grain and extent of outputs) into model development.
-Successful use of an integrated modeling approach will always remain dependent on an intimate understanding of the decision-making process by modelers, emphasizing the importance of close collaboration between modelers and practitioners at all stages of model development \citep{Guisan2013}.
+However, successful use of approaches such as ours will always remain dependent on an intimate understanding of the decision-making process by modelers, emphasizing the importance of close collaboration between modelers and practitioners at all stages of model development \citep{Guisan2013}.
 
 Model uncertainty is another key factor affecting applicability of model outputs \citep{Addison2013}.
-One of the main strengths of our approach is that it allows for a clear, transparent identification of uncertainties and how they are transmitted throughout the modeling process. 
-Transparency in uncertainty can be considered as a sort of sensitivity analysis, in which areas with large uncertainty can be detected and new experimental research or additional data collection can be designed accordingly (e.g., Example 1, Figs.
+One of the main strengths of our approach is that it allows for a transparent identification of uncertainties and how they propagate through the models. 
+Transparency in uncertainty can be considered as a sort of sensitivity analysis, whereby the greatest sources of uncertainty can be detected and further research directed accordingly (e.g., Example 1, Figs.
 \ref{fig:ex1_sampling}, \ref{fig:ex1_precip}).
-The new knowledge resulting from this research can be readily incorporated into the metamodel, allowing for an iterative learning process that will undoubtedly contribute to reduce uncertainty in predictions for a wide range of environmental conditions.
-Moreover, our framework embraces change as a fundamental process and is able to adapt and respond to it. 
-The ease of incorporating new knowledge to the modeling framework (including new theory or the result of management and experimental efforts) will allow for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans \citep{Keith2011}, and the integration of sub-models allows for clear specification of desired model outputs (via the specification of the metamodel) while easily retaining important ecological objectives (via careful specification of sub-models).
-Transparency in the model building process must be accompanied by a clearly documented modeling workflow; the transparency of our approach (and the utility of specifying specific objectives accompanied by appropriate sub-models) will assist in the development of transparent and well documented workflows describing the process of model integration and will ensure reproducibility and applicability (Fig.
+The new knowledge resulting from this research can then be readily incorporated into the metamodel and the model predictions updated to account for the new information. 
+The ease of incorporating new knowledge to the modeling framework allows for a rapid adjustment of the predictions and the incorporation of the most recent available knowledge into management plans \citep{Keith2011}.
+Furthermore, the use of linked sub-models allows for clear specification of desired model outputs (via the specification of the metamodel) while easily retaining important ecological objectives (via careful specification of sub-models).
+Transparency in the model building process must be accompanied by a clearly documented modeling workflow.
+We suggest using the sub-models as a natural proxy for specifying specific objectives, and using this as the basis for developing well documented workflows describing the process of model integration to ensure reproducibility and applicability (Fig.
 \ref{fig:management}).
-In an era of continuous and unprecedented change, adaptive approaches such as the one presented here are often highlighted as a pressing need in order to develop strategies to promote ecosystems that are both feasible and resilient \citep{Seastedt2008, Mori2013}.
+In an era of continuous and unprecedented change, adaptive approaches such as the one presented here are often highlighted as a pressing need in order to develop strategies to promote ecosystems that are both feasible and resilient \citep{Seastedt2008}.
+
+%==================
+% FIGURE
+\begin{figure}[t]
+	\includegraphics{figs/management.pdf}
+	\caption{Sample workflow for applying the models presented in the first example in a management context.
+	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
+	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the metamodel.}
+	\label{fig:management}
+\end{figure}
+%==================
 
@@
 Although our approach is highly flexible and can be applied in a number of situations, there are some challenges to successfully using the framework.
-As with any modeling effort, good model specification with strong links to theory are essential \citep{Austin2007}.
-Poorly specified models will produce outputs that are uninformative or misleading, and model integration is not a cure for these problems.
-In the worst case, constraining a metamodel with a poor sub-model can result in outcomes that are worse in terms of bias and uncertainty than those produced by a naive model.
-We expect model selection will play an important role in applications of our framework, and such schemes can be incorporated relatively easily \citep{Madigan1995, Wasserman2000, Tenan2014}.
-
-In a similar fashion, the quality and availability of data impose a significant constraint on the number and type of models that can be implemented.
-The capacity to implement a model is very low if data requirements are prohibitive. 
-Adequate and unbiased coverage of explanatory variables is a significant obstacle, and exploratory analyses can be a significant aid in understanding how data coverage impacts resulting predictions \citep{Mckenney2002}.
+Data quality and availability of data can present a significant constraint on the number and type of models that can be implemented in our framework.
+One obstacle is a lack of adequate and unbiased coverage of explanatory variables; exploratory analyses can be a significant aid in understanding how data coverage impacts resulting predictions \citep{Mckenney2002}.
 Integration can ameliorate these issues to some extent by using supplemental information (and conceptual advances) in additional sub-models where coverage is weak (e.g., Example 1, Figs.
 \ref{fig:ex1_precip}, \ref{fig:ex1_map}).
-For example, \citet{Freckleton2009} estimated that data are too limited to successfully develop highly mechanistic models predicting weed population numbers. 
 A strong asset of our approach is that it can be used without the full suite of data that would be required to run a fully mechanistic model. 
@@
 
-Finally, determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eqs. \ref{eq:ex1_integrated}, \ref{eq:integrated2}) is a critical point.
+Determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eq. \ref{eq:ex1_integrated}) is a critical point.
 In the context of integrated models, the challenge is three fold: (i) which spatial and temporal scales (i.e.
-which processes), are to be considered, (ii) how to build a scaling function that is consistent with the metamodel and (iii) how error and uncertainty propagate from the sub-models to the metamodel.
+which processes), are to be considered, (ii) how to build a scale the sub-models to be consistent with the metamodels, and (iii) how error and uncertainty propagate from the sub-models to the metamodel.
 Although we argue that our proposed framework is able to easily deal with very different scales and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered, the building of scaling functions deserves further investigation.
-We have shown in our examples that even simple scaling functions can provide reasonable constraints on the metamodel. 
-However, it is likely that the modeling process may include multiple scaling functions operating at different scales when taking all known processes and models into account. 
+It is likely that the modeling process may include multiple scaling functions operating at different scales when taking all known processes and models into account. 
 Indeed, if species distributions are a function of, e.g.
-population growth rate \citep{Sykes1996, Guisan2000}, they will involve processes at the individual (e.g.
+population growth rate \citep{Guisan2000}, they will involve processes at the individual (e.g.
 competition) or cellular (e.g.
@@
 
-In an applied context, the results of models predicting species ranges are increasingly used by resource managers, conservation biologists, and forest ecologists for formulating and adjusting recommendations.
-The utility of such models depends on their ability to help evaluate events beyond the bounds of the available data, including in future situations constrained by climate change. 
-Practitioners may judge model usefulness on two criteria: (1) Does the model provide reliable predictions at the needed time and space scales, and (2) can the model be implemented given the available data and technical expertise?
-Since most decision makers work at local space scales and follow specific time frames, modelers face the obvious challenge of providing detailed outputs while facilitating implementation.
-Our framework is transparent in reporting uncertainty of model output by virtue of providing easily interpretable probability distributions for model parameters and predictions.
-However, developing the models requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming.
+Finally, implementation of the model itself can present an obstacle.
+Developing the models requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming.
 In many cases, off-the-shelf software can adequately express the model likelihoods with minimal programming, but more complicated models will require the development of custom software.
 Thus, this framework will require significant investment in developing customized code for the samplers in order to actually estimate parameters, which may be technically challenging for some practitioners and may limit the adoption of these methods by both practitioners and decision-makers.
-However, the same was true of all statistical techniques or modeling approaches when they were initially developed, and many of the conceptual challenges associated with Bayesian statistics are being reduced as they gain recognition in the scientific literature.
+Furthermore, as with any modeling effort, good model specification with strong links to theory are essential \citep{Austin2007}.
+In the worst case, constraining a metamodel with a poor sub-model can result in outcomes that are worse in terms of bias and uncertainty than those produced by a naive model.
+We expect model selection will play an important role in applications of our framework, and such schemes can be incorporated relatively easily \citep{Wasserman2000, Tenan2014}.
+The flexibility of our approach, and its transparency with respect to the propagation of uncertainty, will often outweigh the implementation challenges.
 We therefore argue that our approach has a strong potential for direct use in our real world where climate is quickly changing and conservation practices must be adjusted accordingly.
-We further argue that forging stronger collaborations between modelers, decision-makers, and practitioners will improve the incorporation of this and other new methods into applied ecology and conservation.
+We further argue that forging stronger collaborations between modelers, decision-makers, and practitioners will overcome implementation challengers and promote the incorporation of this and other new methods into applied ecology and conservation.
+
+%In an applied context, the results of models predicting species ranges are increasingly used by resource managers, conservation biologists, and forest ecologists for formulating and adjusting recommendations.
+%The utility of such models depends on their ability to help evaluate events beyond the bounds of the available data, including in future situations constrained by climate change. 
+%Practitioners may judge model usefulness on two criteria: (1) Does the model provide reliable predictions at the needed time and space scales, and (2) can the model be implemented given the available data and technical expertise?
+%Since most decision makers work at local space scales and follow specific time frames, modelers face the obvious challenge of providing detailed outputs while facilitating implementation.
+%Our framework is transparent in reporting uncertainty of model output by virtue of providing easily interpretable probability distributions for model parameters and predictions.
+%However, developing the models requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming.
+%In many cases, off-the-shelf software can adequately express the model likelihoods with minimal programming, but more complicated models will require the development of custom software.
+%Thus, this framework will require significant investment in developing customized code for the samplers in order to actually estimate parameters, which may be technically challenging for some practitioners and may limit the adoption of these methods by both practitioners and decision-makers.
+%However, the same was true of all statistical techniques or modeling approaches when they were initially developed, and many of the conceptual challenges associated with Bayesian statistics are being reduced as they gain recognition in the scientific literature.
+%
 
@@
 
-\subsection*{Future directions}
-Hierarchical Bayesian modeling requires data at every scale to fit the parameters required to constrain one level with another.
-Moreover, beyond the needs for additional data, more knowledge of the important processes driving species range limits is needed to allow for extrapolation beyond the scope of the original datasets to address future changes due to climate and land-use changes.
-This knowledge can be obtained through a series of experimental manipulations and by substituting space for time data along environmental gradients.
-When experiments are possible, gaps in knowledge can be inferred using surrogate variables, theoretical models and spatially explicit models to explore the range of outcomes possible \citep{Fortin2012}.
-Another way to gain knowledge is to simplify species complexity into a series of states and to use state-transition models (e.g., Markov chain, semi-Markov chain, matrix projection model, integral projection model) which require more theoretical or coarser knowledge of species responses to biotic and abiotic changes.
-Although fewer states allow us to better capture the essence of species responses to environmental conditions, knowledge is still required to model feedback effects across processes and spatio-temporal scales.
-Hence we need to determine which processes are the most important at a given scale and how to weight their effects while scaling-up.
-Our metamodeling approach can act as a springboard for incorporating imperfect knowledge of how multi-scale processes influence species ranges.
-By incorporating this knowledge into species distribution models, we can hopefully both reduce bias and more accurately assess model uncertainty when forecasting future changes to species ranges.
+%\subsection*{Future directions}
+%Hierarchical Bayesian modeling requires data at every scale to fit the parameters required to constrain one level with another.
+%Moreover, beyond the needs for additional data, more knowledge of the important processes driving species range limits is needed to allow for extrapolation beyond the scope of the original datasets to address future changes due to climate and land-use changes.
+%This knowledge can be obtained through a series of experimental manipulations and by substituting space for time data along environmental gradients.
+%When experiments are possible, gaps in knowledge can be inferred using surrogate variables, theoretical models and spatially explicit models to explore the range of outcomes possible \citep{Fortin2012}.
+%Another way to gain knowledge is to simplify species complexity into a series of states and to use state-transition models (e.g., Markov chain, semi-Markov chain, matrix projection model, integral projection model) which require more theoretical or coarser knowledge of species responses to biotic and abiotic changes.
+%Although fewer states allow us to better capture the essence of species responses to environmental conditions, knowledge is still required to model feedback effects across processes and spatio-temporal scales.
+%Hence we need to determine which processes are the most important at a given scale and how to weight their effects while scaling-up.
+%Our metamodeling approach can act as a springboard for incorporating imperfect knowledge of how multi-scale processes influence species ranges.
+%By incorporating this knowledge into species distribution models, we can hopefully both reduce bias and more accurately assess model uncertainty when forecasting future changes to species ranges.
 %---------------------------------------------
@@
 \section*{Data Accessibility}
-All data, as well as all code required to repeat the analyses, have been uploaded as online supporting information in Appendix S4.
+All data, as well as all code required to repeat the analyses, have been uploaded as online supporting information in Appendix S2.
 
@@
 \section*{Supporting information}
-Appendix S1: Complete model description. \\
-Appendix S2: Supporting information on using the code and data for example 1. \\
-Appendix S3: Supporting information on using the code and data for example 2. \\
-Appendix S4: Scripts and data required to reproduce the analyses.
+Appendix S1: Complete model description and model fitting details for the examples. \\
+Appendix S2: Scripts and data required to reproduce the analyses.
 
@@
 
-%% ----------------------------------
-%
-%     FIGURES
-%
-%% ----------------------------------
-
-%TC:break figures
-\newpage
-\section*{Figure Legends}
-\begin{flushleft}
-
-\textbf{Figure 1:} The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
-	The metamodel ($\theta_M$) operates at a single scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
-	The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
-	The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
-	This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
-	This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).
-
-\textbf{Figure 2:} Two simulated datasets used to illustrate the model integration framework.
-	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
-	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., \(-\)1--1).
-	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
-	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.
-
-\textbf{Figure 3:} Results of integration showing the probability of presence (\(\psi\)) when considering only precipitation.
-	(a) Naive model, using only presence-absence data.
-Uncertainty increases dramatically when attempting to project beyond the scope of the source data (represented by the horizontal line above the bottom axis).
-	(b) Mechanistic model, using observations of an experiment to infer probability of presence.
-Predictions are highly precise due to high quality source data, but are likely to be overly optimistic given the simplicity in assuming only precipitation determines probability of presence (see text).
-	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
-	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.
-
-\textbf{Figure 4:} Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
-	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
-	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
-	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.
-
-\textbf{Figure 5:} Results of model integration for sugar maple.
-	All panels show information using predicted climate from 2080 under the HadA2 scenario (see main text).
-	For reference, the present range of the species is outlined in red \citep{Little1971}.
-	The mechanistic sub-model Phenofit (A) predicted fairly small shifts sugar maple range.
-	In contrast, the naive model (B) projected large northward shifts in suitable habitat.
-	Model integration (C) produced predictions that were intermediate between the two models.
-	Uncertainty in the naive model (D) was fairly low, except for some portions of the northeast where uncertainty was extreme.
-	Model integration produced uncertainty (E) that was globally somewhat higher reflecting the disagreement between the naive model (B) and the sub-model (A), but with fewer extremes in prediction uncertainty.
-
-\textbf{Figure 6:} Sample workflow for applying the models presented in the first example in a management context.
-	Critical steps include specifying the metamodel, identifying additional sources of information to be used as constraints on the metamodel, and using the integrated prediction for decision-making.
-	As additional information becomes available from monitoring the results of management, this information can be incorporated in additional sub-models to further refine the metamodel.
-
-\end{flushleft}
-
-%==================
-% FIGURE 1
-%TC:break fig 1
-\newpage
-\begin{figure}[h!]
-\input{diagram}
-\caption{}
-\label{fig:diagram}
-\end{figure}
-
-%==================
-% FIGURE 2
-%TC:break fig 2
-\newpage
-\begin{figure}[h!]
-	\includegraphics{ex1_sampling.pdf}
-	\caption{}
-	\label{fig:ex1_sampling}
-\end{figure}
-
-%==================
-% FIGURE 3
-%TC:break fig 3
-\newpage
-\begin{figure}[h!]
-	\includegraphics{ex1_precip.pdf}
-	\caption{}
-	\label{fig:ex1_precip}
-\end{figure}
-
-%==================
-% FIGURE 4
-%TC:break fig 4
-\newpage
-\begin{figure}[h!]
-	\includegraphics[width=5.25in]{ex1_map.pdf}
-	\caption{}
-	\label{fig:ex1_map}
-\end{figure}
-
-%==================
-% FIGURE 5
-%TC:break fig 5
-\newpage
-\begin{figure}[h!]
-	\includegraphics[width=6in]{ex2.pdf}
-	\caption{}
-	\label{fig:ex2}
-\end{figure}
-
-%==================
-% FIGURE 6
-%TC:break fig 6
-\newpage
-\begin{figure}[h!]
-	\includegraphics{management.pdf}
-	\caption{}
-	\label{fig:management}
-\end{figure}
-
 \end{document}

### 2015-03-24 tex/wordcount.sh

@@ -1,2 +1,4 @@
+word=`texcount -sub=none manuscript.tex | sed -n -E 's/[[:space:]]+([0-9]+).+abstract/\1/p'`
+sed -i backup -E 's/[0-9]+([[:space:]]+%ABS_COUNT)/'"${word}"'\1/g' manuscript.tex 
 word=`texcount -sub=none manuscript.tex | sed -n -E 's/[[:space:]]+([0-9]+).+manuscript/\1/p'`
 sed -i backup -E 's/[0-9]+([[:space:]]+%WORD_COUNT)/'"${word}"'\1/g' manuscript.tex 

### 2015-04-09 tex/manuscript.tex
Worked in most comments

@@
 %% ----------------------------------
-%\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
-%\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
-%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
+\usepackage[top=1in, left=1in, bottom=1in, right=2.0in]{geometry}
+\usepackage[textsize=tiny, backgroundcolor=white, textwidth=1.8in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
+\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
+\newcommand{\ib}[1]{\todo[color=red!40]{\textsuperscript{IB}#1}}
+\newcommand{\fd}[1]{\todo[color=orange!40]{\textsuperscript{FD}#1}}
 
@@
 \author[8]{C. Ronnie Drever (cdrever@TNC.ORG)}
-\author[9]{Marie-Josée Fortin (fortinmj@gmail.com)}
+\author[9]{Marie-Josée Fortin (mariejosee.fortin@utoronto.ca)}
 \author[1]{Tony Franceschini (Tony.Franceschini@uqar.ca)}
@@
 
-	\textbf{Abstract word count: } 258 %ABS_COUNT
+	\textbf{Abstract word count: } 235 %ABS_COUNT
 
-	\textbf{Main text word count: } 4641 %WORD_COUNT %Target 5000
+	\textbf{Main text word count: } 4848 %WORD_COUNT %Target 5000
 
-	\textbf{Authorship:} All authors contributed substantially to an initial outline of the paper and ideas.
+%	\textbf{Authorship:} All authors contributed substantially to an initial outline of the paper and ideas.
 MT, IB, and DG built the model and wrote the first draft of the introduction and model description.
@@
 
-	\textbf{Keywords:} Climate change, decision making, metamodeling, patterns and processes, range dynamics, scaling, spatial ecology, species distribution modeling, uncertainty
+	\textbf{Keywords:} Climate change, decision making, patterns and processes, range dynamics, scaling, spatial ecology, species distribution modeling, uncertainty
 	\end{flushleft}
@@
 Current interest in modeling potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
-However, in general, individual approaches include only a small subset of the available information about a species. 
-Correlative models often predict presence or absence as a function of climate, but ignore  smaller-scale processes such as growth, fecundity, and dispersal.
+However, in general, single approaches include only a small subset of the available information about a species. 
+Correlative models often predict presence or absence as a function of the environment, but ignore smaller-scale processes such as growth, fecundity, and dispersal.
 Furthermore, different approaches often produce divergent predictions, with no simple method to reconcile them.
@@
 Metamodel outputs also incorporate uncertainty from both data and process error from all sub-models, and include all of the information used as input for the original scale-specific sub-models.
-We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
+We illustrate our approach using theoretical and empirical examples, and demonstrate that the framework better characterizes uncertainty and can substantially reduce uncertainty when projecting beyond the range of the original data.
 We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
 The framework has wide potential for use in species distribution modelling, and can drive better integration of multi-scale data into ecological decision-making.
-Although implementation of our method can be technically challenging, we anticipate the results will be of wide interest, and we encourage collaboration between modelers and practitioners in applying our framework.
 
@@
 Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}.
-Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, hybrid zone dynamics \citep{Engler2013}, and forecasting the impacts of climate change on species distributions \citep{Blois2013, Thuiller2014}.
+Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, hybrid zone dynamics \citep{Engler2013}, and climate change impacts on species distributions \citep{Blois2013, Thuiller2014}.
 
 Despite the recognized potential of these models, it can be difficult to produce species distribution models (SDMs) with acceptable levels of precision and bias \citep{Guisan2013}.
-For mechanistic models, two important constraints can present difficulties: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
+For mechanistic models, two important constraints can cause difficulties: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
 In recent decades, however, modeling techniques have proliferated to take advantage of increased data availability. 
@@
 
-Clearly, an approach is needed that unifies the strength of the diversity of modeling approaches and can incorporate multiple data sources and modeling paradigms.
-Toward this end, we present here an application of hierarchical Bayesian methods that uses outputs from multiple model outputs to inform the results of the final model.
+Clearly, an approach is needed to unify the strength of different modeling approaches that can incorporate multiple data sources and modeling paradigms.
+Toward this end, we present here an application of hierarchical Bayesian methods that uses outputs from multiple models to inform the results of the final model.
 Techniques for multimodel inference have proliferated in recent years.
 For example, hybrid models that allow for flexible combinations between mechanistic and phenomenological sub-models are commonly employed in SDMs \citep{Gallien2010, Thuiller2013, Boulangeat2014}.
-Within the hybrid framework, a correlative model might be used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while a more mechanistic approach could include space and time dynamics and therefore account for dispersal \citep{Smolik2010}.
-However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and the sources and magnitudes of uncertainties are difficult to quantify.
+Within the hybrid framework, a correlative model might be used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while a more mechanistic approach could include biotic interactions and space-time dynamics through dispersal \citep{Smolik2010}.
+However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and uncertainties are approximated and their attribution to different sources is difficult to distinguish.
 A commonly-employed alternative to hybrid models is the direct combination of predictions, where predictions from models operating at the same spatio-temporal scales can be combined \citep[e.g., model averaging, ensemble forecasting;][]{Araujo2007}.
-However, the applicability of such techniques is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
+However, the applicability of such techniques is limited; for example, because uncertainty is approximated and may be poorly understood, it is not possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
 
-We propose an alternative to these other approaches to multimodel inference.
-We use the power and generality of a hierarchical Bayesian framework, which provides a number of desirable advantages, including (1) the ability to incorporate multiple modes of inference (e.g., mechanistic, correlative models) \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}, (2) an easy mechanism to include multiple data sources at various scales \citep{Levin1992, Peters2004}, and (3) intuitive and comprehensive reporting of uncertainty in model predictions that  reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}.
-Contrary to hybrid and ensemble methods, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for as much flexibility as possible regarding the type of information included.
+We propose an alternative to these other approaches using a hierarchical Bayesian framework. 
+This approach provides a number of desirable advantages, including (1) the ability to incorporate multiple modes of inference (e.g., mechanistic, correlative models) \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}, (2) an easy mechanism to include multiple data sources at various scales \citep{Levin1992, Peters2004}, and (3) an intuitive and comprehensive reporting of uncertainty in model predictions that  reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}.
+Contrary to hybrid methods, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for more flexibility regarding the type of information included.
 By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Guisan2005, Araujo2006}.
@@
 
-We illustrate our approach with two examples.
+We illustrate our framework with two examples.
 We begin with a hypothetical example, using simulated data, where we define the framework and demonstrate its application to multiple sources of information from different scales.
 In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread foundation tree species from eastern North America, in response to climate change.
-Both examples are presented with complete code and data (in Supporting Information).
-Finally, we provide a more formal mathematical presentation in Appendix S1.
+We provide a more formal mathematical presentation (Appendix S1) along with complete code and data for both examples (Appendix S2).
 
@@
 \section*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
-The key idea of our approach is formulate a metamodel that integrates data at the same ecological scale as the desired predictions, and to constrain the parameters of this model using the output of one or more sub-models
-In this hypothetical example, we build a metamodel relating the distribution (i.e.
-occurrence probability) of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
+The key idea of our approach is to formulate a metamodel that integrates data at the same ecological scale as the desired predictions, and to constrain the parameters of this model using the output of one or more sub-models.
+In this hypothetical example, we build a metamodel relating the distribution of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
 The metamodel attempts to capture the realized distribution of a species; as a correlative model, it implicitly captures the major physiological constraints and ecological processes constraining the distribution of the target species.
@@
 Because these data are often collected at a finer scale than that of rangewide occurence data, we apply a simple scaling function, drawing on ecological theory, to compute the likelihood of a set of metamodel parameters given both the occurrence and the physiological data.
-See Appendix S1 in Supporting Information for procedural details and scripts for executing the model.
+See Appendix S1 for procedural details and scripts for executing the model.
 
@@
 	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
-	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., \(-\)1--1).
+	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., -1.0--1.0).
 	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
@@
 We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation.
-We assume a presence-absence dataset, which we term \(X_M\), where the subscripted \(M\) indicates that the data were collected at the same scale as the metamodel.
-We simulated a relatively high-quality dataset that provides coverage of a variety of ecological conditions (Fig. \ref{fig:ex1_sampling}).
-We desire to model the species' distribution as a function of temperature, \(T_M\), and precipitation, \(P_M\).
+For the metamodel-scale data, we simulate a relatively high-quality presence-absence dataset covering a variety of ecological conditions, which we term \(X_M\), where the subscripted \(M\) indicates that the data were collected at the same scale as the metamodel (Fig.
+\ref{fig:ex1_sampling}).
+We desire to mo0del the species' distribution as a function of temperature, \(T_M\), and precipitation, \(P_M\).
 An initial version of the metamodel \((\theta_M)\) that has no constraints from other datasets will be referred to as the naive model.
@@
 %-----
-where \(theta_M\) is the parameter vector of the model, \(D_M\) is the covariate matrix (i.e., \(T_M, P_M\), and quadratic terms, with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
+where \(\theta_M\) is the parameter vector of the model, \(D_M\) is the covariate matrix (i.e., \(T_M, P_M\)), and \(\text{logit}^{-1}\) is the inverse of the logit function.
 We estimate parameters using a Markov Chain Monte Carlo (MCMC) method using the proportional form of Bayes' theorem:
@@
 This concept gives us a reasonable model to fit a scaling function for our sub-model (Appendix S1).
-If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, then for an observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
+If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed (where the \(S\) subscript indicates information pertaining to the submodel), then for an observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
 \begin{equation}
@@
 \end{equation}
-where \(N\) is the Normal density function.
+where \(N\) is the normal density function.
 We can then estimate the posterior distribution for the sub-model by fitting the relationship between \(\psi_S\) and precipitation \( \left( P_S \right) \) using Bayesian beta regression \citep{Ferrari2004}:
@@
 
-Although the two datasets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation).
-The scaling is quite simplistic; it treats the fundamental niche as the only driver of species distribution, and only considers a single dimension of the niche.
+Although the two datasets were collected at considerably different scales, we have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation).
+The scaling treats the fundamental niche as the only driver of species distribution, and only considers a single dimension of the niche.
 As such, it would be unwise to expect predictions from this model alone to resemble the actual distribution of the species; as a mechanistic model, it is simply too incomplete to predict distribution.
-As we will show, however, the information from this sub-model, when applied as a constraint on the metamodel, can result in improved predictions that incorporate the information within each model.
+However, the information from this sub-model, when applied as a constraint on the metamodel, can result in improved predictions that incorporate the information within each model.
 
@@
 However, these predictions now reflect the presence-absence data \(X_M\) as well as the information from \(\theta_S\), including all of the data sources used to produce this sub-model.
-In other words, with this evaluation, the set of parameters $\theta_M$ is most likely when it agrees with both the original data and predictions from the sub-model. 
 Finally, we note the presence of marginal distributions for both models (i.e., \(p(\theta_M)\) and \(p(\theta_S)\)).
@@
 	\caption{Comparison of the naive model, mechanistic sub-model, and the integrated model showing the probability of presence (\(\psi\)) as a function of precipitation.
+	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.
+	The shaded region shows the calibration range for the naive model.
 	(a) Naive model, using only presence-absence data. Uncertainty increases dramatically when attempting to project beyond the scope of the source data.
@@
 	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
-	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals. The shaded region shows the calibration range for the naive model.}
+	}
 	\label{fig:ex1_precip}
@@
 	\caption{Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
-	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
-	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
-	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.
+	Historical (i.e., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
 }
@@
 These data defined the metamodel scale.
-To these data, we added 4903 recorded presences and 21701 absences derived from permanent forest sample plots located in the United States and Canada.
-We reserved 1/3 of this dataset for validation, and used the remaining records to calibrate the models.
+To these data, we added 4903 recorded presences and 21701 absences derived from permanent forest sample plots located in the United States and Canada (See Appendix S1 for a map of plot locations).
+We reserved 1/3 of this dataset for evaluation, and used the remaining records to calibrate the models.
 We constructed the naive model by using a binomial generalized linear model (GLM) to relate the presence-absence dataset to three climate variables: the number of degree days (ddeg), mean annual precipitation (an\_prcp), and the ratio of annual precipitation to potential evapotranspiration (pToPET).
@@
 We selected a GLM for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., generalized additive models) are compatible with the framework.
-The parameters for the model were estimated using a Metropolis-Hastings algorithm within Markov Chain Monte Carlo (MCMC).
+The parameters for the model were estimated using a Metropolis-Hastings algorithm.
 
@@
 Thus, for our first integration we combined the naive model with the Phenofit predictions for the present; we refer to this model as ``Integrated-Present'' hereafter.
-We also validated this model and compared it against the naive model by computing the area under the receiver operating curve (AUC), a commonly-used metric that measures the rate of correct classification and varies from 0 to 1 (where a value of 1 indicates perfect classification) \citep{Swets1988}.
+We also validated this model and compared it against the naive model by computing the area under the receiver operating curve (AUC), a commonly-used metric that measures the rate of correct classification and varies from 0 to 1 (where a value of 1 indicates perfect classification and 0.5 indicates no difference from a random model) \citep{Swets1988}.
 The second modeling goal is projecting changes to the range of sugar maple following climate change.
 Process-based models and correlative models often differ substantially when projecting beyond the range of the original data (see Example 1).
-Thus, we obtained predictions for 2100 from Phenofit \citep{Morin2009} and used the predictions to condition the metamodel given the predicitons of the naive model under future climate.
+Thus, we obtained predictions for 2100 from Phenofit \citep{Morin2009} and used them to condition the metamodel given the results of the naive model under future climate.
 This procedure provided a consensus view of the future range of the species; we refer to this model as ``Integrated-Future.''
@@
 %% compare parameters
-Model integration resulted substantial reduction in posterior uncertainty in the parameters, and, for the Integrated-Future model, a relatively large revision in the the estimate for the response of sugar maple to temperature (i.e., variable ddeg) (Table \ref{tab:ex2_params}, Fig.
+Model integration resulted in substantial reduction in posterior uncertainty in the parameters, and, for the Integrated-Future model, a relatively large revision in the the estimate for the response of sugar maple to temperature (i.e., variable ddeg) (Table \ref{tab:ex2_params}, Fig.
 \ref{fig:ex2_response}).
 When projecting beyond the calibration range for the naive model, the greater coverage provided by the integrated model produced substantial reductions in uncertainty (Fig. \ref{fig:ex2_response}).
-When considering the species distribution, the naive and Integrated-Present models made very similar predictions (Fig. \ref{fig:ex2_map_pres}).
-Furthermore, both models performed well when validating against reserved data, with AUC values of 0.72 and 0.73 for the naive and integrated models, respectively.
+When considering the present species distribution, the naive and Integrated-Present models made very similar predictions (Fig. \ref{fig:ex2_map_pres}).
+Furthermore, both models performed relatively well when evaluated against reserved data, with AUC values of 0.72 and 0.73 for the naive and integrated models, respectively.
 The small difference between the models indicates that both models effectively predict the probability of presence of sugar maple.
@@
 \subsection*{Comparison with other methods}
-The methods provided here expand upon the motivation of hybrid models to develop a more robust approach for using ecological models for predictoin while overcoming some limitations characteristic of other integrated approaches.
+The methods provided here expand upon the motivation of hybrid models to develop a more robust approach for using ecological models for prediction while overcoming some limitations characteristic of other integrated approaches.
 In particular, it is often difficult in a hybrid model to identify parameters that can be used to connect different modeling frameworks and produce a meaningful response \citep{Thuiller2013}. 
@@
 In contrast with these methods, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of the metamodel.
-%Similar hierarchical approaches have been applied to diverse fields, including engineering \citep{Booth2013}, hydraulic conductivity models \citep{Dostert2009}, and climate and atmospheric modeling \citep{Kang2012}.
-%Although our method as presented is distinct from Bayesian model averaging, it is fully compatible with these methods.
-%Multiple metamodels could be averaged in this way if a consensus prediction from multiple metamodels (containing, e.g., different sub-model sets or different metamodel formulations).
-%Furthermore, the metamodel can easily be incorporated into a broader multimodel inference scheme.
 
@@
 \subsection*{Advantages of Model Integration}
-Species distribution models are important tools that are increasingly used by land managers to face challenges associated to decision-making \citep{Guisan2013}.
-However, the potential for diverse approaches to provide diverging answers as a result of different assumptions and methodologies can create confusion and mistrust towards models, and some managers may be discouraged from incorporating model results in their management plans.
+Species distribution models are important tools that are increasingly used by land managers for science-based decision-making \citep{Guisan2013}.
+However, the possibility that diverse approaches will provide diverging answers as a result of different assumptions and methodologies can create confusion and mistrust towards models, and some managers may be discouraged from incorporating model results in their management plans.
 Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organizations around the world \citep[e.g.,][]{Bernier2013}.
@@
 \ref{fig:management}).
-In an era of continuous and unprecedented change, adaptive approaches such as the one presented here are often highlighted as a pressing need in order to develop strategies to promote ecosystems that are both feasible and resilient \citep{Seastedt2008}.
+Adaptive approaches such as the one presented here are often highlighted as a pressing need in order to develop strategies to promote ecosystems that are both feasible and resilient \citep{Seastedt2008}.
 
@@
 
-In many cases, both the data and theory needed to apply our method already exist, and all that is needed is the development of sub-models and the links between the sub-models and a metamodel.
+In many cases, both the data and theory needed to apply our approach already exist, and all that is needed is the development of sub-models and the links between the sub-models and a metamodel.
 For example, climatic gradients may mediate competitive interactions \citep{Kunstler2011}, which means that simple correlative models that fail to account for competition may be wrong if the climate-competition association changes in the future.
@@
 
+\ib{you say in the intro that the bayesian framework allows a easy and simple scaling-up, but the scale-up is still a problem.
+This part about scale is still a bit unclear.
+Our approach do not solve the scaling-up problem.}
+\fd{I fully agree with Isabellle.
+Adequacy between ecological processes, scales, and meaning for upscaling the information  remain a major challenge.
+It would be important here to say how this new approach helps is disentangling the difficulties listed at lines 35-37 in the introduction.}
+\mt{Although scaling is always a problem, this doesn't invalidate the statement that our method makes it easier.
+Think of the first example.
+This scaling would have been very difficult to do without using Bayesian methods.}
 Determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eq. \ref{eq:ex1_integrated}) is a critical point.
@@
 photosynthesis) scales.
-Such very large differences in spatial scales imply that simple functions will be inadequate, requiring more traditional upscaling methods. 
-Therefore, intermediate models and processes will need to be introduced, and this will require additional parameters, assumptions, and data, and will add uncertainty to the final integrated model.
+Such very large differences in spatial scales imply that simple functions such as those we have used here will be inadequate, requiring more traditional upscaling methods.
 
+\ib{This paragraph is weired. You say that the implementation is very difficult and end the paragraph with decision, practitioners advantage.
+I would end the paragraph with positive points about implementation, saying that we still can do a lot with JAGS for instance.
+And then start a paragraph to briefly talk about challenges in dialogue with practitionners}
 Finally, implementation of the model itself can present an obstacle.
@@
 Thus, this framework will require significant investment in developing customized code for the samplers in order to actually estimate parameters, which may be technically challenging for some practitioners and may limit the adoption of these methods by both practitioners and decision-makers.
+\fd{I am a bit feeling uncomfortable with this section.
+Practionners and decision-makers don't have to use (or even fully understand (as they will never get to that level of complexity) what methods underly the predictions) the models.
+They use the information known as being the best of the actual scientific knowledge.
+As for climate-related variables for CC, the scientific community (as the GIEC for the CC predictions) provide the predictions on SD.
+The scientific community is responsible for indicating the limitations of the use of the generated information.}
 Furthermore, as with any modeling effort, good model specification with strong links to theory are essential \citep{Austin2007}.
 In the worst case, constraining a metamodel with a poor sub-model can result in outcomes that are worse in terms of bias and uncertainty than those produced by a naive model.
-We expect model selection will play an important role in applications of our framework, and such schemes can be incorporated relatively easily \citep{Wasserman2000, Tenan2014}.
 The flexibility of our approach, and its transparency with respect to the propagation of uncertainty, will often outweigh the implementation challenges.
@@
 We further argue that forging stronger collaborations between modelers, decision-makers, and practitioners will overcome implementation challengers and promote the incorporation of this and other new methods into applied ecology and conservation.
-
-%In an applied context, the results of models predicting species ranges are increasingly used by resource managers, conservation biologists, and forest ecologists for formulating and adjusting recommendations.
-%The utility of such models depends on their ability to help evaluate events beyond the bounds of the available data, including in future situations constrained by climate change. 
-%Practitioners may judge model usefulness on two criteria: (1) Does the model provide reliable predictions at the needed time and space scales, and (2) can the model be implemented given the available data and technical expertise?
-%Since most decision makers work at local space scales and follow specific time frames, modelers face the obvious challenge of providing detailed outputs while facilitating implementation.
-%Our framework is transparent in reporting uncertainty of model output by virtue of providing easily interpretable probability distributions for model parameters and predictions.
-%However, developing the models requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming.
-%In many cases, off-the-shelf software can adequately express the model likelihoods with minimal programming, but more complicated models will require the development of custom software.
-%Thus, this framework will require significant investment in developing customized code for the samplers in order to actually estimate parameters, which may be technically challenging for some practitioners and may limit the adoption of these methods by both practitioners and decision-makers.
-%However, the same was true of all statistical techniques or modeling approaches when they were initially developed, and many of the conceptual challenges associated with Bayesian statistics are being reduced as they gain recognition in the scientific literature.
-%
-
-%----------------------------
-%----------------------------
-%----------------------------
-
-%\subsection*{Future directions}
-%Hierarchical Bayesian modeling requires data at every scale to fit the parameters required to constrain one level with another.
-%Moreover, beyond the needs for additional data, more knowledge of the important processes driving species range limits is needed to allow for extrapolation beyond the scope of the original datasets to address future changes due to climate and land-use changes.
-%This knowledge can be obtained through a series of experimental manipulations and by substituting space for time data along environmental gradients.
-%When experiments are possible, gaps in knowledge can be inferred using surrogate variables, theoretical models and spatially explicit models to explore the range of outcomes possible \citep{Fortin2012}.
-%Another way to gain knowledge is to simplify species complexity into a series of states and to use state-transition models (e.g., Markov chain, semi-Markov chain, matrix projection model, integral projection model) which require more theoretical or coarser knowledge of species responses to biotic and abiotic changes.
-%Although fewer states allow us to better capture the essence of species responses to environmental conditions, knowledge is still required to model feedback effects across processes and spatio-temporal scales.
-%Hence we need to determine which processes are the most important at a given scale and how to weight their effects while scaling-up.
-%Our metamodeling approach can act as a springboard for incorporating imperfect knowledge of how multi-scale processes influence species ranges.
-%By incorporating this knowledge into species distribution models, we can hopefully both reduce bias and more accurately assess model uncertainty when forecasting future changes to species ranges.
-%---------------------------------------------
-%---------------------------------------------
+\ib{I would add somewhere that this approach is not just a new methodological tool but is a framework for forecasting species distributions which is fundamentally designed to make the link between modellers and pratitionners, with (1) correctly estimating uncertainties and (2) being adaptable/updatable to new data or theoretical advances.}
 

### 2015-04-09 tex/manuscript.tex
isa work on discussion

@@
 photosynthesis) scales.
-Such very large differences in spatial scales imply that simple functions such as those we have used here will be inadequate, requiring more traditional upscaling methods.
+Such very large differences in spatial scales would require more sophisticated upscaling methods than simple functions such as those we have used here.
+Our framework is still applicable whatever the chosen upscaling approach and is able to propagate uncertainties from sub-models to the metamodel.
 
@@
 And then start a paragraph to briefly talk about challenges in dialogue with practitionners}
-Finally, implementation of the model itself can present an obstacle.
-Developing the models requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming.
-In many cases, off-the-shelf software can adequately express the model likelihoods with minimal programming, but more complicated models will require the development of custom software.
-Thus, this framework will require significant investment in developing customized code for the samplers in order to actually estimate parameters, which may be technically challenging for some practitioners and may limit the adoption of these methods by both practitioners and decision-makers.
+
+The implementation of the model itself can present an obstacle when model complexity increases.
+In many cases, off-the-shelf software can adequately express the model likelihoods with minimal programming, but more complicated models will require the development of custom programs.
+Developing such customized code would requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming.
+However, the flexibility of our approach, and its transparency with respect to the propagation of uncertainty, will often outweigh the implementation challenges.
+
+Finally, this approach is not just a new methodological tool but is a framework for forecasting species distributions which is fundamentally designed to make the link between modellers and pratitionners, with (1) correctly estimating uncertainties and (2) being updatable to new data or theoretical advances.
+A scientific approach such as presented here is particularly adapted to make a synthesis of available knowledge and provide robust species distribution forecasts based on information known as being the best of the actual scientific knowledge.
+It is actually able both to use large databases, valuing the efforts of data collection, and to include models based on the latest theoretical advances, which is essential to decrease the errors due to model specification \citep{Austin2007}.
+In addition, it would provides practionners and decision-makers with the best possible estimation of uncertainties, with direct applications to risk assessment or to guide the choice where investigating new research and accumulating new data.
+We further argue that the adaptability of our approach is particularly appropriate in our world where collected data and theoretical knowledge is changing as quickly as climate and conservation practices must be adjusted accordingly.
+
 \fd{I am a bit feeling uncomfortable with this section.
@@
 The scientific community is responsible for indicating the limitations of the use of the generated information.}
-Furthermore, as with any modeling effort, good model specification with strong links to theory are essential \citep{Austin2007}.
-In the worst case, constraining a metamodel with a poor sub-model can result in outcomes that are worse in terms of bias and uncertainty than those produced by a naive model.
-The flexibility of our approach, and its transparency with respect to the propagation of uncertainty, will often outweigh the implementation challenges.
-We therefore argue that our approach has a strong potential for direct use in our real world where climate is quickly changing and conservation practices must be adjusted accordingly.
-We further argue that forging stronger collaborations between modelers, decision-makers, and practitioners will overcome implementation challengers and promote the incorporation of this and other new methods into applied ecology and conservation.
+
 \ib{I would add somewhere that this approach is not just a new methodological tool but is a framework for forecasting species distributions which is fundamentally designed to make the link between modellers and pratitionners, with (1) correctly estimating uncertainties and (2) being adaptable/updatable to new data or theoretical advances.}

### 2015-04-13 tex/manuscript.tex
Minor edits from DG

@@
 \author[1,2]{Alyssa Butler (ca.butler10@gmail.com)} 
-\author[6,7]{Frédérik Doyon (Frederik.Doyon@uqo.ca)}
-\author[8]{C. Ronnie Drever (cdrever@TNC.ORG)}
-\author[9]{Marie-Josée Fortin (mariejosee.fortin@utoronto.ca)}
-\author[1]{Tony Franceschini (Tony.Franceschini@uqar.ca)}
-\author[10]{Jean Liénard (jean.lienard@gmail.com)}
-\author[4]{Dan McKenney (Dan.McKenney@nrcan-rncan.gc.ca)}
-\author[2,3]{Kevin A. Solarik (kevinsolarik@hotmail.com)}
-\author[10]{Nikolay Strigul (nick.strigul@vancouver.wsu.edu)}
-\author[11, 12]{Wilfried Thuiller (wilfried.thuiller@ujf-grenoble.fr)}
-\author[1,2]{Dominique Gravel (dominique\_gravel@uqar.ca)}
-\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
-\affil[2]{Quebec Centre for Biodiversity Science, Montreal, Quebec, Canada}
-\affil[3]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
+\dg{I think we should drop her, now that she is not in science anymore}
 \affil[4]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
@@
 
-Current interest in modeling potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
+Current interests in forecasting potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
 However, in general, single approaches include only a small subset of the available information about a species. 
@@
 Metamodel outputs also incorporate uncertainty from both data and process error from all sub-models, and include all of the information used as input for the original scale-specific sub-models.
-We illustrate our approach using theoretical and empirical examples, and demonstrate that the framework better characterizes uncertainty and can substantially reduce uncertainty when projecting beyond the range of the original data.
+We illustrate our approach using theoretical and empirical examples, and demonstrate that the framework better characterizes uncertainty and can substantially reduce it when projecting beyond the range of the original data.
 We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
@@
 
-Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
+Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-support tools in biodiversity management \citep{Guisan2013}.
 Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}.
@@
 For instance, biotic interactions are usually not modeled mechanistically at the scale of species distributions because they are poorly known or have not been recorded, despite being considered a key determinant of range limits \citep{Pigot2013}.
-In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
-However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}.
+In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to associated distributions of other variables have the advantage of indirectly accounting for all underlying processes \citep{Guisan2000}.
+However, their predictions rely on the stationarity of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}.
 Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
@@
 Within the hybrid framework, a correlative model might be used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while a more mechanistic approach could include biotic interactions and space-time dynamics through dispersal \citep{Smolik2010}.
-However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and uncertainties are approximated and their attribution to different sources is difficult to distinguish.
+However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, uncertainties are approximated and their attribution to different sources is difficult to distinguish.
 A commonly-employed alternative to hybrid models is the direct combination of predictions, where predictions from models operating at the same spatio-temporal scales can be combined \citep[e.g., model averaging, ensemble forecasting;][]{Araujo2007}.
@@
 This approach provides a number of desirable advantages, including (1) the ability to incorporate multiple modes of inference (e.g., mechanistic, correlative models) \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}, (2) an easy mechanism to include multiple data sources at various scales \citep{Levin1992, Peters2004}, and (3) an intuitive and comprehensive reporting of uncertainty in model predictions that reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}.
-Contrary to hybrid methods, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for more flexibility regarding the type of information included.
+Contrary to hybrid methods, the aim is not to link different sub-models into a single one, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for more flexibility regarding the type of information included.
 By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Guisan2005, Araujo2006}.
@@
 We begin with a hypothetical example, using simulated data, where we define the framework and demonstrate its application to multiple sources of information from different scales.
-In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread foundation tree species from eastern North America, in response to climate change.
-We provide a more formal mathematical presentation (Appendix S1) along with complete code and data for both examples (Appendix S2).
+In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread and dominant tree species from eastern North America, in response to climate change.
+We provide a more formal mathematical presentation in the Appendix S1, along with complete code and data for both examples (Appendix S2).
 
@@
 \ref{fig:ex1_sampling}).
-We desire to mo0del the species' distribution as a function of temperature, \(T_M\), and precipitation, \(P_M\).
+We desire to model the species' distribution as a function of temperature, \(T_M\), and precipitation, \(P_M\).
 An initial version of the metamodel \((\theta_M)\) that has no constraints from other datasets will be referred to as the naive model.
-This naive model uses a simple Bayesian logistic regression to estimate the naive probability of occurrence \((\psi_N)\) as a function of second order effects of temperature and precipitation:
+This naive model uses a simple logistic regression to estimate the naive probability of occurrence \((\psi_N)\) as a function of second order effects of temperature and precipitation:
+\dg{I dropped the 'bayesian' from this sentence because I was looking at the posterior in the following equation and only found the likelihood. It comes in the following equation.}
+
 %-----
@@
 We estimate parameters using a Markov Chain Monte Carlo (MCMC) method using the proportional form of Bayes' theorem:
+\dg{you specify the algorithm of the sampler below, but not here}
+
 %-----
@@
 
+\dg{There is something wrong with the following equation, or I missed the reasonning somewhere in the last year.
+In any case, I tried to understand the logic from the text and haven't succeed at it.
+Either it needs to be fixed, or it needs to be better explained.
+If I don't get it, my prior tells me nobody will.
+This equation is central to the paper now that the maths have been put in the appendix.
+I was expecting to see the likelihood of X_M given the parameters theta_M.
+The prior of theta_M being the posterior of the model theta_M applied to the data psi_S.
+The actual likelihood gives the probability of observing the prediction of the sub-model given the parameters of the sub-model.
+There is not integration on the conditions, it only results from the product of the probabilities.}
+
 We accomplish model integration by treating the posterior of \(\theta_S\) as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
@@
 For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modeling process (due to the lack of additional mechanisms in the second model, as mentioned above).
+
+\dg{This is a vague statement. Could you give just an extra sentence specifying how you did it, and then pointing to the Appendix?}
 
@@
 %==================
+
+\dg{might worth increasing font size on the figures. And make sure it is the same for all figures.}
 
@@
 Using both datasets to develop a range model for the species has a number of advantages.
-Assuming we have chosen climate variables that well represent the constraints on the species, including Phenofit in our model is likely to reduce bias in our estimate of the bioclimatic niche.
-Second, the posterior predictions of the model (that is, the probability of presence projected into geographic space) will incorporate uncertainty from all sources, including the modeling process, data, parameterization, and the ecological processes included.
+Assuming we have chosen climate variables that well represent the constraints on the species, including Phenofit in our model is likely to reduce bias in our estimate of the fundamental niche.
+The posterior predictions of the model (that is, the probability of presence projected into geographic space) will incorporate uncertainty from all sources, including the modeling process, data, parameterization, and the ecological processes included.
 This can provide a much more accurate estimate of the uncertainty of our predictions.
@@
 The difficulty of including information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Smolik2010, Thuiller2014a}. 
-Bayesian methods such as ours are able to overcome some of these limitations, as they provide a natural framework for the incorporation of multiple sources of information.
+Bayesian methods such as our are able to overcome some of these limitations, as they provide a natural framework for the incorporation of multiple sources of information.
 Because the metamodel is expressed as a series of conditional probabilities (see Appendix S1), this information can be included simply as long as the probability of the metamodel can be expressed mathematically.
@@
 Species distribution models are important tools that are increasingly used by land managers for science-based decision-making \citep{Guisan2013}.
-However, the possibility that diverse approaches will provide diverging answers as a result of different assumptions and methodologies can create confusion and mistrust towards models, and some managers may be discouraged from incorporating model results in their management plans.
+However, the possibility that diverse approaches will provide diverging answers as a result of different assumptions and methodologies can create confusion and mistrust towards models, and some managers may be discouraged from incorporating their results in their management plans.
 Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organizations around the world \citep[e.g.,][]{Bernier2013}.
@@
 Such approaches are needed in formulating management plans for vulnerable species and ecosystems to avoid basing decisions on too-narrow subsets of the available information \citep{Dawson2011}.
-However, successful use of approaches such as ours will always remain dependent on an intimate understanding of the decision-making process by modelers, emphasizing the importance of close collaboration between modelers and practitioners at all stages of model development \citep{Guisan2013}.
+However, successful use of approaches such as ours will always remain dependent on an intimate understanding of the decision-making process by modelers, emphasizing the importance of close collaboration between them and practitioners at all stages of model development \citep{Guisan2013}.
 
@@
 
-In many cases, both the data and theory needed to apply our approach already exist, and all that is needed is the development of sub-models and the links between the sub-models and a metamodel.
+In many cases, both the data and theory needed to apply our approach already exist, and all that is needed is the development of sub-models and their integration into a metamodel.
 For example, climatic gradients may mediate competitive interactions \citep{Kunstler2011}, which means that simple correlative models that fail to account for competition may be wrong if the climate-competition association changes in the future.
@@
 Although our approach is highly flexible and can be applied in a number of situations, there are some challenges to successfully using the framework.
-Data quality and availability of data can present a significant constraint on the number and type of models that can be implemented in our framework.
+Data quality and availability can present a significant constraint on the number and type of models that can be implemented in our framework.
 One obstacle is a lack of adequate and unbiased coverage of explanatory variables; exploratory analyses can be a significant aid in understanding how data coverage impacts resulting predictions \citep{Mckenney2002}.
-Integration can ameliorate these issues to some extent by using supplemental information (and conceptual advances) in additional sub-models where coverage is weak (e.g., Example 1, Figs.
+Integration can solve these issues to some extent by using supplemental information (and conceptual advances) in additional sub-models where coverage is weak (e.g., Example 1, Figs.
 \ref{fig:ex1_precip}, \ref{fig:ex1_map}).
@@
 This scaling would have been very difficult to do without using Bayesian methods.}
+\dg{Do agree with mt}
+
 Determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eq. \ref{eq:ex1_integrated}) is a critical point.
-In the context of integrated models, the challenge is three fold: (i) which spatial and temporal scales (i.e.
+The challenge is three fold: (i) which spatial and temporal scales (i.e.
 which processes), are to be considered, (ii) how to build a scale the sub-models to be consistent with the metamodels, and (iii) how error and uncertainty propagate from the sub-models to the metamodel.
-Although we argue that our proposed framework is able to easily deal with very different scales and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered, the building of scaling functions deserves further investigation.
-It is likely that the modeling process may include multiple scaling functions operating at different scales when taking all known processes and models into account. 
+Although we argue that our proposed framework is able to easily deal with very different scales and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered, the building of scaling functions is an object of investigation on its own.
+It is likely that the modeling process may include multiple functions operating at different scales when taking all known processes and models into account. 
 Indeed, if species distributions are a function of, e.g.
@@
 Our framework is still applicable whatever the chosen upscaling approach and is able to propagate uncertainties from sub-models to the metamodel.
+
+\dg{I also think that there is the challenge inherent to complex systems that is overlooked to a certain extent here.
+Small scale processes might generate organization at higher scales.
+Model integration might be limited by this problem in some cases, while in others it might solve it by putting bounds on the higher-order organization.
+Think about the evaluation of model parameters with observations at the scale of the operating process.
+There are some parameter values, within the posterior distribution, that could result in non-desirable properties at the larger scale.
+Model integration could buffer this propagation.}
 
@@
 \section*{Biosketch}
-The Theoretical Ecosystem Ecology lab group at the Université du Québec à Rimouski focuses on interactions between species distributions, community structure and ecosystem functioning.
+The Canada Research Chair in Biogeography and metacommunity ecology at the Université du Québec à Rimouski focuses on interactions between species distributions, community structure and ecosystem functioning.
 We apply principles of spatial ecology to a variety of organisms and systems, from bacteria to entire forests.

### 2015-07-27 tex/manuscript.tex
Commit changes for first revision to GEB (submitted in Apr 2015

@@
 \usepackage[utf8]{inputenc}
+\usepackage{fixltx2e}
 
@@
 \usepackage{graphicx}
-%\usepackage[nomarkers,nolists,heads]{endfloat}
+\usepackage[nomarkers, tablesfirst]{endfloat} % for final
 \usepackage{caption}
-\captionsetup{labelformat=simple}
+\captionsetup{labelsep=none,textformat=empty} % for final
+% \captionsetup{labelformat=simple} % for drafts
 \usepackage{booktabs}
@@
 %% ----------------------------------
-\usepackage[top=1in, left=1in, bottom=1in, right=2.0in]{geometry}
-\usepackage[textsize=tiny, backgroundcolor=white, textwidth=1.8in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
-\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
-\newcommand{\ib}[1]{\todo[color=red!40]{\textsuperscript{IB}#1}}
-\newcommand{\fd}[1]{\todo[color=orange!40]{\textsuperscript{FD}#1}}
+%\usepackage[top=1in, left=1in, bottom=1in, right=2.0in]{geometry}
+%\usepackage[textsize=tiny, backgroundcolor=white, textwidth=1.8in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
+%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
+%\newcommand{\ib}[1]{\todo[color=red!40]{\textsuperscript{IB}#1}}
+%\newcommand{\fd}[1]{\todo[color=orange!40]{\textsuperscript{FD}#1}}
+%\newcommand{\dg}[1]{\todo[color=green!40]{\textsuperscript{DG}#1}}
 
@@
 \author[1,2]{Alyssa Butler (ca.butler10@gmail.com)} 
-\dg{I think we should drop her, now that she is not in science anymore}
+\author[6,7]{Frédérik Doyon (Frederik.Doyon@uqo.ca)}
+\author[8]{C. Ronnie Drever (cdrever@tnc.org)}
+\author[9]{Marie-Josée Fortin (mariejosee.fortin@utoronto.ca)}
+\author[1]{Tony Franceschini (Tony.Franceschini@uqar.ca)}
+\author[10]{Jean Liénard (jean.lienard@gmail.com)}
+\author[4]{Dan McKenney (Dan.McKenney@nrcan-rncan.gc.ca)}
+\author[2,3]{Kevin A. Solarik (kevinsolarik@hotmail.com)}
+\author[10]{Nikolay Strigul (nick.strigul@vancouver.wsu.edu)}
+\author[11,12]{Wilfried Thuiller (wilfried.thuiller@ujf-grenoble.fr)}
+\author[1,2]{Dominique Gravel (dominique\_gravel@uqar.ca)}
+\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
+\affil[2]{Quebec Centre for Biodiversity Science, Montreal, Quebec, Canada}
+\affil[3]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
 \affil[4]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
@@
 
-	\textbf{Main text word count: } 4848 %WORD_COUNT %Target 5000
+	\textbf{Main text word count: } 4595 %WORD_COUNT %Target 5000
 
@@
 
-Current interests in forecasting potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
+Current interest in forecasting potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
 However, in general, single approaches include only a small subset of the available information about a species. 
@@
 In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread and dominant tree species from eastern North America, in response to climate change.
-We provide a more formal mathematical presentation in the Appendix S1, along with complete code and data for both examples (Appendix S2).
+We provide a more formal mathematical presentation in Appendix S1, along with complete code and data for both examples (Appendix S2).
 
@@
 This naive model uses a simple logistic regression to estimate the naive probability of occurrence \((\psi_N)\) as a function of second order effects of temperature and precipitation:
-\dg{I dropped the 'bayesian' from this sentence because I was looking at the posterior in the following equation and only found the likelihood. It comes in the following equation.}
 
@@
 where \(\theta_M\) is the parameter vector of the model, \(D_M\) is the covariate matrix (i.e., \(T_M, P_M\)), and \(\text{logit}^{-1}\) is the inverse of the logit function.
-We estimate parameters using a Markov Chain Monte Carlo (MCMC) method using the proportional form of Bayes' theorem:
-\dg{you specify the algorithm of the sampler below, but not here}
+We estimate parameters using a Metropolis-Hastings algorithm within a Markov Chain Monte Carlo (MCMC) method using the proportional form of Bayes' theorem:
 
@@
 
-\dg{There is something wrong with the following equation, or I missed the reasonning somewhere in the last year.
+We accomplish model integration by treating the \(\psi_S\), the posterior predictions of the submodel\(\theta_S\), as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
+%-----
+\begin{equation}
+	\label{eq:ex1_integrated}
+	\overbrace{p(\theta_M \mid X_M, T_M, P_M, \theta_S, \psi_S)}^\text{integrated posterior}
+	\propto
+	\overbrace{p\left (\psi_S \mid \theta_M,P_M \right )}^{\substack{\text{new information} \\ \text{from sub-model}}}
+	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) p \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
+	\overbrace{p \left(\theta_S \right)}^{\substack{\text{prior for} \\ \text{sub-model}}}	
+\end{equation}
+%-----
+%\dg{There is something wrong with the following equation, or I missed the reasonning somewhere in the last year.
 In any case, I tried to understand the logic from the text and haven't succeed at it.
@@
 This equation is central to the paper now that the maths have been put in the appendix.
-I was expecting to see the likelihood of X_M given the parameters theta_M.
-The prior of theta_M being the posterior of the model theta_M applied to the data psi_S.
+I was expecting to see the likelihood of X\_M given the parameters theta\_M.
+The prior of theta\_M being the posterior of the model theta\_M applied to the data psi\_S.
 The actual likelihood gives the probability of observing the prediction of the sub-model given the parameters of the sub-model.
 There is not integration on the conditions, it only results from the product of the probabilities.}
-
-We accomplish model integration by treating the posterior of \(\theta_S\) as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
-%-----
-\begin{equation}
-	\label{eq:ex1_integrated}
-	\overbrace{p(\theta_M \mid X_M, T_M, P_M, \theta_S, \psi_S, P_S)}^\text{integrated posterior}
-	\propto
-	\overbrace{p\left (\psi_S \mid \theta_S,P_S \right )}^{\substack{\text{new information} \\ \text{from sub-model}}}
-	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) p \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
-	\overbrace{p \left(\theta_S \right)}^{\substack{\text{prior for} \\ \text{sub-model}}}	
-\end{equation}
-%-----
 As before, the metamodel \(\theta_M\) can be used to predict probability of occurrence \((\psi_I)\).
@@
 These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
-For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modeling process (due to the lack of additional mechanisms in the second model, as mentioned above).
-
-\dg{This is a vague statement. Could you give just an extra sentence specifying how you did it, and then pointing to the Appendix?}
+For purposes of this example, we applied prior weights of 20 and 1 to the correlative and mechanistic models, respectively, reflecting the increased generality and much larger sample size of the correlative data.
+This procedure has the effect of increasing the variance of the model and prevents biasing the parameter estimation in favor of the mechanistic model.
 
@@
 	\includegraphics{figs/ex1_precip.pdf}
-	\caption{Comparison of the naive model, mechanistic sub-model, and the integrated model showing the probability of presence (\(\psi\)) as a function of precipitation.
+	\caption{Comparison of the naive model, mechanistic sub-model, and the integrated model showing the probability of presence $(\psi)$ as a function of precipitation.
 	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.
@@
 %==================
-
-\dg{might worth increasing font size on the figures. And make sure it is the same for all figures.}
 
@@
 We selected a GLM for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., generalized additive models) are compatible with the framework.
-The parameters for the model were estimated using a Metropolis-Hastings algorithm.
 
@@
 \ref{fig:ex1_precip}, \ref{fig:ex1_map}).
-A strong asset of our approach is that it can be used without the full suite of data that would be required to run a fully mechanistic model. 
+A strength of our approach is that it can be used without the full suite of data that would be required to run a fully mechanistic model. 
 Given that the metamodel is correlative, it can be effectively implemented with, e.g., only presence-absence data, or, in the case where true absences are difficult to obtain, with presences and pseudo absences (provided sufficient care is used in interpreting the results of such a model).
@@
 
-\ib{you say in the intro that the bayesian framework allows a easy and simple scaling-up, but the scale-up is still a problem.
-This part about scale is still a bit unclear.
-Our approach do not solve the scaling-up problem.}
-\fd{I fully agree with Isabellle.
-Adequacy between ecological processes, scales, and meaning for upscaling the information  remain a major challenge.
-It would be important here to say how this new approach helps is disentangling the difficulties listed at lines 35-37 in the introduction.}
-\mt{Although scaling is always a problem, this doesn't invalidate the statement that our method makes it easier.
-Think of the first example.
-This scaling would have been very difficult to do without using Bayesian methods.}
-\dg{Do agree with mt}
-
 Determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eq. \ref{eq:ex1_integrated}) is a critical point.
-The challenge is three fold: (i) which spatial and temporal scales (i.e.
-which processes), are to be considered, (ii) how to build a scale the sub-models to be consistent with the metamodels, and (iii) how error and uncertainty propagate from the sub-models to the metamodel.
+The challenge is three fold: (1) which spatial and temporal scales (i.e.
+which processes), are to be considered, (2) how to build a scale the sub-models to be consistent with the metamodels, and (3) how error and uncertainty propagate from the sub-models to the metamodel.
 Although we argue that our proposed framework is able to easily deal with very different scales and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered, the building of scaling functions is an object of investigation on its own.
@@
 
-\dg{I also think that there is the challenge inherent to complex systems that is overlooked to a certain extent here.
-Small scale processes might generate organization at higher scales.
-Model integration might be limited by this problem in some cases, while in others it might solve it by putting bounds on the higher-order organization.
-Think about the evaluation of model parameters with observations at the scale of the operating process.
-There are some parameter values, within the posterior distribution, that could result in non-desirable properties at the larger scale.
-Model integration could buffer this propagation.}
-
-\ib{This paragraph is weired. You say that the implementation is very difficult and end the paragraph with decision, practitioners advantage.
-I would end the paragraph with positive points about implementation, saying that we still can do a lot with JAGS for instance.
-And then start a paragraph to briefly talk about challenges in dialogue with practitionners}
-
 The implementation of the model itself can present an obstacle when model complexity increases. 
@@
 
-Finally, this approach is not just a new methodological tool but is a framework for forecasting species distributions which is fundamentally designed to make the link between modellers and pratitionners, with (1) correctly estimating uncertainties and (2) being updatable to new data or theoretical advances.
-A scientific approach such as presented here is particularly adapted to make a synthesis of available knowledge and provide robust species distribution forecasts based on information known as being the best of the actual scientific knowledge.
-It is actually able both to use large databases, valuing the efforts of data collection, and to include models based on the latest theoretical advances, which is essential to decrease the errors due to model specification \citep{Austin2007}.
-In addition, it would provides practionners and decision-makers with the best possible estimation of uncertainties, with direct applications to risk assessment or to guide the choice where investigating new research and accumulating new data.
-We further argue that the adaptability of our approach is particularly appropriate in our world where collected data and theoretical knowledge is changing as quickly as climate and conservation practices must be adjusted accordingly.
-
-\fd{I am a bit feeling uncomfortable with this section.
-Practionners and decision-makers don't have to use (or even fully understand (as they will never get to that level of complexity) what methods underly the predictions) the models.
-They use the information known as being the best of the actual scientific knowledge.
-As for climate-related variables for CC, the scientific community (as the GIEC for the CC predictions) provide the predictions on SD.
-The scientific community is responsible for indicating the limitations of the use of the generated information.}
-
-\ib{I would add somewhere that this approach is not just a new methodological tool but is a framework for forecasting species distributions which is fundamentally designed to make the link between modellers and pratitionners, with (1) correctly estimating uncertainties and (2) being adaptable/updatable to new data or theoretical advances.}
+Finally, this approach is not just a new methodological tool but is a framework for forecasting species distributions which is fundamentally designed to make the link between modelers and practitioners while correctly estimating uncertainties and being updatable with new data and theoretical advances.
+A scientific approach such as presented here is particularly adapted to make a synthesis of available information and provide robust species distribution forecasts based on information known to be the best available scientific knowledge.
+It can incorporate large databases, valuing the efforts of data collection, and include models based on the latest theoretical advances, which is essential to decrease errors due to model specification \citep{Austin2007}.
+In addition, it provides practitioners and decision-makers with the best possible estimation of uncertainties, with direct applications to risk assessment or to guide the choice where investigating new research and accumulating new data.
+Finally, we argue that the adaptability of our approach is particularly appropriate in a world where collected data and theoretical knowledge is changing as quickly as climate and conservation practices must be adjusted accordingly.
 

### 2015-07-27 tex/appendix_s1.tex
Performed revisions for the EIC

@@
 \usepackage[margin=1in, left=1.2in]{geometry}
+
+\usepackage[dvipsnames]{xcolor}
+\newcommand{\rev}[1]{{\color{RoyalBlue}#1}}
 
@@
 See \citet{Morin2009} for a complete description of these variables.
-For projection, we used predictions for 2080 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the models to forecast suitability into the future.
+For projection, we used predictions for \rev{2100} from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the models to forecast suitability into the future.
 Exploratory analysis revealed that there were collinearities within the predictor variables.

### 2015-07-27 tex/manuscript.tex
Performed revisions for the EIC

@@
 \usepackage[margin=1in]{geometry}
+
+\usepackage[dvipsnames]{xcolor}
+\newcommand{\rev}[1]{{\color{RoyalBlue}#1}}
 
@@
 	\caption{Two simulated datasets used to illustrate the model integration framework.
-	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
+	(a) Presences (circles) or absences (x's) of the species in ecological space, \rev{where the range of precipitation values sampled was 0.1--1}.
 	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., -1.0--1.0).
@@
 We used our framework to constrain a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
-Here we describe briefly the dataset, methods, and the results of the analysis (see Appendix S1 for data and scripts to reproduce the analysis).
+Here we describe briefly the dataset, methods, and the results of the analysis (see \rev{Appendix S2} for data and scripts to reproduce the analysis).
 
-We obtained climate variables, occurrence data, and Phenofit projections at 0.5-degree resolution for both the present and for 2080 for sugar maple, an economically and ecologically important species occurring in eastern North America \citep{Morin2009}.
+We obtained climate variables, occurrence data, and Phenofit projections at 0.5-degree resolution for both the present and for \rev{2100} for sugar maple, an economically and ecologically important species occurring in eastern North America \citep{Morin2009}.
 These data defined the metamodel scale.
@@
 We constructed the naive model by using a binomial generalized linear model (GLM) to relate the presence-absence dataset to three climate variables: the number of degree days (ddeg), mean annual precipitation (an\_prcp), and the ratio of annual precipitation to potential evapotranspiration (pToPET).
-These variables were selected from an initial set of 6 variables \citep[see Appendix S1 and ][for details on the climate variables]{Morin2009}.
+These variables were selected from an initial set of \rev{seven} variables \citep[see Appendix S1 and ][for details on the climate variables]{Morin2009}.
 We selected a GLM for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., generalized additive models) are compatible with the framework.
@@
 	\footnotesize
-	\caption{Standardized logit-scale parameter estimates and 95\% credible intervals for the three models.}
+	\caption{\rev{Parameter estimates and 95\% credible intervals for the naive model (i.e., a species distribution model fit with occurrence data only), and two integrated models combining predictions from the naive model and either the present or future predictions from a mechanistic model.
+All models were fit on predictor variables standardized to mean=0 and unit variance, and all estimates are on the logit scale.
+Climate variables included the number of degree days (ddeg), mean annual precipitation (an\_prcp), and the ratio of annual precipitation to potential evapotranspiration (pToPET).}}
 	\begin{tabular}{lcccccccc}

### 2015-07-31 tex/appendix_s1.tex
Added nicer plot map to appendix
Additional revisions for Reviewer 1

@@
 \begin{figure}[t]
-\includegraphics{figs/plotmap.pdf}
+\includegraphics[width=4.5in]{figs/plotmap.png}
 \caption{Map of permanent forest plot locations}
@@
 The search scope included all models with linear, quadratic, and cubic terms for all three variables.
-Because Phenofit provided predictions in the form of a probability of presence (rather than directly in terms of the metamodel parameters as in the previous example), it was necessary to develop a method for relating these predictions to the metamodel.
+\rev{Thus, the naive model (which also defined the scope of the metamodel), can be formulated as a simple Binomial GLM, as in Example 1:
+\begin{equation}
+\psi_N =\text{logit}^{-1}\left( \theta_M D_M \right)
+\end{equation}
+where \(\psi_N\) is the probability of presence in the naive model, \(\theta_M\) is the parameter vector of the metamodel (presently unconstrained by any additional information), and \(D_M\) is the macroclimate covariate matrix at the metamodel scale.
+The parameter vector is estimated following equation 2 in the main text:
+\begin{equation}
+p\left (\theta_M \mid X_M,D_M \right ) \propto 
+p \left(X_M \mid \theta_M, D_M \right)
+p \left(\theta_M \right)
+\end{equation}
+
+Because Phenofit provided predictions in the form of a probability of presence (rather than directly in terms of the metamodel parameters as in the previous example), it was necessary to relate these predictions to the metamodel.
 We opted to do this via simulation because it is a natural method to use in an MCMC scheme.
-For each MCMC iteration, we generated a simulated dataset by drawing \(N_x\) samples with probability \(p_x\) from a Binomial distribution, where \(N_x\) is the number of observations in the occurrence dataset for a given grid cell \(x\), and \(p_x\) is the predicted probability of presence from Phenofit for that cell.
-The metamodel was then fit to these simulated data (with prior conditioning on the naive model).
+We thus treated the Phenofit predictions (\(\psi_P\)) as an observed result of a true underlying set of occurrences (\(X_P\)).
+Because these occurrences were unobserved (and in the case of the future distribution, unobservable), it was necessary to infer them using MCMC.
+For each MCMC iteration, we generated a simulated occurrence dataset \(\hat{X}_P\) by drawing from a Binomial distribution with probabilities equal to \(\psi_P\).
+In order to give both the occurrence dataset and Phenofit predictions equal weight, we drew the same number of samples as were present in the occurrence dataset.
+The metamodel was then fit to these simulated data (with conditioning on the naive model as well):
+\begin{equation} 
+\overbrace{p(\theta_M \mid X_M, D_M, X_P, \psi_P)}^\text{integrated posterior}
+\propto
+\overbrace{p\left (X_P \mid \theta_M,D_M,\psi_P \right )}^{\substack{\text{new information} \\ \text{from Phenofit}}}
+\overbrace{p \left(X_M \mid \theta_M, D_M \right) p \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
+\overbrace{p \left(\psi_P \right)}^{\substack{\text{prior for} \\ \text{Phenofit}}}	
+\end{equation}
 This procedure, by generating a random dataset with each iteration, incorporated the variance in the Phenofit predictions.
+Finally, because we were interested in two separate questions with respect to Phenofit integration (one regarding how integration affects uncertainty in our present estimate of the species distribution and a one regarding the effect of integration on projections into the future), we performed two separate integrations with the Phenofit information.
+To address the first question, we used the Phenofit predictions for the present as \(\psi_P\) and present climate as \(D_M\), producing the Integrated-Present model.
+To address the second, we used the future predictions from Phenofit and future climate, resulting in the Integrated-Future model.
+In both cases, the naive model was a GLM relating present occurrences to present climate (and thus the prior response to the environment was static, similarly to how one might use the same conventional SDM to both model present distribution and project into the future).
 Convergence for all models was assessed by comparing the results of multiple chains with overdispersed starting points.
-Based on the results of these initial runs, we thinned all posterior samples by recording only every 50th sample, and we obtained 25000 posterior samples for each model following a burn-in period of 20000 samples.
+Based on the results of these initial runs, we thinned all posterior samples by recording only every 50th sample, and we obtained 25000 posterior samples for each model following a burn-in period of 20000 samples.}
+
+\section*{Implementation}
+\subsection*{Model weights}
+\rev{As a start, users may consider weighting models based on sample size or number of independent treatments, as we have done in example 1, where the weight for the sub-model of 0.05 was chosen based on the 5 precipitation levels relative to 100 independent data points in the correlative dataset.
+Because we are integrating model predictions, a highly precise model model with a small sample size (e.g., an experiment with a small number of treatment levels but strong effects) may provide too strong of a constraint on the metamodel.}
 

### 2015-07-31 tex/manuscript.tex
Additional revisions for Reviewer 1

@@
 
-	\textbf{Main text word count: } 4595 %WORD_COUNT %Target 5000
+	\textbf{Main text word count: } 5083 %WORD_COUNT %Target 5000
 
@@
 
-We accomplish model integration by treating the \(\psi_S\), the posterior predictions of the submodel\(\theta_S\), as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
+We accomplish model integration by treating \(\psi_S\), the posterior predictions of the submodel \(\theta_S\), as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
 %-----
@@
 These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
-For purposes of this example, we applied prior weights of 20 and 1 to the correlative and mechanistic models, respectively, reflecting the increased generality and much larger sample size of the correlative data.
+For purposes of this example, we applied prior weights of 1 and 0.05 to the correlative and mechanistic models, respectively, reflecting the increased generality and much larger sample size of the correlative data.
 This procedure has the effect of increasing the variance of the model and prevents biasing the parameter estimation in favor of the mechanistic model.
@@
 We used our framework to constrain a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
-Here we describe briefly the dataset, methods, and the results of the analysis (see \rev{Appendix S2} for data and scripts to reproduce the analysis).
+Here we describe briefly the dataset, methods, and the results of the analysis (see \rev{ Appendix S1 for implementation details and Appendix S2} for data and scripts to reproduce the analysis).
 
@@
 This procedure provided a consensus view of the future range of the species; we refer to this model as ``Integrated-Future.''
+\rev{In both cases, it was necessary to scale the Phenofit predictions, which were probabilistic and not directly derived from the macroclimatic variables used in to fit the naive model, to make them compatible with the naive model, which was based on occurrence data.
+We used a latent variable approach, where an unobserved occurrence dataset similar to the one used for the naive model was generated at each MCMC iteration by drawing occurrences from a Binomial distribution with probabilities equal to the probability of occurrence from Phenofit (see Appendix S1 for a full statistical presentation of the model).
+This has several advantages.
+In addition to expressing the information from Phenofit in a way that is compatible with the naive model, the uncertainty in the predictions is automatically propagated (by simulating a new pseudo-dataset at each iteration), and it becomes possible to address future distribution via simulated future occurrences, (which, by nature, are unobservable).}
 
@@
 The Integrated-Future model presents a different interpretation of the response of sugar maple to warmer temperatures. 
-The naive model predicted a substantial northward migration; in other words, the expectation under the naive model is that the present occurrence data is an accurate estimation of the niche, and thus the species should track its niche northward as the climate warms.
+The naive model predicted a substantial northward migration; in other words, the expectation under the naive model \rev{is that the present estimate of the realized niche (obtained using occurrence data) is an unbiased reflection of the fundamental niche; thus the species should track temperature northward as the climate warms}.
 The Integrated-Future model, in contrast, predicted substantially more tolerance to warmer temperatures, reflecting similar predictions from Phenofit (Figs.
 \ref{fig:ex2_response}, \ref{fig:ex2_map_fut}).
+\rev{This is because Phenofit estimates different aspects of the realized niche.}
 Although both models predicted a northward range shift, the change under Phenofit was limited to approximately 200 km north of the present range limit of the species, compared with more than 900 km for the naive model.
@@
 \ref{fig:ex2_map_fut}).
+\rev{The metamodel thus presents a consensus view of the niche of the species with respect to the macroclimatic variables included in the model (Fig.
+\ref{fig:ex2_map_pres}), incorporating the present range of the species (using the occurrence data) and information from Phenofit on what conditions will be tolerable in the future.}
 
@@
 The difficulty of including information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Smolik2010, Thuiller2014a}. 
-Bayesian methods such as our are able to overcome some of these limitations, as they provide a natural framework for the incorporation of multiple sources of information.
+\rev{Bayesian methods provide a natural framework for the incorporation of multiple sources of information, making them an attractive alternative for SDMs.
+Hierarchical models in particular have the potential to capture many of the intricacies necessary for implementing hybrid models \citep{Latimer2006}.
+\citet{Pagel2012} developed a hybrid approach to species distributions via a dynamic range model.
+Similar to our approach, their model integrated demographic information, abundances, and presence/absence data within a hierarchical Bayesian framework to predict species ranges.
+However, their approach explicitly links the modelled processes to occurrences/abundances via a detailed demographic model, requiring data that may not be available for many species.
+In contrast, our approach allows for the inclusion of more incomplete datasets because integration is performed via the separate predictions of each model (following the application of the scaling function where necessary).}
 Because the metamodel is expressed as a series of conditional probabilities (see Appendix S1), this information can be included simply as long as the probability of the metamodel can be expressed mathematically.
@@
 Our framework is still applicable whatever the chosen upscaling approach and is able to propagate uncertainties from sub-models to the metamodel.
+\rev{Indeed, if a sub-model provides poor information (due, for example, to cross-scale nonlinearities in the response to the environment), the resulting metamodel predictions may be worse than the pre-integration naive model.
+In general, we advise users of this framework to carefully consider the scaling in their models with respect to the biology of the organism studied, and to use prior model weights to downweight the scaled models when there is uncertainty about the applicability of a sub-model at the metamodel scale.
+For instance, in the first example, we applied a model weight to decrease the influence of the mechanistic model on the metamodel (while still retaining some information contained therein), recognizing that such a simplistic model may not scale well.
+Finally, as always when modelling ecological systems, we urge humility in the interpretation of model results and suggest the use of model evaluation and validation tools whenever possible.
+We provide additional discussion on the implementation of model weights in Appendix S1.}
 

### 2015-08-05 tex/appendix_s1.tex
Almost done with revisions

@@
 \section*{Implementation}
+\rev{Here, we provide a brief guide on implementing our framework.
+We focus on our two examples, providing a brief outline of the steps to reproduce the analyses using the code and data provided in Appendix S2.
+In the section for each example, we provide some references to software packages that may be of assistance to beginning users.
+As a full discussion of the practicalities of Bayesian modelling is beyond the scope of this paper, we direct the reader to some of the many excellent resources for Bayesian and hierarchical modelling in ecology \citep[e.g., we suggest][]{Bolker2007, Royle2009, Link2010}.
+We also provide some additional information on applying model weights.
+Our intent is that, with this guide and the accompanying code, users familiar with R and with some experience with Bayesian modelling will be able to implement our method.}
+
+\subsection*{Example 1}
+\rev{
+Example 1 is implemented fully using off-the shelf statistical software.
+Data preparation, exploration, and visualization is all performed in R \citep{R}.
+Parameterization of all models is performed using JAGS, which is a standalone implementation of a Metropolis-Hastings and Gibbs sampler, as well as RJAGS, the R interface to the JAGS program \citep{RJAGS}.
+Within the {\tt example\_1} directory in Appendix S2, we provide 6 numbered scripts.
+These are designed to be run from the command line using the RScript program that ships with most standard R installations (e.g., {\tt Rscript 1-m1.r} to run the first script).
+However, we encourage users to review the code rather than simply running the examples from the command line.
+Users of the R console (or the RStudio environment) will be able to run the code one line at a time, provided the working directory is set to the {\tt example\_1} directory beforehand.
+
+There are three main steps to running the analysis, corresponding to the first three numbered scripts.
+The first (file {\tt 1-m1.r}) generates the simulated presence-absence dataset, loads a JAGS model representing the metamodel (from the file {\tt ex1\_metamodel.jags}), and runs JAGS on the model to produce posterior samples for the naive model.
+Note that, because the naive model and the metamodel are at the same scale and differ only in whether they include additional information from the sub-model, the JAGS model file is the same for both.
+All data and results from this step are saved to the {\tt dat/ex1\_m1.rdata} file.
+Step two ({\tt 2-m2.r}) creates the dataset for the sub-model, loads the JAGS model ({\tt ex1\_model2.jags}), and runs the model, saving all results to {\tt dat/ex1\_m2.rdata}.
+Finally, the third step ({\tt 3-mm.r}) simply reads the estimates from the sub-model, incorporates them into the information passed to JAGS for estimating the metamodel, and runs JAGS to obtain the integrated posterior samples and save them to {\tt dat/ex1\_mm.rdata}.
+The posterior samples are saved as MCMC objects from package coda; users should refer to the documentation for that package for information on using these objects.
+The remaining three scripts contain code for reproducing the figures contained in the main manuscript.
+Users may find them useful for guidance on working with posterior samples generated with JAGS.
+}
+
+\subsection*{Example 2}
+\rev{
+For performance reasons, we have implemented Example 2 as a standalone C++ program, designed to be run multi-threaded in high-performance computing applications.
+Due to the size of the dataset and the increased complexity of the model (relative to Example 1), the computational load is considerably higher.
+The procedure to compile and run the model is also somewhat more complex.
+Therefore, to facilitate exploration without requiring users to actually build and run the model, we have included the posterior samples we obtained with Appendix S2.
+This will allow users to build the figures and explore the results of the naive and integrated models.
+For the portions that can be run in R, as in Example 1, we have included numbered scripts that can be run with {\tt Rscript}.
+There are three scripts included in the {\tt example\_2/r/} directory, numbered 1, 3, and 5 (steps 2 and 4 are running the C++ program to obtain samples for the naive and integrated models, respectively).
+Step 1 ({\tt 1\_prepare\_naive\_data.r}) prepares the dataset for the naive model to be used with the sampler, resulting in the files {\tt dat/mcmc/naiveData.csv} (for use with the C++ program) and {\tt dat/mcmc/naiveData.rds} (for use with R, can be loaded using the {\tt readRDS} function).
+Step 2 (see following paragraph) creates the file {\tt results/mcmc/naivePosterior.csv}, which contains the posterior samples for the naive model and which we have included for users that do not wish to build and run the C++ program themselves.
+Step 3 reads the results of the naive model estimation and uses it to prepare the files needed to estimate the integrated model.
+As with step 2, step 4 depends on the C++ program and produces posterior samples for both the Integrated-Present and Integrated-Future models (files {\tt integratedPresent.csv} and {\tt integratedFuture.csv}, respectively, in the {\tt results/mcmc/} directory).
+Finally, step 5 reads all of the CSV files containing posterior samples and processes them into MCMC objects (saved in {\tt results/posteriors.rdata}) that can be easily manipulated in R with the coda package.
+We also include code for reproducing the figures contained in the manuscript (all scripts starting with {\tt fig}), and a brief interactive script for performing model evaluation ({\tt evaluation.r})
+
+For users wishing to build and run the C++ program to estimate the posterior distributions, some knowledge of C++ compilers will be required.
+A familiarity with GNU Make will also be of assistance, because the build procedure is fully automated and documented within the included makefile.
+All C++ code is contained within the {\tt src} subdirectory and should be fully compatible with any platform capable of compiling and running standard C++89 programs.
+The only external dependency is the GNU Scientific Library (freely available at http://www.gnu.org/software/gsl/), an open-source C library which we have used for random number generation and other mathematical operations.
+Additionally, for high-perfomance computing applications, we have included compiler commands for using OpenMP.
+When enabled, this will significantly improve the speed at which the likelihood is computed, and thus reducing the runtime of the model.
+All compiler commands, as well as the commands used to run the model and generate posterior samples, are included in the makefile.
+Additionally, once the program is compiled, help with program options can be obtained using the {\tt -h} switch (i.e., {\tt integrated\_model -h} at the command line).
+}
+
 \subsection*{Model weights}
-\rev{As a start, users may consider weighting models based on sample size or number of independent treatments, as we have done in example 1, where the weight for the sub-model of 0.05 was chosen based on the 5 precipitation levels relative to 100 independent data points in the correlative dataset.
-Because we are integrating model predictions, a highly precise model model with a small sample size (e.g., an experiment with a small number of treatment levels but strong effects) may provide too strong of a constraint on the metamodel.}
+\rev{Model weights, expressed mathematically as the prior probability of each individual sub-model, allow the user to express prior belief that a particular sub-model is effective at describing the range of a species.
+This is particularly useful, as in Example 1, when a sub-model is highly statistically precise but, due to limited scope, is not effective alone at describing the range of a species.
+The hypothetical experiment in Example 1 included only 5 treatments in a single variable, and thus misses many important processes that generate species ranges.
+Selecting the specific weighting to use will, of necessity, be somewhat subjective, as it is often difficult to express prior belief of this sort in precise mathematical terms.
+As a start, users may consider weighting models based on sample size or number of independent treatments, as we have done in example 1, where the weight for the sub-model of 0.05 was chosen based on the 5 precipitation levels relative to 100 independent data points in the correlative dataset.
+Implementation of the weights within the model code itself will depend on the implementation.
+In the case of Example 1, the weight was applied by decreasing the precision of all terms that were informed by the submodel (i.e., making the prior that resulted from the sub-model less informative).
+}
 

### 2015-08-05 tex/manuscript.tex
Almost done with revisions

@@
 
-	\textbf{Main text word count: } 5083 %WORD_COUNT %Target 5000
+	\textbf{Main text word count: } 5104 %WORD_COUNT %Target 5000
 
@@
 \begin{figure}[tb]
-	\includegraphics[width=5.25in]{figs/ex1_map.pdf}
+	\includegraphics[width=6.5in]{figs/ex1_map.png}
 	\caption{Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
@@
 \begin{figure}[t]
-\includegraphics[width=6.5in]{figs/ex2_pres_map.pdf}
+\includegraphics[width=6.5in]{figs/ex2_pres_map.png}
 \caption{Range predictions for present climate for Phenofit (A), the naive model (B), the Integrated-Present model (C), and the difference between posterior prediction standard errors of the Integrated-Present and Naive models.
@@
 \begin{figure}[t]
-\includegraphics[width=6.5in]{figs/ex2_fut_map.pdf}
+\includegraphics[width=6.5in]{figs/ex2_fut_map.png}
 \caption{Range predictions for future climate for Phenofit (A), the naive model (B), the Integrated-Future model (C), and the difference between posterior prediction standard errors of the Integrated-Future and Naive models.
@@
 We acknowledge funding from the Quebec Centre for Biodiversity Science as well as NSERC strategic grant 430393-12.
-Comments from Charles Canham, William Godsoe, and Tamara Münkemüller improved a previous version of this manuscript.
+We thank editors David Currie and Arndt Hampe, as well as two anonymous reviewers, whose feedback greatly improved the manuscript.
+Additionally, comments from Charles Canham, William Godsoe, and Tamara Münkemüller improved a previous version of this manuscript.
 WT received support funding from the European Research Council under the European Community's Seven Framework Programme FP7/2007-2013 Grant Agreement no.

### 2015-08-07 tex/appendix_s1.tex

@@
 \usepackage[margin=1in, left=1.2in]{geometry}
-
-\usepackage[dvipsnames]{xcolor}
-\newcommand{\rev}[1]{{\color{RoyalBlue}#1}}
 
@@
 % FIGURE 1
-\begin{figure}[t]
+\begin{figure}[p!]
 \input{diagram}
@@
 This allowed the sub-model to inform the integrated model without dominating the results.
+For temperature-related parameters, the posterior distributions for the naive and integrated models were indistinguishable, whereas parameters related to precipitation were substantially revised, including large reductions in uncertainty for the integrated model (Fig.
+\ref{fig:ex1params}).
+
+\begin{figure}
+\includegraphics[width=6.5in]{figs/ex1_params.png}
+\caption{Relative posterior density of all example one parameters for the naive model, mechanistic sub-model, and the integrated model.
+Parameters related to temperature ($b_1$ and $b_2$) were not informed with additional data, and showed little change from the naive model.
+The intercept ($b_0$ and parameters related to precipitation ($b_3, b_4$) were informed by integration and showed substantial revision between the naive and integrated models.
+Boxplots show the median, interquartile range, and the range.}
+\label{fig:ex1params}
+\end{figure}
 
@@
 
-\begin{figure}[t]
+\begin{figure}
 \includegraphics[width=4.5in]{figs/plotmap.png}
@@
 See \citet{Morin2009} for a complete description of these variables.
-For projection, we used predictions for \rev{2100} from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the models to forecast suitability into the future.
+For projection, we used predictions for 2100 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the models to forecast suitability into the future.
 Exploratory analysis revealed that there were collinearities within the predictor variables.
@@
 The search scope included all models with linear, quadratic, and cubic terms for all three variables.
-\rev{Thus, the naive model (which also defined the scope of the metamodel), can be formulated as a simple Binomial GLM, as in Example 1:
+Thus, the naive model (which also defined the scope of the metamodel), can be formulated as a simple Binomial GLM, as in Example 1:
 \begin{equation}
@@
 Convergence for all models was assessed by comparing the results of multiple chains with overdispersed starting points.
-Based on the results of these initial runs, we thinned all posterior samples by recording only every 50th sample, and we obtained 25000 posterior samples for each model following a burn-in period of 20000 samples.}
+Based on the results of these initial runs, we thinned all posterior samples by recording only every 50th sample, and we obtained 25000 posterior samples for each model following a burn-in period of 20000 samples.
+Finally, we evaluated the Naive and Integrated-Present models by computing the area under the receiver operating curve (AUC).
+AUC measures the ability of the model to discriminate presences and absences, where a value of 1 indicates perfect discrimination and 0.5 is random discrimination.
+Because we obtained joint posterior distributions for all models, it is not possible to evaluate the performance of a `best' model; rather, we computed AUC for every posterior sample of both models, resulting in posterior distributions of AUC (Fig.
+\ref{fig:auc}).
+Median AUC values were 0.802 and 0.797 for the Naive and Integrated-Present models, respectively, and the overlap between the two distributions suggests little difference in the performance of each model.
+
+\begin{figure}
+\includegraphics[width=5.5in]{figs/ex2_auc.png}
+\caption{Posterior distributions of the area under the receiver operating curve (AUC) for the Naive and Integrated-Present models.
+For comparison, the y-values have been scaled by the maximum densities.
+Boxplots in the upper panel show the median, interquartile range and \(1.5 \times\) the interquartile range, along with more extreme values plotted individually.}
+\label{fig:auc}
+\end{figure}
 
 \section*{Implementation}
-\rev{Here, we provide a brief guide on implementing our framework.
+Here, we provide a brief guide on implementing our framework.
 We focus on our two examples, providing a brief outline of the steps to reproduce the analyses using the code and data provided in Appendix S2.
@@
 We also provide some additional information on applying model weights.
-Our intent is that, with this guide and the accompanying code, users familiar with R and with some experience with Bayesian modelling will be able to implement our method.}
+Our intent is that, with this guide and the accompanying code, users familiar with R and with some experience with Bayesian modelling will be able to implement our method.
 
 \subsection*{Example 1}
-\rev{
 Example 1 is implemented fully using off-the shelf statistical software.
@@
 Users may find them useful for guidance on working with posterior samples generated with JAGS.
-}
 
 \subsection*{Example 2}
-\rev{
 For performance reasons, we have implemented Example 2 as a standalone C++ program, designed to be run multi-threaded in high-performance computing applications.
@@
 A familiarity with GNU Make will also be of assistance, because the build procedure is fully automated and documented within the included makefile.
-All C++ code is contained within the {\tt src} subdirectory and should be fully compatible with any platform capable of compiling and running standard C++89 programs.
+All C++ code is contained within the {\tt src} subdirectory and should be fully compatible with any platform capable of compiling and running standard C++98 programs.
 The only external dependency is the GNU Scientific Library (freely available at http://www.gnu.org/software/gsl/), an open-source C library which we have used for random number generation and other mathematical operations.
@@
 Additionally, once the program is compiled, help with program options can be obtained using the {\tt -h} switch (i.e., {\tt integrated\_model -h} at the command line).
-}
 
 \subsection*{Model weights}
-\rev{Model weights, expressed mathematically as the prior probability of each individual sub-model, allow the user to express prior belief that a particular sub-model is effective at describing the range of a species.
+Model weights, expressed mathematically as the prior probability of each individual sub-model, allow the user to express prior belief that a particular sub-model is effective at describing the range of a species.
 This is particularly useful, as in Example 1, when a sub-model is highly statistically precise but, due to limited scope, is not effective alone at describing the range of a species.
@@
 In the case of Example 1, the weight was applied by decreasing the precision of all terms that were informed by the submodel (i.e., making the prior that resulted from the sub-model less informative).
-}
 

### 2015-08-07 tex/manuscript.tex

@@
 
-	\textbf{Abstract word count: } 235 %ABS_COUNT
+	\textbf{Abstract word count: } 299 %ABS_COUNT
 
-	\textbf{Main text word count: } 5104 %WORD_COUNT %Target 5000
+	\textbf{Main text word count: } 5039 %WORD_COUNT %Target 5000
 
@@
 \noindent
+\paragraph{Aim} 
+Current interest in forecasting changes to species ranges have resulted in a great diversity of approaches to species distribution models (SDMs).
+However, most approaches include only a small subset of the available information, and many ignore smaller-scale processes such as growth, fecundity, and dispersal.
+Furthermore, different approaches often produce divergent predictions, with no simple method to reconcile them.
+Here, we present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods.
 
-Current interest in forecasting potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
-However, in general, single approaches include only a small subset of the available information about a species. 
-Correlative models often predict presence or absence as a function of the environment, but ignore smaller-scale processes such as growth, fecundity, and dispersal.
-Furthermore, different approaches often produce divergent predictions, with no simple method to reconcile them.
-We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
-Our framework uses a metamodel, estimated in a hierarchical Bayesian framework, that is constrained by both data directly applied to the metamodel and by the results of one or more sub-models.
-The output of the metamodel is in the form of probabilistic estimates of species presence.
-Metamodel outputs also incorporate uncertainty from both data and process error from all sub-models, and include all of the information used as input for the original scale-specific sub-models.
-We illustrate our approach using theoretical and empirical examples, and demonstrate that the framework better characterizes uncertainty and can substantially reduce it when projecting beyond the range of the original data.
-We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
-The framework has wide potential for use in species distribution modelling, and can drive better integration of multi-scale data into ecological decision-making.
+\paragraph{Location} 
+Eastern North America (as an example).
+
+\paragraph{Methods}
+\rev{Our framework builds a metamodel that is constrained by the results of multiple sub-models and results in probabilistic estimates of species presence.
+We applied our approach to a simulated dataset to demonstrate the integration of a correlative SDM with a theoretical model.
+In a second example, we built an integrated model combining the results of a physiological model with presence-absence data for sugar maple \emph{Acer saccharum}, an abundant tree native to eastern North America.}
+
+\paragraph{Results}
+\rev{For both examples, the integrated models included information from all data sources and substantially improved uncertainty when making projections.
+For the second example, we found that the integrated model outperformed the source models with respect to uncertainty when modelling the present range of the species.
+When projecting into the future, the model provided a consensus view of two models that differed substantially in their predictions.
+Uncertainty was reduced where the models agreed and was greater where they diverged, providing a more realistic view of the state of knowledge than either source model.}
+
+\paragraph{Main conclusions}
+We conclude by discussing the application of our method and its accessibility to applied ecologists.
+In ideal cases, our framework can be easily implemented using off the shelf software.
+The framework has wide potential for use in species distribution modelling, and can drive better integration of multi-source and multi-scale data into ecological decision-making.
 
@@
 Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-support tools in biodiversity management \citep{Guisan2013}.
-Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}.
+Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the processes generating species range limits \citep{Wu1995, Levin1998}.
 Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, hybrid zone dynamics \citep{Engler2013}, and climate change impacts on species distributions \citep{Blois2013, Thuiller2014}.
@@
 A growing body of theory, reflecting the diversity of processes generating species ranges, also has contributed to model diversification \citep{Boulangeat2012}.
-Of these model types, fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, perform poorly when applied at the scale of species ranges.
-For instance, biotic interactions are usually not modeled mechanistically at the scale of species distributions because they are poorly known or have not been recorded, despite being considered a key determinant of range limits \citep{Pigot2013}.
-In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to associated distributions of other variables have the advantage of indirectly accounting for all underlying processes \citep{Guisan2000}.
+Of these model types, fine-scale mechanistic models often capture important ecological processes quite well, but may perform poorly when applied at the scale of species ranges.
+For instance, biotic interactions are usually not modeled mechanistically at regional or continental scales because they are poorly known or unrecorded, despite being considered a key determinant of range limits \citep{Pigot2013}.
+In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
 However, their predictions rely on the stationarity of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}.
@@
 Clearly, an approach is needed to unify the strength of different modeling approaches that can incorporate multiple data sources and modeling paradigms.
-Toward this end, we present here an application of hierarchical Bayesian methods that uses outputs from multiple models to inform the results of the final model.
+Toward this end, we present an application of hierarchical Bayesian methods that uses outputs from multiple models to inform the results of the final model.
 Techniques for multimodel inference have proliferated in recent years.
-For example, hybrid models that allow for flexible combinations between mechanistic and phenomenological sub-models are commonly employed in SDMs \citep{Gallien2010, Thuiller2013, Boulangeat2014}.
-Within the hybrid framework, a correlative model might be used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while a more mechanistic approach could include biotic interactions and space-time dynamics through dispersal \citep{Smolik2010}.
+For example, hybrid models that allow for combinations between mechanistic and phenomenological sub-models are commonly employed in SDMs \citep{Gallien2010, Thuiller2013, Boulangeat2014}.
+Within the hybrid framework, a correlative model might be used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while a more mechanistic approach could include biotic interactions and space-time dynamics \citep{Smolik2010}.
 However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, uncertainties are approximated and their attribution to different sources is difficult to distinguish.
-A commonly-employed alternative to hybrid models is the direct combination of predictions, where predictions from models operating at the same spatio-temporal scales can be combined \citep[e.g., model averaging, ensemble forecasting;][]{Araujo2007}.
-However, the applicability of such techniques is limited; for example, because uncertainty is approximated and may be poorly understood, it is not possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
+An alternative to hybrid models is the direct combination of predictions, allowing models operating at the same spatio-temporal scales to be combined \citep[e.g., model averaging, ensemble forecasting;][]{Araujo2007}.
+However, because uncertainty is approximated and may be poorly understood, it is not possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
 
 We propose an alternative to these other approaches using a hierarchical Bayesian framework. 
-This approach provides a number of desirable advantages, including (1) the ability to incorporate multiple modes of inference (e.g., mechanistic, correlative models) \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}, (2) an easy mechanism to include multiple data sources at various scales \citep{Levin1992, Peters2004}, and (3) an intuitive and comprehensive reporting of uncertainty in model predictions that reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}.
+This approach provides a number of advantages, including (1) the ability to incorporate multiple modes of inference (e.g., mechanistic, correlative models) \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}, (2) an easy mechanism to include multiple data sources at various scales \citep{Levin1992, Peters2004}, and (3) an intuitive and comprehensive reporting of uncertainty in model predictions that reflects variation at all levels of organization \citep{Cressie2009, Hobbs2011}.
 Contrary to hybrid methods, the aim is not to link different sub-models into a single one, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for more flexibility regarding the type of information included.
@@
 %-----
-%\dg{There is something wrong with the following equation, or I missed the reasonning somewhere in the last year.
-In any case, I tried to understand the logic from the text and haven't succeed at it.
-Either it needs to be fixed, or it needs to be better explained.
-If I don't get it, my prior tells me nobody will.
-This equation is central to the paper now that the maths have been put in the appendix.
-I was expecting to see the likelihood of X\_M given the parameters theta\_M.
-The prior of theta\_M being the posterior of the model theta\_M applied to the data psi\_S.
-The actual likelihood gives the probability of observing the prediction of the sub-model given the parameters of the sub-model.
-There is not integration on the conditions, it only results from the product of the probabilities.}
 As before, the metamodel \(\theta_M\) can be used to predict probability of occurrence \((\psi_I)\).
@@
 Assuming we have chosen climate variables that well represent the constraints on the species, including Phenofit in our model is likely to reduce bias in our estimate of the fundamental niche.
-The posterior predictions of the model (that is, the probability of presence projected into geographic space) will incorporate uncertainty from all sources, including the modeling process, data, parameterization, and the ecological processes included.
+The posterior predictions of the model (that is, the probability of presence in geographic space) will incorporate uncertainty from all sources.
 This can provide a much more accurate estimate of the uncertainty of our predictions.
 Thus, for our first integration we combined the naive model with the Phenofit predictions for the present; we refer to this model as ``Integrated-Present'' hereafter.
-We also validated this model and compared it against the naive model by computing the area under the receiver operating curve (AUC), a commonly-used metric that measures the rate of correct classification and varies from 0 to 1 (where a value of 1 indicates perfect classification and 0.5 indicates no difference from a random model) \citep{Swets1988}.
+\rev{We also evaluated this model by constructing calibration curves and by computing the area under the receiver operating curve (AUC; seeAppendix S1), a metric that evaluates classification ability where 0 indicates perfect classification and 0.5 indicates no difference from a random model \citep{Swets1988}.}
 The second modeling goal is projecting changes to the range of sugar maple following climate change.
@@
 This procedure provided a consensus view of the future range of the species; we refer to this model as ``Integrated-Future.''
-\rev{In both cases, it was necessary to scale the Phenofit predictions, which were probabilistic and not directly derived from the macroclimatic variables used in to fit the naive model, to make them compatible with the naive model, which was based on occurrence data.
-We used a latent variable approach, where an unobserved occurrence dataset similar to the one used for the naive model was generated at each MCMC iteration by drawing occurrences from a Binomial distribution with probabilities equal to the probability of occurrence from Phenofit (see Appendix S1 for a full statistical presentation of the model).
-This has several advantages.
-In addition to expressing the information from Phenofit in a way that is compatible with the naive model, the uncertainty in the predictions is automatically propagated (by simulating a new pseudo-dataset at each iteration), and it becomes possible to address future distribution via simulated future occurrences, (which, by nature, are unobservable).}
+\rev{In both cases, it was necessary to scale the Phenofit predictions, which were probabilistic, to make them compatible with the naive model, which was fit using occurrence data.
+We used a latent variable approach, which posits an unobserved ``true'' presence-absence dataset from which the Phenofit probabilities are derived. 
+Similar to any other unknown parameter, we can generate a posterior distribution of this dataset by drawing samples during the MCMC procedure; thus, at each iteration, a dataset similar to the one used for the naive model was generated using the Phenofit probabilities (see Appendix S1 for a full statistical presentation of the model).
+This procedure expresses the information from Phenofit in a way that is compatible with the naive model, and also propagates the uncertainty in the Phenofit predictions.
+Additionally, it becomes possible to address future distribution via simulated future occurrences (which, by nature, are unobservable).}
 
@@
 %% compare parameters
-Model integration resulted in substantial reduction in posterior uncertainty in the parameters, and, for the Integrated-Future model, a relatively large revision in the the estimate for the response of sugar maple to temperature (i.e., variable ddeg) (Table \ref{tab:ex2_params}, Fig.
+Model integration resulted in substantial reduction in posterior uncertainty in the parameters, and, for the Integrated-Future model, a large revision in the the estimate for the response of sugar maple to temperature (i.e., variable ddeg) (Table \ref{tab:ex2_params}, Fig.
 \ref{fig:ex2_response}).
@@
 When considering the present species distribution, the naive and Integrated-Present models made very similar predictions (Fig. \ref{fig:ex2_map_pres}).
-Furthermore, both models performed relatively well when evaluated against reserved data, with AUC values of 0.72 and 0.73 for the naive and integrated models, respectively.
+Furthermore, both models performed well when evaluated against reserved data, with \rev{median} AUC values of 0.802 and 0.797 for the naive and integrated models, respectively (see Appendix S1).
 The small difference between the models indicates that both models effectively predict the probability of presence of sugar maple.
 The major advantage to integration is the improved understanding of uncertainty in the predictions, with greater uncertainty in southern portions of the range (Fig. \ref{fig:ex2_map_pres}).
-It is important to note that the naive model was the basis for the integrated model; thus the increased uncertainty present in the integrated model is not the result of a ``worse'' model, but rather should be viewed as a correction to overfitting in the naive model.
-The uncertainty in the integrated model includes uncertainty arising from the processes that we have incorporated into the metamodel.
+It is important to note that the naive model was the basis for the integrated model; thus the increased uncertainty present in the integrated model is not the result of a ``worse'' model, but rather should be viewed as a correction to overfitting in the naive model that incorporates uncertainty arising from the processes that we have included into the metamodel.
 Phenofit predicts fitness based on how climate affects phenological timings, frost injury, reproduction, and survival \citep{Chuine2001, Morin2009}.
-Thus, climatic factors that ultimately limit species distribution might be quite different under Phenofit, as illustrated by the substantial differences in the potential future distributions predicted by Phenofit and the naive model.
+Thus, climatic factors that ultimately limit species distribution might be quite different under Phenofit, as illustrated by the differences in the potential future distributions predicted by Phenofit and the naive model.
 

### 2015-08-28 tex/appendix_s1.tex
Nearly final draft

@@
 Based on the results of these initial runs, we thinned all posterior samples by recording only every 50th sample, and we obtained 25000 posterior samples for each model following a burn-in period of 20000 samples.
-Finally, we evaluated the Naive and Integrated-Present models by computing the area under the receiver operating curve (AUC).
+Finally, we evaluated the Naive and Integrated-Present models by computing the area under the receiver operating curve (AUC) and by computing calibration curves using reserved data.
 AUC measures the ability of the model to discriminate presences and absences, where a value of 1 indicates perfect discrimination and 0.5 is random discrimination.
@@
 Median AUC values were 0.802 and 0.797 for the Naive and Integrated-Present models, respectively, and the overlap between the two distributions suggests little difference in the performance of each model.
+Calibration curves suggested that the Integrated-Present model overpredicted somewhat compared to the Naive model (Fig. \ref{fig:calib}).
+This is unsurprising, as the validation data were a reserved subset of the same data used to fit the naive model, while the Integrated-Present model included additional information.
+Furthermore, fitted values from Phenofit were generally greater in the same geographic region than the Naive model (see Fig.
+6, main manuscript), explaining the overprediction of the Integrated-Present model and suggesting that values from the two models may not be exactly numerically equivalent, despite the interpretation of both as a probability of presence.
+No model evaluation was performed for the Integrated-Future model, because no independent dataset is available for future predictions.
 
@@
 \label{fig:auc}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=6.5in]{figs/ex2_calib.png}
+\caption{Calibration curve for the Naive (left panel) and Integrated-Present (right panel) models.
+The calibration curve compares the predicted probability of presence from the model (horizontal axis) to the observed proportion of presences (left axis); values were computed by classifying grid cells into 20 equal-width bins and performing a linear regression (weighted by the sample size in each bin) of predicted values against observed frequencies.
+Axes end at 0.7 because neither model predicted probabilities of presence greater than this value.
+Solid lines are posterior mean values, shaded regions are 95\% credible intervals.
+The dashed line indicates perfect calibration.}
+\label{fig:calib}
 \end{figure}

### 2015-08-28 tex/manuscript.tex
Nearly final draft

@@
 \date{}
-\author[1,2,13]{Matthew V. Talluto (mtalluto@gmail.com)}
+\author[1,2,3,4,13]{Matthew V. Talluto (mtalluto@gmail.com)}
 \author[1,2]{Isabelle Boulangeat (isabelle.boulangeat@gmail.com)}
-\author[3]{Aitor Ameztegui (ameztegui@gmail.com)}
-\author[4]{Isabelle Aubin (Isabelle.Aubin@NRCan-RNCan.gc.ca)}
-\author[1,2,5]{Dominique Berteaux (Dominique\_Berteaux@uqar.ca)}
+\author[5]{Aitor Ameztegui (ameztegui@gmail.com)}
+\author[6]{Isabelle Aubin (Isabelle.Aubin@NRCan-RNCan.gc.ca)}
+\author[1,2,7]{Dominique Berteaux (Dominique\_Berteaux@uqar.ca)}
 \author[1,2]{Alyssa Butler (ca.butler10@gmail.com)} 
-\author[6,7]{Frédérik Doyon (Frederik.Doyon@uqo.ca)}
-\author[8]{C. Ronnie Drever (cdrever@tnc.org)}
-\author[9]{Marie-Josée Fortin (mariejosee.fortin@utoronto.ca)}
+\author[8,9]{Frédérik Doyon (Frederik.Doyon@uqo.ca)}
+\author[10]{C. Ronnie Drever (cdrever@tnc.org)}
+\author[11]{Marie-Josée Fortin (mariejosee.fortin@utoronto.ca)}
 \author[1]{Tony Franceschini (Tony.Franceschini@uqar.ca)}
-\author[10]{Jean Liénard (jean.lienard@gmail.com)}
-\author[4]{Dan McKenney (Dan.McKenney@nrcan-rncan.gc.ca)}
+\author[12]{Jean Liénard (jean.lienard@gmail.com)}
+\author[6]{Dan McKenney (Dan.McKenney@nrcan-rncan.gc.ca)}
 \author[2,3]{Kevin A. Solarik (kevinsolarik@hotmail.com)}
-\author[10]{Nikolay Strigul (nick.strigul@vancouver.wsu.edu)}
-\author[11,12]{Wilfried Thuiller (wilfried.thuiller@ujf-grenoble.fr)}
+\author[12]{Nikolay Strigul (nick.strigul@vancouver.wsu.edu)}
+\author[3,4]{Wilfried Thuiller (wilfried.thuiller@ujf-grenoble.fr)}
 \author[1,2]{Dominique Gravel (dominique\_gravel@uqar.ca)}
@@
 \affil[2]{Quebec Centre for Biodiversity Science, Montreal, Quebec, Canada}
-\affil[3]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
-\affil[4]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
-\affil[5]{Centre for Northern Studies, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
-\affil[6]{Université du Québec en Outaouais, Gatineau, Quebec, Canada}
-\affil[7]{Institut des Sciences de la Forêt Tempérée (ISFORT), Ripon, Quebec, Canada}
-\affil[8]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
-\affil[9]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
-\affil[10]{Department of Mathematics, Washington State University, Vancouver, Washington, USA}
-\affil[11]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
-\affil[12]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
-\affil[13]{Author for correspondance. Address: Departament de Biologie, chimie, et geographie, 300, Allée des Ursulines, Rimouski, Quebec G5L 3A1, Canada}
+\affil[3]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
+\affil[4]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
+\affil[5]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
+\affil[6]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
+\affil[7]{Centre for Northern Studies, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
+\affil[8]{Université du Québec en Outaouais, Gatineau, Quebec, Canada}
+\affil[9]{Institut des Sciences de la Forêt Tempérée (ISFORT), Ripon, Quebec, Canada}
+\affil[10]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
+\affil[11]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
+\affil[12]{Department of Mathematics, Washington State University, Vancouver, Washington, USA}
+\affil[13]{Author for correspondance. Address: LECA, BP 53, 2233 Rue De La Piscine
+38041 Grenoble Cedex 9, France}
 
@@
 
-	\textbf{Abstract word count: } 299 %ABS_COUNT
+	\textbf{Abstract word count: } 295 %ABS_COUNT
 
-	\textbf{Main text word count: } 5039 %WORD_COUNT %Target 5000
+	\textbf{Main text word count: } 5010 %WORD_COUNT %Target 5000
 
@@
 \paragraph{Aim} 
-Current interest in forecasting changes to species ranges have resulted in a great diversity of approaches to species distribution models (SDMs).
+Current interest in forecasting changes to species ranges have resulted in a wide range of approaches to species distribution models (SDMs).
 However, most approaches include only a small subset of the available information, and many ignore smaller-scale processes such as growth, fecundity, and dispersal.
-Furthermore, different approaches often produce divergent predictions, with no simple method to reconcile them.
+Furthermore, different approaches often produce divergent predictions with no simple method to reconcile them.
 Here, we present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods.
@@
 \paragraph{Methods}
-\rev{Our framework builds a metamodel that is constrained by the results of multiple sub-models and results in probabilistic estimates of species presence.
+\rev{Our framework builds a metamodel that is constrained by the results of multiple sub-models and provides probabilistic estimates of species presence.
 We applied our approach to a simulated dataset to demonstrate the integration of a correlative SDM with a theoretical model.
@@
 \paragraph{Results}
-\rev{For both examples, the integrated models included information from all data sources and substantially improved uncertainty when making projections.
-For the second example, we found that the integrated model outperformed the source models with respect to uncertainty when modelling the present range of the species.
+\rev{For both examples, the integrated models successfully included information from all data sources and substantially improved the characterization of uncertainty.
+For the second example, the integrated model outperformed the source models with respect to uncertainty when modelling the present range of the species.
 When projecting into the future, the model provided a consensus view of two models that differed substantially in their predictions.
@@
 \paragraph{Main conclusions}
-We conclude by discussing the application of our method and its accessibility to applied ecologists.
-In ideal cases, our framework can be easily implemented using off the shelf software.
-The framework has wide potential for use in species distribution modelling, and can drive better integration of multi-source and multi-scale data into ecological decision-making.
+We conclude by discussing the potential applications of our method and its accessibility to applied ecologists.
+In ideal cases, our framework can be easily implemented using off-the-shelf software.
+The framework has wide potential for use in species distribution modelling and can drive better integration of multi-source and multi-scale data into ecological decision-making.
 
@@
 
-Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-support tools in biodiversity management \citep{Guisan2013}.
-Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the processes generating species range limits \citep{Wu1995, Levin1998}.
+Models of species range limits have wide applications, particularly in conservation biology where they can be used as decision-support tools in biodiversity management \citep{Guisan2013}.
+Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that adequately explore the processes generating species range limits \citep{Wu1995, Levin1998}.
 Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, hybrid zone dynamics \citep{Engler2013}, and climate change impacts on species distributions \citep{Blois2013, Thuiller2014}.
@@
 Despite the recognized potential of these models, it can be difficult to produce species distribution models (SDMs) with acceptable levels of precision and bias \citep{Guisan2013}.
-For mechanistic models, two important constraints can cause difficulties: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
+For mechanistic models, two important constraints can be problematic: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
 In recent decades, however, modeling techniques have proliferated to take advantage of increased data availability. 
-A growing body of theory, reflecting the diversity of processes generating species ranges, also has contributed to model diversification \citep{Boulangeat2012}.
+A growing body of theory, reflecting the diversity of processes generating species ranges, has also contributed to model diversification \citep{Boulangeat2012}.
 Of these model types, fine-scale mechanistic models often capture important ecological processes quite well, but may perform poorly when applied at the scale of species ranges.
@@
 However, their predictions rely on the stationarity of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}.
-Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
+Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinear responses to novel combinations of explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
 %% ----------------------------------
@@
 
-Clearly, an approach is needed to unify the strength of different modeling approaches that can incorporate multiple data sources and modeling paradigms.
-Toward this end, we present an application of hierarchical Bayesian methods that uses outputs from multiple models to inform the results of the final model.
+Clearly, an approach is needed to unify the strength of different modeling approaches that can also incorporate multiple data sources.
+To this end, we present an application of hierarchical Bayesian methods that uses outputs from multiple models to inform the results of the final model.
 Techniques for multimodel inference have proliferated in recent years.
@@
 Within the hybrid framework, a correlative model might be used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while a more mechanistic approach could include biotic interactions and space-time dynamics \citep{Smolik2010}.
-However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, uncertainties are approximated and their attribution to different sources is difficult to distinguish.
+However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and as such uncertainties are approximated and their attribution to different sources is difficult to distinguish.
 An alternative to hybrid models is the direct combination of predictions, allowing models operating at the same spatio-temporal scales to be combined \citep[e.g., model averaging, ensemble forecasting;][]{Araujo2007}.
@@
 
-We propose an alternative to these other approaches using a hierarchical Bayesian framework. 
+We propose an alternative to these approaches using a hierarchical Bayesian framework. 
 This approach provides a number of advantages, including (1) the ability to incorporate multiple modes of inference (e.g., mechanistic, correlative models) \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}, (2) an easy mechanism to include multiple data sources at various scales \citep{Levin1992, Peters2004}, and (3) an intuitive and comprehensive reporting of uncertainty in model predictions that reflects variation at all levels of organization \citep{Cressie2009, Hobbs2011}.
 Contrary to hybrid methods, the aim is not to link different sub-models into a single one, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for more flexibility regarding the type of information included.
-By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Guisan2005, Araujo2006}.
-Moreover, a more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying the parameters that contribute most to variance in the model predictions \citep{McMahon2011}.
+By integrating all available knowledge into a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Guisan2005, Araujo2006}.
+Moreover, a more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying parameters that contribute most to variance in the model predictions \citep{McMahon2011}.
 
@@
 In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread and dominant tree species from eastern North America, in response to climate change.
-We provide a more formal mathematical presentation in Appendix S1, along with complete code and data for both examples (Appendix S2).
+We provide a more formal mathematical presentation in Appendix S1, along with complete code and data for both examples in Appendix S2.
 
@@
 An initial version of the metamodel \((\theta_M)\) that has no constraints from other datasets will be referred to as the naive model.
-This naive model uses a simple logistic regression to estimate the naive probability of occurrence \((\psi_N)\) as a function of second order effects of temperature and precipitation:
+This naive model uses a simple logistic regression to estimate the probability of occurrence \((\psi_N)\) as a function of second order effects of temperature and precipitation:
 
@@
 Furthermore, no information is available regarding the temperature regime for the experiment.
-Transplant experiments that evaluate performance beyond the range of a species are common, and represent a plausible scenario for model integration \citep{Hargreaves2014}.
+Transplant experiments that evaluate performance beyond the range of a species are common and represent a plausible scenario for model integration \citep{Hargreaves2014}.
 According to niche theory \citep{Holt2009}, the fundamental niche corresponds to the set of environmental conditions where the per capita intrinsic growth rate $r$ is positive.
@@
 
-Although the two datasets were collected at considerably different scales, we have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation).
+Although the two datasets were collected at considerably different scales, we have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation regime).
 The scaling treats the fundamental niche as the only driver of species distribution, and only considers a single dimension of the niche.
@@
 \subsection*{Example 2: Constraining an SDM using phenological information}
-For the second example, we consider the problem of forecasting the potential distribution of a species following climate change.
+For the second example, we consider the problem of forecasting a species' potential distribution following climate change.
 There is considerable interest in comparing correlative and mechanistic projections with respect to climate change \citep{Morin2009}, and correctly characterizing uncertainty is a critical aspect of this problem \citep{Cheaib2012}.
@@
 To these data, we added 4903 recorded presences and 21701 absences derived from permanent forest sample plots located in the United States and Canada (See Appendix S1 for a map of plot locations).
-We reserved 1/3 of this dataset for evaluation, and used the remaining records to calibrate the models.
+We reserved 1/3 of this dataset for evaluation and used the remaining records to calibrate the models.
 We constructed the naive model by using a binomial generalized linear model (GLM) to relate the presence-absence dataset to three climate variables: the number of degree days (ddeg), mean annual precipitation (an\_prcp), and the ratio of annual precipitation to potential evapotranspiration (pToPET).
@@
 
-To perform the integration, we constrained the estimates of the naive model with the additional information from Phenofit while considering two different potential modeling goals.
+To perform the integration, we constrained the estimates of the naive model with the additional information from Phenofit while considering two different modeling objectives.
 The first is improving our model of the present range of the species.
@@
 This can provide a much more accurate estimate of the uncertainty of our predictions.
-Thus, for our first integration we combined the naive model with the Phenofit predictions for the present; we refer to this model as ``Integrated-Present'' hereafter.
-\rev{We also evaluated this model by constructing calibration curves and by computing the area under the receiver operating curve (AUC; seeAppendix S1), a metric that evaluates classification ability where 0 indicates perfect classification and 0.5 indicates no difference from a random model \citep{Swets1988}.}
-The second modeling goal is projecting changes to the range of sugar maple following climate change.
-Process-based models and correlative models often differ substantially when projecting beyond the range of the original data (see Example 1).
+Thus, for our first integration we combined the naive model with the Phenofit predictions for the present; we refer to this model as ``Integrated-Present''.
+\rev{We also evaluated this model by constructing calibration curves and by computing the area under the receiver operating curve (AUC; seeAppendix S1), which evaluates classification ability where 1 indicates perfect classification and 0.5 indicates no difference from a random model \citep{Swets1988}.}
+The second modeling goal is to project changes to the range of sugar maple following climate change.
+Process-based and correlative models often differ substantially when projecting beyond the range of the original data (see Example 1).
 Thus, we obtained predictions for 2100 from Phenofit \citep{Morin2009} and used them to condition the metamodel given the results of the naive model under future climate.
@@
 %% compare parameters
-Model integration resulted in substantial reduction in posterior uncertainty in the parameters, and, for the Integrated-Future model, a large revision in the the estimate for the response of sugar maple to temperature (i.e., variable ddeg) (Table \ref{tab:ex2_params}, Fig.
+Model integration resulted in substantial reduction in posterior uncertainty in the parameters, and, for the Integrated-Future model, a large revision in the estimate of the response of sugar maple to temperature (i.e., variable ddeg) (Table \ref{tab:ex2_params}, Fig.
 \ref{fig:ex2_response}).
@@
 Furthermore, both models performed well when evaluated against reserved data, with \rev{median} AUC values of 0.802 and 0.797 for the naive and integrated models, respectively (see Appendix S1).
-The small difference between the models indicates that both models effectively predict the probability of presence of sugar maple.
+The small difference between the models indicates that both models adequately predict the probability of presence of sugar maple.
 The major advantage to integration is the improved understanding of uncertainty in the predictions, with greater uncertainty in southern portions of the range (Fig. \ref{fig:ex2_map_pres}).
-It is important to note that the naive model was the basis for the integrated model; thus the increased uncertainty present in the integrated model is not the result of a ``worse'' model, but rather should be viewed as a correction to overfitting in the naive model that incorporates uncertainty arising from the processes that we have included into the metamodel.
+It is important to note that the naive model was the basis for the integrated model; thus the increased uncertainty present in the integrated model is not the result of a ``worse'' model, but rather should be viewed as a correction to overfitting in the naive model that incorporates uncertainty arising from the processes included in the metamodel.
 Phenofit predicts fitness based on how climate affects phenological timings, frost injury, reproduction, and survival \citep{Chuine2001, Morin2009}.
-Thus, climatic factors that ultimately limit species distribution might be quite different under Phenofit, as illustrated by the differences in the potential future distributions predicted by Phenofit and the naive model.
+Thus, climatic factors that ultimately limit species distribution might be quite different between the two source models, as illustrated by the differences in the potential future distributions predicted by Phenofit and the naive model.
 
@@
 \subsection*{Comparison with other methods}
-The methods provided here expand upon the motivation of hybrid models to develop a more robust approach for using ecological models for prediction while overcoming some limitations characteristic of other integrated approaches.
+The methods provided here expand upon the motivation of hybrid models to develop more robust approaches for using ecological models for prediction while overcoming some limitations characteristic of other integrated approaches.
 In particular, it is often difficult in a hybrid model to identify parameters that can be used to connect different modeling frameworks and produce a meaningful response \citep{Thuiller2013}. 
 The difficulty of including information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Smolik2010, Thuiller2014a}. 
-\rev{Bayesian methods provide a natural framework for the incorporation of multiple sources of information, making them an attractive alternative for SDMs.
+\rev{Bayesian methods provide a natural framework for the incorporation of multiple sources of information, making them an attractive alternative to SDMs.
 Hierarchical models in particular have the potential to capture many of the intricacies necessary for implementing hybrid models \citep{Latimer2006}.
 \citet{Pagel2012} developed a hybrid approach to species distributions via a dynamic range model.
-Similar to our approach, their model integrated demographic information, abundances, and presence/absence data within a hierarchical Bayesian framework to predict species ranges.
+Similar to our approach, their model integrated demographic information, abundance, and presence/absence data within a hierarchical Bayesian framework to predict species ranges.
 However, their approach explicitly links the modelled processes to occurrences/abundances via a detailed demographic model, requiring data that may not be available for many species.
-In contrast, our approach allows for the inclusion of more incomplete datasets because integration is performed via the separate predictions of each model (following the application of the scaling function where necessary).}
+In contrast, our approach allows for the inclusion of less complete datasets because integration is performed via the separate predictions of each model (following the application of the scaling function where necessary).}
 Because the metamodel is expressed as a series of conditional probabilities (see Appendix S1), this information can be included simply as long as the probability of the metamodel can be expressed mathematically.
-Furthermore, Bayesian methods produce posterior distributions of parameters and predictions, rather than point estimates, allowing for a comprehensive understanding of the uncertainty. 
-Finally, Bayesian methods inherently allow for feedbacks or interactions between sub models, which may be a more realistic representation of ecological dynamics where many factors may simultaneously influence the system.
+Furthermore, Bayesian methods produce posterior distributions of parameters and predictions rather than point estimates, allowing for a comprehensive understanding of the uncertainty. 
+Finally, Bayesian methods inherently allow for feedbacks or interactions between sub-models, which may be a more realistic representation of ecological dynamics where many factors may simultaneously influence the system.
 
@@
 In particular, it has certain similarities with Bayesian model averaging and Bayesian calibration of process-based models. 
-Bayesian model averaging aims to combine several alternative models that operate at the same scale to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}, and has been applied numerous times in ecology where the mechanisms underlying a complex phenomenon are often unknown \citep[e.g.,][]{Wintle2003, Link2006}.
+Bayesian model averaging aims to combine several alternative models that operate at the same scale to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}.
+This is of particular interest in ecology, where the mechanisms underlying complex phenomena are often unknown \citep[e.g.,][]{Wintle2003, Link2006}.
 Bayesian calibration of process-based models focuses on uncertainty of the parameter values, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, Hartig2012}.
-In contrast with these methods, our approach handles data and models operating at different hierarchical scales, and uses process-based models to constrain the shape of the metamodel.
+In contrast with these methods, our approach handles data and models operating at different hierarchical scales and uses process-based models to constrain the shape of the metamodel.
 
@@
 Species distribution models are important tools that are increasingly used by land managers for science-based decision-making \citep{Guisan2013}.
-However, the possibility that diverse approaches will provide diverging answers as a result of different assumptions and methodologies can create confusion and mistrust towards models, and some managers may be discouraged from incorporating their results in their management plans.
+However, the possibility that diverse approaches will provide contrasting answers as a result of different assumptions and methodologies can create confusion and mistrust towards models, and some managers may be discouraged from incorporating their results in management plans.
 Integrated approaches have gained momentum in recent years, with integrative science being featured as a central theme for several science-based governmental organizations around the world \citep[e.g.,][]{Bernier2013}.
@@
 \ref{fig:management}]{Rehme2011}.
-Such approaches are needed in formulating management plans for vulnerable species and ecosystems to avoid basing decisions on too-narrow subsets of the available information \citep{Dawson2011}.
-However, successful use of approaches such as ours will always remain dependent on an intimate understanding of the decision-making process by modelers, emphasizing the importance of close collaboration between them and practitioners at all stages of model development \citep{Guisan2013}.
+Such approaches are needed in designing management plans for vulnerable species and ecosystems to avoid basing decisions on too-narrow subsets of the available information \citep{Dawson2011}.
+However, successful use of approaches such as ours will always remain dependent on an intimate understanding of the decision-making process, emphasizing the importance of close collaboration between modelers and practitioners at all stages of model development \citep{Guisan2013}.
 
@@
 Furthermore, the use of linked sub-models allows for clear specification of desired model outputs (via the specification of the metamodel) while easily retaining important ecological objectives (via careful specification of sub-models).
-Transparency in the model building process must be accompanied by a clearly documented modeling workflow.
-We suggest using the sub-models as a natural proxy for specifying specific objectives, and using this as the basis for developing well documented workflows describing the process of model integration to ensure reproducibility and applicability (Fig.
+Transparency in the model-building process must be accompanied by a clearly documented  workflow.
+We suggest using the sub-models as a natural proxy for specifying specific objectives, and using this as the basis for developing workflows describing the process of model integration to ensure reproducibility and applicability (Fig.
 \ref{fig:management}).
@@
 For example, climatic gradients may mediate competitive interactions \citep{Kunstler2011}, which means that simple correlative models that fail to account for competition may be wrong if the climate-competition association changes in the future.
-In North America, the US Forest Service maintains a long term Forest Inventory Analysis database that could be utilized to parameterize a competition model.
+In North America, the US Forest Service maintains a long-term Forest Inventory Analysis database that could be utilized to parameterize a competition model.
 Such a model need not explicitly predict occurrence limits; rather it could be integrated with a larger-scale model to include information about competition in a distribution model.
@@
 Given that the metamodel is correlative, it can be effectively implemented with, e.g., only presence-absence data, or, in the case where true absences are difficult to obtain, with presences and pseudo absences (provided sufficient care is used in interpreting the results of such a model).
-Then any additional mechanistic data that are available will enhance predictions by constraining outputs of the metamodel.
+Consequently, any additional mechanistic data that become available will enhance predictions by constraining outputs of the metamodel.
 
 Determining the functions to use to express the likelihood of the sub-models given the metamodel (i.e., Eq. \ref{eq:ex1_integrated}) is a critical point.
-The challenge is three fold: (1) which spatial and temporal scales (i.e.
-which processes), are to be considered, (2) how to build a scale the sub-models to be consistent with the metamodels, and (3) how error and uncertainty propagate from the sub-models to the metamodel.
-Although we argue that our proposed framework is able to easily deal with very different scales and that the Bayesian framework allows for an efficient integration of uncertainty throughout all scales considered, the building of scaling functions is an object of investigation on its own.
+The challenge is three-fold: (1) determining which spatial and temporal scales (i.e.
+which processes), are to be considered, (2) selecting how to build a scale the sub-models to be consistent with the metamodels, and (3) understanding how error and uncertainty propagate from the sub-models to the metamodel.
+Although we argue that our proposed framework is able to easily deal with different scales and that the Bayesian framework allows for an efficient integration of uncertainty across all scales considered, the building of scaling functions is an object of investigation on its own.
 It is likely that the modeling process may include multiple functions operating at different scales when taking all known processes and models into account. 
@@
 In many cases, off-the-shelf software can adequately express the model likelihoods with minimal programming, but more complicated models will require the development of custom programs. 
-Developing such customized code would requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming. 
-However, the flexibility of our approach, and its transparency with respect to the propagation of uncertainty, will often outweigh the implementation challenges.
+Developing such customized code requires careful model specification, understanding of applied Bayesian methods, and, in some cases, extensive programming. 
+However, the flexibility of our approach and its transparency with respect to the propagation of uncertainty will often outweigh the implementation challenges.
 
-Finally, this approach is not just a new methodological tool but is a framework for forecasting species distributions which is fundamentally designed to make the link between modelers and practitioners while correctly estimating uncertainties and being updatable with new data and theoretical advances.
+Finally, this approach is not just a new methodological tool but a framework for forecasting species distributions which is fundamentally designed to make the link between modelers and practitioners while correctly estimating uncertainties and being updatable with new data and theoretical advances.
 A scientific approach such as presented here is particularly adapted to make a synthesis of available information and provide robust species distribution forecasts based on information known to be the best available scientific knowledge.
@@
 In addition, it provides practitioners and decision-makers with the best possible estimation of uncertainties, with direct applications to risk assessment or to guide the choice where investigating new research and accumulating new data.
-Finally, we argue that the adaptability of our approach is particularly appropriate in a world where collected data and theoretical knowledge is changing as quickly as climate and conservation practices must be adjusted accordingly.
+Finally, we argue that the adaptability of our approach is particularly appropriate in a world where collected data and theoretical knowledge are changing as quickly as climate and conservation practices must be adjusted accordingly.
 

### 2015-09-05 tex/manuscript.tex

@@
 
-	\textbf{Abstract word count: } 295 %ABS_COUNT
+	\textbf{Abstract word count: } 294 %ABS_COUNT
 
@@
 \paragraph{Aim} 
-Current interest in forecasting changes to species ranges have resulted in a wide range of approaches to species distribution models (SDMs).
+Current interest in forecasting changes to species ranges have resulted in a multitude of approaches to species distribution models (SDMs).
 However, most approaches include only a small subset of the available information, and many ignore smaller-scale processes such as growth, fecundity, and dispersal.
@@
 We applied our approach to a simulated dataset to demonstrate the integration of a correlative SDM with a theoretical model.
-In a second example, we built an integrated model combining the results of a physiological model with presence-absence data for sugar maple \emph{Acer saccharum}, an abundant tree native to eastern North America.}
+In a second example, we built an integrated model combining the results of a physiological model with presence-absence data for sugar maple (\emph{Acer saccharum}), an abundant tree native to eastern North America.}
 

### 2015-09-14 tex/manuscript.tex
Final revision; submitted ms 14 Sept 2015

@@
 \usepackage[dvipsnames]{xcolor}
-\newcommand{\rev}[1]{{\color{RoyalBlue}#1}}
 
@@
 
-	\textbf{Main text word count: } 5010 %WORD_COUNT %Target 5000
+	\textbf{Main text word count: } 5001 %WORD_COUNT %Target 5000
+
+	\textbf{Reference count: } 50
 
@@
 \paragraph{Methods}
-\rev{Our framework builds a metamodel that is constrained by the results of multiple sub-models and provides probabilistic estimates of species presence.
+Our framework builds a metamodel that is constrained by the results of multiple sub-models and provides probabilistic estimates of species presence.
 We applied our approach to a simulated dataset to demonstrate the integration of a correlative SDM with a theoretical model.
-In a second example, we built an integrated model combining the results of a physiological model with presence-absence data for sugar maple (\emph{Acer saccharum}), an abundant tree native to eastern North America.}
+In a second example, we built an integrated model combining the results of a physiological model with presence-absence data for sugar maple (\emph{Acer saccharum}), an abundant tree native to eastern North America.
 
 \paragraph{Results}
-\rev{For both examples, the integrated models successfully included information from all data sources and substantially improved the characterization of uncertainty.
+For both examples, the integrated models successfully included information from all data sources and substantially improved the characterization of uncertainty.
 For the second example, the integrated model outperformed the source models with respect to uncertainty when modelling the present range of the species.
 When projecting into the future, the model provided a consensus view of two models that differed substantially in their predictions.
-Uncertainty was reduced where the models agreed and was greater where they diverged, providing a more realistic view of the state of knowledge than either source model.}
+Uncertainty was reduced where the models agreed and was greater where they diverged, providing a more realistic view of the state of knowledge than either source model.
 
@@
 	\caption{Two simulated datasets used to illustrate the model integration framework.
-	(a) Presences (circles) or absences (x's) of the species in ecological space, \rev{where the range of precipitation values sampled was 0.1--1}.
+	(a) Presences (circles) or absences (x's) of the species in ecological space, where the range of precipitation values sampled was 0.1--1.
 	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., -1.0--1.0).
@@
 We used our framework to constrain a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
-Here we describe briefly the dataset, methods, and the results of the analysis (see \rev{ Appendix S1 for implementation details and Appendix S2} for data and scripts to reproduce the analysis).
+Here we describe briefly the dataset, methods, and the results of the analysis (see Appendix S1 for implementation details and Appendix S2 for data and scripts to reproduce the analysis).
 
-We obtained climate variables, occurrence data, and Phenofit projections at 0.5-degree resolution for both the present and for \rev{2100} for sugar maple, an economically and ecologically important species occurring in eastern North America \citep{Morin2009}.
+We obtained climate variables, occurrence data, and Phenofit projections at 0.5-degree resolution for both the present and for 2100 for sugar maple, an economically and ecologically important species occurring in eastern North America \citep{Morin2009}.
 These data defined the metamodel scale.
@@
 We constructed the naive model by using a binomial generalized linear model (GLM) to relate the presence-absence dataset to three climate variables: the number of degree days (ddeg), mean annual precipitation (an\_prcp), and the ratio of annual precipitation to potential evapotranspiration (pToPET).
-These variables were selected from an initial set of \rev{seven} variables \citep[see Appendix S1 and ][for details on the climate variables]{Morin2009}.
+These variables were selected from an initial set of seven variables \citep[see Appendix S1 and ][for details on the climate variables]{Morin2009}.
 We selected a GLM for its simplicity and interpretability because our focus was on demonstrating the framework, but more complex methods (e.g., generalized additive models) are compatible with the framework.
@@
 Thus, for our first integration we combined the naive model with the Phenofit predictions for the present; we refer to this model as ``Integrated-Present''.
-\rev{We also evaluated this model by constructing calibration curves and by computing the area under the receiver operating curve (AUC; seeAppendix S1), which evaluates classification ability where 1 indicates perfect classification and 0.5 indicates no difference from a random model \citep{Swets1988}.}
+We also evaluated this model by constructing calibration curves and by computing the area under the receiver operating curve (AUC; see Appendix S1), which evaluates classification ability where 1 indicates perfect classification and 0.5 indicates no difference from a random model \citep{Swets1988}.
 The second modeling goal is to project changes to the range of sugar maple following climate change.
@@
 This procedure provided a consensus view of the future range of the species; we refer to this model as ``Integrated-Future.''
-\rev{In both cases, it was necessary to scale the Phenofit predictions, which were probabilistic, to make them compatible with the naive model, which was fit using occurrence data.
+In both cases, it was necessary to scale the Phenofit predictions, which were probabilistic, to make them compatible with the naive model, which was fit using occurrence data.
 We used a latent variable approach, which posits an unobserved ``true'' presence-absence dataset from which the Phenofit probabilities are derived. 
@@
 This procedure expresses the information from Phenofit in a way that is compatible with the naive model, and also propagates the uncertainty in the Phenofit predictions.
-Additionally, it becomes possible to address future distribution via simulated future occurrences (which, by nature, are unobservable).}
+Additionally, it becomes possible to address future distribution via simulated future occurrences (which, by nature, are unobservable).
 
@@
 	\footnotesize
-	\caption{\rev{Parameter estimates and 95\% credible intervals for the naive model (i.e., a species distribution model fit with occurrence data only), and two integrated models combining predictions from the naive model and either the present or future predictions from a mechanistic model.
+	\caption{Parameter estimates and 95\% credible intervals for the naive model (i.e., a Binomial GLM relating sugar maple presence-absence data to climatic variables), and two integrated models combining the naive model and either the present or future predictions from the mechanistic model Phenofit.
 All models were fit on predictor variables standardized to mean=0 and unit variance, and all estimates are on the logit scale.
-Climate variables included the number of degree days (ddeg), mean annual precipitation (an\_prcp), and the ratio of annual precipitation to potential evapotranspiration (pToPET).}}
+Climate variables included the number of degree days (ddeg), mean annual precipitation (an\_prcp), and the ratio of annual precipitation to potential evapotranspiration (pToPET).
+Area under the receiver operating curve (AUC) values measure model classification ability, with values of 0.5 indicating no improvement over random classification and 1 indicating perfect classification.}
 	\begin{tabular}{lcccccccc}
@@
 	& (-0.74,-0.37) & (-0.15,-0.03) & (-0.070,-0.073) \\
+	\midrule
+	\multirow{2}{*}{AUC} & 0.802 & 0.797 & \\
+	& (0.78,0.82) & (0.78,0.81) & --\textsuperscript{*}\\
 	\bottomrule
+	\multicolumn{4}{l}{\footnotesize \textsuperscript{*}AUC is unavailable for the Integrated-future model because} \\ 
+	\multicolumn{4}{l}{\footnotesize independent validation data are not available for future predictions.}
 	\end{tabular}
@@
 	Integration reduced prediction uncertainty for all three variables, particularly for domains outside the naive calibration range.
-	Single-variable predictions were computed with the other variables set to their medians.}
+	Single-variable predictions were computed with the other variables set to their medians.
+	Uncertainty is represented by colored shaded regions showing 95\% Bayesian credible intervals.
+	The gray shaded region shows the calibration range for the naive model.
+}
 	\label{fig:ex2_response}
@@
 \includegraphics[width=6.5in]{figs/ex2_pres_map.png}
-\caption{Range predictions for present climate for Phenofit (A), the naive model (B), the Integrated-Present model (C), and the difference between posterior prediction standard errors of the Integrated-Present and Naive models.
+\caption{Range predictions for present climate for Phenofit (A), the naive model (B), the Integrated-Present model (C), and the difference between posterior prediction standard errors (SE) of the Integrated-Present and Naive models (D; Integrated SE -- Naive SE).
 Model predictions were quite similar, with reduced uncertainty where the models were in strongest agreement, and increased uncertainty near the range boundary where the models disagreed.
@@
 When considering the present species distribution, the naive and Integrated-Present models made very similar predictions (Fig. \ref{fig:ex2_map_pres}).
-Furthermore, both models performed well when evaluated against reserved data, with \rev{median} AUC values of 0.802 and 0.797 for the naive and integrated models, respectively (see Appendix S1).
+Furthermore, both models performed well when evaluated against reserved data, with median AUC values of 0.802 and 0.797 for the naive and integrated models, respectively (see Appendix S1).
 The small difference between the models indicates that both models adequately predict the probability of presence of sugar maple.
@@
 \includegraphics[width=6.5in]{figs/ex2_fut_map.png}
-\caption{Range predictions for future climate for Phenofit (A), the naive model (B), the Integrated-Future model (C), and the difference between posterior prediction standard errors of the Integrated-Future and Naive models.
+\caption{Range predictions for future climate for Phenofit (A), the naive model (B), the Integrated-Future model (C), and the difference between posterior prediction standard errors (SE) of the Integrated-Future and Naive models (D; Integrated SE -- Naive SE).
 The mechanistic sub-model Phenofit (A) predicted small shifts sugar maple range.
@@
 The Integrated-Future model presents a different interpretation of the response of sugar maple to warmer temperatures. 
-The naive model predicted a substantial northward migration; in other words, the expectation under the naive model \rev{is that the present estimate of the realized niche (obtained using occurrence data) is an unbiased reflection of the fundamental niche; thus the species should track temperature northward as the climate warms}.
+The naive model predicted a substantial northward migration; in other words, the expectation under the naive model is that the present estimate of the realized niche (obtained using occurrence data) is an unbiased reflection of the fundamental niche; thus the species should track temperature northward as the climate warms.
 The Integrated-Future model, in contrast, predicted substantially more tolerance to warmer temperatures, reflecting similar predictions from Phenofit (Figs.
 \ref{fig:ex2_response}, \ref{fig:ex2_map_fut}).
-\rev{This is because Phenofit estimates different aspects of the realized niche.}
+This is because Phenofit estimates different aspects of the realized niche.
 Although both models predicted a northward range shift, the change under Phenofit was limited to approximately 200 km north of the present range limit of the species, compared with more than 900 km for the naive model.
@@
 \ref{fig:ex2_map_fut}).
-\rev{The metamodel thus presents a consensus view of the niche of the species with respect to the macroclimatic variables included in the model (Fig.
-\ref{fig:ex2_map_pres}), incorporating the present range of the species (using the occurrence data) and information from Phenofit on what conditions will be tolerable in the future.}
+The metamodel thus presents a consensus view of the niche of the species with respect to the macroclimatic variables included in the model (Fig.
+\ref{fig:ex2_map_pres}), incorporating the present range of the species (using the occurrence data) and information from Phenofit on what conditions will be tolerable in the future.
 
@@
 The difficulty of including information from experimental studies or ecological processes at lower scales can be a possible drawback of hybrid models \citep{Smolik2010, Thuiller2014a}. 
-\rev{Bayesian methods provide a natural framework for the incorporation of multiple sources of information, making them an attractive alternative to SDMs.
+Bayesian methods provide a natural framework for the incorporation of multiple sources of information, making them an attractive alternative to SDMs.
 Hierarchical models in particular have the potential to capture many of the intricacies necessary for implementing hybrid models \citep{Latimer2006}.
@@
 However, their approach explicitly links the modelled processes to occurrences/abundances via a detailed demographic model, requiring data that may not be available for many species.
-In contrast, our approach allows for the inclusion of less complete datasets because integration is performed via the separate predictions of each model (following the application of the scaling function where necessary).}
+In contrast, our approach allows for the inclusion of less complete datasets because integration is performed via the separate predictions of each model (following the application of the scaling function where necessary).
 Because the metamodel is expressed as a series of conditional probabilities (see Appendix S1), this information can be included simply as long as the probability of the metamodel can be expressed mathematically.
@@
 Bayesian model averaging aims to combine several alternative models that operate at the same scale to obtain better predictions while taking into account parameter uncertainties \citep{Hoeting1999}.
-This is of particular interest in ecology, where the mechanisms underlying complex phenomena are often unknown \citep[e.g.,][]{Wintle2003, Link2006}.
+This is of particular interest in ecology, where the mechanisms underlying complex phenomena are often unknown \citep[e.g.,][]{Link2006}.
 Bayesian calibration of process-based models focuses on uncertainty of the parameter values, in this case the values of the parameters are calibrated by the model output \citep{VanOijen2005, Hartig2012}.
@@
 Our framework is still applicable whatever the chosen upscaling approach and is able to propagate uncertainties from sub-models to the metamodel.
-\rev{Indeed, if a sub-model provides poor information (due, for example, to cross-scale nonlinearities in the response to the environment), the resulting metamodel predictions may be worse than the pre-integration naive model.
+Indeed, if a sub-model provides poor information (due, for example, to cross-scale nonlinearities in the response to the environment), the resulting metamodel predictions may be worse than the pre-integration naive model.
 In general, we advise users of this framework to carefully consider the scaling in their models with respect to the biology of the organism studied, and to use prior model weights to downweight the scaled models when there is uncertainty about the applicability of a sub-model at the metamodel scale.
@@
 Finally, as always when modelling ecological systems, we urge humility in the interpretation of model results and suggest the use of model evaluation and validation tools whenever possible.
-We provide additional discussion on the implementation of model weights in Appendix S1.}
+We provide additional discussion on the implementation of model weights in Appendix S1.
 
@@
 All data, as well as all code required to repeat the analyses, have been uploaded as online supporting information in Appendix S2.
-
-%% ----------------------------------
-%
-%     Supporting information
-%
-%% ----------------------------------
-\section*{Supporting information}
-Appendix S1: Complete model description and model fitting details for the examples. \\
-Appendix S2: Scripts and data required to reproduce the analyses.
 
@@
 %
+%     Supporting information
+%
+%% ----------------------------------
+\section*{Supporting information}
+Appendix S1: Model description and methodology \\
+Appendix S2: Model code and data
+
+%% ----------------------------------
+%
 %     BIOSKETCH (GEB)
